### 2012-07-30 text/ptf_microlensing_paper.tex
add paper to repository

\documentclass[12pt,preprint]{aastex}
\newcounter{address}
\setcounter{address}{1}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{subfigure}
\usepackage{natbib}

\newcommand{\apwsim}{\raisebox{0.2ex}{\scriptsize$\sim$\normalsize}}

\begin{document}

\title{Title Placeholder}
\author{
  Adrian~M.~Price-Whelan\altaffilmark{\ref{col}}, Marcel~Ag\"ueros\altaffilmark{\ref{col}}, Amanda Fournier\altaffilmark{\ref{ucsb}}, Eran Ofek\altaffilmark{\ref{weiz}}, Rachel Street\altaffilmark{\ref{lcogt}}
}

\altaffiltext{\theaddress}{\stepcounter{address}\label{col} Department of Astronomy, Columbia University, 550 W 120th St., New York, NY 10027, USA}
\altaffiltext{\theaddress}{\stepcounter{address}\label{ucsb} Department of Physics, Broida Hall, University of California, Santa Barbara, CA 93106, USA}
\altaffiltext{\theaddress}{\stepcounter{address}\label{weiz} Benoziyo Center for Astrophysics, Weizmann Institute of Science, 76100 Rehovot, Israel}
\altaffiltext{\theaddress}{\stepcounter{address}\label{lcogt} Las Cumbres Observatory Global Telescope Network, Inc., 6740 Cortona Dr.\ Suite 102, Santa Barbara, CA 93117, USA}

\begin{abstract}
	Many current photometric, time-domain surveys are driven by specific goals, such as supernova searches, transiting exoplanet discoveries, or stellar variability studies, which set the cadence with which individual fields get re-imaged. In the case of the Palomar Transient Factory (PTF), several such sub-surveys are being conducted in parallel, leading to an extremely non-uniform sampling gradient over the survey footprint of nearly 20,000 deg$^2$: while the typical 7.26~deg$^2$ PTF field has been imaged 15 times, \apwsim1000~deg$^2$ of the survey has been observed more than 150 times. We use the existing PTF data to study the trade-off between a large survey footprint and irregular sampling when searching for microlensing events, and to examine the probability that such events can be recovered in these data. We conduct Monte Carlo simulations to evaluate our detection efficiency in a hypothetical survey field as a function of both the baseline and number of observations. We also apply variability statistics to systematically differentiate between periodic, transient, and flat light curves. Preliminary results suggest that both recovery and discovery of microlensing events are possible with a careful consideration of photometric systematics. This work can help inform predictions about the observability of microlensing signals in future wide-field time-domain surveys such as that of LSST.

\end{abstract}

\keywords{
  survey science
  ---
  gravitational microlensing
  ---
  time-domain astrophysics
}

\section{Introduction}

\section{PTF Observations and Data}

\section{Simulated microlensing event recovery}

\section{Searching for events in full PTF data set}

\section{Comparing results to model predictions (?)}

\section{Conclusions and Future Work}

%\bibliographystyle{apj}
%\bibliography{apj-jour,refs}
%
%\begin{figure}
%	\centering
%	\caption{Computed detection efficiency as a function of number of observations for a fixed baseline of 365 days. This plot uses entirely simulated light curves.}
%    \includegraphics[width=1.0\textwidth]{figures/detection_efficiency_figure.pdf}
%    \label{fig:detection_eff}
%\end{figure}	
%
%\begin{figure}
%	\centering
%	\caption{A visualization of the different sampling regimes of six different PTF fields over the same one year baseline. Darker points mean more data (more exposures). Some fields have roughly uniform coverage, where others are very dense during some weeks and not others. }
%    \includegraphics[width=1.0\textwidth]{figures/ptf_sampling_figure.pdf}
%    \label{fig:light_curves}
%\end{figure}
%
%\begin{figure}
%	\centering
%	\caption{Example of how the variability indices may be used to separate microlensing events from periodic or flat light curves.}
%    \includegraphics[width=1.0\textwidth]{figures/J_vs_K_Con_figure.png}
%    \label{fig:var_idx1}
%\end{figure}
%
%\begin{figure}[htbp]
%	\centering
%	\caption{Two neighboring light curves. The time scale is the same on both plots. Note the strong similarities in the overall shapes of the data.} 
%	 \includegraphics[width=1.0\textwidth]{figures/bad_data_figure.pdf}
%	\label{fig:lc_correlated}
%\end{figure}

\end{document}

### 2012-07-31 detectionefficiency.py
detection efficiency code ready to run on fields

@@ -9,19 +9,21 @@ import logging
 import multiprocessing
 import os
 import sys
-import time
+import cPickle as pickle
 
 # Third-party
-import apwlib.geometry as g
-import apwlib.convert as c
 from apwlib.globals import greenText
 import numpy as np
 from scipy.stats import scoreatpercentile
+import matplotlib
+matplotlib.use("Agg")
+import matplotlib.pyplot as plt
 
 # Project
 from ptf.simulation.simulatedlightcurve import SimulatedLightCurve
 import ptf.photometricdatabase as pdb
 import ptf.analyze.analyze as analyze
+from ptf.globals import index_to_label
 
 # Create logger
 logger = logging.getLogger(__name__)
@@ -30,7 +32,67 @@ formatter = logging.Formatter("%(name)s / %(levelname)s / %(message)s")
 ch.setFormatter(formatter)
 logger.addHandler(ch)
 
-def variability_indices_detection_efficiency(light_curves, events_per_light_curve=100, indices=["j","k","eta","sigma_mu","delta_chi_squared"]):
+def detection_efficiency(statistic, indices_with_events, lower_cut=None, upper_cut=None):
+    """ Compute the detection efficiency of a specific variability statistic.
+        
+        This function computes the detection efficiency by accepting a column
+        of variability statistic values, and upper and lower bounds, and computes
+        the fraction of events detected. 
+    """
+    
+    and_it = False
+    if lower_cut != None:
+        lower_idx = indices_with_events[statistic] < lower_cut
+    else:
+        lower_idx = np.array([True]*len(indices_with_events))
+        and_it = True
+    
+    if upper_cut != None:
+        upper_idx = indices_with_events[statistic] > upper_cut
+    else:
+        upper_idx = np.array([True]*len(indices_with_events))
+        and_it = True
+    
+    if and_it:
+        idx = lower_idx & upper_idx
+    else:
+        idx = lower_idx | upper_idx
+        
+    selected = indices_with_events[idx]
+    
+    efficiency = sum(selected["event_added"] == True) / sum(indices_with_events["event_added"])
+    N_false_positives = sum(selected["event_added"] == False)
+    
+    return (efficiency, N_false_positives)
+
+def simulate_events_compute_indices(light_curves, events_per_light_curve=100, indices=["j","k","eta","sigma_mu","delta_chi_squared"]):
+    """ """
+    var_indices_with_events = []
+    
+    # For each light curve, add 100 different microlensing events and recompute indices
[…]

### 2012-07-31 ptf/analyze/analyze.py

@@ -375,7 +375,7 @@ def compute_variability_indices(light_curve, indices=[], return_tuple=False):
         idx_dict["continuum"] = continuum_mag
     
     if "sigma_mu" in indices:
-        idx_dict["sigma_mu"] = np.std(light_curve.mag) / continuum_mag
+        idx_dict["sigma_mu"] = np.std(light_curve.mag) / continuum_mag[0]
     
     if return_tuple:
         return tuple([idx_dict[idx] for idx in indices])

### 2012-07-31 ptf/photometricdatabase.py

@@ -12,7 +12,6 @@ import logging
 import numpy as np
 import apwlib.geometry as g
 from apwlib.globals import redText, greenText, yellowText
-import pytest
 
 try:
     import tables
@@ -82,6 +81,8 @@ class Field(object):
         # Validate filter_id
         if isinstance(filter, Filter):
             self.filter = filter
+        elif isinstance(filter, str):
+            self.filter = Filter(filter)
         else:
             raise ValueError(redText("filter") + " parameter must be Filter object")
         
@@ -313,7 +314,7 @@ if __name__ == "__main__":
     #time_compute_var_indices()
     #select_most_observed_fields()
     
-    logging.basicConfig(level=logging.INFO)
+    #logging.basicConfig(level=logging.INFO)
     #test_filter()
     #test_field()
     test_select_most_observed_fields()

### 2012-08-05 text/ptf_microlensing_paper.tex
worked on paper

@@
 \section{Introduction}
+Since Einstein first studied gravitational microlensing (cite), the phenomena has been used to study dark objects, probe galactic structure and kinematics, and discover extrasolar planets.
+Microlensing events were once relatively rare, but with advances in CCD technology and the help of several dedicated microlensing surveys (), $>500$ lensing events are now detected per observing season.
+
+[MORE MICROLENSING INTRODUCTION]
+
+[QUICK INTRO TO THEORY AND JARGON]
+
+[MOTIVATION FOR PROJECT -- Han, Di Stefano, Di Stefano et al. -- mesolensing and near-field microlensing]
 
 \section{PTF Observations and Data}
+[TODO: Marcel]
 
-\section{Simulated microlensing event recovery}
+\section{Microlensing event recovery}
+Microlensing surveys typically use difference image analysis (cite Lupton?) to initially identify transient events and then analyze the candidate light curves to search for microlensing events (cite MACHO, OGLE).
+This method is essential for studying crowded fields such as the Galactic bulge, and also limits the number of light curves that must be searched for microlensing events, thus saving on computation time.
+This [method] also provides real-time detection of events so surveys can utilize networks of telescopes to provide follow-up observations of these transients.
+Our approach is to instead take the preexisting PTF light curve database over the entire survey footprint and develop fast, robust search algorithms to pick out microlensing event candidates from this massive data archive.
+
+In order to develop predictions for event rates and then eventually use observed event rates to infer [] we must understand our microlensing event recovery rate, or \textit{detection efficiency}, $\varepsilon$.
+The detection efficiency is commonly expressed as a function of microlensing event timescale and is therefore a function of survey parameters, such as cadence and baseline [FIGURE simulated detection efficiency].
+[What did MACHO do exactly?
+What does OGLE do?] The nature of the PTF survey means that different fields are observed with very different sampling patterns [FIGURE showing different sampling].
+We must compute the detection efficiency for each field individually.
+
+We compute the detection efficiency with a Monte Carlo simulation over real data.
+For a particular field of interest, we randomly sample 11,000 light curves from the database with $>25$ ``good'' observations.
+[define ``good''].
+We compute a set of variability statistics for each light curve, we then iterate and add 100 different simulated microlensing events to each light curve and recompute the variability statistics after each iteration [Figures here?].
+The parameters for the random microlensing events are chosen as follows:
+\begin{align}
+	t_0&\sim\mathcal{U}[t_{min}, t_{max}] \\
+	u_0&\sim\mathcal{U}[0, 1] \\
+	\log_{10}t_E&\sim\mathcal{U}[0, 3] 
+\end{align}
+where $\mathcal{U}$ is a uniform distribution.
+
+[How to compute detection efficiency?] [Specifics of our method, e.g. details about my code]
+
+[Parts of the above text will get absorbed into sections below]
+
+\subsection{Selecting candidate microlensing events}
+- Proper statistical modeling is too time consuming, so we first have to define some selection criteria to limit the sample (number of good observations, quality cuts, etc.)
+
+- Discuss $\Delta \chi^2$ method, downfalls (speed, not robust -- how do you define the threshold?)
+
+- Shin et al. paper
+
+- Introduce variability statistics: define them, what they pick out, what they have been used for before
+
+\subsection{Detection efficiency}
+- Previous surveys had fairly uniform time sampling, and smaller areas
+
+- With PTF, fields have vastly different sampling patterns, baselines, and cadences
+
+- We have to consider the efficiency on a field-by-field basis
+
+- Show detection efficiency of each variability statistic
 
 \section{Searching for events in full PTF data set}
+- Describe pipeline for events ... ?
 
@@
 \section{Conclusions and Future Work}
+\subsection{Globular Clusters and Other Overdensities}
+- Do they contribute to the lensing rate?
 

### 2012-08-07 ptf/analyze/analyze.py
fallback python code for C extension

@@ -18,29 +18,31 @@ try:
 except ImportError:
     raise ImportError("apwlib not found! \nDo: 'git clone git@github.com:adrn/apwlib.git' and run 'python setup.py install' to install.")
 
+
+
 try:
     import error_functions
-except ImportError:
-    print redText("**Error**: ") + "C extension error_functions.so not found! Make sure to do 'python setup.py build_ext' before running."
-    raise
+    constant_error_func = error_functions.constant_error_func
+    linear_error_func = error_functions.linear_error_func
+    gaussian_error_func = error_functions.gaussian_error_func
+except ImportError, RuntimeError:
+    print redText("**Error**: ") + "C extension error_functions.so not found or unable to import it! Make sure to do 'python setup.py build_ext' before running."
+    
+    constant_model = lambda p, x: p[0] + np.zeros(len(x))
+    constant_error_func = lambda p, x, mag, sigma: (mag - constant_model(p, x)) / sigma
+    
+    linear_model = lambda p, x: p[0]*x + p[1]
+    linear_error_func = lambda p, x, mag, sigma: (mag - linear_model(p, x)) / sigma
+    
+    gaussian_model = lambda p, x: p[0]*np.exp(-(x - p[1])**2 / (2*p[2]**2)) + p[3]
+    gaussian_error_func = lambda p, x, mag, sigma: (mag - gaussian_model(p, x)) / sigma
+    
 
 # ------
 # Models
 # ------
 def error_function(p, x, y, sigma_y, model):
     return (y - model(p,x)) / sigma_y
-    
-constant_model = lambda p, x: p[0] + np.zeros(len(x))
-constant_error_func = error_functions.constant_error_func
-#constant_error_func = lambda p, x, mag, sigma: (mag - constant_model(p, x)) / sigma
-
-linear_model = lambda p, x: p[0]*x + p[1]
-linear_error_func = error_functions.linear_error_func
-#linear_error_func = lambda p, x, mag, sigma: (mag - linear_model(p, x)) / sigma
-
-gaussian_model = lambda p, x: p[0]*np.exp(-(x - p[1])**2 / (2*p[2]**2)) + p[3]
-gaussian_error_func = error_functions.gaussian_error_func
-#gaussian_error_func = lambda p, x, mag, sigma: (mag - gaussian_model(p, x)) / sigma
 
 def u_t(p, t):
     return np.sqrt(p[0]**2 + ((t - p[1])/p[2])**2)

### 2012-08-07 text/ptf_microlensing_paper.tex

@@
+%\documentclass{emulateapj}
 \documentclass[12pt,preprint]{aastex}
@@
 \setcounter{address}{1}
-
+\usepackage{lscape, longtable}
+\usepackage{lmodern}
+\usepackage[T1]{fontenc}
 \usepackage{graphicx}
@@
 \usepackage{natbib}
+\newcommand{\Msun}{\ifmmode {M_{\odot}}\else${M_{\odot}}$\fi}
+\newcommand{\Rsun}{\ifmmode {R_{\odot}}\else${R_{\odot}}$\fi}
+\newcommand{\lapprox }{{\lower0.8ex\hbox{$\buildrel <\over\sim$}}}
+\newcommand{\gapprox }{{\lower0.8ex\hbox{$\buildrel >\over\sim$}}}
+\def\amin{\ifmmode^{\prime}\else$^{\prime}$\fi}
+\def\asec{\ifmmode^{\prime\prime}\else$^{\prime\prime}$\fi}
+\newcommand{\apwsim}{\raisebox{0.2ex}{\scriptsize$\sim$\normalsize}}
 
-\newcommand{\apwsim}{\raisebox{0.2ex}{\scriptsize$\sim$\normalsize}}
+\slugcomment{DRAFT \today}
+\shorttitle{Microlensing \& PTF}
+\shortauthors{Price-Whelan et al.}
+
+\bibliographystyle{apj}
 
@@
 
-\title{Title Placeholder}
-\author{
-  Adrian~M.~Price-Whelan\altaffilmark{\ref{col}}, Marcel~Ag\"ueros\altaffilmark{\ref{col}}, Amanda Fournier\altaffilmark{\ref{ucsb}}, Eran Ofek\altaffilmark{\ref{weiz}}, Rachel Street\altaffilmark{\ref{lcogt}}
-}
+\title{Identifying Microlenses in Large, Non-Uniformly Sampled Surveys: The Case
+ of the Palomar Transient Factory}
+\author{Adrian~M.~Price-Whelan\altaffilmark{\ref{col}}, Marcel~A.~Ag\"ueros\altaffilmark{\ref{col}}, Amanda Fournier\altaffilmark{\ref{ucsb}}, Rachel Street\altaffilmark{\ref{lcogt}}, Eran Ofek\altaffilmark{\ref{weiz},\ref{eins}}, David Levitan\altaffilmark{\ref{calt}}, Joshua S.\ Bloom\altaffilmark{\ref{cal}}, S.\ Bradley Cenko\altaffilmark{\ref{cal}}, Mansi M.\ Kasliwal\altaffilmark{\ref{calt}}, Shrinivas R.\ Kulkarni\altaffilmark{\ref{calt}}, Nicholas~M.~Law\altaffilmark{\ref{to},\ref{dun}}, Peter Nugent\altaffilmark{\ref{lbnl}}, Dovi Poznanski\altaffilmark{\ref{cal},\ref{calt},\ref{eins}}, Robert M.\ Quimby\altaffilmark{\ref{calt}}}
 
@@
 \altaffiltext{\theaddress}{\stepcounter{address}\label{ucsb} Department of Physics, Broida Hall, University of California, Santa Barbara, CA 93106, USA}
+\altaffiltext{\theaddress}{\stepcounter{address}\label{lcogt} Las Cumbres Observatory Global Telescope Network, Inc., 6740 Cortona Dr.\ Suite 102, Santa Barbara, CA 93117, USA}
 \altaffiltext{\theaddress}{\stepcounter{address}\label{weiz} Benoziyo Center for Astrophysics, Weizmann Institute of Science, 76100 Rehovot, Israel}
-\altaffiltext{\theaddress}{\stepcounter{address}\label{lcogt} Las Cumbres Observatory Global Telescope Network, Inc., 6740 Cortona Dr.\ Suite 102, Santa Barbara, CA 93117, USA}
+\altaffiltext{\theaddress}{\stepcounter{address}\label{calt} Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
+\altaffiltext{\theaddress}{\stepcounter{address}\label{cal} Department of Astronomy, University of California, Berkeley, CA  94720, USA}
+\altaffiltext{\theaddress}{\stepcounter{address}\label{to} Dunlap Institute for Astronomy and Astrophysics, University of Toronto, 50 St.\ George Street, Toronto M5S 3H4, Ontario, Canada}
+\altaffiltext{\theaddress}{\stepcounter{address}\label{lbnl} Computational Cosmology Center, Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA}
+\altaffiltext{\theaddress}{\stepcounter{address}\label{dun} Dunlap Fellow}
+\altaffiltext{\theaddress}{\stepcounter{address}\label{eins} Einstein Fellow}
 
@@
 \section{PTF Observations and Data}
-[TODO: Marcel]
+The Palomar Transient Factory is a transient detection system that begins with a wide-field survey camera mounted on the automated 48 inch Oschin Schmidt telescope at Palomar Observatory, CA, uses a real-time data reduction pipeline to identify transients of interest, passes these to a dedicated photometric follow-up telescope, and generates an archive of all detected sources \citep{nick2009,rau2009}.
+
+The PTF camera is the 12K$\times$8K mosaic camera formerly at the Canada-France-Hawaii Telescope.
+The camera has 11 working chips, 10$^7$ pixels, and a 7.26 deg$^2$ field-of-view \citep{rahmer2008}.
+Under median seeing conditions (1.1$\arcsec$), observations in Mould $R$ or Sloan Digital Sky Survey (SDSS) $g$ achieve 2.0$\arcsec$ full-width half-maximum images and reach 5$\sigma$ magnitudes of $R \approx 21.0$ and $g \approx 21.3$ mag in a standard 60~s exposure \citep{nick2010}.
+As of 2012 May, the PTF footprint included $\sim$9800 (2500) deg$^2$ imaged $>$25 ($>$100) times in $R$ and $\sim$2400 (250) deg$^2$ imaged that often in $g$.
+
+[UPDATE NUMBERS FOR FOOTPRINT, INCLUDE FIGURE?]
 
@@
 
+\acknowledgments
+This paper is based on observations obtained with the Samuel Oschin Telescope as part of the Palomar Transient Factory project, a scientific collaboration between the California Institute of Technology, Columbia University, Las Cumbres Observatory, the Lawrence Berkeley National Laboratory, the National Energy Research Scientific Computing Center, the University of Oxford and the Weizmann Institute of Science.
+
+\clearpage
+\setlength{\baselineskip}{0.6\baselineskip}
+\bibliography{references}
+\setlength{\baselineskip}{1.667\baselineskip}
+
 %\bibliographystyle{apj}

### 2012-08-07 ptf/ptflightcurve.py
add properties field_id,ccd_id,source_id back to lightcurve object

@@ -9,16 +9,14 @@ import logging
 
 # Third-party dependencies
 import numpy as np
-
-# Package dependences
-import ptf.analyze.analyze as analyze
+import matplotlib.pyplot as plt
 
 # ~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=~=
 
 class PTFLightCurve:
     """ Represents a PTF Light Curve """
     
-    def __init__(self, mjd, mag, error, metadata=None):
+    def __init__(self, mjd, mag, error, metadata=None, exposures=None):
         """ Create a PTFLightCurve by passing equal-length arrays of mjd, magnitude, and
             magnitude errors. This object also accepts an optional metadata parameter, 
             which is any numpy recarray that contains extra information about the light
@@ -29,7 +27,43 @@ class PTFLightCurve:
         self.amjd = self.mjd = np.array(mjd)[idx].astype(np.float64)
         self.amag = self.mag = np.array(mag)[idx].astype(np.float64)
         self.error = np.array(error)[idx].astype(np.float64)
-        self.metadata = np.array(metadata)
+        
+        if metadata != None:
+            self.metadata = np.array(metadata)
+        else:
+            self.metadata = None
+    
+        if exposures != None:
+            self.exposures = np.array(exposures)
+        else:
+            self.exposures = None
+            
+    @property
+    def field_id(self):
+        if self.exposures != None:
+            fid = int(self.exposures["fieldID"][0])
+        else:
+            fid = None
+        
+        return fid
+        
+    @property
+    def ccd_id(self):
+        if self.exposures != None:
+            cid = int(self.exposures["ccdID"][0])
+        else:
+            cid = None
+        
+        return cid
+    
+    @property
+    def source_id(self):
+        if self.metadata != None:
+            sid = int(self.metadata["matchedSourceID"][0])
+        else:
+            sid = None
+        
+        return sid
     
     def plot(self, ax=None, **kwargs):
         """ Either plots the light curve and show()'s it to the display, or plots it on 
@@ -46,7 +80,6 @@ class PTFLightCurve:
             kwargs["c"] = 'k'
         
         if ax == None:
-            import matplotlib.pyplot as plt
             fig = plt.figure()
             ax = fig.add_subplot(111)
             ax.errorbar(self.mjd, self.mag, self.error, ecolor='0.7', capsize=0, **kwargs)

### 2012-08-07 detectionefficiency.py
add ability to specify fixed impact parameter

@@ -65,7 +65,7 @@ def detection_efficiency(statistic, indices_with_events, lower_cut=None, upper_c
     
     return (efficiency, N_false_positives)
 
-def simulate_events_compute_indices(light_curves, events_per_light_curve=100, indices=["j","k","eta","sigma_mu","delta_chi_squared"]):
+def simulate_events_compute_indices(light_curves, events_per_light_curve=100, indices=["j","k","eta","sigma_mu","delta_chi_squared"], u0=None):
     """ """
     var_indices_with_events = []
     
@@ -80,7 +80,7 @@ def simulate_events_compute_indices(light_curves, events_per_light_curve=100, in
             # Only add events 50% of the time, to allow for false positives
             if np.random.uniform() > 0.5:
                 tE = 10**np.random.uniform(0., 3.)
-                light_curve.addMicrolensingEvent(tE=tE)
+                light_curve.addMicrolensingEvent(tE=tE, u0=u0)
                 event_added = True
             
             try:
@@ -95,7 +95,7 @@ def simulate_events_compute_indices(light_curves, events_per_light_curve=100, in
     
     return var_indices_with_events
     
-def compare_detection_efficiencies(light_curves, events_per_light_curve=100, indices=["j","k","eta","sigma_mu","delta_chi_squared"]):
+def compare_detection_efficiencies(light_curves, events_per_light_curve=100, indices=["j","k","eta","sigma_mu","delta_chi_squared"], u0=None):
     """ This figure should show the detection efficiency curve for all indices by
         injecting events into a random sampling of light_curves.
         
@@ -117,7 +117,7 @@ def compare_detection_efficiencies(light_curves, events_per_light_curve=100, ind
             pass
             
     selection_var_indices = np.array(var_list, dtype=zip(indices, [float]*len(indices)))
-    simulated_var_indices = simulate_events_compute_indices(light_curves, events_per_light_curve, indices=indices)
+    simulated_var_indices = simulate_events_compute_indices(light_curves, events_per_light_curve, indices=indices, u0=u0)
     
     timescale_bins = np.logspace(0, 3, 100) # from 1 day to 1000 days
     total_counts_per_bin, tE_bin_edges = np.histogram(simulated_var_indices[simulated_var_indices["event_added"]]["tE"], bins=timescale_bins)
@@ -138,7 +138,7 @@ def compare_detection_efficiencies(light_curves, events_per_light_curve=100, ind
     
     return data
 
[…]

### 2012-08-09 ptf/photometricdatabase.py
revise candidate selection, but still not final

@@ -204,8 +204,7 @@ class CCD(object):
     
     def light_curve(self, source_id, mag_type="relative", clean=False):
         """ Get a light curve for a given source ID from this chip """
-        # TODO: should this be here?
-
+        
         chip = self.read()
         
         if clean:

### 2012-08-09 detectionefficiency.py
last

@@ -148,7 +148,10 @@ def compare_detection_efficiencies_on_field(field, light_curves_per_ccd, events_
     logger.debug(greenText("/// compare_detection_efficiencies_on_field ///"))
     logger.debug("Computing indices {} for {} light curves on each CCD of field {}".format(",".join(indices), light_curves_per_ccd, field))
     
-    file_base = "field{:06d}_Nperccd{}_Nevents{}".format(field.id, light_curves_per_ccd, events_per_light_curve)
+    if u0 == None:
+        file_base = "field{:06d}_Nperccd{}_Nevents{}".format(field.id, light_curves_per_ccd, events_per_light_curve)
+    else:
+        file_base = "field{:06d}_Nperccd{}_Nevents{}_u0{:.3f}".format(field.id, light_curves_per_ccd, events_per_light_curve, u0)
     
     if overwrite and os.path.exists("data/detectionefficiency/{}.pickle".format(file_base)):
         logger.debug("File {} already exists, and you want to overwrite it -- so I'm going to remove it, fine!".format("data/detectionefficiency/{}.pickle".format(file_base)))
@@ -198,17 +201,17 @@ def compare_detection_efficiencies_on_field(field, light_curves_per_ccd, events_
     fig1 = plt.figure(figsize=(20,15))
     ax1 = fig1.add_subplot(111)
     
-    # Styling for lines
-    line_styles = [{"lw" : 3, "ls" : "-.", "color" : "c"}, \
-                   {"lw" : 3, "ls" : ":", "color" : "m"}, \
-                   {"lw" : 3, "ls" : "--", "color" : "g"}, \
+    # Styling for lines: J, K, eta, sigma_mu, delta_chi_squared
+    line_styles = [{"lw" : 2, "ls" : "-.", "color" : "c"}, \
+                   {"lw" : 2, "ls" : ":", "color" : "m"}, \
+                   {"lw" : 3, "ls" : "-", "color" : "k"}, \
                    {"lw" : 1.5, "ls" : "--", "color" : "y"}, \
-                   {"lw" : 2, "ls" : "-", "color" : "k"}]
+                   {"lw" : 3, "ls" : "--", "color" : "k"}]
     
     for ii,index_name in enumerate(indices):
         ax1.semilogx((data["bin_edges"][1:]+data["bin_edges"][:-1])/2, \
                      data[index_name]["detections_per_bin"] / data["total_counts_per_bin"], \
[…]

### 2012-08-09 ptf/photometricdatabase.py
added relPhotFlags and ipacflags to quality cuts

@@ -43,6 +43,49 @@ def quality_cut(sourcedata, source_id=None, ccd_edge_cutoff=25):
         ccd_edge_cutoff : int
             Define the cutoff for sources near the edge of a CCD. The cut will remove
             all data points where the source is nearer than this limit to the edge.
+            
+        IPAC Flags:
+            # 0 	aircraft/satellite track
+            # 1 	detection by SExtractor (before flattening)
+            # 2 	high dark current
+            # 3 	reserved
+            # 4 	noisy/hot pixel
+            # 5 	ghost
+            # 6 	CCD-bleed
+            # 7 	rad hit
+            # 8 	saturated
+            # 9 	dead/bad pixel
+            # 10 	not-a-number pixel
+            # 11 	dirt on optics (pixel Nsigma below coarse local median)
+            # 12 	halo
+            
+        relPhotFlags:
+            # bit 0 = calibration detection
+            # bit 1 = no relative photometry generated
+            # bit 2 = exposure with high systematic error
+            # bit 3 = saturated
+        
+        SExtractor flags:
+        ----------------
+        
+        1     The object has neighbors, bright and close enough to 
+              significantly bias the photometry, or bad pixels 
+              (more than 10% of the integrated area affected).
+        
+        2     The object was originally blended with another one.
+        
+        4     At least one pixel of the object is saturated 
+              (or very close to).
+        
+        8     The object is truncates (to close to an image boundary).
+        
+        16    Object's aperture data are incomplete or corrupted.
+        
+        32    Object's isophotal data are incomplete or corrupted.
+        
+        64    A memory overflow occurred during deblending.
+        
+        128   A memory overflow occurred during extraction.
     """    
     x_cut1, x_cut2 = ccd_edge_cutoff, globals.ccd_size[0] - ccd_edge_cutoff
     y_cut1, y_cut2 = ccd_edge_cutoff, globals.ccd_size[1] - ccd_edge_cutoff
@@ -52,13 +95,18 @@ def quality_cut(sourcedata, source_id=None, ccd_edge_cutoff=25):
     else:
         src = "(matchedSourceID == {}) &".format(source_id)
     
+    # Saturation limit, 14.3, based on email conversation with David Levitan
+    
     sourcedata = sourcedata.readWhere(src + '(sextractorFlags < 8) & \
                                        (x_image > {}) & (x_image < {}) & \
[…]

### 2012-08-10 detectionefficiency.py
reorder detection efficiency code to allow for multiple u0 and magnitudes. also fixed quality cut

@@ -69,8 +69,9 @@ def simulate_events_compute_indices(light_curves, events_per_light_curve=100, in
     """ """
     var_indices_with_events = []
     
-    # For each light curve, add 100 different microlensing events and recompute indices
+    # For each light curve, add [100] different microlensing events and recompute indices
     for light_curve in (SimulatedLightCurve(lc.mjd, mag=lc.mag, error=lc.error) for lc in light_curves):
+        reference_mag = np.median(light_curve.mag)
         for event_id in range(events_per_light_curve):
             # Reset the simulated light curve back to the original data, e.g. erase any previously
             #   added microlensing events
@@ -87,10 +88,10 @@ def simulate_events_compute_indices(light_curves, events_per_light_curve=100, in
                 lc_var_indices = analyze.compute_variability_indices(light_curve, indices, return_tuple=True)
             except:
                 break
-            var_indices_with_events.append(lc_var_indices + (light_curve.tE, event_added))
+            var_indices_with_events.append(lc_var_indices + (light_curve.tE, light_curve.u0, reference_mag, event_added))
     
-    names = indices + ["tE", "event_added"]
-    dtypes = [float]*len(indices) + [float,bool]
+    names = indices + ["tE", "u0", "m", "event_added"]
+    dtypes = [float]*len(indices) + [float,float,float,bool]
     var_indices_with_events = np.array(var_indices_with_events, dtype=zip(names,dtypes))
     
     return var_indices_with_events
@@ -138,7 +139,7 @@ def compare_detection_efficiencies(light_curves, events_per_light_curve=100, ind
     
     return data
 
-def compare_detection_efficiencies_on_field(field, light_curves_per_ccd, events_per_light_curve, overwrite=False, indices=["j","k","eta","sigma_mu","delta_chi_squared"], u0=None):
+def compare_detection_efficiencies_on_field(field, light_curves_per_ccd, events_per_light_curve, overwrite=False, indices=["j","k","eta","sigma_mu","delta_chi_squared"], u0s=[], limiting_mags=[]):
     """ Compare the detection efficiencies of various variability statistics using the
         light curves from an entire PTF field.
     """
@@ -148,77 +149,125 @@ def compare_detection_efficiencies_on_field(field, light_curves_per_ccd, events_
     logger.debug(greenText("/// compare_detection_efficiencies_on_field ///"))
[…]

### 2012-08-10 ptf/photometricdatabase.py

@@ -30,7 +30,7 @@ pytable_base_string = os.path.join(match_path, "match_{filter.id:02d}_{field.id:
 filter_map = {"R" : 2, "g" : 1}
 inv_filter_map = {2 : "R", 1 : "g"}
 
-def quality_cut(sourcedata, source_id=None, ccd_edge_cutoff=25):
+def quality_cut(sourcedata, source_id=None, ccd_edge_cutoff=15):
     """ This function accepts a Pytables table object (from the 'sourcedata' table)
         and returns only the rows that pass the given quality cuts.
         
@@ -66,26 +66,17 @@ def quality_cut(sourcedata, source_id=None, ccd_edge_cutoff=25):
             # bit 3 = saturated
         
         SExtractor flags:
-        ----------------
-        
-        1     The object has neighbors, bright and close enough to 
-              significantly bias the photometry, or bad pixels 
-              (more than 10% of the integrated area affected).
-        
-        2     The object was originally blended with another one.
-        
-        4     At least one pixel of the object is saturated 
-              (or very close to).
-        
-        8     The object is truncates (to close to an image boundary).
-        
-        16    Object's aperture data are incomplete or corrupted.
-        
-        32    Object's isophotal data are incomplete or corrupted.
-        
-        64    A memory overflow occurred during deblending.
-        
-        128   A memory overflow occurred during extraction.
+            1     The object has neighbors, bright and close enough to 
+                  significantly bias the photometry, or bad pixels 
+                  (more than 10% of the integrated area affected).
+            2     The object was originally blended with another one.
+            4     At least one pixel of the object is saturated 
+                  (or very close to).
+            8     The object is truncates (to close to an image boundary).
+            16    Object's aperture data are incomplete or corrupted.
+            32    Object's isophotal data are incomplete or corrupted.
+            64    A memory overflow occurred during deblending.
+            128   A memory overflow occurred during extraction.
     """    
     x_cut1, x_cut2 = ccd_edge_cutoff, globals.ccd_size[0] - ccd_edge_cutoff
     y_cut1, y_cut2 = ccd_edge_cutoff, globals.ccd_size[1] - ccd_edge_cutoff
@@ -96,20 +87,17 @@ def quality_cut(sourcedata, source_id=None, ccd_edge_cutoff=25):
         src = "(matchedSourceID == {}) &".format(source_id)
     
[…]

### 2012-08-10 text/ptf_microlensing_paper.tex

@@
 \def\asec{\ifmmode^{\prime\prime}\else$^{\prime\prime}$\fi}
+
 \newcommand{\apwsim}{\raisebox{0.2ex}{\scriptsize$\sim$\normalsize}} 
+\newcommand{\inlinecode}{\texttt}
 
@@
 Under median seeing conditions (1.1$\arcsec$), observations in Mould $R$ or Sloan Digital Sky Survey (SDSS) $g$ achieve 2.0$\arcsec$ full-width half-maximum images and reach 5$\sigma$ magnitudes of $R \approx 21.0$ and $g \approx 21.3$ mag in a standard 60~s exposure \citep{nick2010}.
-As of 2012 May, the PTF footprint included $\sim$9800 (2500) deg$^2$ imaged $>$25 ($>$100) times in $R$ and $\sim$2400 (250) deg$^2$ imaged that often in $g$.
+As of August 9, 2012, the PTF footprint included $\apwsim$10,100 (2600) deg$^2$ imaged $>$25 ($>$100) times in $R$ and $\apwsim$2200 (170) deg$^2$ imaged that often in $g$.
 
@@
 \section{Microlensing event recovery}
-Microlensing surveys typically use difference image analysis (cite Lupton?) to initially identify transient events and then analyze the candidate light curves to search for microlensing events (cite MACHO, OGLE).
-This method is essential for studying crowded fields such as the Galactic bulge, and also limits the number of light curves that must be searched for microlensing events, thus saving on computation time.
-This [method] also provides real-time detection of events so surveys can utilize networks of telescopes to provide follow-up observations of these transients.
+Microlensing surveys typically use difference image analysis (DIA;cite Lupton?) to identify transient events and then analyze light curves of these transient sources to search for microlensing events (cite MACHO, OGLE).
+This method is essential for studying crowded fields such as the Galactic bulge where photometry is difficult (cite or figure?), but also limits the number of light curves that must be searched for microlensing events, thus saving on computation time.
+Pre-selection with DIA also enables fast, real-time detection of events so dedicated microlensing surveys can obtain high-cadence follow-up observations of transient sources.
 Our approach is to instead take the preexisting PTF light curve database over the entire survey footprint and develop fast, robust search algorithms to pick out microlensing event candidates from this massive data archive.
+
+The average number of light curves with $>$25 ``good'' observations is $\apwsim 50,000$ per PTF field, with $\apwsim64,000,000$  ``good'' light curves in total.
+By our estimates, utilizing all cores on an 8-core 2.0 GHz compute, it would take [TODO: estimate this number] to fit even the simplest 5-parameter microlensing event model to each of these light curves using \inlinecode{ANSI-C} wrapped in \inlinecode{Python}.
+Previous studies have had the advantage of pre-selecting candidates via difference image analysis, but have utilized a variety of selection techniques for flagging candidate microlensing light curves after this step.
+However, most of these methods still rely on fitting a point-source, point-lens microlensing model to each light curve and performing some selection based on the $\chi^2$ value of this fit (cite Sumi et al.
+2006, MACHO work, ?).
+More recent work has begun to apply machine learning algorithms to the general light curve classification problem (cite Joey Richards), however this is still a work in progress.
+We set out to find a fast, robust selection method that [TODO:].
 
@@
 where $\mathcal{U}$ is a uniform distribution.
+
+% Talk about hypothetical selection criteria, and the things we have to consider for detection efficiency (possible selection criteria, optimizing selection method, selecting candidates)
 

### 2012-08-12 detectionefficiency.py
major revision of detection efficiency code, but NOT IN WORKING STATE

@@ -11,6 +11,13 @@ import os
 import sys
 import cPickle as pickle
 
+# Create logger
+logger = logging.getLogger(__name__)
+#ch = logging.StreamHandler()
+#formatter = logging.Formatter("%(name)s / %(levelname)s / %(message)s")
+#ch.setFormatter(formatter)
+#logger.addHandler(ch)
+
 # Third-party
 from apwlib.globals import greenText
 import numpy as np
@@ -24,13 +31,6 @@ from ptf.simulation.simulatedlightcurve import SimulatedLightCurve
 import ptf.photometricdatabase as pdb
 import ptf.analyze.analyze as analyze
 from ptf.globals import index_to_label
-
-# Create logger
-logger = logging.getLogger(__name__)
-ch = logging.StreamHandler()
-formatter = logging.Formatter("%(name)s / %(levelname)s / %(message)s")
-ch.setFormatter(formatter)
-logger.addHandler(ch)
 
 def detection_efficiency(statistic, indices_with_events, lower_cut=None, upper_cut=None):
     """ Compute the detection efficiency of a specific variability statistic.
@@ -65,6 +65,7 @@ def detection_efficiency(statistic, indices_with_events, lower_cut=None, upper_c
     
     return (efficiency, N_false_positives)
 
+'''
 def simulate_events_compute_indices(light_curves, events_per_light_curve=100, indices=["j","k","eta","sigma_mu","delta_chi_squared"], u0=None):
     """ """
     var_indices_with_events = []
@@ -95,7 +96,88 @@ def simulate_events_compute_indices(light_curves, events_per_light_curve=100, in
     var_indices_with_events = np.array(var_indices_with_events, dtype=zip(names,dtypes))
     
     return var_indices_with_events
+'''
+
+def simulate_events_worker(sim_light_curve, tE, u0, reference_mag, indices):
+    """ Only to be used in the multiprocessing pool below! """
     
+    light_curve = copy.copy(sim_light_curve)
+    
+    # Reset the simulated light curve back to the original data, e.g. erase any previously
+    #   added microlensing events
+    light_curve.reset()
+    
+    light_curve.addMicrolensingEvent(tE=tE, u0=u0)
+    
+    try:
+        lc_var_indices = analyze.compute_variability_indices(light_curve, indices, return_tuple=True) + (light_curve.tE, light_curve.u0, reference_mag, True)
+        return lc_var_indices
+    except:
[…]

### 2012-08-12 ptf/photometricdatabase.py

@@ -30,7 +30,7 @@ pytable_base_string = os.path.join(match_path, "match_{filter.id:02d}_{field.id:
 filter_map = {"R" : 2, "g" : 1}
 inv_filter_map = {2 : "R", 1 : "g"}
 
-def quality_cut(sourcedata, source_id=None, ccd_edge_cutoff=15):
+def quality_cut(sourcedata, source_id=None, ccd_edge_cutoff=15, where=[]):
     """ This function accepts a Pytables table object (from the 'sourcedata' table)
         and returns only the rows that pass the given quality cuts.
         
@@ -43,6 +43,8 @@ def quality_cut(sourcedata, source_id=None, ccd_edge_cutoff=15):
         ccd_edge_cutoff : int
             Define the cutoff for sources near the edge of a CCD. The cut will remove
             all data points where the source is nearer than this limit to the edge.
+        where : list
+            Additional where conditions for the search
             
         IPAC Flags:
             # 0 	aircraft/satellite track
@@ -86,13 +88,18 @@ def quality_cut(sourcedata, source_id=None, ccd_edge_cutoff=15):
     else:
         src = "(matchedSourceID == {}) &".format(source_id)
     
+    if len(where) > 0:
+        where_string = " & " + " & ".join(where)
+    else:
+        where_string = ""
+    
     # Saturation limit, 14.3, based on email conversation with David Levitan
-    sourcedata = sourcedata.readWhere(src + '(sextractorFlags < 8) & \
+    sourcedata = sourcedata.readWhere(src + '(sextractorFlags == 0) & \
                                        (x_image > {}) & (x_image < {}) & \
                                        (y_image > {}) & (y_image < {}) & \
                                        (relPhotFlags < 4) & \
                                        ( (ipacFlags == 2) | (ipacFlags == 0) ) & \
-                                       (mag > 14.3) & (mag < 22)'.format(x_cut1, x_cut2, y_cut1, y_cut2))
+                                       (mag > 14.3) & (mag < 22)'.format(x_cut1, x_cut2, y_cut1, y_cut2) + where_string)
     
     # implied...
     #sourcedata = sourcedata[np.isfinite(sourcedata["mag"]) & \
@@ -237,16 +244,21 @@ class CCD(object):
     
     def close(self):
         self._file.close()
+        self._file = None
     
-    def light_curve(self, source_id, mag_type="relative", clean=False):
+    def light_curve(self, source_id, mag_type="relative", clean=False, where=[], barebones=False):
[…]

### 2012-08-12 ptf/ptflightcurve.py

@@ -100,14 +100,15 @@ class PTFLightCurve:
         elif os.path.exists(filename) and overwrite:
             os.remove(filename)
         
-        xx, ext = os.path.split(filename)
+        # TODO: This is broken!
+        xx, ext = os.path.splitext(filename)
         
         if ext == ".txt":
             np.savetxt(filename, np.transpose((self.mjd, self.mag, self.error)), fmt="%.5f\t%.5f\t%.5f")
         elif ext == ".pickle":
             import cPickle as pickle
             
-            f = open(filename)
+            f = open(filename, "w")
             pickle.dump(self, f)
             f.close()
         else:

### 2012-08-14 detectionefficiency.py
uses multiprocessing to reduce the detection efficienct runtime

@@ -65,39 +65,6 @@ def detection_efficiency(statistic, indices_with_events, lower_cut=None, upper_c
     
     return (efficiency, N_false_positives)
 
-'''
-def simulate_events_compute_indices(light_curves, events_per_light_curve=100, indices=["j","k","eta","sigma_mu","delta_chi_squared"], u0=None):
-    """ """
-    var_indices_with_events = []
-    
-    # For each light curve, add [100] different microlensing events and recompute indices
-    for light_curve in (SimulatedLightCurve(lc.mjd, mag=lc.mag, error=lc.error) for lc in light_curves):
-        reference_mag = np.median(light_curve.mag)
-        for event_id in range(events_per_light_curve):
-            # Reset the simulated light curve back to the original data, e.g. erase any previously
-            #   added microlensing events
-            light_curve.reset()
-            event_added = False
-            
-            # Only add events 50% of the time, to allow for false positives
-            if np.random.uniform() > 0.5:
-                tE = 10**np.random.uniform(0., 3.)
-                light_curve.addMicrolensingEvent(tE=tE, u0=u0)
-                event_added = True
-            
-            try:
-                lc_var_indices = analyze.compute_variability_indices(light_curve, indices, return_tuple=True)
-            except:
-                break
-            var_indices_with_events.append(lc_var_indices + (light_curve.tE, light_curve.u0, reference_mag, event_added))
-    
-    names = indices + ["tE", "u0", "m", "event_added"]
-    dtypes = [float]*len(indices) + [float,float,float,bool]
-    var_indices_with_events = np.array(var_indices_with_events, dtype=zip(names,dtypes))
-    
-    return var_indices_with_events
-'''
-
 def simulate_events_worker(sim_light_curve, tE, u0, reference_mag, indices):
     """ Only to be used in the multiprocessing pool below! """
     
@@ -142,28 +109,6 @@ def simulate_events_compute_indices(light_curve, events_per_light_curve=100, ind
             pool.apply_async(simulate_events_worker, args=(sim_light_curve, tE, u0, reference_mag, indices), callback=callback)
         else:
             var_indices_with_events.append(vanilla_var_indices + (sim_light_curve.tE, sim_light_curve.u0, reference_mag, False))
-        
-        """
-        # Reset the simulated light curve back to the original data, e.g. erase any previously
-        #   added microlensing events
-        sim_light_curve.reset()
[…]

### 2012-08-14 ptf/ptflightcurve.py

@@ -37,7 +37,19 @@ class PTFLightCurve:
             self.exposures = np.array(exposures)
         else:
             self.exposures = None
-            
+    
+    def shuffle(self):
+        """ Randomly move around the magnitude values in the light curve to shuffle up
+            the data points, but keep the mjd values fixed
+        """
+        idx = np.arange(len(self.mjd), dtype=int)
+        
+        np.random.shuffle(idx)
+        self.mag = self.mag[idx]
+        
+        np.random.shuffle(idx)
+        self.error = self.mag[idx]
+    
     @property
     def field_id(self):
         if self.exposures != None:

### 2012-08-14 text/paper/ptf_microlensing_paper.tex
move paper into paper directory, add figures directory

--- 2012-08-10 text/ptf_microlensing_paper.tex
@@
 
-\title{Identifying Microlenses in Large, Non-Uniformly Sampled Surveys: The Case
+\title{Identifying Microlensing Events in Large, Non-Uniformly Sampled Surveys: The Case
  of the Palomar Transient Factory}
@@
 Under median seeing conditions (1.1$\arcsec$), observations in Mould $R$ or Sloan Digital Sky Survey (SDSS) $g$ achieve 2.0$\arcsec$ full-width half-maximum images and reach 5$\sigma$ magnitudes of $R \approx 21.0$ and $g \approx 21.3$ mag in a standard 60~s exposure \citep{nick2010}.
-As of August 9, 2012, the PTF footprint included $\apwsim$10,100 (2600) deg$^2$ imaged $>$25 ($>$100) times in $R$ and $\apwsim$2200 (170) deg$^2$ imaged that often in $g$.
-
-[UPDATE NUMBERS FOR FOOTPRINT, INCLUDE FIGURE?]
+As of August 9, 2012, the PTF footprint included $\apwsim$10,100 (2600) deg$^2$ imaged $>$25 ($>$100) times in $R$ and $\apwsim$2200 (170) deg$^2$ imaged that often in $g$ (\figurename~\ref{fig:survey_footprint}).
 
@@
 
-The average number of light curves with $>$25 ``good'' observations is $\apwsim 50,000$ per PTF field, with $\apwsim64,000,000$  ``good'' light curves in total.
+The average number of light curves per PTF field with $>$25 ``good'' observations is $\apwsim 50,000$, with $\apwsim64,000,000$  ``good'' light curves in total.
 By our estimates, utilizing all cores on an 8-core 2.0 GHz compute, it would take [TODO: estimate this number] to fit even the simplest 5-parameter microlensing event model to each of these light curves using \inlinecode{ANSI-C} wrapped in \inlinecode{Python}.
-Previous studies have had the advantage of pre-selecting candidates via difference image analysis, but have utilized a variety of selection techniques for flagging candidate microlensing light curves after this step.
-However, most of these methods still rely on fitting a point-source, point-lens microlensing model to each light curve and performing some selection based on the $\chi^2$ value of this fit (cite Sumi et al.
+Previous studies have had the advantage of pre-selecting candidates via difference image analysis, automatically limiting their sample to $\apwsim$thousands of light curves.
+Past work has also utilized a variety of selection techniques for flagging candidate microlensing light curves after these steps, however most of these methods still rely on fitting a point-source, point-lens microlensing model to each light curve and performing some selection based on the $\chi^2$ value of this fit (cite Sumi et al.
 2006, MACHO work, ?).
 More recent work has begun to apply machine learning algorithms to the general light curve classification problem (cite Joey Richards), however this is still a work in progress.
-We set out to find a fast, robust selection method that [TODO:].
+We set out to find a fast, robust selection method for identifying microlensing events in big data archives by using a set of variability statistics previously applied to finding variable stars in general (Shin M.-S.
+et al.
+).
+
+[Describe variability indices?]
 
@@
 
-- Discuss $\Delta \chi^2$ method, downfalls (speed, not robust -- how do you define the threshold?)
+- Discuss $\Delta \chi^2$ method, downfalls (speed, not robust -- how do you define the threshold?).
+Show a plot of distribution of $\Delta \chi^2$ with and without simulated events, e.g.
+the distribution doesn't separate well, then show $\eta$ and how the peak of the distribution moves.
 
@@
 %
-%\begin{figure}
-%	\centering
-%	\caption{Computed detection efficiency as a function of number of observations for a fixed baseline of 365 days. This plot uses entirely simulated light curves.}
-%    \includegraphics[width=1.0\textwidth]{figures/detection_efficiency_figure.pdf}
-%    \label{fig:detection_eff}
-%\end{figure}	
+\begin{figure}
+	\centering
+	\caption{Computed detection efficiency as a function of number of observations for a fixed baseline of 365 days. This plot uses entirely simulated light curves.}
+    \includegraphics[width=1.0\textwidth]{figures/detection_efficiency_figure.pdf}
+    \label{fig:survey_footprint}
+\end{figure}	
 %

### 2012-08-14 ptf/photometricdatabase.py
Field now knows its position, added new script for dealign with survey coverage

@@ -101,10 +101,9 @@ def quality_cut(sourcedata, source_id=None, ccd_edge_cutoff=15, where=[]):
                                        ( (ipacFlags == 2) | (ipacFlags == 0) ) & \
                                        (mag > 14.3) & (mag < 22)'.format(x_cut1, x_cut2, y_cut1, y_cut2) + where_string)
     
-    # implied...
-    #sourcedata = sourcedata[np.isfinite(sourcedata["mag"]) & \
-    #                        np.isfinite(sourcedata["mjd"]) & \
-    #                        np.isfinite(sourcedata["magErr"])]
+    sourcedata = sourcedata[np.isfinite(sourcedata["mag"]) & \
+                            np.isfinite(sourcedata["mjd"]) & \
+                            np.isfinite(sourcedata["magErr"])]
                             
     return sourcedata
 
@@ -130,6 +129,9 @@ class Filter(object):
         
     def __repr__(self):
         return "<Filter: id={}, name={}>".format(self.id, self.name)
+    
+    def __str__(self):
+        return self.name
 
 class Field(object):
     """ Represents PTF Field """
@@ -175,6 +177,10 @@ class Field(object):
         
         if len(self.ccds) == 0:
             logging.debug("No CCD data found for: {}".format(self))
+        
+        this_field = all_fields[all_fields["id"] == self.id]
+        self.ra = this_field["ra"]
+        self.dec = this_field["dec"]
     
     @property
     def number_of_exposures(self):
@@ -294,52 +300,7 @@ def convert_all_fields_txt(txt_filename="ptf_allfields.txt", npy_filename="all_f
 
     np.save(npy_filename, all_fields)
     
-    return True    
-
-def select_most_observed_fields(minimum_num_observations=25, limit=0):
-    """ Return a list of PTFField objects for fields with more than 
-        minimum_num_observations observations.
-        
-        There are 16598 fields in total.
-    """
-    
-    filter_R = Filter("R")
-    field = None
-    fields = []
-    for row in all_fields:
-        if field is not None: field.close()
-        field = Field(row["id"], filter=filter_R)
-
-        if sum(np.array(field.number_of_exposures.values()) > minimum_num_observations) > 0:
-            fields.append(field)
-        
-        if limit > 0:
-            if len(fields) >= limit: break
-    
-    return fields
-
-'''
-def get_raw_light_curve(source_id, filter_id=2, field_id=None, ccd_id=None):
-    """ Given a source_id, return a PTFLightCurve object for the 
[…]

### 2012-08-14 survey_coverage.py

@@ -1,0 +1,177 @@
+# coding: utf-8
+from __future__ import division
+
+""" This module contains utilities for computing PTF's survey coverage, both in time and position. """
+
+__author__ = "adrn <adrn@astro.columbia.edu>"
+
+# Standard library
+import os
+import datetime
+import re
+import glob
+import cPickle as pickle
+import logging
+import time, datetime
+
+# Third-party
+import numpy as np
+import tables
+
+# Project
+# TODO: add some kind of catch here -- if this import fails, it's probably because I'm trying to
+#           run on the wrong machine...
+import ptf.photometricdatabase as pdb
+
+# Create logger for this module
+logger = logging.getLogger(__name__)
+ch = logging.StreamHandler()
+formatter = logging.Formatter("%(name)s / %(levelname)s / %(message)s")
+ch.setFormatter(formatter)
+logger.addHandler(ch)
+
+class SurveyInfo(object):
+    
+    def __init__(self, filter, overwrite=False):
+        """ Convenience class for getting information about the PTF survey
+            
+            Parameters
+            ----------
+            filter : int, str, ptf.photometricdatabase.Filter
+                Any parseable, unambiguous filter representation.
+            overwrite : bool (optional)
+                Overwrite any cache files if they already exist
+        """
+        
+        if not isinstance(filter, pdb.Filter):
+            self.filter = pdb.Filter(filter)
+        
+        cache_filename = os.path.join("data", "survey_coverage", "fields_observations_{}.npy".format(str(self.filter)))
+        self._fields_exposures = get_fields_exposures(self.filter, filename=cache_filename, overwrite=overwrite)
+        self.timestamp = datetime.datetime.strptime(time.ctime(os.path.getmtime(cache_filename)), "%a %b %d %H:%M:%S %Y")
+        
+    def fields(self, min_num_observations):
+        """ Return a list of fields with more than the above number of observations """
+        
+        field_ids = self._fields_exposures[self._fields_exposures["num_exposures"] > min_num_observations]["field"]
+        return [pdb.Field(x, self.filter) for x in field_ids]
+
+def get_fields_exposures(filter, filename=None, overwrite=False):
+    """ Given a filter, go to the PTF photometric database and get information about all PTF
+        fields for that filter, and the number of good exposures per field.
+        
+        Parameters
+        ----------
[…]

### 2012-08-15 detectionefficiency.py
finish survey coverage plotting and add more failsafes to detection efficiency code

@@ -13,10 +13,10 @@ import cPickle as pickle
 
 # Create logger
 logger = logging.getLogger(__name__)
-#ch = logging.StreamHandler()
-#formatter = logging.Formatter("%(name)s / %(levelname)s / %(message)s")
-#ch.setFormatter(formatter)
-#logger.addHandler(ch)
+ch = logging.StreamHandler()
+formatter = logging.Formatter("%(name)s / %(levelname)s / %(message)s")
+ch.setFormatter(formatter)
+logger.addHandler(ch)
 
 # Third-party
 from apwlib.globals import greenText
@@ -232,6 +232,10 @@ def compare_detection_efficiencies_on_field(field, light_curves_per_ccd, events_
                     good_source_ids.append(source_id)
                     count += 1
                     if count >= light_curves_per_ccd: break
+                
+                if len(good_source_ids) == 0:
+                    logger.error("No good sources selected from this CCD for mag range {:.2f}-{:.2f}!".format(limiting_mag1, limiting_mag2))
+                    continue
                 
                 # HACK: This is super hacky...
                 # ----------------------------------------------

### 2012-08-15 ptf/photometricdatabase.py

@@ -13,10 +13,17 @@ import numpy as np
 import apwlib.geometry as g
 from apwlib.globals import redText, greenText, yellowText
 
+# Create logger for this module
+logger = logging.getLogger(__name__)
+ch = logging.StreamHandler()
+formatter = logging.Formatter("%(name)s / %(levelname)s / %(message)s")
+ch.setFormatter(formatter)
+logger.addHandler(ch)
+
 try:
     import tables
 except ImportError:
-    logging.warn("PyTables not found! Some functionality won't work.\nTry running with: /scr4/dlevitan/sw/epd-7.1-1-rh5-x86_64/bin/python instead.")
+    logger.warning("PyTables not found! Some functionality won't work.\nTry running with: /scr4/dlevitan/sw/epd-7.1-1-rh5-x86_64/bin/python instead.")
 
 # PTF
 import globals
@@ -139,7 +146,7 @@ class Field(object):
     def __repr__(self):
         return "<Field: id={}, filter={}>".format(self.id, self.filter.id)
     
-    def __init__(self, field_id, filter):
+    def __init__(self, field_id, filter, number_of_exposures=None):
         """ Create a field object given a PTF Field ID
             
             Parameters
@@ -148,7 +155,8 @@ class Field(object):
                 The PTF Field ID for a field.
             filter : Filter
                 A PTF Filter object (e.g. R = 2, g = 1)
-            
+            number_of_exposures : int (optional)
+                The number of exposures this field has in the specified filter.
         """
         
         # Validate Field ID
@@ -166,6 +174,12 @@ class Field(object):
         else:
             raise ValueError(redText("filter") + " parameter must be Filter object")
         
+        # Validate number_of_exposures
+        if number_of_exposures != None:
+            self._num_exp = int(number_of_exposures)
+        else:
+            self._num_exp = None
+        
         self.ccds = dict()
         
         # Create new CCD objects for this field
@@ -173,26 +187,45 @@ class Field(object):
             try:
                 self.ccds[ccd_id] = CCD(ccd_id, field=self, filter=self.filter)
             except ValueError:
-                logging.debug("CCD {} not found for Field {}.".format(ccd_id, self))
+                logger.debug("CCD {} not found for Field {}.".format(ccd_id, self))
         
         if len(self.ccds) == 0:
-            logging.debug("No CCD data found for: {}".format(self))
+            logger.debug("No CCD data found for: {}".format(self))
         
         this_field = all_fields[all_fields["id"] == self.id]
[…]

### 2012-08-15 survey_coverage.py

@@ -16,12 +16,19 @@ import time, datetime
 
 # Third-party
 import numpy as np
+import matplotlib
+matplotlib.use("Agg")
+import matplotlib.cm as cm
+import matplotlib.colors as mc
+import matplotlib.pyplot as plt
+from matplotlib.patches import Rectangle
 import tables
 
 # Project
 # TODO: add some kind of catch here -- if this import fails, it's probably because I'm trying to
 #           run on the wrong machine...
 import ptf.photometricdatabase as pdb
+from ptf.globals import camera_size_degrees
 
 # Create logger for this module
 logger = logging.getLogger(__name__)
@@ -31,7 +38,11 @@ ch.setFormatter(formatter)
 logger.addHandler(ch)
 
 class SurveyInfo(object):
+    """ Some example use cases for this object:
+        
+        TODO: Fill this in!
     
+    """
     def __init__(self, filter, overwrite=False):
         """ Convenience class for getting information about the PTF survey
             
@@ -53,8 +64,8 @@ class SurveyInfo(object):
     def fields(self, min_num_observations):
         """ Return a list of fields with more than the above number of observations """
         
-        field_ids = self._fields_exposures[self._fields_exposures["num_exposures"] > min_num_observations]["field"]
-        return [pdb.Field(x, self.filter) for x in field_ids]
+        rows = self._fields_exposures[self._fields_exposures["num_exposures"] >= min_num_observations]
+        return [pdb.Field(row["field"], self.filter, number_of_exposures=row["num_exposures"]) for row in rows]
 
 def get_fields_exposures(filter, filename=None, overwrite=False):
     """ Given a filter, go to the PTF photometric database and get information about all PTF
@@ -82,7 +93,7 @@ def get_fields_exposures(filter, filename=None, overwrite=False):
         os.remove(filename)
         logger.debug("File {} deleted".format(filename))
     elif os.path.exists(filename) and not overwrite:
-        logger.info("Data file already exists.")
+        logger.info("Data file already exists: {}".format(filename))
     
     if not os.path.exists(filename):
         logger.info("Data file doesn't exist -- it could take some time to create it!")
@@ -91,7 +102,7 @@ def get_fields_exposures(filter, filename=None, overwrite=False):
         exposures = []
         
         pattr = re.compile(".*match_(\d+)_(\d+)_(\d+)")
-        for match_filename in glob.glob("/scr4/dlevitan/matches/match_{:02d}_*".format(filter.id)):
[…]

### 2012-08-15 detectionefficiency.py
add con as an index

@@ -83,7 +83,7 @@ def simulate_events_worker(sim_light_curve, tE, u0, reference_mag, indices):
         logger.warning("Failed to compute variability indices for simulated light curve!")
     
 
-def simulate_events_compute_indices(light_curve, events_per_light_curve=100, indices=["j","k","eta","sigma_mu","delta_chi_squared"], u0=None):
+def simulate_events_compute_indices(light_curve, events_per_light_curve, indices, u0=None):
     """ """
     #logger.debug("\t\tsimulate_events_compute_indices")
     
@@ -131,8 +131,12 @@ def compute_detection_efficiency(var_indices, var_indices_with_events, indices):
                 bin_edges=tE_bin_edges)
     
     for index_name in indices:
-        mu, sigma = np.mean(var_indices[index_name]), np.std(var_indices[index_name])
-        idx = (var_indices_with_events[index_name] > (mu+2.*sigma)) | (var_indices_with_events[index_name] < (mu-2.*sigma))
+        if index_name == "con":
+            idx = var_indices_with_events[index_name] >= 1
+        else:
+            mu, sigma = np.mean(var_indices[index_name]), np.std(var_indices[index_name])
+            idx = (var_indices_with_events[index_name] > (mu+2.*sigma)) | (var_indices_with_events[index_name] < (mu-2.*sigma))
+            
         detections_per_bin, tE_bin_edges = np.histogram(var_indices_with_events[idx]["tE"], bins=timescale_bins)
         
         total_efficiency = sum(var_indices_with_events[idx & events_added_idx]["event_added"]) / sum(events_added_idx)
@@ -237,6 +241,8 @@ def compare_detection_efficiencies_on_field(field, light_curves_per_ccd, events_
                     logger.error("No good sources selected from this CCD for mag range {:.2f}-{:.2f}!".format(limiting_mag1, limiting_mag2))
                     continue
                 
+                logger.info("\t\t{} good light curves selected".format(count))
+                
                 # HACK: This is super hacky...
                 # ----------------------------------------------
                 while count < light_curves_per_ccd:
@@ -264,8 +270,6 @@ def compare_detection_efficiencies_on_field(field, light_curves_per_ccd, events_
                 # ----------------------------------------------   
                     
                 
-                logger.info("\t\t{} good light curves used".format(count))
-                
             ccd.close()
         
[…]

### 2012-08-15 ptf/analyze/analyze.py
work on survey coverage, and re-adding con as a var. index

@@ -374,6 +374,19 @@ def test_eta():
     
     print eta(light_curve)
 
+def consecutive_brighter(light_curve):
+    """ Compute the number of consecutive points brighter than 3sigma from the baseline """
+    
+    continuum_mag, noise_sigma = estimate_continuum(light_curve)
+    where_condition = np.where(light_curve.mag < (continuum_mag - (3.*noise_sigma)))[0]
+    
+    n_groups = 0
+    for group in np.array_split(where_condition, np.where(np.diff(where_condition)!=1)[0]+1):
+        if len(group) > 3:
+            n_groups += 1
+    
+    return n_groups
+
 def compute_variability_indices(light_curve, indices=[], return_tuple=False):
     """ Computes variability statistics, such as those explained in M.-S. Shin et al. 2009.
         Valid indices:
@@ -383,6 +396,7 @@ def compute_variability_indices(light_curve, indices=[], return_tuple=False):
             eta : Ratio of mean square successive difference to the sample variance, von Neumann 1941
             continuum : the continuum magnitude
             sigma_mu : the ratio of the root-sample variance to the continuum magnitude
+            con : the number of clusters of 3 or more points brighter than 3-sigma from the baseline
         
         Parameters
         ----------
@@ -421,6 +435,9 @@ def compute_variability_indices(light_curve, indices=[], return_tuple=False):
     
     if "sigma_mu" in indices:
         idx_dict["sigma_mu"] = np.std(light_curve.mag) / continuum_mag[0]
+    
+    if "con" in indices:
+        idx_dict["con"] = consecutive_brighter(light_curve)
     
     if return_tuple:
         return tuple([idx_dict[idx] for idx in indices])

### 2012-08-15 survey_coverage.py

@@ -25,8 +25,6 @@ from matplotlib.patches import Rectangle
 import tables
 
 # Project
-# TODO: add some kind of catch here -- if this import fails, it's probably because I'm trying to
-#           run on the wrong machine...
 import ptf.photometricdatabase as pdb
 from ptf.globals import camera_size_degrees
 
@@ -351,20 +349,20 @@ def fields_to_coverage_plot(fields):
     for field in fields:
         kwargs = dict()
         if field.number_of_exposures >= 100:
-            kwargs["color"] = (252/255., 141/255., 89/255.)
-            kwargs["alpha"] = 0.75
+            kwargs["color"] = (136/255., 86/255., 167/255.)
+            kwargs["alpha"] = 0.8
             if r and field.id > 1000: 
                 kwargs["label"] = r"$\geq$100 observations"
                 r = False
         elif field.number_of_exposures >= 25:
-            kwargs["color"] = (145/255., 191/255., 219/255.)
-            kwargs["alpha"] = 0.75
+            kwargs["color"] = (158/255., 188/255., 218/255.)
+            kwargs["alpha"] = 0.85
             if b and field.id > 1000:
                 kwargs["label"] = r"$\geq$25 observations"
                 b = False
         else:
-            kwargs["color"] = "k"
-            kwargs["alpha"] = 0.25
+            kwargs["color"] = (0.8,0.8,0.8)
+            kwargs["alpha"] = 0.75
             if k and field.id > 1000:
                 kwargs["label"] = "<25 observations"
                 k = False
@@ -417,9 +415,11 @@ if __name__ == "__main__":
         # Make survey coverage plots
         R_cov_plot = fields_to_coverage_plot(R_info.fields(1))
         R_cov_plot.save("plots/survey_coverage/R_coverage.pdf", title="PTF R-band survey coverage")
+        R_cov_plot.save("plots/survey_coverage/R_coverage.png", title="PTF R-band survey coverage")
         
         g_cov_plot = fields_to_coverage_plot(g_info.fields(1))
         g_cov_plot.save("plots/survey_coverage/g_coverage.pdf", title="PTF g-band survey coverage")
+        g_cov_plot.save("plots/survey_coverage/g_coverage.png", title="PTF g-band survey coverage")
     
     if args.dump_json:
         # dump json file for survey coverage site

### 2012-08-15 text/paper/ptf_microlensing_paper.tex
more work on paper, added some new references

--- 2012-08-07 text/ptf_microlensing_paper.tex
@@
 Under median seeing conditions (1.1$\arcsec$), observations in Mould $R$ or Sloan Digital Sky Survey (SDSS) $g$ achieve 2.0$\arcsec$ full-width half-maximum images and reach 5$\sigma$ magnitudes of $R \approx 21.0$ and $g \approx 21.3$ mag in a standard 60~s exposure \citep{nick2010}.
-As of 2012 May, the PTF footprint included $\sim$9800 (2500) deg$^2$ imaged $>$25 ($>$100) times in $R$ and $\sim$2400 (250) deg$^2$ imaged that often in $g$.
-
-[UPDATE NUMBERS FOR FOOTPRINT, INCLUDE FIGURE?]
+As of August 9, 2012, the PTF footprint included $\apwsim$10,100 (2600) deg$^2$ imaged $>$25 ($>$100) times in $R$ and $\apwsim$2200 (170) deg$^2$ imaged that often in $g$ (\figurename~\ref{fig:survey_footprint})\footnote{For an interactive visualization of the PTF survey coverage, see: [TODO: url]}.
 
 \section{Microlensing event recovery}
-Microlensing surveys typically use difference image analysis (cite Lupton?) to initially identify transient events and then analyze the candidate light curves to search for microlensing events (cite MACHO, OGLE).
-This method is essential for studying crowded fields such as the Galactic bulge, and also limits the number of light curves that must be searched for microlensing events, thus saving on computation time.
-This [method] also provides real-time detection of events so surveys can utilize networks of telescopes to provide follow-up observations of these transients.
-Our approach is to instead take the preexisting PTF light curve database over the entire survey footprint and develop fast, robust search algorithms to pick out microlensing event candidates from this massive data archive.
+Microlensing surveys typically use difference image analysis (cite Lupton?) to identify transient events in raw imaging data.
+Light curves of any transient sources are then analyzed and further pruned using a variety of selection methods to search for microlensing event candidates (cite MACHO, OGLE).
+Using difference imaging is essential for studying crowded fields such as the Galactic bulge where photometry is difficult (cite or figure?), and also limits the number of light curves that must be searched for microlensing events, thus saving on computation time.
+Pre-selection with difference imaging also enables fast, real-time detection of events, allowing dedicated microlensing surveys to obtain high-cadence follow-up observations of interesting sources.
+Such surveys also have the advantage of relatively uniform time sampling of the survey footprint over an observing season.
 
-In order to develop predictions for event rates and then eventually use observed event rates to infer [] we must understand our microlensing event recovery rate, or \textit{detection efficiency}, $\varepsilon$.
-The detection efficiency is commonly expressed as a function of microlensing event timescale and is therefore a function of survey parameters, such as cadence and baseline [FIGURE simulated detection efficiency].
-[What did MACHO do exactly?
-What does OGLE do?] The nature of the PTF survey means that different fields are observed with very different sampling patterns [FIGURE showing different sampling].
-We must compute the detection efficiency for each field individually.
+The PTF survey footprint (\figurename~\ref{fig:survey_footprint}) is not uniformly sampled -- spatially or temporally.
+Each field has a unique sampling pattern determined by (1) the sub-surveys within PTF, (2) the time of year, as some low-declination fields can only be imaged during part of the year, or (3) the priority of the field, which could  change at any time.
+As a result, PTF light curves often contain gaps, regions of high-cadence observations, and regions of low-cadence observations, leading to a massive data set of irregularly-sampled, time-domain photometry.
+\figurename~\ref{fig:sampling} shows six randomly-selected light curves (with their magnitudes removed) and demonstrates the varying cadences and coverage fractions that different fields may have over the same one-year baseline.
 
-We compute the detection efficiency with a Monte Carlo simulation over real data.
-For a particular field of interest, we randomly sample 11,000 light curves from the database with $>25$ ``good'' observations.
-[define ``good''].
-We compute a set of variability statistics for each light curve, we then iterate and add 100 different simulated microlensing events to each light curve and recompute the variability statistics after each iteration [Figures here?].
-The parameters for the random microlensing events are chosen as follows:
-\begin{align}
-	t_0&\sim\mathcal{U}[t_{min}, t_{max}] \\
-	u_0&\sim\mathcal{U}[0, 1] \\
-	\log_{10}t_E&\sim\mathcal{U}[0, 3] 
-\end{align}
-where $\mathcal{U}$ is a uniform distribution.
+--
 
-[How to compute detection efficiency?] [Specifics of our method, e.g. details about my code]
+Previous work has used a variety of survey-specific selection methods (as described in \citealt{alcock2000, wyrzykowski2009, hamadache2009, sumi2011}), but the general idea is to \begin{enumerate}
+	\item detect transient events using difference image analysis, 
+	\item require that any selected light curves have some number of consecutive points brighter than some threshold,
+	\item require that a microlensing model fit best describes the data using some cut on $\chi^2$,
+	\item require that the model event parameters have reasonable values.
+\end{enumerate}
+Applying the second cut above to irregularly sampled data is more challenging and more computationally expensive because we must account for gaps in our data; it isn't obvious that this same selection procedure will be as successful for the PTF data.
+Motivated by work on identifying and classifying general variability in light curve data (Shin M.-S.
+et al., other?), we set out to compare the \textit{detection efficiency} of performing microlensing event selection with a set of variability statistics.
 
-[Parts of the above text will get absorbed into sections below]
+\subsection{Event Selection} \label{sec:event_selection}
+[Describe the different selection methods]
 
-\subsection{Selecting candidate microlensing events}
 - Proper statistical modeling is too time consuming, so we first have to define some selection criteria to limit the sample (number of good observations, quality cuts, etc.)
 
-- Discuss $\Delta \chi^2$ method, downfalls (speed, not robust -- how do you define the threshold?)
+- Discuss $\Delta \chi^2$ method, downfalls (speed, not robust -- how do you define the threshold?).
+Show a plot of distribution of $\Delta \chi^2$ with and without simulated events, e.g.
+the distribution doesn't separate well, then show $\eta$ and how the peak of the distribution moves.
 
@@
 
-\subsection{Detection efficiency}
+\subsection{Detection Efficiency}
+The probability of detecting any given microlensing event depends on the parameters of the event, the survey properties (observing strategy, limiting magnitude, weather), and the nature of the stellar field (crowding).
+To compute the detection efficiency, $\varepsilon$, for a given selection criterium, we run Monte Carlo simulations to artificially add microlensing events to real data and evaluate how often these events are recovered by the cut.
+This analysis must be done on a field-by-field basis, however we can compare our selection methods on a sub-sample of representative cases.
+
+For a particular field, we randomly sample 1,000 light curves from each chip with $>25$ ``good'' observations (good observations are those that have no IPAC or SExtractor flags set [is there some IPAC paper to cite here?]).
+With this sample, we compute the set of variability statistics (\S\ref{sec:event_selection}), then iterate to add different simulated microlensing events to each light curve.
+We use a random number generator to only add events during $\apwsim 50\%$ of the iterations, which allows us to later quantify the false positive rate of our selection methods.
+For each iteration we recompute the variability statistics and store the event parameters.
+For each simulated event, the peak time, $t_0$, is drawn from a uniform distribution between the minimum and maximum observation times, the impact parameter, $u_0$, is drawn from a uniform distribution between 0 and 1, and the event timescale, $t_E$, is drawn from a log-uniform distribution between 1 and 1000 days.
+
+We apply our selection criteria to each simulated light curve using the variability statistics computed for the initial light curves (from \textit{before} we add simulated events), and then bin the selected data by the event timescale.
+
 - Previous surveys had fairly uniform time sampling, and smaller areas
@@
 - Show detection efficiency of each variability statistic
+
+%The parameters for the random microlensing events are chosen as follows: 
+%\begin{align}
+%	t_0&\sim\mathcal{U}[t_{min}, t_{max}] \\
+%	u_0&\sim\mathcal{U}[0, 1] \\
+%	\log_{10}t_E&\sim\mathcal{U}[0, 3] 
+%\end{align}
+%where $\mathcal{U}$ is a uniform distribution.
+
+% Below here is old text that needs a home!
+
+%The PTF database (as of August 2012) contains $\apwsim64,000,000$  ``good'' light curves (light curves with $>25$ exposures with low systematics), or an average of $\apwsim60,000$ light curves per field with $>25$ exposures.
+By our estimates, utilizing all cores on an 8-core 2.0 GHz computer, it would take $\apwsim 2$ weeks to perform this cut on the entire dataset using \inlinecode{ANSI-C} code wrapped in \inlinecode{Python}.
+%
+%- Previous work did this (describe OGLE paper's selection)
+%- When we applied this to our data, we found the selection criteria performed even worse
+%- Motivated by statistical work on identifying variability in light curve databases, we sought to find a better selection method
+%
+%Our approach for finding microlensing events is to use the preexisting PTF light curve database over the entire survey footprint and develop fast, robust search algorithms to pick out candidates from this data archive.
+The average number of light curves per PTF field with $>$25 ``good'' observations is $\apwsim 50,000$, with $\apwsim64,000,000$  ``good'' light curves in total.
+By our estimates, utilizing all cores on an 8-core 2.0 GHz compute, it would take [TODO: estimate this number] to fit even the simplest 5-parameter microlensing event model to each of these light curves using \inlinecode{ANSI-C} wrapped in \inlinecode{Python}.
+Previous studies have had the advantage of pre-selecting candidates via difference image analysis, automatically limiting their sample to $\apwsim$thousands of light curves.
+Past work has also utilized a variety of selection techniques for flagging candidate microlensing light curves after these steps, however most of these methods still rely on fitting a point-source, point-lens microlensing model to each light curve and performing some selection based on the $\chi^2$ value of this fit (cite Sumi et al.
+2006, MACHO work, ?).
+More recent work has begun to apply machine learning algorithms to the general light curve classification problem (cite Joey Richards), however this is still a work in progress.
+We set out to find a fast, robust selection method for identifying microlensing events in big data archives by using a set of variability statistics previously applied to finding variable stars in general (Shin M.-S.
+et al.
+).
+%
+%[Describe variability indices?]
+%
+%In order to develop predictions for event rates and then eventually use observed event rates to infer [] we must understand our microlensing event recovery rate, or \textit{detection efficiency}, $\varepsilon$.
+The detection efficiency is commonly expressed as a function of microlensing event timescale and is therefore a function of survey parameters, such as cadence and baseline [FIGURE simulated detection efficiency].
+[What did MACHO do exactly?
+What does OGLE do?] The nature of the PTF survey means that different fields are observed with very different sampling patterns [FIGURE showing different sampling].
+We must compute the detection efficiency for each field individually.
+%
+%We compute the detection efficiency with a Monte Carlo simulation over real data.
+For a particular field of interest, we randomly sample 11,000 light curves from the database with $>25$ ``good'' observations.
+[define ``good''].
+We compute a set of variability statistics for each light curve, we then iterate and add 100 different simulated microlensing events to each light curve and recompute the variability statistics after each iteration [Figures here?].
+The parameters for the random microlensing events are chosen as follows:
+%\begin{align}
+%	t_0&\sim\mathcal{U}[t_{min}, t_{max}] \\
+%	u_0&\sim\mathcal{U}[0, 1] \\
+%	\log_{10}t_E&\sim\mathcal{U}[0, 3] 
+%\end{align}
+%where $\mathcal{U}$ is a uniform distribution.
+%
+%% Talk about hypothetical selection criteria, and the things we have to consider for detection efficiency (possible selection criteria, optimizing selection method, selecting candidates)
+%
+%[How to compute detection efficiency?] [Specifics of our method, e.g. details about my code]
+
+[Parts of the above text will get absorbed into sections below]
 
@@
 This paper is based on observations obtained with the Samuel Oschin Telescope as part of the Palomar Transient Factory project, a scientific collaboration between the California Institute of Technology, Columbia University, Las Cumbres Observatory, the Lawrence Berkeley National Laboratory, the National Energy Research Scientific Computing Center, the University of Oxford and the Weizmann Institute of Science.
+Work by A.
+M.
+Price-Whelan is supported by a National Science Foundation Graduate Research Fellowship under grant No.
+<Grant Number>.
 
@@
 %\bibliography{apj-jour,refs}
-%
-%\begin{figure}
-%	\centering
-%	\caption{Computed detection efficiency as a function of number of observations for a fixed baseline of 365 days. This plot uses entirely simulated light curves.}
-%    \includegraphics[width=1.0\textwidth]{figures/detection_efficiency_figure.pdf}
-%    \label{fig:detection_eff}
-%\end{figure}	
-%
-%\begin{figure}
-%	\centering
-%	\caption{A visualization of the different sampling regimes of six different PTF fields over the same one year baseline.
+
+\begin{figure}
+	\centering
+	\caption{The PTF survey footprint shown in equatorial coordinates. Each field's color corresponds to some minimum number of observations of that field. }
+    \includegraphics[angle=270, scale=0.55]{figures/R_coverage.pdf}
+    \label{fig:survey_footprint}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\caption{A visualization of the different sampling regimes of six different PTF fields over the same one year baseline.
 Darker points mean more data (more exposures).
@@
 }
-%    \includegraphics[width=1.0\textwidth]{figures/ptf_sampling_figure.pdf}
-%    \label{fig:light_curves}
-%\end{figure}
+    \includegraphics[width=1.0\textwidth]{figures/sampling_figure.pdf}
+    \label{fig:sampling}
+\end{figure}
 %

### 2012-08-17 var_indices_figure.py
add var indices script, not complete

@@ -1,0 +1,413 @@
+# coding: utf-8
+from __future__ import division
+
+""" This module contains routines for generating the variability indices figures, with and 
+    without simulated events
+"""
+
+__author__ = "adrn <adrn@astro.columbia.edu>"
+
+# Standard library
+import os, sys
+import logging
+import time
+import cPickle as pickle
+import json
+
+# Third-party
+import matplotlib
+matplotlib.use("Agg")
+import matplotlib.cm as cm
+import matplotlib.colors as mc
+import matplotlib.pyplot as plt
+import numpy as np
+from apwlib.globals import greenText, redText
+
+# Project
+import ptf.photometricdatabase as pdb
+import detectionefficiency as de
+import ptf.analyze.analyze as analyze
+from ptf.globals import index_to_label
+
+# Create logger for this module
+logger = logging.getLogger(__name__)
+ch = logging.StreamHandler()
+formatter = logging.Formatter("%(name)s / %(levelname)s / %(message)s")
+ch.setFormatter(formatter)
+logger.addHandler(ch)
+
+def get_var_indices(field, light_curves_per_ccd, events_per_light_curve, indices, overwrite=True):
+    """ This function will create a len(indices) x len(indices) plot grid and plot
+        distributions of all of the variability indices
+    """
+
+    limiting_mags = [14.3, 21]
+    
+    file_base = "field{:06d}_Nperccd{}_Nevents{}".format(field.id, light_curves_per_ccd, events_per_light_curve) + ".{ext}"
+    pickle_filename = os.path.join("data", "var_indices", file_base.format(ext="pickle"))
+    plot_filename = os.path.join("plots", "var_indices", file_base.format(ext="png"))
+    
+    if not os.path.exists(os.path.dirname(pickle_filename)):
+        os.mkdir(os.path.dirname(pickle_filename))
+    
+    if not os.path.exists(os.path.dirname(plot_filename)):
+        os.mkdir(os.path.dirname(plot_filename))
+
+    if os.path.exists(pickle_filename) and overwrite:
+        logger.debug("Data file exists, but you want to overwrite it!")
+        os.remove(pickle_filename)
+        logger.debug("Data file deleted...")
+
+    # If the cache pickle file doesn't exist, generate the data
+    if not os.path.exists(pickle_filename):
+        logger.info("Data file {} not found. Generating data...".format(pickle_filename))
+        
+        # Conditions for reading from the 'sources' table
[…]

### 2012-08-17 text/paper/ptf_microlensing_paper.tex

@@
 
-\section{Microlensing event recovery}
+\section{Microlensing event recovery} \label{sec:event_recovery}
 Microlensing surveys typically use difference image analysis (cite Lupton?) to identify transient events in raw imaging data.
@@
 Applying the second cut above to irregularly sampled data is more challenging and more computationally expensive because we must account for gaps in our data; it isn't obvious that this same selection procedure will be as successful for the PTF data.
-Motivated by work on identifying and classifying general variability in light curve data (Shin M.-S.
-et al., other?), we set out to compare the \textit{detection efficiency} of performing microlensing event selection with a set of variability statistics.
+Motivated by work on identifying and classifying general variability in light curve data \citep{shin2009}[others?], we set out to compare the \textit{detection efficiency} of performing microlensing event selection with a set of variability statistics to previously used methods.
 
 \subsection{Event Selection} \label{sec:event_selection}
+We choose five statistical measures of variability, as compiled by \cite{shin2009}, that have been previously applied to classification of periodic variables, and evaluate their effectiveness in identifying microlensing events.
+The statistics, or \emph{variability indices}, are $\sigma/\mu$, $Con$, $\eta$, $J$, and $K$ (we choose not to implement the sixth, $AoVM$, because it mainly helps with periodic sources).
+$\sigma/\mu$ is the ratio of the root-variance to the sample mean,
+\begin{align}
+	\frac{\sigma}{\mu} = \frac{\sqrt{\sum^N_i (x_i - \mu)^2 / (N-1)}}{\sum^N_i x_i/N},
+\end{align}
+where N is the total number of observations in the light curve.
+We modify their definition of $Con$ such that it is the number of clusters of three or more consecutive observations that are more than $3\sigma$ brighter than the reference magnitude of the source (e.g., for a single microlensing event in an otherwise flat light curve this value should be $Con=1$).
+This change allows us to use the performance of $Con$ as a measure of Cut 2 (\S\ref{sec:event_recovery}).
+$\eta$ is the von Neumann ratio, originally proposed in \cite{von_neumann1941}, and is a ratio of the mean square successive difference over the sample variance:
+\begin{align}
+	\eta = \frac{\delta^2}{\sigma^2} = \frac{\sum^{N-1}_i(x_{i+1} - x_i)^2/(N-1)}{\sigma^2}
+\end{align}
+$\eta$ is small, but still greater than zero, when there is strong serial correlation between successive data points. $J$ and $K$ were suggested by \cite{stetson1996}:
+\begin{align}
+	\delta_i &= \sqrt{\frac{N}{N-1}}\frac{x_i-\mu}{e_i}\\
+	J &= \sum^{N-1}_i sign(\delta_i \delta_{i+1})\sqrt{|\delta_i \delta_{i+1}|}\\
+	K &= \frac{1/N\sum^N_i |\delta_i|}{\sqrt{1/N\sum^N_i\delta_i^2}}
+\end{align}
+where $e_i$ is the photometric error of each data point.
+$J$ tends to 0 for non-variable stars, but is large when there are significant differences between successive data points in a light curve.
+$K$ is a measure of the kurtosis or ``peakedness'' of the distribution of points in the light curve.
+We add one more variability index motivated by Cut 3 (\S\ref{sec:event_recovery}): $\Delta \chi^2$ is defined as the difference in chi-squared between fitting a five-parameter Gaussian model and fitting a linear model to the light curve.
+
 [Describe the different selection methods]

### 2012-08-17 var_indices_figure.py
add var indices script, not complete

@@ -256,6 +256,8 @@ class VIFigure(object):
     def save(self, filename):
         """ Save the figure to the specified filename """
 
+
+
 def make_var_indices_plots(var_indices, var_indices_with_events, indices, filename_base):
     """ """
     
@@ -397,17 +399,5 @@ if __name__ == "__main__":
                                                light_curves_per_ccd=args.limit,
                                                overwrite=args.overwrite)
     
-    for ii,mag in enumerate(args.limiting_mag[:-1]):
-        filename_base = "plots/var_indices/field{}_mag{}".format(field.id, mag)
-        
-        idx1 = np.ones_like(var_indices["m"]).astype(bool)
-        idx2 = np.ones_like(var_indices_with_events["m"]).astype(bool)
-        
-        if mag != None:
-            idx1 = idx1 & (var_indices["m"] > mag) & (var_indices["m"] < args.limiting_mag[ii+1])
-            idx2 = idx2 & (var_indices_with_events["m"] > mag) & (var_indices_with_events["m"] < args.limiting_mag[ii+1])
-            
-        make_var_indices_plots(var_indices[idx1], 
-                               var_indices_with_events[idx2],
-                               indices=args.indices,
-                               filename_base=filename_base)
+    vifigure = VIFigure(var_indices, indices=args.indices, scatter=False, figsize=(15,15))
+    plt.show()

### 2012-08-17 var_indices_figure.py
finished variability index plot script

@@ -16,7 +16,7 @@ import json
 
 # Third-party
 import matplotlib
-matplotlib.use("Agg")
+#matplotlib.use("Agg")
 import matplotlib.cm as cm
 import matplotlib.colors as mc
 import matplotlib.pyplot as plt
@@ -184,43 +184,46 @@ def var_indices_to_json(var_indices, var_indices_with_events, filename=None):
         return blob
 ##############
 
+import scipy.optimize as so
+def find_confidence_interval(x, pdf, confidence_level):
+    return pdf[pdf > x].sum() - confidence_level
+
 class VIFigure(object):
     """ This class represents a variability index (VI) figure where we plot the distributions
         of the variability indices against each other.
     """
     
-    def __init__(self, variability_indices, indices, scatter=True, **figure_kwargs):
-        """ Given a numpy structured array containing variability index values, 
-            initialize a new variability index figure
-            
-            Parameters
-            ----------
-            variability_indices : numpy structured array
-                This must be a numpy structured array with at minimum the following columns:
-                    - one column for each index in 'indices',
-                    - m (the reference magnitude of the source),
-                    - u0 (the impact parameter of any microlensing event),
-                    - tE (the microlensing event timescale)
-            indices : iterable
-                Probably a list of strings, e.g. ['eta', 'j', 'k'] etc.
-            scatter: bool
-                Make a scatter plot. If False, it will make a contour plot instead.
-            
-            figure_kwargs are any keyword arguments that will get passed through to the matplotlib figure function.
-        """
+    def _validate_variability_indices(self, variability_indices):
+        """ Validate any variability indices given to this object """
         
         # Validate input
-        for var_idx in indices:
+        for var_idx in self.indices:
             try:
                 variability_indices[var_idx]
             except ValueError: # e.g. field not found
                 print redText("Variability statistic '{}' not found in the data you provided.".format(var_idx))
                 raise
         
-        self.variability_indices = variability_indices
+        return True
+    
+    def __init__(self, indices, **figure_kwargs):
+        """ Given a numpy structured array containing variability index values, 
+            initialize a new variability index figure
+            
+            Parameters
+            ----------
+            indices : iterable
[…]

### 2012-08-17 detectionefficiency.py
fix typos

@@ -149,10 +149,7 @@ def compute_detection_efficiency(var_indices, var_indices_with_events, indices):
     return data
 
 def compare_detection_efficiencies_on_field(field, light_curves_per_ccd, events_per_light_curve, overwrite=False, indices=["j","k","eta","sigma_mu","delta_chi_squared"], u0s=[], limiting_mags=[]):
-    
-    # TODO:
-    #   - Document the new code!
-    #   - Clean up / delete old stuff
+    """ TODO: document """   
     
     if u0s == None or len(u0s) == 0:
         u0s = [None]
@@ -163,6 +160,12 @@ def compare_detection_efficiencies_on_field(field, light_curves_per_ccd, events_
     file_base = "field{:06d}_Nperccd{}_Nevents{}_u0_{}_m{}".format(field.id, light_curves_per_ccd, events_per_light_curve, "-".join(map(str,u0s)), "-".join(map(str,limiting_mags))) + ".{ext}"
     pickle_filename = os.path.join("data", "detectionefficiency", file_base.format(ext="pickle"))
     plot_filename = os.path.join("plots", "detectionefficiency", file_base.format(ext="png"))
+    
+    if not os.path.exists(os.path.dirname(pickle_filename)):
+        os.mkdir(os.path.dirname(pickle_filename))
+    
+    if not os.path.exists(os.path.dirname(plot_filename)):
+        os.mkdir(os.path.dirname(plot_filename))
     
     if os.path.exists(pickle_filename) and overwrite:
         logger.debug("Data file exists, but you want to overwrite it!")
@@ -295,14 +298,18 @@ def compare_detection_efficiencies_on_field(field, light_curves_per_ccd, events_
     
     num_u0_bins = len(u0s)
     num_mag_bins = len(limiting_mags)-1
-    fig, axes = plt.subplots(num_u0_bins, num_mag_bins, sharex=True, sharey=True, figsize=(30,30))
+    fig, axes = plt.subplots(num_u0_bins, num_mag_bins, sharex=True, sharey=True, figsize=(15,15))
     
     for ii, limiting_mag_pair in enumerate(sorted(var_indices.keys())):
         selection_indices = var_indices[limiting_mag_pair]
         for jj, u0 in enumerate(sorted(var_indices_with_events[limiting_mag_pair].keys())):
             data = compute_detection_efficiency(selection_indices, var_indices_with_events[limiting_mag_pair][u0], indices)
             
-            ax = axes[ii, jj]
+            try:
[…]

### 2012-08-27 paper_figures.py
added script for making figures that dont fit elsewhere

@@ -1,0 +1,106 @@
+# coding: utf-8
+from __future__ import division
+
+""" This module contains routines for generating figures for the paper that don't really
+    belong anywhere else
+"""
+
+__author__ = "adrn <adrn@astro.columbia.edu>"
+
+# Standard library
+import sys
+import logging
+import time
+
+# Third-party
+import matplotlib
+#matplotlib.use("TkAgg")
+import matplotlib.cm as cm
+import matplotlib.colors as mc
+import matplotlib.pyplot as plt
+import numpy as np
+
+# Project
+import ptf.photometricdatabase as pdb
+import ptf.analyze.analyze as pa
+import survey_coverage
+
+# Create logger for this module
+logger = logging.getLogger(__name__)
+ch = logging.StreamHandler()
+formatter = logging.Formatter("%(name)s / %(levelname)s / %(message)s")
+ch.setFormatter(formatter)
+logger.addHandler(ch)
+
+def make_survey_sampling_figure(N=10):
+    """ Generate the survey sampling figure that shows how different the
+        time sampling is for different light curves in PTF.
+    """
+    
+    np.random.seed(42)
+    
+    """
+    R_info = survey_coverage.SurveyInfo("R")
+    fields = R_info.fields(100)
+    for ii in range(N):
+        idx = np.random.randint(len(fields))
+        rand_fields.append(fields[idx])
+    """
+    field_ids = [100031, 2794, 4252, 100085, 110002, 3685]
+    rand_fields = [pdb.Field(f, "R") for f in field_ids]
+    
+    fig = plt.figure(figsize=(10,10))
+    ax = fig.add_subplot(111)
+    
+    min_mjd = 55250
+    max_mjd = min_mjd+365
+    for ii,field in enumerate(rand_fields):
+        ccd_idx = field.ccds.keys()[np.random.randint(len(field.ccds))]
+        ccd = field.ccds[ccd_idx]
+        chip = ccd.read()
+        
+        source_ids = chip.sources.readWhere("nobs > 25")["matchedSourceID"]
+        np.random.shuffle(source_ids)
+        source_id = source_ids[0]
+        
+        logger.info("Reading source {} on {}, {}".format(source_id, field, ccd))
+        light_curve = ccd.light_curve(source_id, clean=False, barebones=True, where=["(mjd >= {})".format(min_mjd), "(mjd <= {})".format(max_mjd)])
+    
[…]

### 2012-08-28 event_fitter.py
modified emcee based event fitter

@@ -139,7 +139,7 @@ def ln_prior(p, mag, mjd):
     m0, u0, t0, tE = p
     
     p_u0 = ln_p_u0(u0)
-    p_t0 = ln_p_t0(t0, mjd.min(), mag.min())
+    p_t0 = ln_p_t0(t0, mjd.min(), mjd.max())
     p_tE = ln_p_tE(tE)
     p_m0 = ln_p_m0(m0, np.median(mag), np.std(mag))
     
@@ -175,6 +175,49 @@ def test_ln_prior():
     print "Bad tE:", ln_prior([15, 0.3, 50., 0.2], light_curve.mag, light_curve.mjd)
     print "Bad tE:", ln_prior([15, 0.3, 50., 2000.], light_curve.mag, light_curve.mjd)
 
+def fit_model_to_light_curve(light_curve, nwalkers=200, nsamples=250):
+    """ Fit a microlensing model to a given light curve using Emcee
+        
+    """
+    
+    p0 = np.array([[np.random.normal(np.median(light_curve.mag), np.std(light_curve.mag)),
+                    np.random.uniform(0., 1.34),
+                    np.random.normal(light_curve.mjd[np.argmin(light_curve.mag)], 50),
+                    10**np.random.uniform(0., 3.)] for ii in range(nwalkers)])
+    ndim = p0.shape[1]
+    
+    sampler = emcee.EnsembleSampler(nwalkers, ndim, ln_posterior, 
+                                    args=[light_curve.mag, light_curve.mjd, light_curve.error], 
+                                    threads=4)
+    pos, prob, state = sampler.run_mcmc(p0, 100)
+    sampler.reset()
+    
+    sampler.run_mcmc(pos, nsamples)
+    
+    logger.info("Mean acceptance fraction: {0:.3f}".format(np.mean(sampler.acceptance_fraction)))
+    
+    plt.figure()
+    for i,name in enumerate(["m0", "u0", "t0", "tE"]):
+        plt.subplot(2,2,i+1)
+        plt.hist(sampler.flatchain[:,i], 100, color="k", histtype="step")
+        if name != "m0": plt.axvline(getattr(light_curve, name), color='r')
+        plt.title("{}".format(name))
+        plt.ylabel("Posterior Probability")
+    
+    fig = plt.figure()
+    ax = fig.add_subplot(111)
[…]

### 2012-08-28 paper_figures.py
new intro figures for paper

@@ -10,6 +10,7 @@ __author__ = "adrn <adrn@astro.columbia.edu>"
 # Standard library
 import sys
 import logging
+import cPickle as pickle
 import time
 
 # Third-party
@@ -20,17 +21,21 @@ import matplotlib.colors as mc
 import matplotlib.pyplot as plt
 import numpy as np
 
-# Project
-import ptf.photometricdatabase as pdb
-import ptf.analyze.analyze as pa
-import survey_coverage
-
 # Create logger for this module
 logger = logging.getLogger(__name__)
 ch = logging.StreamHandler()
 formatter = logging.Formatter("%(name)s / %(levelname)s / %(message)s")
 ch.setFormatter(formatter)
 logger.addHandler(ch)
+
+# Project
+from ptf.simulation.simulatedlightcurve import SimulatedLightCurve
+try:
+    import ptf.photometricdatabase as pdb
+    import ptf.analyze.analyze as pa
+    import survey_coverage
+except ImportError:
+    logger.warning("photometric database modules failed to load! If this is on Navtara, you made a boo-boo.")
 
 def make_survey_sampling_figure(N=10):
     """ Generate the survey sampling figure that shows how different the
@@ -82,6 +87,102 @@ def make_survey_sampling_figure(N=10):
     ax.set_xlabel("Time [day]", fontsize=20)
     fig.savefig("plots/sampling_figure.pdf", facecolor="white", bbox_inches="tight")
 
+def microlensing_event_sim():
+    """ Create the multi-panel figure with simulated microlensing events for a single
+        'typical' PTF light curve.
+    """
+    with open("data/paper_figures/sample_lightcurve.pickle") as f:
+        light_curve = pickle.load(f)
+    
+    num = 4
+    fig, axes = plt.subplots(num,1, sharex=True, figsize=(11,15))
+    
+    sim_light_curve = SimulatedLightCurve(mjd=light_curve.mjd, mag=light_curve.mag, error=light_curve.error)
+    
+    t0 = sim_light_curve.mjd[int(len(sim_light_curve.mjd)/2)]
+    
+    kwarg_list = [None, {"u0" : 1.0, "t0" : t0, "tE" : 20},
+                        {"u0" : 0.5, "t0" : t0, "tE" : 20}, 
+                        {"u0" : 0.01, "t0" : t0, "tE" : 20}]
+    
+    args_list = [(17.3, "a)"), (17., "b)"), (16.6, "c)"), (13, "d)")]
[…]

### 2012-08-28 ptf/simulation/simulatedlightcurve.py

@@ -1,3 +1,4 @@
+import logging
 import copy
 import numpy as np
 from ..ptflightcurve import PTFLightCurve
@@ -15,6 +16,13 @@ def u_t(t, u_0, t_0, t_E):
 
 def A_u(u):
     return (u**2 + 2) / (u*np.sqrt(u**2 + 4))
+    
+def A(t, u0, t0, tE):
+    u = u_t(t, u0, t0, tE)
+    return (u**2 + 2) / (u*np.sqrt(u**2 + 4))
+
+def magnitude_model(t, mag, **p):
+    return mag - 2.5*np.log10(A(t, p["u0"], p["t0"], p["tE"]))
 
 def fluxModel(t, **p):
     return p["F0"]*A_u(u_t(t, p["u0"], p["t0"], p["tE"]))
@@ -42,12 +50,19 @@ class SimulatedLightCurve(PTFLightCurve):
         
         if error != None:
             self.error = np.array(error)
+            if len(self.error) == 1:
+                self.error = np.zeros_like(self.mjd) + error
+            elif len(self.error) != len(self.mjd):
+                raise ValueError("Error array should have same shape as mjd")
         else:
             # TODO: Implement this
             raise NotImplementedError("TODO!")
         
-        if mag != None:
+        if isinstance(mag, np.ndarray) or isinstance(mag, list):
             self.amag = self.mag = np.array(mag)
+        elif isinstance(mag, int) or isinstance(mag, float):
+            self.amag = self.mag = np.zeros(len(mjd)) + mag
+            self._addNoise()
         else:
             if outliers:
                 # Add ~1% outliers
@@ -106,11 +121,10 @@ class SimulatedLightCurve(PTFLightCurve):
         if tE == None: self.tE = 10**np.random.normal(1.3, 0.5)
         else: self.tE = float(tE)
         
-        flux = fluxModel(self.mjd, u0=self.u0, t0=self.t0, tE=self.tE, F0=1.)#self.F0)
-        self.mag = FluxToRMag(flux*RMagToFlux(self.mag))
-        
-        self.amag = self.mag
+        #flux = fluxModel(self.mjd, u0=self.u0, t0=self.t0, tE=self.tE, F0=1.)#self.F0)
+        self.amag = self.mag = magnitude_model(self.mjd, self.mag, u0=self.u0, t0=self.t0, tE=self.tE)
     
     def _addNoise(self):
         """ Add scatter to the light curve """
[…]

### 2012-08-28 text/paper/ptf_microlensing_paper.tex
text updates and new referenceS

@@
 \begin{abstract}
+[APW: update the abstract..]
 Many current photometric, time-domain surveys are driven by specific goals, such as supernova searches, transiting exoplanet discoveries, or stellar variability studies, which set the cadence with which individual fields get re-imaged.
@@
 \section{Introduction}
-Since Einstein first studied gravitational microlensing (cite), the phenomena has been used to study dark objects, probe galactic structure and kinematics, and discover extrasolar planets.
-Microlensing events were once relatively rare, but with advances in CCD technology and the help of several dedicated microlensing surveys (), $>500$ lensing events are now detected per observing season.
 
-[MORE MICROLENSING INTRODUCTION]
+\subsection{Overview}
+General relativity tells us that mass alters the propagation of light.
+In particular, light from a background source passing some mass distribution will be deflected by an angle that depends on the distance between the source and mass distribution, as well as the structure and kinematics of the mass distribution itself.
+For an observer aligned with a source and a point lensing mass, the relevant angular scale for the observed deflection is given by the Einstein ring radius:
+\begin{equation} \label{eq:einstein_radius}
+	\theta_E = \sqrt{\frac{4GM}{c^2}\frac{D_{S} - D_L}{D_L D_S}}
+\end{equation}
+where $D_L$ is the distance to the lensing object, $D_S$ is the distance to the source, and $M$ is the mass of the lens.
 
-[QUICK INTRO TO THEORY AND JARGON]
+Microlensing designates any unresolvable gravitational lensing, where the images formed by the lens just contribute to an increase in flux.
+For Galactic distance and mass scales, the Einstein radius (Eq.~\ref{eq:einstein_radius}) can be expressed as:
+\begin{align}
+	\theta_E &\approx 550\mathrm{\mu as}~\sqrt{\frac{1-D_L/D_S}{D_L/D_S}}~\Big(\frac{D_S}{8~\mathrm{kpc}}\Big)^{-1/2} \Big(\frac{M}{0.3~M_\sun}\Big)^{1/2}
+\end{align}
+\citep{paczynski1986, gaudi2011}.
+A solar-mass star in the bulge lensing another star in the bulge would have an Einstein radius of about $\theta_E \sim 0.3~\mathrm{mas}$, well below the resolution limit of even our most powerful telescopes.
+
+Since Einstein first mentioned gravitational microlensing \citep{einstein1936}, it has been applied to the search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, galactic structure and kinematics \citep{binney2000}, and extrasolar planets \citep{}[APW: cite].
+Microlensing events were once a relatively rare phenomena, but with advances in CCD technology and the help of several dedicated microlensing surveys [APW: cite], more than [APW: check number]500 events are detected each year by several microlensing collaborations and surveys.
+Within the next decade, this number is expected to [] as the next generation of photometric, time-domain surveys come online [APW: cite LCOGT, LSST, ??].
+
+[APW: Some brief description of what is PTF, why it is the bridge between what has been done and what will be done, etc...]
+
+\subsection{Something??}
+Assuming that the source is an un-blended point source and the lens is a foreground, dim object, a microlensing event is fully described by three parameters: the angular impact parameter $u_0$, the peak time of the event $t_0$, and the timescale of the event (Einstein crossing time) $t_E$ such that the increase in flux due to source magnification is given by:
+\begin{align}
+	F(t) &= A(t)\times F_{source} \\
+	A(u) &= \frac{u^2 + 2}{u\sqrt{u^2 + 4}}\\
+	u(t) &= \sqrt{u_0^2 + 2\Big(\frac{t-t_0}{t_E}\Big)}
+\end{align}
+where $A(t)$ is the amplification factor.
+
+The microlensing perturbation can also be expressed in terms of magnitudes as:
+\begin{align}
+	m(t) &= m_0(t) - 2.5\log A(t)
+\end{align}
+where $m_0(t)$ is the unperturbed, possibly time-variable magnitude of the source.
+
+%However, if the event is \textit{too} bright, the pixels will saturate and there could be photometric errors flagging the most interesting data as bad.
+When $u_0$ is small ($u_0<<1$), the amplification is large.
+If $t_E$ is short ($t_E \lesssim\mathrm{days}$), surveys may miss or poorly sample the event, but if $t_E$ is long ($t_E \gtrsim 1~\mathrm{year}$) the event may be confused with other forms of long-duration variability.
+\figurename~\ref{fig:microlensing_sim} shows a ``typical'' PTF light curve and demonstrates what microlensing events with $t_E=20~\mathrm{days}$ and varying $u_0$ may look like in the data.
+
+The microlensing event rate in some particular region of sky depends on the space and velocity distributions of sources and on the space, velocity, and mass distribution of lenses.
+To maximize the event rate, microlensing surveys have typically focused their observations towards high stellar density regions such as the Galactic Bulge, M31, and the Magellanic clouds.
+Microlensing events do occur away from stellar population centers, but without a large-area, time-domain search of these regions, only one such event has been recorded ([APW: cite Fukui 2007, Gaudi 2008 RE: Tago event).
+[APW] Gaudi 2008 were able to establish probabilistic limits on the mass, distance, proper motion, and magnitude of this lens using existing observations near the location of the event.
+With JHK magnitudes of the lens itself or a direct measurement of the proper velocity of the lens, these limits could be turned into precise measurements.
+Such a detailed analysis is not normally possible for events in the previously mentioned high-density regions where the lenses are much farther away, and blending makes precise photometry difficult.
+
+Not all events away from the bulge will be as fortunate as the Tago event, but lensing events away from stellar population centers are generally expected to have closer sources, closer lenses, and less crowded background fields compared to  those found by modern microlensing surveys ([APW: cite Di Stefano mesolensing]).
+Thus, the properties of the lenses are more likely to be well-constrained by direct observation, possibly enabling mass measurements of objects not bound in binary pairs.
+The event rate in low stellar density regions will certainly be smaller than the event rate near the Bulge \citep[e.g.,][]{wood_optical_depth, ogle_optical_depth, macho_optical_depth, eros_optical_depth}, but the opportunity for precise, direct observations of lenses would be extremely valuable for studying low-mass stars, substellar objects, and stellar remnants such as isolated neutron stars [APW: cite NS work?].
 
@@
 Light curves of any transient sources are then analyzed and further pruned using a variety of selection methods to search for microlensing event candidates (cite MACHO, OGLE).
-Using difference imaging is essential for studying crowded fields such as the Galactic bulge where photometry is difficult (cite or figure?), and also limits the number of light curves that must be searched for microlensing events, thus saving on computation time.
+Using difference imaging is essential for studying crowded fields such as the Galactic bulge where photometry is difficult (cite or figure?), and also limits the number of light curves that must be analyzed, thus saving on computation time.
 Pre-selection with difference imaging also enables fast, real-time detection of events, allowing dedicated microlensing surveys to obtain high-cadence follow-up observations of interesting sources.
@@
 	\item detect transient events using difference image analysis, 
-	\item require that any selected light curves have some number of consecutive points brighter than some threshold,
+	\item require that any selected light curves have some number of consecutive datapoints brighter than some threshold,
 	\item require that a microlensing model fit best describes the data using some cut on $\chi^2$,
@@
 \end{enumerate}
-Applying the second cut above to irregularly sampled data is more challenging and more computationally expensive because we must account for gaps in our data; it isn't obvious that this same selection procedure will be as successful for the PTF data.
-Motivated by work on identifying and classifying general variability in light curve data \citep{shin2009}[others?], we set out to compare the \textit{detection efficiency} of performing microlensing event selection with a set of variability statistics to previously used methods.
+Applying the second cut above to irregularly sampled data is more challenging and more computationally expensive because we must account for gaps in our data.
+Furthermore, it isn't obvious that this same selection procedure will be as successful for the PTF data.
+Motivated by work on identifying and classifying general variability in light curve data \citep{shin2009}[others?], we set out to compare the performance of microlensing event selection using a set of variability statistics as an initial cut in our candidate detection pipeline.
 
 \subsection{Event Selection} \label{sec:event_selection}
-We choose five statistical measures of variability, as compiled by \cite{shin2009}, that have been previously applied to classification of periodic variables, and evaluate their effectiveness in identifying microlensing events.
-The statistics, or \emph{variability indices}, are $\sigma/\mu$, $Con$, $\eta$, $J$, and $K$ (we choose not to implement the sixth, $AoVM$, because it mainly helps with periodic sources).
+We choose five statistical measures of variability, as compiled by \cite{shin2009}, that have been previously applied to classification of periodic variables, and evaluate their effectiveness in recovering simulated microlensing events in the data.
+These statistics, or \emph{variability indices}, are $\sigma/\mu$, $Con$, $\eta$, $J$, and $K$ (we choose not to implement the sixth, $AoVM$, because it mainly helps with periodic sources).
 $\sigma/\mu$ is the ratio of the root-variance to the sample mean,
@@
 We modify their definition of $Con$ such that it is the number of clusters of three or more consecutive observations that are more than $3\sigma$ brighter than the reference magnitude of the source (e.g., for a single microlensing event in an otherwise flat light curve this value should be $Con=1$).
-This change allows us to use the performance of $Con$ as a measure of Cut 2 (\S\ref{sec:event_recovery}).
+This change allows us to use the performance of $Con$ as a proxy for Cut 2 (\S\ref{sec:event_recovery}).
 $\eta$ is the von Neumann ratio, originally proposed in \cite{von_neumann1941}, and is a ratio of the mean square successive difference over the sample variance:
@@
 We add one more variability index motivated by Cut 3 (\S\ref{sec:event_recovery}): $\Delta \chi^2$ is defined as the difference in chi-squared between fitting a five-parameter Gaussian model and fitting a linear model to the light curve.
+\figurename~\ref{fig:indices_examples} shows examples of a median light curve for each variability statistic (left panels) along with the most extreme outliers (right panels) for XXX light curves on a single CCD on PTF field YYY.
+
+\figurename~\ref{fig:scatter_plot_matrix} shows a scatter-plot matrix of all six of our variability statistics.
+In this figure, we take 10,000 light curves from a single PTF field (field id = ), compute the variability indices for each light curve, and then plot $1\sigma$ (red) and $2\sigma$ (green) contours in each panel for the resultant distributions.
+We then add 10 simulated microlensing events to each light curve, recompute the variability indices, and plot these values as black points.
 
@@
 This analysis must be done on a field-by-field basis, however we can compare our selection methods on a sub-sample of representative cases.
+Performing these simulations with real data allows us to compute the ...
+[APW: explain difference between two different detection efficiencies (of selection method vs.
+of survey parameters), and how we marginalize over both].
 
-For a particular field, we randomly sample 1,000 light curves from each chip with $>25$ ``good'' observations (good observations are those that have no IPAC or SExtractor flags set [is there some IPAC paper to cite here?]).
-With this sample, we compute the set of variability statistics (\S\ref{sec:event_selection}), then iterate to add different simulated microlensing events to each light curve.
+For a particular field, we randomly sample 1,000 light curves from each chip with $>25$ ``good'' observations (good observations are those that have no IPAC or SExtractor flags set [Marcel: is there some IPAC paper to cite here?]).
+With this sample, we compute the set of variability statistics (\S\ref{sec:event_selection}), then iterate to add [APW: N] different simulated microlensing events to each light curve.
 We use a random number generator to only add events during $\apwsim 50\%$ of the iterations, which allows us to later quantify the false positive rate of our selection methods.
 For each iteration we recompute the variability statistics and store the event parameters.
-For each simulated event, the peak time, $t_0$, is drawn from a uniform distribution between the minimum and maximum observation times, the impact parameter, $u_0$, is drawn from a uniform distribution between 0 and 1, and the event timescale, $t_E$, is drawn from a log-uniform distribution between 1 and 1000 days.
+[The parameters for each simulated event, the peak time, $t_0$, is drawn from a uniform distribution between the minimum and maximum observation times, the impact parameter, $u_0$, is drawn from a uniform distribution between 0 and 1, and the event timescale, $t_E$, is drawn from a log-uniform distribution between 1 and 1000 days.]
 
-We apply our selection criteria to each simulated light curve using the variability statistics computed for the initial light curves (from \textit{before} we add simulated events), and then bin the selected data by the event timescale.
-
-- Previous surveys had fairly uniform time sampling, and smaller areas
-
-- With PTF, fields have vastly different sampling patterns, baselines, and cadences
-
-- We have to consider the efficiency on a field-by-field basis
-
-- Show detection efficiency of each variability statistic
+We apply our selection criteria to each simulated light curve using the variability statistics computed for the initial light curves (from \textit{before} we add simulated events), and then bin the selected data by the event timescale [APW: figure reference].
 
@@
 	\centering
+	\caption{(a) the light curve plotted as is, (b) light curve with a simulated microlensing event with $u_0=1.0$, $t_E=20~\mathrm{days}$, (c) same as (b) with $u_0=0.5$, (d) same as (b) with $u_0=0.01$.}
+    \includegraphics[width=1.0\textwidth]{figures/simulated_events.pdf}
+    \label{fig:microlensing_sim}
+\end{figure}
+
+\begin{figure}
+	\centering
 	\caption{The PTF survey footprint shown in equatorial coordinates. Each field's color corresponds to some minimum number of observations of that field. }
@@
 \end{figure}
+
 %

### 2012-08-28 paper_figures.py
import os

@@ -8,7 +8,7 @@ from __future__ import division
 __author__ = "adrn <adrn@astro.columbia.edu>"
 
 # Standard library
-import sys
+import os, sys
 import logging
 import cPickle as pickle
 import time
@@ -153,11 +153,11 @@ def median_maximum_indices_plot(field_id, ccd_id=5):
         chip = ccd.read()
         
         source_ids = chip.sources.readWhere("ngoodobs > 25")["matchedSourceID"]
+        logger.info("{} source ids selected".format(len(source_ids)))
         
         var_indices = []
-        
-        count = 0
         for source_id in source_ids:
+            logger.debug("Source ID: {}".format(source_id))
             light_curve = ccd.light_curve(source_id, clean=True, barebones=True)
             
             if len(light_curve.mjd) > 25:
@@ -168,7 +168,6 @@ def median_maximum_indices_plot(field_id, ccd_id=5):
                     continue
                 
                 var_indices.append(lc_var_indices)
-                count += 1 
         
         var_indices = np.array(var_indices, dtype=zip(indices, [float]*len(indices)))
         

### 2012-09-25 ptf/db/mongodb.py
function to load all fields into mongodb

@@ -27,8 +27,9 @@ import pymongo
 
 # Project 
 from ..ptflightcurve import PDBLightCurve
-from ..globals import config
+from ..globals import config, _base_path
 from ..analyze import analyze as pa
+from ..photometricdatabase import Field
 
 # Create logger for this module
 logger = logging.getLogger(__name__)
@@ -132,24 +133,6 @@ def save_light_curve_document_to_collection(light_curve_document, collection, ov
     
     return True
 
-'''
-def add_tag_to_light_curve_document(light_curve_document, tag, light_curve_collection):
-    """ Add a tag to the light curve object """
-    
-    update = light_curve_collection.update({"_id" : light_curve_document["_id"]},\
-                                           {"$addToSet": { "tags" : tag.lower() } });
-
-    return update
-
-def remove_tag_from_light_curve_document(light_curve_document, tag, light_curve_collection):
-    """ Remove a tag from the light curve object """
-    
-    update = light_curve_collection.update({"_id" : light_curve_document["_id"]},\
-                                           {"$pull": { "tags" : tag.lower() } });
-
-    return update
-'''
-
 def update_light_curve_document_tags(light_curve_document, tags, light_curve_collection):
     """ Add a tag to the light curve object """
     
@@ -192,3 +175,26 @@ def field_to_document(field, **kwargs):
             continue
     
     return new_field
+
+def load_all_fields(field_collection):
+    """ Load all of the field information into mongodb """
+    
+    R_fields = np.load("data/survey_coverage/fields_observations_R.npy")
+    loaded_field_ids = [a["_id"] for a in field_collection.find(fields=["_id"])]
+    
+    field_ids_to_load = list(set(R_fields["field"]).difference(set(loaded_field_ids)))
+    
+    for field_id in field_ids_to_load:
+        field = Field(field_id)
+        
+        # Double check that it's not already in there!
+        if field_collection.find_one({"_id" : field.id}) != None:
+            logger.info("Field is already loaded! What happened?! {}".format(field.id))
+            continue
+            
+        field_doc = field_to_document(field)
+        field_collection.insert(field_doc)
+        logger.debug("Field loaded into mongodb {}".format(field.id))
+    
+    return True
+    

### 2012-09-25 ptf/globals.py

@@ -1,4 +1,5 @@
 import os
+import numpy as np
 
 # PTF Stats
 pix_scale = 1.01 #arcsec / pixel
@@ -39,3 +40,5 @@ def source_index_name_to_pdb_index(source, index_name):
         return source[pdb_index_name[index_name][0]] / source[pdb_index_name[index_name][1]]            
     else:
         return source[pdb_index_name[index_name]]
+
+all_fields = np.load(os.path.join(os.path.split(_base_path)[0], "data", "all_fields.npy"))

### 2012-09-25 survey_coverage.py

@@ -39,7 +39,7 @@ formatter = logging.Formatter("%(name)s / %(levelname)s / %(message)s")
 ch.setFormatter(formatter)
 logger.addHandler(ch)
 
-all_fields = np.load(os.path.join("data", "all_fields.npy"))
+all_fields = np.load(os.path.join(os.path.split(pg._base_path)[0], "data", "all_fields.npy"))
 
 class SurveyInfo(object):
     """ Some example use cases for this object:
@@ -61,7 +61,7 @@ class SurveyInfo(object):
         if not isinstance(filter, pdb.Filter):
             self.filter = pdb.Filter(filter)
         
-        cache_filename = os.path.join("data", "survey_coverage", "fields_observations_{}.npy".format(str(self.filter)))
+        cache_filename = os.path.join(os.path.split(pg._base_path)[0], "data", "survey_coverage", "fields_observations_{}.npy".format(str(self.filter)))
         self._fields_exposures = get_fields_exposures(self.filter, filename=cache_filename, overwrite=overwrite)
         self.timestamp = datetime.datetime.strptime(time.ctime(os.path.getmtime(cache_filename)), "%a %b %d %H:%M:%S %Y")
         

### 2012-09-27 text/paper/ptf_microlensing_paper.tex
pre-marcel changes

@@
  of the Palomar Transient Factory}
-\author{Adrian~M.~Price-Whelan\altaffilmark{\ref{col}}, Marcel~A.~Ag\"ueros\altaffilmark{\ref{col}}, Amanda Fournier\altaffilmark{\ref{ucsb}}, Rachel Street\altaffilmark{\ref{lcogt}}, Eran Ofek\altaffilmark{\ref{weiz},\ref{eins}}, David Levitan\altaffilmark{\ref{calt}}, Joshua S.\ Bloom\altaffilmark{\ref{cal}}, S.\ Bradley Cenko\altaffilmark{\ref{cal}}, Mansi M.\ Kasliwal\altaffilmark{\ref{calt}}, Shrinivas R.\ Kulkarni\altaffilmark{\ref{calt}}, Nicholas~M.~Law\altaffilmark{\ref{to},\ref{dun}}, Peter Nugent\altaffilmark{\ref{lbnl}}, Dovi Poznanski\altaffilmark{\ref{cal},\ref{calt},\ref{eins}}, Robert M.\ Quimby\altaffilmark{\ref{calt}}}
+\author{Adrian~M.~Price-Whelan\altaffilmark{\ref{col}}, Marcel~A.~Ag\"ueros\altaffilmark{\ref{col}}, Amanda Fournier\altaffilmark{\ref{ucsb}}, Rachel Street\altaffilmark{\ref{lcogt}}, Eran Ofek\altaffilmark{\ref{weiz},\ref{eins}}, David Levitan\altaffilmark{\ref{calt}}, and the PTF Collaboration}
+%, Joshua S.\ Bloom\altaffilmark{\ref{cal}}, S.\ Bradley Cenko\altaffilmark{\ref{cal}}, Mansi M.\ Kasliwal\altaffilmark{\ref{calt}}, Shrinivas R.\ Kulkarni\altaffilmark{\ref{calt}}, Nicholas~M.~Law\altaffilmark{\ref{to},\ref{dun}}, Peter Nugent\altaffilmark{\ref{lbnl}}, Dovi Poznanski\altaffilmark{\ref{cal},\ref{calt},\ref{eins}}, Robert M.\ Quimby\altaffilmark{\ref{calt}}}
 
@@
 In particular, light from a background source passing some mass distribution will be deflected by an angle that depends on the distance between the source and mass distribution, as well as the structure and kinematics of the mass distribution itself.
-For an observer aligned with a source and a point lensing mass, the relevant angular scale for the observed deflection is given by the Einstein ring radius:
+For an observer aligned with a background source and a point lensing mass, the relevant angular scale for the observed deflection is given by the Einstein ring radius:
 \begin{equation} \label{eq:einstein_radius}
@@
 
-Microlensing designates any unresolvable gravitational lensing, where the images formed by the lens just contribute to an increase in flux.
+Microlensing refers to any unresolved gravitational lensing, where the images formed simply contribute to a transient increase in flux from the source.
 For Galactic distance and mass scales, the Einstein radius (Eq.~\ref{eq:einstein_radius}) can be expressed as:
@@
 \citep{paczynski1986, gaudi2011}. 
-A solar-mass star in the bulge lensing another star in the bulge would have an Einstein radius of about $\theta_E \sim 0.3~\mathrm{mas}$, well below the resolution limit of even our most powerful telescopes.
+%A solar-mass star in the bulge lensing another star in the bulge would have an Einstein radius of about $\theta_E \sim 0.3~\mathrm{mas}$, well below the resolution limit of even our most powerful telescopes.
 
-Since Einstein first mentioned gravitational microlensing \citep{einstein1936}, it has been applied to the search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, galactic structure and kinematics \citep{binney2000}, and extrasolar planets \citep{}[APW: cite].
+Since Einstein first mentioned gravitational microlensing \citep{einstein1936}, it has been applied to the search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, galactic structure and kinematics \citep{binney2000}, and extrasolar planets \citep{gaudi2011}.
 Microlensing events were once a relatively rare phenomena, but with advances in CCD technology and the help of several dedicated microlensing surveys [APW: cite], more than [APW: check number]500 events are detected each year by several microlensing collaborations and surveys.
-Within the next decade, this number is expected to [] as the next generation of photometric, time-domain surveys come online [APW: cite LCOGT, LSST, ??].
+Within the next decade, this number is expected to rise as the next generation of photometric, time-domain surveys come online.
 
-[APW: Some brief description of what is PTF, why it is the bridge between what has been done and what will be done, etc...]
-
-\subsection{Something??}
-Assuming that the source is an un-blended point source and the lens is a foreground, dim object, a microlensing event is fully described by three parameters: the angular impact parameter $u_0$, the peak time of the event $t_0$, and the timescale of the event (Einstein crossing time) $t_E$ such that the increase in flux due to source magnification is given by:
+\subsection{Signatures}
+Assuming that the source is an un-blended point source and the lens is a foreground, dim object, a microlensing event is fully described by three parameters: the angular impact parameter $u_0$, the peak time of the event $t_0$, and the timescale of the event (Einstein crossing time) $t_E$ such that the flux is given by:
 \begin{align}
@@
 \end{align}
-where $A(t)$ is the amplification factor.
+where $A(u(t))$ is the amplification factor.
 
@@
 If $t_E$ is short ($t_E \lesssim\mathrm{days}$), surveys may miss or poorly sample the event, but if $t_E$ is long ($t_E \gtrsim 1~\mathrm{year}$) the event may be confused with other forms of long-duration variability.
-\figurename~\ref{fig:microlensing_sim} shows a ``typical'' PTF light curve and demonstrates what microlensing events with $t_E=20~\mathrm{days}$ and varying $u_0$ may look like in the data.
+\figurename~\ref{fig:microlensing_sim} shows a well-sampled PTF light curve and demonstrates what a microlensing event with fixed $t_E=20~\mathrm{days}$ and varied $u_0$ may look like in the data.
 
@@
 Thus, the properties of the lenses are more likely to be well-constrained by direct observation, possibly enabling mass measurements of objects not bound in binary pairs.
-The event rate in low stellar density regions will certainly be smaller than the event rate near the Bulge \citep[e.g.,][]{wood_optical_depth, ogle_optical_depth, macho_optical_depth, eros_optical_depth}, but the opportunity for precise, direct observations of lenses would be extremely valuable for studying low-mass stars, substellar objects, and stellar remnants such as isolated neutron stars [APW: cite NS work?].
-
-[MOTIVATION FOR PROJECT -- Han, Di Stefano, Di Stefano et al. -- mesolensing and near-field microlensing]
+The event rate in low stellar density regions will certainly be smaller than the event rate near the Bulge \citep[e.g.,][]{wood_optical_depth, ogle_optical_depth, macho_optical_depth, eros_optical_depth}, but the opportunity for the precise study of lenses would be extremely valuable for studying low-mass stars, substellar objects, and stellar remnants such as isolated neutron stars [APW: cite NS work?].
 
@@
 
-\section{Microlensing event recovery} \label{sec:event_recovery}
-Microlensing surveys typically use difference image analysis (cite Lupton?) to identify transient events in raw imaging data.
-Light curves of any transient sources are then analyzed and further pruned using a variety of selection methods to search for microlensing event candidates (cite MACHO, OGLE).
-Using difference imaging is essential for studying crowded fields such as the Galactic bulge where photometry is difficult (cite or figure?), and also limits the number of light curves that must be analyzed, thus saving on computation time.
-Pre-selection with difference imaging also enables fast, real-time detection of events, allowing dedicated microlensing surveys to obtain high-cadence follow-up observations of interesting sources.
-Such surveys also have the advantage of relatively uniform time sampling of the survey footprint over an observing season.
-
 The PTF survey footprint (\figurename~\ref{fig:survey_footprint}) is not uniformly sampled -- spatially or temporally.
@@
 As a result, PTF light curves often contain gaps, regions of high-cadence observations, and regions of low-cadence observations, leading to a massive data set of irregularly-sampled, time-domain photometry.
-\figurename~\ref{fig:sampling} shows six randomly-selected light curves (with their magnitudes removed) and demonstrates the varying cadences and coverage fractions that different fields may have over the same one-year baseline.
+\figurename~\ref{fig:sampling} shows six randomly-selected light curves (with their magnitudes removed) and demonstrates the varying cadences and coverage that different fields may have over the same one-year baseline.
 
---
+\section{Microlensing event recovery} \label{sec:event_recovery}
+Microlensing surveys typically use difference image analysis \citep{alard1998} to identify transient events in raw imaging data.
+Light curves of any transient sources are then analyzed and further pruned using a variety of selection methods to search for microlensing event candidates (cite MACHO, OGLE).
+Using difference imaging is essential for studying crowded fields such as the Galactic bulge where photometry is difficult and also limits the number of light curves that must be analyzed, thus saving on computation time.
+Pre-selection with difference imaging also enables fast, real-time detection of events, allowing dedicated microlensing surveys to obtain high-cadence follow-up observations of interesting sources.
+Such surveys also have the advantage of relatively uniform time sampling of their survey footprint over an observing season.
 
-Previous work has used a variety of survey-specific selection methods (as described in \citealt{alcock2000, wyrzykowski2009, hamadache2009, sumi2011}), but the general idea is to \begin{enumerate}
+After identifying transients, the light curves must be further analyzed to distinguish microlensing event candidates from variable stars, outbursting systems, and novae.
+Previous efforts have used a variety of survey-specific selection methods (as described in \citealt{alcock2000, wyrzykowski2009, hamadache2009, sumi2011}), but the general idea is to \begin{enumerate}
 	\item detect transient events using difference image analysis, 
@@
 Furthermore, it isn't obvious that this same selection procedure will be as successful for the PTF data.
-Motivated by work on identifying and classifying general variability in light curve data \citep{shin2009}[others?], we set out to compare the performance of microlensing event selection using a set of variability statistics as an initial cut in our candidate detection pipeline.
+Motivated by work on identifying and classifying general variability in light curve data \citep{shin2009}[others?], we set out to compare the performance of microlensing event selection using a set of variability statistics.
 
 \subsection{Event Selection} \label{sec:event_selection}
-We choose five statistical measures of variability, as compiled by \cite{shin2009}, that have been previously applied to classification of periodic variables, and evaluate their effectiveness in recovering simulated microlensing events in the data.
+We choose five statistical measures of variability, as compiled by \cite{shin2009}, that have been previously applied to classification and discovery of periodic variables, and evaluate their effectiveness in recovering simulated microlensing events in the data.
 These statistics, or \emph{variability indices}, are $\sigma/\mu$, $Con$, $\eta$, $J$, and $K$ (we choose not to implement the sixth, $AoVM$, because it mainly helps with periodic sources).
@@
 \end{align}
-$\eta$ is small, but still greater than zero, when there is strong serial correlation between successive data points. $J$ and $K$ were suggested by \cite{stetson1996}:
+$\eta$ is small when there is strong positive serial correlation between successive data points. $J$ and $K$ were suggested by \cite{stetson1996}:
 \begin{align}
@@
 $J$ tends to 0 for non-variable stars, but is large when there are significant differences between successive data points in a light curve.
-$K$ is a measure of the kurtosis or ``peakedness'' of the distribution of points in the light curve.
-We add one more variability index motivated by Cut 3 (\S\ref{sec:event_recovery}): $\Delta \chi^2$ is defined as the difference in chi-squared between fitting a five-parameter Gaussian model and fitting a linear model to the light curve.
-\figurename~\ref{fig:indices_examples} shows examples of a median light curve for each variability statistic (left panels) along with the most extreme outliers (right panels) for XXX light curves on a single CCD on PTF field YYY.
+$K$ is a measure of the kurtosis of the distribution of data points in the light curve.
+We add one more variability index, $\Delta \chi^2$, motivated by Cut 3 (\S\ref{sec:event_recovery}): $\Delta \chi^2$ is defined as the difference in chi-squared between fitting a Gaussian model and fitting a linear model to the light curve data.
+\figurename~\ref{fig:indices_examples} shows maximal outlier light curves for each variability statistic (except $Con$) out of \apwsim 20,000 light curves on a single CCD on PTF field 100049.
+[The indices are sensitive to different aspects of variability in the data...]
 
-\figurename~\ref{fig:scatter_plot_matrix} shows a scatter-plot matrix of all six of our variability statistics.
-In this figure, we take 10,000 light curves from a single PTF field (field id = ), compute the variability indices for each light curve, and then plot $1\sigma$ (red) and $2\sigma$ (green) contours in each panel for the resultant distributions.
+- Here discuss a figure that shows variance of the distribution of the indices for simulated light curves w/ varied noise sigma
+
+%\figurename~\ref{fig:scatter_plot_matrix} shows a scatter-plot matrix of the variability statistics.
+In this figure, we take 1,000 light curves from each CCD on a single PTF field, compute the variability indices for each light curve, and then plot $1\sigma$ (red) and $2\sigma$ (green) contours in each panel for the resultant distributions.
 We then add 10 simulated microlensing events to each light curve, recompute the variability indices, and plot these values as black points.
 
-[Describe the different selection methods]
-
-- Proper statistical modeling is too time consuming, so we first have to define some selection criteria to limit the sample (number of good observations, quality cuts, etc.)
-
-- Discuss $\Delta \chi^2$ method, downfalls (speed, not robust -- how do you define the threshold?).
+%- Discuss $\Delta \chi^2$ method, downfalls (speed, not robust -- how do you define the threshold?).
 Show a plot of distribution of $\Delta \chi^2$ with and without simulated events, e.g.
@@
 
-- Shin et al. paper
-
-- Introduce variability statistics: define them, what they pick out, what they have been used for before
-
 \subsection{Detection Efficiency}
 The probability of detecting any given microlensing event depends on the parameters of the event, the survey properties (observing strategy, limiting magnitude, weather), and the nature of the stellar field (crowding).
-To compute the detection efficiency, $\varepsilon$, for a given selection criterium, we run Monte Carlo simulations to artificially add microlensing events to real data and evaluate how often these events are recovered by the cut.
-This analysis must be done on a field-by-field basis, however we can compare our selection methods on a sub-sample of representative cases.
-Performing these simulations with real data allows us to compute the ...
-[APW: explain difference between two different detection efficiencies (of selection method vs.
-of survey parameters), and how we marginalize over both].
+To compute the detection efficiency, $\varepsilon$, for a given selection method, we must first determine our selection criteria for each variability index.
+For a given PTF field, we randomly sample 1000 light curves from the database with good observations and for each of these we use the time and error information to simulate 100 light curves with Gaussian scatter.
+We find the value of each variability index such that the selection returns 1\% of these simulated data -- that is, we choose the limiting value of each variability index so our false positive rate is 1\%.
 
-For a particular field, we randomly sample 1,000 light curves from each chip with $>25$ ``good'' observations (good observations are those that have no IPAC or SExtractor flags set [Marcel: is there some IPAC paper to cite here?]).
-With this sample, we compute the set of variability statistics (\S\ref{sec:event_selection}), then iterate to add [APW: N] different simulated microlensing events to each light curve.
-We use a random number generator to only add events during $\apwsim 50\%$ of the iterations, which allows us to later quantify the false positive rate of our selection methods.
+Once we have figured out our selection boundaries for each variability index, we run more Monte Carlo simulations to artificially add microlensing events to real data from a given field and evaluate how often these events are recovered by the procedure.
+For a particular field, we randomly sample 1,000 light curves from each CCD with $>25$ ``good'' observations (observations that have no IPAC or SExtractor flags set [Marcel: is there some IPAC paper to cite here?]).
+With this sample, we compute the set of variability statistics (\S\ref{sec:event_selection}), then iterate to add 100 different simulated microlensing events to each light curve.
+The parameters for each simulated event are chosen as follows: the peak time, $t_0$, is drawn from a uniform distribution between the minimum and maximum observation times, the impact parameter, $u_0$, is drawn from a uniform distribution between 0 and 1.34 (cite?), finally, the event timescale, $t_E$, is drawn from a log-uniform distribution between 1 and 1000 days.
 For each iteration we recompute the variability statistics and store the event parameters.
-[The parameters for each simulated event, the peak time, $t_0$, is drawn from a uniform distribution between the minimum and maximum observation times, the impact parameter, $u_0$, is drawn from a uniform distribution between 0 and 1, and the event timescale, $t_E$, is drawn from a log-uniform distribution between 1 and 1000 days.]
+\figurename~\ref{fig:detection_efficiency} shows the detection efficiency for each variability statistic as functions of microlensing impact parameter and source limiting magnitude [APW: TODO, create this figure].
+These detection efficiency curves are affected by the structure of the data, thus with a non-uniformly sampled field the efficiency will never reach 100\%.
 
-We apply our selection criteria to each simulated light curve using the variability statistics computed for the initial light curves (from \textit{before} we add simulated events), and then bin the selected data by the event timescale [APW: figure reference].
+%Possibly..
+Figures XX-YY show detection efficiency curves for three representative fields chosen for their different sampling patterns.
+For each of these cases we find that ZZZ consistently performs better than the other indices at recovering microlensing events.
 
@@
 	\centering
-	\caption{(a) the light curve plotted as is, (b) light curve with a simulated microlensing event with $u_0=1.0$, $t_E=20~\mathrm{days}$, (c) same as (b) with $u_0=0.5$, (d) same as (b) with $u_0=0.01$.}
+	\caption{(a) the light curve plotted as is, (b) light curve with a simulated microlensing event with $u_0=1.0$, $t_E=20~\mathrm{days}$, (c) same as (b) with $u_0=0.5$, (d) same as (b) with $u_0=0.01$.
+The red dashed line shows the approximate saturation limit of the PTF camera in R-band.}
     \includegraphics[width=1.0\textwidth]{figures/simulated_events.pdf}
@@
 
+\begin{figure}
+	\centering
+	\caption{The maximum outlier light curves for each variability index on PTF field 100049.}
+    \includegraphics[width=1.0\textwidth]{figures/max_outlier_light_curves.pdf}
+    \label{fig:indices_examples}
+\end{figure}
+
 %

### 2012-10-03 text/paper/first_yr_report_6.tex
vastly revised

\documentclass[iop]{emulateapj}
%\documentclass[12pt,preprint]{aastex}
\newcounter{address}
\setcounter{address}{1}
%\usepackage{lscape, longtable}
\usepackage{lmodern}
%\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{subfigure}
\usepackage{natbib}

% for flowchart
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}

\newcommand{\Msun}{\ifmmode {M_{\odot}}\else${M_{\odot}}$\fi}
\newcommand{\Rsun}{\ifmmode {R_{\odot}}\else${R_{\odot}}$\fi}
\newcommand{\lapprox }{{\lower0.8ex\hbox{$\buildrel <\over\sim$}}}
\newcommand{\gapprox }{{\lower0.8ex\hbox{$\buildrel >\over\sim$}}}
\def\amin{\ifmmode^{\prime}\else$^{\prime}$\fi}
\def\asec{\ifmmode^{\prime\prime}\else$^{\prime\prime}$\fi}

\newcommand{\apwsim}{\raisebox{0.2ex}{\scriptsize$\sim$\normalsize}} 
\newcommand{\inlinecode}{\texttt}

\slugcomment{DRAFT \today}
\shorttitle{Microlensing \& PTF}
\shortauthors{Price-Whelan et al.}

\bibliographystyle{apj}

\begin{document}

\title{Identifying Microlensing Events in Large, Non-Uniformly Sampled Surveys: The Case
 of the Palomar Transient Factory}
\author{Adrian~M.~Price-Whelan\altaffilmark{\ref{col}}, Marcel~A.~Ag\"ueros\altaffilmark{\ref{col}}, Amanda Fournier\altaffilmark{\ref{ucsb}}, Rachel Street\altaffilmark{\ref{lcogt}}, Eran Ofek\altaffilmark{\ref{weiz},\ref{eins}}, David Levitan\altaffilmark{\ref{calt}}, and the PTF Collaboration}
%, Joshua S.\ Bloom\altaffilmark{\ref{cal}}, S.\ Bradley Cenko\altaffilmark{\ref{cal}}, Mansi M.\ Kasliwal\altaffilmark{\ref{calt}}, Shrinivas R.\ Kulkarni\altaffilmark{\ref{calt}}, Nicholas~M.~Law\altaffilmark{\ref{to},\ref{dun}}, Peter Nugent\altaffilmark{\ref{lbnl}}, Dovi Poznanski\altaffilmark{\ref{cal},\ref{calt},\ref{eins}}, Robert M.\ Quimby\altaffilmark{\ref{calt}}}

\altaffiltext{\theaddress}{\stepcounter{address}\label{col} Department of Astronomy, Columbia University, 550 W 120th St., New York, NY 10027, USA}
\altaffiltext{\theaddress}{\stepcounter{address}\label{ucsb} Department of Physics, Broida Hall, University of California, Santa Barbara, CA 93106, USA}
\altaffiltext{\theaddress}{\stepcounter{address}\label{lcogt} Las Cumbres Observatory Global Telescope Network, Inc., 6740 Cortona Dr.\ Suite 102, Santa Barbara, CA 93117, USA}
\altaffiltext{\theaddress}{\stepcounter{address}\label{weiz} Benoziyo Center for Astrophysics, Weizmann Institute of Science, 76100 Rehovot, Israel}
\altaffiltext{\theaddress}{\stepcounter{address}\label{calt} Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
%\altaffiltext{\theaddress}{\stepcounter{address}\label{cal} Department of Astronomy, University of California, Berkeley, CA  94720, USA}
%\altaffiltext{\theaddress}{\stepcounter{address}\label{to} Dunlap Institute for Astronomy and Astrophysics, University of Toronto, 50 St.\ George Street, Toronto M5S 3H4, Ontario, Canada}
%\altaffiltext{\theaddress}{\stepcounter{address}\label{lbnl} Computational Cosmology Center, Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA}
%\altaffiltext{\theaddress}{\stepcounter{address}\label{dun} Dunlap Fellow}
\altaffiltext{\theaddress}{\stepcounter{address}\label{eins} Einstein Fellow}

\begin{abstract}
Many current photometric, time-domain surveys are driven by specific goals such as searches for supernovae or transiting exoplanets, or studies of stellar variability. These goals in turn set the cadence with which individual fields are re-imaged. In the case of the Palomar Transient Factory (PTF), several such sub-surveys are being conducted in parallel, leading to extremely non-uniform sampling over the survey's nearly $20,000~\mathrm{deg}^2$ footprint. While the typical $7.26~\mathrm{deg}^2$ PTF field has been imaged 20 times in \textit{R}-band, \apwsim$2300~\mathrm{deg}^2$ have been observed more than 100 times. We use the existing PTF data (\apwsim$6.4\times10^7$ light curves) to study the trade-off that occurs when searching for microlensing events when one has access to a large survey footprint with irregular sampling. To examine the probability that microlensing events can be recovered in these data, we also test previous statistics used on uniformly sampled data to identify variables and transients. We find that one such statistic, the von Neumann ratio, performs best for identifying simulated microlensing events. We develop a selection method using this statistic and apply it to data from all PTF fields with $>100$ observations to uncover a number of interesting candidate events. This work can help constrain all-sky event rate predictions and tests microlensing signal recovery in large datasets, both of which will be useful to future wide-field, time-domain surveys such as the LSST.

\end{abstract}

\keywords{
  survey science
  ---
  gravitational microlensing
  ---
  time-domain astrophysics
}

\section{Introduction}
Over the past 20 years, microlensing, in which gravitational lensing results in a transient increase in flux from a background source, has been applied to the search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, to study Galactic structure and kinematics \citep{binney2000}, and, most recently. to identify extrasolar planets \citep{gaudi2011}. Such studies were once limited by the small number of detected events, but with advances in CCD technology and the development dedicated microlensing surveys \citep{original_ogle, original_macho, eros_original}, more than 1000 events are now detected each year. In the next decade, this number is expected to rise as the next generation of photometric, time-domain surveys come online, providing an opportunity for the precise study of otherwise hard-to-characterize objects such as low-mass stars and substellar objects, and stellar remnants such as isolated neutron stars.

The microlensing event rate in some particular region of sky depends on the space and velocity distributions of sources and on the space, velocity, and mass distribution of lenses. To maximize the event rate, microlensing surveys have typically focused their observations towards high-density stellar regions such as the Galactic Bulge, M31, and the Magellanic clouds \citep[e.g.][]{original_ogle, original_macho, eros_original, crotts1996}. Microlensing events are of course not limited to dense stellar fields, but only one event outside of such fields has been recorded: the so-called Tago event \citep{fukui2007, gaudi2008}. \cite{gaudi2008} were able to establish probabilistic limits on the mass, distance, proper motion, and magnitude of this lens using existing observations near the location of the event. With \textit{JHK} magnitudes of the lens itself or a direct measurement of the proper velocity of the lens, as will be possible within the next year, these limits can be turned into precise measurements. Such a detailed analysis is not normally possible for events in the previously mentioned high-density regions where the lenses are much farther away, and blending makes precise photometry difficult.

Not all events away from the Bulge will be as fortuitous as the Tago event, but lensing events away from high density fields are generally expected to have closer sources, closer lenses, and less crowded background fields compared to those found by modern microlensing surveys \citep{mesolensing}. Thus, the properties of the lenses are more likely to be well-constrained by direct observation, possibly enabling mass measurements of objects not bound in binary pairs. The event rate in low-density stellar regions will certainly be smaller than that near the Bulge \citep[e.g.,][]{wood_optical_depth, ogle_optical_depth, macho_optical_depth, eros_optical_depth}, but the opportunity for the precise study of even a handful of lenses would be extremely valuable.

In this paper, we use the Palomar Transient Factory (PTF) survey data (described in \S2) as a test case for the frequency and detectability of microlensing events in all-sky, irregularly sampled time-domain surveys. We apply previous selection methods for identifying microlensing event signatures in light curves to the data and compare these methods with an event identification procedure based on a set of variability statistics (\S\ref{sec:event_recovery}). We then present the detection efficiency for these methods as computed for a representative set of PTF fields. In \S4 we apply a set of criteria that use the results of these tests to all PTF light curves with $>$100 high-quality observations to identify candidate microlensing events and present three high confidence candidates. We conclude in \S5.

%General relativity tells us that mass alters the propagation of light. In particular, light from a background source passing some mass distribution will be deflected by an angle that depends on the distance between the source and mass distribution, as well as the structure and kinematics of the mass distribution itself. Microlensing refers to any unresolved gravitational lensing, where the images formed simply contribute to a transient increase in flux from the source. Since Einstein first mentioned gravitational microlensing \citep{einstein1936}, it has been applied to the search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, Galactic structure and kinematics \citep{binney2000}, and extrasolar planets \citep{gaudi2011}. Microlensing events were once relatively rare, but with advances in CCD technology and the help of several dedicated microlensing surveys \citep{original_ogle, original_macho, eros_original}, more than 1000 events are detected each year by several microlensing collaborations and surveys. Within the next decade, this number is expected to rise as the next generation of photometric, time-domain surveys come online.

\section{PTF } % Better section header?
The Palomar Transient Factory is a transient detection system incorporating three elements. The first is a wide-field survey camera mounted on the automated 48 inch Oschin Schmidt telescope (the P48) at Palomar Observatory, CA. This is the former Canada-France-Hawaii Telescope 12K$\times$8K mosaic camera, which has 11 working chips, 10$^7$ pixels, and a 7.26 deg$^2$ field-of-view \citep{rahmer2008}. Under median seeing conditions (1.1$\arcsec$), observations in Mould $R$ or Sloan Digital Sky Survey \citep[SDSS;][]{york00} $g$ achieve 2.0$\arcsec$ full-width half-maximum images and reach 5$\sigma$ magnitudes of $R \approx 21.0$ and $g \approx 21.3$ mag in a standard 60~s exposure \citep{nick2010}. As of September 2012, the PTF footprint included $\apwsim$10,100 (2300) deg$^2$ imaged $>$25 ($>$100) times in $R$ and $\apwsim$2200 (170) deg$^2$ imaged that often in $g$ (\figurename~\ref{fig:survey_footprint})\footnote{For an interactive visualization of the PTF survey coverage, see: \url{http://bit.ly/PTFSky}}. PTF is also a real-time data reduction pipeline that identifies transients of interest, passes these to a dedicated photometric follow-up telescope, and generates an archive of all detected sources \citep{nick2009,rau2009}. %[CHECK NUMBERS]

%  [WE ALSO PROBABLY WANT TO CITE A FEW OF THE ACTUAL RESULT PAPERS; WE CAN WORRY ABOUT THAT LATER]
The PTF survey footprint is not uniformly sampled either spatially or temporally. Each field has a unique sampling pattern determined by (1) which of the PTF sub-surveys it belongs to, (2) which time of the year it is visible from Palomar, and (3) how high a priority it is given by the scheduler. 92\% of the P48 time is distributed between three  surveys: a 5-day cadence experiment covering 8000~deg$^2$ at $|b|>30^\circ$ to search for extragalactic transients, a dynamic cadence experiment to identify short duration ($<$5~days) transients, and a high-cadence imaging campaign of a single field in Orion to find transiting Jupiter-mass planets around young stars \citep{nick2009}. As a result, PTF light curves often contain gaps, regions of high-cadence observations, and regions of low-cadence observations, leading to a massive data set of irregularly-sampled, time-domain photometry. \figurename~\ref{fig:sampling} shows six randomly-selected light curves (with their magnitudes removed) and demonstrates the varying cadences and coverage that different fields may have over the same one-year baseline. We restrict our sample to the \apwsim$6.4\times10^7$ light curves with $>$25 \textit{R}-band observations.

What would a microlensing event look like in a typical PTF light curve? Assuming that the source is an un-blended point source and the lens is a foreground, dim object, a microlensing event is fully described by three parameters: the angular impact parameter $u_0$, the peak time of the event $t_0$, and the timescale of the event (Einstein crossing time) $t_E$ such that the flux is given by:
\begin{align}
	F(t) &= A(t)\times F_{source} \\
	A(u) &= \frac{u^2 + 2}{u\sqrt{u^2 + 4}}\\
	u(t) &= \sqrt{u_0^2 + 2\Big(\frac{t-t_0}{t_E}\Big)}
\end{align}
where $A(u(t))$ is the amplification factor. The microlensing perturbation can also be expressed in terms of magnitudes as:
\begin{align}
	m(t) &= m_0(t) - 2.5\log A(t)
\end{align}
where $m_0(t)$ is the unperturbed, possibly time-variable magnitude of the source.

%However, if the event is \textit{too} bright, the pixels will saturate and there could be photometric errors flagging the most interesting data as bad.
When $u_0$ is small ($u_0<<1$), the amplification is large. But even in cases of high amplification, a survey may miss or poorly sample an event if $t_E$ is short ($t_E \lesssim\mathrm{days}$); if $t_E$ is long ($t_E \gtrsim 1~\mathrm{year}$) the event may be confused with other forms of long-duration variability. \figurename~\ref{fig:microlensing_sim} shows a well-sampled PTF light curve and illustrates what a microlensing event with fixed $t_E=20~\mathrm{days}$ and varied $u_0$ might look like in the data.

\section{Microlensing event recovery} \label{sec:event_recovery}
Microlensing surveys typically use difference difference image analysis \citep{alard1998} to identify transient events in raw imaging data. Light curves of any transient sources are then analyzed and vetted using a variety of selection methods to search for microlensing event candidates and distinguish them from e.g., variable stars, outbursting systems, and novae. Different surveys approach this problem differently (as described in \citealt{ogle_optical_depth, con_idx, alcock2000, macho_detection_efficiency, hamadache2009, wyrzykowski2009, sumi2011}), but the general idea is to:

\begin{enumerate}
	\item require that any selected light curve have some number of consecutive datapoints brighter than some threshold;
	\item require that a microlensing model best describe the data, as determined by using a $\chi^2$ cut;
	\item require that the model event parameters have reasonable values.
\end{enumerate}

This process is ideal for studying crowded fields such as the Galactic Bulge where photometry is difficult. It also limits the number of light curves that must be fit with a model, thus saving on computation time and allowing for real-time detection of events, so that high-cadence follow-up observations of interesting sources can be triggered. Microlensing surveys also have the advantage of relatively uniform time sampling of their survey footprint over an observing season.

Applying this prescription to the PTF data presents some obvious challenges. In particular, the consecutive points requirement would seem hard to use when there can be significant (and irregular) gaps in the data. In order to develop the most successful prescription for PTF data, we set out to examine the relative performance of a set of variability statistics in selecting simulated microlensing events.

\subsection{Event Selection} \label{sec:event_selection}
We choose five statistical measures of variability compiled by \cite{shin2009} that have been previously applied to classification and discovery of periodic variables, and evaluate their effectiveness in recovering simulated microlensing events in the PTF data. These variability indices are $\sigma/\mu$, $Con$, $\eta$, $J$, and $K$.\footnote{We do not to implement the sixth index described by \cite{shin2009}, $AoVM$, because it mainly helps with periodic sources.} $\sigma/\mu$ is the ratio of the root-variance to the sample mean, 
\begin{align}
	\frac{\sigma}{\mu} = \frac{\sqrt{\sum^N_i (x_i - \mu)^2 / (N-1)}}{\sum^N_i x_i/N},
\end{align}
where N is the total number of observations in the light curve.

We modify the definition of $Con$ such that it is the number of clusters of three or more consecutive observations that are more than $3\sigma$ brighter than the reference magnitude of the source (e.g., for a single microlensing event in an otherwise flat light curve $Con=1$). This change allows us to use the performance of $Con$ as a proxy for the consecutive point requirement described above.

$\eta$, the von Neumann ratio \citep{von_neumann1941}, is the mean square successive difference over the sample variance:
\begin{align}
	\eta = \frac{\delta^2}{\sigma^2} = \frac{\sum^{N-1}_i(x_{i+1} - x_i)^2/(N-1)}{\sigma^2}.
\end{align}
$\eta$ is small when there is strong positive serial correlation between successive data points.

$J$ and $K$ were suggested by \cite{stetson1996}:
\begin{align}
	\delta_i &= \sqrt{\frac{N}{N-1}}\frac{x_i-\mu}{e_i}\\
	J &= \sum^{N-1}_i sign(\delta_i \delta_{i+1})\sqrt{|\delta_i \delta_{i+1}|}\\
	K &= \frac{1/N\sum^N_i |\delta_i|}{\sqrt{1/N\sum^N_i\delta_i^2}}
\end{align}
where $e_i$ is the photometric error of each data point. $J$ tends to 0 for non-variable stars, but is large when there are significant differences between successive data points in a light curve. $K$ is a measure of the kurtosis of the distribution of data points in the light curve.

We add one more variability index, $\Delta \chi^2$, defined as the difference in $\chi^2$ between fitting a Gaussian model and fitting a linear model to the light curve data. This is the standard statistical test used by microlensing surveys, and allows us to compare the relative performance of the (slightly modified) \cite{shin2009} indices and of this approach.

\figurename~\ref{fig:indices_examples} shows maximal outlier light curves for each variability statistic selected from \apwsim20,000 light curves for objects on a single CCD in PTF field 100049. Clearly, the indices are sensitive to different aspects of variability in the data, although in this case $\Delta \chi^2$, $Con$, $K$, and $\sigma/\mu$ all finger the same object as being maximally variable. We would expect $\sigma/\mu$ to be more useful for discerning periodic or semi-periodic variability where the variance is large; however, the expectations are not so clear for other indices.

To determine the regions of parameter space where microlensing light curves fall, we conduct Monte Carlo simulations and inject artificial microlensing events into real PTF light curves and compare the distributions of variability indices for a set of light curves both with and without these simulated events. \figurename~\ref{fig:var_indices} shows projections of the distribution for various combinations of variability indices. These simulations do define regions to search for microlensing events, but there is still a large amount of overlap between the distributions with and without microlensing events. In the next section, we describe our method for defining selection boundaries for each variability statistic.

% ^ discuss how the distributions change with injecting microlensing events
%  Here discuss a figure that shows variance of the distribution of the indices for simulated light curves w/ varied noise sigma

\subsection{Detection Efficiency}
The probability of detecting any given microlensing event depends on the parameters of the event, the survey properties (i.e., observing strategy, limiting magnitude, weather), and the nature of the stellar field (i.e., amount of crowding). In order to compute the detection efficiency, $\varepsilon$, for a given selection method, we must first define the selection criteria for each variability index. For a given PTF field, we randomly sample 1000 light curves with good observations\citep[these have no photometric flags; see description of processing pipeline in][]{nick2009}. For each of these light curves, we then use the time and error information to simulate 100 light curves with purely Gaussian scatter. We find the value of each variability index such that the selection returns 1\% of these simulated data --- i.e., we set the limiting value of each variability index such that the false positive rate (FPR) is 1\%.

We then run additional Monte Carlo simulations in which we artificially add microlensing events to real data from the same field and evaluate how often these events are recovered given the cuts as defined above. For a given field, we randomly sample 1000 light curves from each CCD with $>$25 good observations. We compute the set of variability statistics for these light curves, then iterate to add 100 different simulated microlensing events to each light curve.

The parameters for each simulated event are chosen as follows: the peak time, $t_0$, is drawn from a uniform distribution between the minimum and maximum observation times, the impact parameter, $u_0$, is drawn from a uniform distribution between 0 and 1.34 (cite?), and the event timescale, $t_E$, is drawn from a log-uniform distribution between 1 and 1000 days. For each iteration we recompute the variability statistics and store the event parameters. Figures~\ref{fig:detection_efficiency_4327}-\ref{fig:detection_efficiency_100152} show detection efficiency curves (bottom panels) computed using this simulation for three representative fields chosen for their different sampling patterns (top panels). We ignore the indices \textit{K} and $\sigma/\mu$ because their integrated detection efficiencies are each below 1\%. For each of these fields we find that $\eta$ consistently performs better than the other indices at recovering microlensing events. This is especially interesting because computing $\eta$ is \apwsim100 times faster per light curve as compared to $\Delta\chi^2$, and that identifying candidate microlensing events in the full PTF database using this statistic is therefore computationally plausible.

%\figurename~\ref{fig:detection_efficiency} shows the detection efficiency for each variability statistic as functions of microlensing impact parameter and source limiting magnitude [APW: TODO, create this figure].

%The parameters for the random microlensing events are chosen as follows: 
%\begin{align}
%	t_0&\sim\mathcal{U}[t_{min}, t_{max}] \\
%	u_0&\sim\mathcal{U}[0, 1] \\
%	\log_{10}t_E&\sim\mathcal{U}[0, 3] 
%\end{align}
%where $\mathcal{U}$ is a uniform distribution.

\section{Searching for events in the full PTF data set}
As an initial test, we applied the 1\% FPR selection on the 325 PTF fields (\apwsim2400 deg$^2$) with $>$100 $R$-band observations using the variability index $\eta$. Our procedure is as follows:
\begin{enumerate}
	\item select a field from the sample of fields with $>100$ good $R$-band observations, and run Monte Carlo simulations to compute the 1\% FPR limiting values for the variability indices;
	\item compute $\eta$ for each light curve, and select all light curves that pass the cut determined from the previous simulations;
	\item for each selected light curve, fit a microlensing event model and subtract the model, then recompute $\eta$;
	\item if the new value of $\eta$ still passes the cut, we reject the light curve (it is probably a periodic variable);
	\item any remaining light curves are loaded into a local database to be vetted by hand.
\end{enumerate}
% [THE ABOVE WOULD MAKE A VERY NICE FLOWCHART] --> for the actual paper..

From the \apwsim$10^7$ light curves in the 325 field footprint, this procedure selected 6306 candidates. For each candidate, we look at the SDSS colors (if the PTF field is in the SDSS footprint), search SIMBAD for any nearby sources, and visually inspect the PTF images to vet the data. We require that the sources are not associated with SDSS galaxies so we may eliminate supernovae from our sample, but acknowledge the possibility of contamination from supernovae in faint galaxies. Of the 6306 candidates, \apwsim1/3 of the light curves had bad or poorly calibrated data that mimicked a transient increase in flux. Approximately 2/3 of the light curves were quasars, AGN, or objects with unknown long-term variability. We have not yet searched this sample to identify stars because it is often hard to distinguish long duration microlensing events from other forms of long timescale stellar variability. A small number (four) of the selected light curves are variable stars that made it past our periodic variability cut, and 97 light curves were unidentified transients (novae, flares, etc.).

% In paper, explain MCMC fit in more detail...cit Emcee
From this entire sample, we identified four microlensing event candidates. \figurename~\ref{fig:candidates} and \figurename~\ref{fig:candidate_probs} show the light curve data, microlensing model fits using MCMC, and the posterior probability distributions for each parameter in the model. We also find a number of ambiguous transient events not associated with galaxies, but will have to further study these sources to properly classify them.

\section{Conclusions and Future Work}
Microlensing event rate predictions away from the Galactic Plane are not well constrained and only a single microlensing event has been observed in a sparse stellar field, despite the great scientific interest in observing such events. Using nearly all-sky, time-domain data from the Palomar Transient Factory, we (1) develop new methods for identifying interesting transients in massive light curve datasets, (2) search for sparse-field microlensing events in the data, and (3) will estimate the microlensing event rate over a large area of sky away from the Galactic plane.

We examine the detection efficiency of recovering simulated microlensing events using a set of variability statistics. We first determine selection criteria for each statistic using Monte Carlo simulations to simulate flat light curves with Gaussian noise and choose the selection boundaries such that our cuts achieve a 1\% false positive rate. We then simulate microlensing events in real PTF data and compute the detection efficiency for each selection method. We find that the von Neumann ratio, $\eta$, performs better than previously used statistics in recovering injected microlensing events in non-uniformly sampled data. We use $\eta$ to develop a selection procedure for extracting microlensing event candidates from the PTF light curve database and identify four microlensing event candidates, along with numerous unclassified variable objects.

For this document, we have restricted our sample to only light curves with $>100$ \textit{R}-band observations, but we will run our pipeline on the entire dataset. A large fraction of the contaminants in our selected sample of ``candidates'' are AGN or quasars exhibiting long-term, peaked variability. For much of the PTF footprint, we will be able to incorporate SDSS ($ugriz$) colors and spectroscopy into our pipeline to reject many of these objects. We also recovered a large number of transient events such as novae, outbursts, and flares. To remove these from our candidate list, we may be able to cross-reference the source positions with the PTF follow-up marshal to see if the transient has been classified. For light curves that survive these cuts, we will do a detailed model comparison between a microlensing event signal and asymmetric classes of transients.

Many of the light curves we reject present interesting forms of variability, but classification and cataloging such sources is outside the scope of this project. Still, it would be advantageous to both the PTF survey and community to keep a record of these flagged objects. We will work to incorporate many of our data visualization and analysis tools that have enabled us to quickly browse, classify, and vet our sample of data into the PTF data web interfaces\footnote{For example: \url{http://bit.ly/PTFSky}, \url{http://bit.ly/PTFDSky}, \url{http://bit.ly/PTFCandidates}}.

\acknowledgments
This paper is based on observations obtained with the Samuel Oschin Telescope as part of the Palomar Transient Factory project, a scientific collaboration between the California Institute of Technology, Columbia University, Las Cumbres Observatory, the Lawrence Berkeley National Laboratory, the National Energy Research Scientific Computing Center, the University of Oxford and the Weizmann Institute of Science. Work by A. M. Price-Whelan is supported by a National Science Foundation Graduate Research Fellowship under grant No. $<$Grant Number$>$.

\clearpage
\setlength{\baselineskip}{0.6\baselineskip}
\bibliography{references_2}
\setlength{\baselineskip}{1.667\baselineskip}

%\bibliographystyle{apj}
%\bibliography{apj-jour,refs}

\begin{figure}
	\centering
	\caption{The PTF survey footprint shown in equatorial coordinates. Each field's color corresponds to some minimum number of observations of that field. }
    \includegraphics[angle=270, scale=0.55]{figures/R_coverage.pdf}
    \label{fig:survey_footprint}
\end{figure}

\begin{figure}
	\centering
	\caption{A visualization of the different sampling regimes of six different PTF fields over the same one year baseline. Darker points mean more data (more exposures). Some fields have roughly uniform coverage, where others are very dense during some weeks and not others. }
    \includegraphics[width=1.0\textwidth]{figures/sampling_figure.pdf}
    \label{fig:sampling}
\end{figure}

\begin{figure}
	\centering
	\caption{(a) the light curve plotted as is, (b) light curve with a simulated microlensing event with $u_0=1.0$, $t_E=20~\mathrm{days}$, (c) same as (b) with $u_0=0.5$, (d) same as (b) with $u_0=0.01$. The red dashed line shows the approximate saturation limit of the PTF camera in R-band.}
    \includegraphics[width=1.0\textwidth]{figures/simulated_events.pdf}
    \label{fig:microlensing_sim}
\end{figure}

\begin{figure}
	\centering
	\caption{The maximum outlier light curves for each variability index on PTF field 100049.}
    \includegraphics[width=1.0\textwidth]{figures/max_outlier_light_curves.pdf}
    \label{fig:indices_examples}
\end{figure}

\begin{figure}
        \centering
        \caption{Two-dimensional density histograms for projections of the six-dimensional variability statistic distribution for 10,000 light curves sampled from PTF field 100018 (right) and the equivalent with the same light curves with simulated microlensing events (left).}
        \subfigure{
	   \includegraphics[height=0.3\textheight] {figures/eta_vs_delta_chi_squared.pdf}
	 }

	 \subfigure{
	   \includegraphics[height=0.3\textheight] {figures/eta_vs_j.pdf}
	 }

	 \subfigure{
	   \includegraphics[height=0.3\textheight] {figures/delta_chi_squared_vs_j.pdf}
	 }
          \label{fig:var_indices}
\end{figure}

\begin{figure}
        \centering
        \caption{Sample light curve (top) and detection efficiency curves (bottom) for PTF field 4327.}
        \subfigure[A randomly selected light curve from this field. Note the sampling pattern.]{
	   \includegraphics[height=0.2\textheight] {figures/example_light_curve_f4327_ccd2.pdf}
	   \label{fig:subfig1}
	 }

	 \subfigure[Detection efficiency curves for $\eta$, $\Delta\chi^2$, and $J$ as functions of microlensing event parameters $t_E$, $u_0$, and $m_0$. Black (dashed) lines show input distributions, red (solid) lines show the recovered distributions.]{
	   \includegraphics[width=\textwidth] {figures/field004327_Nperccd1000_Nevents100.pdf}
	   \label{fig:subfig2}
	 }
         \label{fig:detection_efficiency_4327}
\end{figure}

\begin{figure}
        \centering
        \caption{Same as \figurename~\ref{fig:detection_efficiency_4327} for PTF field 4588.}
        \subfigure{
	   \includegraphics[height=0.2\textheight] {figures/example_light_curve_f4588_ccd2.pdf}
	   \label{fig:subfig1}
	 }

	 \subfigure{
	   \includegraphics[width=\textwidth] {figures/field004588_Nperccd1000_Nevents100.pdf}
	   \label{fig:subfig2}
	 }
         \label{fig:detection_efficiency_4588}
\end{figure}

\begin{figure}
        \centering
        \caption{Same as \figurename~\ref{fig:detection_efficiency_4327} for PTF field 100152.}
        \subfigure{
	   \includegraphics[height=0.2\textheight] {figures/example_light_curve_f100152_ccd2.pdf}
	   \label{fig:subfig1}
	 }

	 \subfigure{
	   \includegraphics[width=\textwidth] {figures/field100152_Nperccd1000_Nevents100.pdf}
	   \label{fig:subfig2}
	 }
         \label{fig:detection_efficiency_100152}
\end{figure}

\begin{figure}
        \centering
        \caption{Zoomed-in (top) and full (bottom) light curves for four microlensing event candidates found in the PTF light curve database. Lines (red) show models with parameters sampled from the posterior probability distribution for the model fits.}
        \subfigure{
	   \includegraphics[width=0.42\textwidth] {figures/field4721_ccd8_source3208.pdf}
	   \label{fig:candidate1}
	 }
	 \subfigure{
	   \includegraphics[width=0.42\textwidth] {figures/field4789_ccd6_source11457.pdf}
	   \label{fig:candidate2}
	 }

	 \subfigure{
	   \includegraphics[width=0.42\textwidth] {figures/field100003_ccd6_source10741.pdf}
	   \label{fig:candidate3}
	 }
	 \subfigure{
	   \includegraphics[width=0.42\textwidth] {figures/field100400_ccd1_source49501.pdf}
	   \label{fig:candidate4}
	 }
         \label{fig:candidates}
\end{figure}

\begin{figure}
        \centering
        \caption{Posterior probability distributions for each microlensing event model parameter for each event candidate shown in \figurename~\ref{fig:candidates} }
        \subfigure{
	   \includegraphics[width=0.45\textwidth] {figures/field4721_ccd8_source3208_posterior.pdf}
	   \label{fig:candidate1_probs}
	 }
	 \subfigure{
	   \includegraphics[width=0.45\textwidth] {figures/field4789_ccd6_source11457_posterior.pdf}
	   \label{fig:candidate2_probs}
	 }

	 \subfigure{
	   \includegraphics[width=0.45\textwidth] {figures/field100003_ccd6_source10741_posterior.pdf}
	   \label{fig:candidate3_probs}
	 }
	 \subfigure{
	   \includegraphics[width=0.45\textwidth] {figures/field100400_ccd1_source49501_posterior.pdf}
	   \label{fig:candidate4_probs}
	 }
         \label{fig:candidate_probs}
\end{figure}

%
%\begin{figure}
%	\centering
%	\caption{Example of how the variability indices may be used to separate microlensing events from periodic or flat light curves.}
%    \includegraphics[width=1.0\textwidth]{figures/J_vs_K_Con_figure.png}
%    \label{fig:var_idx1}
%\end{figure}
%
%\begin{figure}[htbp]
%	\centering
%	\caption{Two neighboring light curves. The time scale is the same on both plots. Note the strong similarities in the overall shapes of the data.} 
%	 \includegraphics[width=1.0\textwidth]{figures/bad_data_figure.pdf}
%	\label{fig:lc_correlated}
%\end{figure}

\end{document}

### 2012-10-15 text/paper/ptf_microlensing.tex
new tex documents

--- 2012-10-03 text/paper/first_yr_report_6.tex
@@
-\documentclass[iop]{emulateapj}
-%\documentclass[12pt,preprint]{aastex}
+%\documentclass{emulateapj}
+\documentclass[12pt,preprint]{aastex}
 \newcounter{address}
@@
  of the Palomar Transient Factory}
-\author{Adrian~M.~Price-Whelan\altaffilmark{\ref{col}}, Marcel~A.~Ag\"ueros\altaffilmark{\ref{col}}, Amanda Fournier\altaffilmark{\ref{ucsb}}, Rachel Street\altaffilmark{\ref{lcogt}}, Eran Ofek\altaffilmark{\ref{weiz},\ref{eins}}, David Levitan\altaffilmark{\ref{calt}}, and the PTF Collaboration}
+\author{Adrian~M.~Price-Whelan\altaffilmark{\ref{col},\ref{nsf}}, Marcel~A.~Ag\"ueros\altaffilmark{\ref{col}}, Amanda Fournier\altaffilmark{\ref{ucsb}}, Rachel Street\altaffilmark{\ref{lcogt}}, Eran Ofek\altaffilmark{\ref{weiz},\ref{eins}}, David Levitan\altaffilmark{\ref{calt}}, and the PTF Collaboration}
 %, Joshua S.\ Bloom\altaffilmark{\ref{cal}}, S.\ Bradley Cenko\altaffilmark{\ref{cal}}, Mansi M.\ Kasliwal\altaffilmark{\ref{calt}}, Shrinivas R.\ Kulkarni\altaffilmark{\ref{calt}}, Nicholas~M.~Law\altaffilmark{\ref{to},\ref{dun}}, Peter Nugent\altaffilmark{\ref{lbnl}}, Dovi Poznanski\altaffilmark{\ref{cal},\ref{calt},\ref{eins}}, Robert M.\ Quimby\altaffilmark{\ref{calt}}}
@@
 \altaffiltext{\theaddress}{\stepcounter{address}\label{col} Department of Astronomy, Columbia University, 550 W 120th St., New York, NY 10027, USA}
+\altaffiltext{\theaddress}{\stepcounter{address}\label{nsf} NSF graduate fellow}
 \altaffiltext{\theaddress}{\stepcounter{address}\label{ucsb} Department of Physics, Broida Hall, University of California, Santa Barbara, CA 93106, USA}
@@
 \begin{abstract}
-Many current photometric, time-domain surveys are driven by specific goals such as searches for supernovae or transiting exoplanets, or studies of stellar variability.
-These goals in turn set the cadence with which individual fields are re-imaged.
+Many current photometric, time-domain surveys are driven by specific goals such as searches for supernovae or transiting exoplanets which set the cadence with which individual fields are re-imaged.
 In the case of the Palomar Transient Factory (PTF), several such sub-surveys are being conducted in parallel, leading to extremely non-uniform sampling over the survey's nearly $20,000~\mathrm{deg}^2$ footprint.

### 2012-10-15 candidate_pipeline.py
remove old docs

@@ -6,10 +6,6 @@ from __future__ import division
         For each field, we run simulations to determine the selection criteria for "interesting"
         variable sources, store those values in mongodb, then we use these criteria to go back
         to the PDB and select out candidates. 
-        
-    NEW IDEA! Today is 9/16/12. It is now 4:58 PM
-    - When doing delta chi-squared, subtract off the fit, recompute eta and see if it's still an outlier.
-        if it is, throw it away!
 """
 
 __author__ = "adrn <adrn@astro.columbia.edu>"

### 2012-10-15 text/paper/ptf_microlensing.tex
paper updates

@@
 In the case of the Palomar Transient Factory (PTF), several such sub-surveys are being conducted in parallel, leading to extremely non-uniform sampling over the survey's nearly $20,000~\mathrm{deg}^2$ footprint.
-While the typical $7.26~\mathrm{deg}^2$ PTF field has been imaged 20 times in \textit{R}-band, \apwsim$2300~\mathrm{deg}^2$ have been observed more than 100 times.
-We use the existing PTF data (\apwsim$6.4\times10^7$ light curves) to study the trade-off that occurs when searching for microlensing events when one has access to a large survey footprint with irregular sampling.
-To examine the probability that microlensing events can be recovered in these data, we also test previous statistics used on uniformly sampled data to identify variables and transients.
+While the median $7.26~\mathrm{deg}^2$ PTF field has been imaged 20 times in \textit{R}-band, \apwsim$2300~\mathrm{deg}^2$ have been observed more than 100 times.
+[We use the existing PTF data (\apwsim$6.4\times10^7$ light curves as of September 2012) to study the trade-off between searching for microlensing events in a survey footprint that is much larger than most microlensing searches, but with far-from-optimal sampling.] To examine the probability that microlensing events can be recovered in these data, we test previous statistics used on uniformly sampled data to identify variables and transients.
 We find that one such statistic, the von Neumann ratio, performs best for identifying simulated microlensing events.
-We develop a selection method using this statistic and apply it to data from all PTF fields with $>100$ observations to uncover a number of interesting candidate events.
-This work can help constrain all-sky event rate predictions and tests microlensing signal recovery in large datasets, both of which will be useful to future wide-field, time-domain surveys such as the LSST.
+[We develop a selection method using this statistic and apply it to data from all PTF fields with $>$100 observations to uncover a number of interesting candidate events.] This work can help constrain all-sky event rate predictions and tests microlensing signal recovery in large datasets, both of which will be useful to future wide-field, time-domain surveys such as the LSST.
+
+% Second to last sentence will change when I run on all fields with >25 observations
 
@@
 \section{Introduction}
-Over the past 20 years, microlensing, in which gravitational lensing results in a transient increase in flux from a background source, has been applied to the search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, to study Galactic structure and kinematics \citep{binney2000}, and, most recently.
+Over the past 20 years, microlensing, in which gravitational lensing results in a transient increase in the flux from a background source, has been used to search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, to study Galactic structure and kinematics \citep{binney2000}, and, most recently.
 to identify extrasolar planets \citep{gaudi2011}.
-Such studies were once limited by the small number of detected events, but with advances in CCD technology and the development dedicated microlensing surveys \citep{original_ogle, original_macho, eros_original}, more than 1000 events are now detected each year.
-In the next decade, this number is expected to rise as the next generation of photometric, time-domain surveys come online, providing an opportunity for the precise study of otherwise hard-to-characterize objects such as low-mass stars and substellar objects, and stellar remnants such as isolated neutron stars.
+Such studies were once limited by the small number of detected events, but with advances in CCD technology and the development of dedicated microlensing surveys \citep{original_ogle, original_macho, eros_original}, $>$1000 events are now detected each year.
+In the coming decade, this number is expected to rise as the next generation of photometric, time-domain surveys come online, providing an opportunity for the precise study of otherwise hard-to-characterize objects such as low-mass stars, substellar objects, and isolated neutron stars.
 
-The microlensing event rate in some particular region of sky depends on the space and velocity distributions of sources and on the space, velocity, and mass distribution of lenses.
-To maximize the event rate, microlensing surveys have typically focused their observations towards high-density stellar regions such as the Galactic Bulge, M31, and the Magellanic clouds \citep[e.g.][]{original_ogle, original_macho, eros_original, crotts1996}.
+The microlensing event rate in a given region of the sky depends on the space and velocity distributions of the sources and on the space, velocity, and mass distribution of the lenses.
+To maximize the event rate, microlensing surveys have typically focused on high-density stellar regions such as the Galactic Bulge, M31, and the Magellanic Clouds \citep[e.g.][]{original_ogle, original_macho, eros_original, crotts1996}.
 Microlensing events are of course not limited to dense stellar fields, but only one event outside of such fields has been recorded: the so-called Tago event \citep{fukui2007, gaudi2008}.
 \cite{gaudi2008} were able to establish probabilistic limits on the mass, distance, proper motion, and magnitude of this lens using existing observations near the location of the event.
-With \textit{JHK} magnitudes of the lens itself or a direct measurement of the proper velocity of the lens, as will be possible within the next year, these limits can be turned into precise measurements.
+With \textit{JHK} magnitudes of the lens itself or a direct measurement of the proper motion of the lens, as will be possible within the next year, these limits can be turned into precise measurements.
 Such a detailed analysis is not normally possible for events in the previously mentioned high-density regions where the lenses are much farther away, and blending makes precise photometry difficult.
 
-Not all events away from the Bulge will be as fortuitous as the Tago event, but lensing events away from high density fields are generally expected to have closer sources, closer lenses, and less crowded background fields compared to those found by modern microlensing surveys \citep{mesolensing}.
+Lensing events away from high density fields are generally expected to involve closer sources, closer lenses, and less crowded background fields compared to those found by modern microlensing surveys \citep{mesolensing}.
 Thus, the properties of the lenses are more likely to be well-constrained by direct observation, possibly enabling mass measurements of objects not bound in binary pairs.
@@
 We apply previous selection methods for identifying microlensing event signatures in light curves to the data and compare these methods with an event identification procedure based on a set of variability statistics (\S\ref{sec:event_recovery}).
-We then present the detection efficiency for these methods as computed for a representative set of PTF fields.
-In \S4 we apply a set of criteria that use the results of these tests to all PTF light curves with $>$100 high-quality observations to identify candidate microlensing events and present three high confidence candidates.
-We conclude in \S5.
+[APW: We then present the detection efficiency for these methods as computed for a representative set of PTF fields.
+In \S4 we apply a set of criteria that use the results of these tests to all PTF light curves with $>$100 high-quality observations to identify candidate microlensing events and present three high confidence candidates.] We conclude in \S5.
 
@@
 Each field has a unique sampling pattern determined by (1) which of the PTF sub-surveys it belongs to, (2) which time of the year it is visible from Palomar, and (3) how high a priority it is given by the scheduler.
-92\% of the P48 time is distributed between three  surveys: a 5-day cadence experiment covering 8000~deg$^2$ at $|b|>30^\circ$ to search for extragalactic transients, a dynamic cadence experiment to identify short duration ($<$5~days) transients, and a high-cadence imaging campaign of a single field in Orion to find transiting Jupiter-mass planets around young stars \citep{nick2009}.
+The survey design strategy was to allocate 92\% of the P48 time between three surveys: a 5-day cadence experiment covering 8000~deg$^2$ at $|b|>30^\circ$ to search for extragalactic transients, a dynamic cadence experiment to identify  short duration ($<$5~days) transients, and a high-cadence imaging campaign of a single field in Orion to find transiting Jupiter-mass planets around young stars \citep{nick2009}.
 As a result, PTF light curves often contain gaps, regions of high-cadence observations, and regions of low-cadence observations, leading to a massive data set of irregularly-sampled, time-domain photometry.
-\figurename~\ref{fig:sampling} shows six randomly-selected light curves (with their magnitudes removed) and demonstrates the varying cadences and coverage that different fields may have over the same one-year baseline.
+\figurename~\ref{fig:sampling} shows six randomly-selected light curves and demonstrates the varying cadences and coverage that different fields may have over the same one-year baseline.
 We restrict our sample to the \apwsim$6.4\times10^7$ light curves with $>$25 \textit{R}-band observations.
@@
 What would a microlensing event look like in a typical PTF light curve?
-Assuming that the source is an un-blended point source and the lens is a foreground, dim object, a microlensing event is fully described by three parameters: the angular impact parameter $u_0$, the peak time of the event $t_0$, and the timescale of the event (Einstein crossing time) $t_E$ such that the flux is given by:
+If the source is an un-blended point source and the lens is a foreground, dim object, a microlensing event is fully described by three parameters: the angular impact parameter $u_0$, the peak time of the event $t_0$, and the timescale of the event (Einstein crossing time) $t_E$ such that the flux is given by:
 \begin{align}
@@
 \end{align}
-where $A(u(t))$ is the amplification factor. The microlensing perturbation can also be expressed in terms of magnitudes as:
+where $A$ is the amplification factor. The microlensing perturbation can also be expressed in terms of magnitudes as:
 \begin{align}
@@
 \end{align}
-where $m_0(t)$ is the unperturbed, possibly time-variable magnitude of the source.
+where $m_0(t)$ is the unperturbed (but possibly time-variable) magnitude of the source.
 
@@
 When $u_0$ is small ($u_0<<1$), the amplification is large.
-But even in cases of high amplification, a survey may miss or poorly sample an event if $t_E$ is short ($t_E \lesssim\mathrm{days}$); if $t_E$ is long ($t_E \gtrsim 1~\mathrm{year}$) the event may be confused with other forms of long-duration variability.
-\figurename~\ref{fig:microlensing_sim} shows a well-sampled PTF light curve and illustrates what a microlensing event with fixed $t_E=20~\mathrm{days}$ and varied $u_0$ might look like in the data.
+But even in cases of high amplification, a survey may miss or poorly sample an event if $t_E$ is short ($t_E \lesssim\mathrm{days}$), while if $t_E$ is long ($t_E \gtrsim 1~\mathrm{year}$) the event may be confused with other forms of long-duration variability.
+\figurename~\ref{fig:microlensing_sim} shows a well-sampled PTF light curve and illustrates what a microlensing event with a fixed $t_E=20~\mathrm{days}$ but different $u_0$ might look like in the data.
 
@@
 Light curves of any transient sources are then analyzed and vetted using a variety of selection methods to search for microlensing event candidates and distinguish them from e.g., variable stars, outbursting systems, and novae.
-Different surveys approach this problem differently (as described in \citealt{ogle_optical_depth, con_idx, alcock2000, macho_detection_efficiency, hamadache2009, wyrzykowski2009, sumi2011}), but the general idea is to:
+Different surveys approach this problem differently (as described in \citealt{ogle_optical_depth, con_idx, alcock2000, macho_detection_efficiency, hamadache2009, wyrzykowski2009, sumi2011}), but the general idea is to require that:
 
 \begin{enumerate}
-	\item require that any selected light curve have some number of consecutive datapoints brighter than some threshold;
-	\item require that a microlensing model best describe the data, as determined by using a $\chi^2$ cut;
-	\item require that the model event parameters have reasonable values.
+	\item any selected light curve have some number of consecutive datapoints brighter than some threshold,
+	\item a microlensing model best describe the data, as determined by using a $\chi^2$ cut, and
+	\item the model event parameters have reasonable values.
 \end{enumerate}
 
-This process is ideal for studying crowded fields such as the Galactic Bulge where photometry is difficult.
+This process is ideal for studying crowded fields where photometry is difficult.
 It also limits the number of light curves that must be fit with a model, thus saving on computation time and allowing for real-time detection of events, so that high-cadence follow-up observations of interesting sources can be triggered.
-Microlensing surveys also have the advantage of relatively uniform time sampling of their survey footprint over an observing season.
 
 Applying this prescription to the PTF data presents some obvious challenges.
-In particular, the consecutive points requirement would seem hard to use when there can be significant (and irregular) gaps in the data.
+Microlensing surveys have the advantage of relatively uniform time sampling of their survey footprint over an observing season, but it is hard to apply a cut on consecutive points on survey data with significant (and irregular) gaps.
 In order to develop the most successful prescription for PTF data, we set out to examine the relative performance of a set of variability statistics in selecting simulated microlensing events.
@@
 \subsection{Event Selection} \label{sec:event_selection}
-We choose five statistical measures of variability compiled by \cite{shin2009} that have been previously applied to classification and discovery of periodic variables, and evaluate their effectiveness in recovering simulated microlensing events in the PTF data.
+We choose five statistical measures of variability compiled by \cite{shin2009} that have been previously applied to the classification and discovery of periodic variables, and evaluate their effectiveness in recovering simulated microlensing events in the PTF data.
 These variability indices are $\sigma/\mu$, $Con$, $\eta$, $J$, and $K$.\footnote{We do not to implement the sixth index described by \cite{shin2009}, $AoVM$, because it mainly helps with periodic sources.} $\sigma/\mu$ is the ratio of the root-variance to the sample mean,
@@
 
-We modify the definition of $Con$ such that it is the number of clusters of three or more consecutive observations that are more than $3\sigma$ brighter than the reference magnitude of the source (e.g., for a single microlensing event in an otherwise flat light curve $Con=1$).
+We modify the definition of $Con$ such that it is the number of clusters of three or more consecutive observations that are more than $3\sigma$ brighter than the reference magnitude of the source (e.g., for a single microlensing event in an otherwise flat light curve, $Con=1$).
 This change allows us to use the performance of $Con$ as a proxy for the consecutive point requirement described above.
 
-$\eta$, the von Neumann ratio \citep{von_neumann1941}, is the mean square successive difference over the sample variance:
+$\eta$, the von Neumann ratio \citep{von_neumann1941}, is the mean square successive difference divided by the sample variance:
 \begin{align}
@@
 
-\figurename~\ref{fig:indices_examples} shows maximal outlier light curves for each variability statistic selected from \apwsim20,000 light curves for objects on a single CCD in PTF field 100049.
-Clearly, the indices are sensitive to different aspects of variability in the data, although in this case $\Delta \chi^2$, $Con$, $K$, and $\sigma/\mu$ all finger the same object as being maximally variable.
+\figurename~\ref{fig:indices_examples} shows maximal outlier light curves for each variability statistic selected from \apwsim20,000 light curves for objects on a single CCD in PTF field 100101.
+Clearly, the indices are sensitive to different aspects of variability in the data, though for some fields a single object's variability is so large that multiple indices select out the same object.
 We would expect $\sigma/\mu$ to be more useful for discerning periodic or semi-periodic variability where the variance is large; however, the expectations are not so clear for other indices.
@@
 To determine the regions of parameter space where microlensing light curves fall, we conduct Monte Carlo simulations and inject artificial microlensing events into real PTF light curves and compare the distributions of variability indices for a set of light curves both with and without these simulated events.
-\figurename~\ref{fig:var_indices} shows projections of the distribution for various combinations of variability indices.
+\figurename~\ref{fig:var_indices} shows projections of the distribution for various combinations of the variability indices.
 These simulations do define regions to search for microlensing events, but there is still a large amount of overlap between the distributions with and without microlensing events.
@@
 In order to compute the detection efficiency, $\varepsilon$, for a given selection method, we must first define the selection criteria for each variability index.
-For a given PTF field, we randomly sample 1000 light curves with good observations\citep[these have no photometric flags; see description of processing pipeline in][]{nick2009}.
-For each of these light curves, we then use the time and error information to simulate 100 light curves with purely Gaussian scatter.
-We find the value of each variability index such that the selection returns 1\% of these simulated data --- i.e., we set the limiting value of each variability index such that the false positive rate (FPR) is 1\%.
+For a given PTF field, we randomly sample 1000 light curves with good observations \citep[these have no photometric flags; see description of processing pipeline in][]{nick2009}.
+For each of these light curves, we then use the time and magnitude error information to simulate 100 light curves with purely Gaussian scatter.
+We find the value of each variability index such that the selection using that index to identify interesting light curves returns 1\% of these simulated data --- i.e., we set the limiting value of each variability index such that the false positive rate (FPR) is 1\%.
 
-We then run additional Monte Carlo simulations in which we artificially add microlensing events to real data from the same field and evaluate how often these events are recovered given the cuts as defined above.
+We then run additional Monte Carlo simulations in which we artificially add microlensing events to real data from the same field and evaluate how often these events are recovered given the cuts defined above.
 For a given field, we randomly sample 1000 light curves from each CCD with $>$25 good observations.
@@
 
-The parameters for each simulated event are chosen as follows: the peak time, $t_0$, is drawn from a uniform distribution between the minimum and maximum observation times, the impact parameter, $u_0$, is drawn from a uniform distribution between 0 and 1.34 (cite?), and the event timescale, $t_E$, is drawn from a log-uniform distribution between 1 and 1000 days.
+The parameters for each simulated event are chosen as follows: $t_0$ is drawn from a uniform distribution between the minimum and maximum observation times, $u_0$ is drawn from a uniform distribution between 0 and 1.34 ([TODO:] cite?), and $t_E$ is drawn from a log-uniform distribution between 1 and 1000 days.
 For each iteration we recompute the variability statistics and store the event parameters.
@@
 \section{Searching for events in the full PTF data set}
+% FLUSH OUT AND UPDATE THIS SECTION
 As an initial test, we applied the 1\% FPR selection on the 325 PTF fields (\apwsim2400 deg$^2$) with $>$100 $R$-band observations using the variability index $\eta$. Our procedure is as follows:
@@
 	\item compute $\eta$ for each light curve, and select all light curves that pass the cut determined from the previous simulations;
-	\item for each selected light curve, fit a microlensing event model and subtract the model, then recompute $\eta$;
-	\item if the new value of $\eta$ still passes the cut, we reject the light curve (it is probably a periodic variable);
+	\item fit a microlensing event model and subtract the model, then recompute $\eta$ for each selected light curve;
+	\item we reject the light curve if the new value of $\eta$ still passes the cut (it is probably a periodic variable);
 	\item any remaining light curves are loaded into a local database to be vetted by hand.
@@
 From the \apwsim$10^7$ light curves in the 325 field footprint, this procedure selected 6306 candidates.
-For each candidate, we look at the SDSS colors (if the PTF field is in the SDSS footprint), search SIMBAD for any nearby sources, and visually inspect the PTF images to vet the data.
+For each candidate, we look at the SDSS colors (if the PTF field is in the SDSS footprint), search SIMBAD for any nearby sources within 10 arcseconds, and visually inspect the PTF images to vet the data.
 We require that the sources are not associated with SDSS galaxies so we may eliminate supernovae from our sample, but acknowledge the possibility of contamination from supernovae in faint galaxies.

### 2012-10-15 candidate_pipeline.py
remove logging

@@ -111,10 +111,10 @@ def compute_selection_criteria_for_field(field, fpr=0.01, number_of_fpr_light_cu
         for source_id in source_ids:
             light_curve = ccd.light_curve(source_id, barebones=True, clean=True)
             if len(light_curve.mjd) < min_number_of_good_observations: 
-                logger.debug("\tRejected source {}".format(source_id))
+                #logger.debug("\tRejected source {}".format(source_id))
                 continue
                 
-            logger.debug("\tSelected source {}".format(source_id))
+            #logger.debug("\tSelected source {}".format(source_id))
             these_indices = vi.simulate_light_curves_compute_indices(light_curve, num_simulated=number_of_fpr_simulations_per_light_curve, indices=indices)
             try:
                 simulated_light_curve_statistics = np.hstack((simulated_light_curve_statistics, these_indices))
@@ -246,6 +246,10 @@ def select_candidates(field, selection_criteria):
                (round(light_curve.u0, 2) < 1.34) and (round(light_curve.tE) > 2.) and \
                (indices["delta_chi_squared"] > 250.) and (light_curve.tE < 2.*baseline):
                 candidates.append(light_curve)
+            
+            elif (indices["eta"] <= lower_cut) and (new_eta <= lower_cut) and (indices["delta_chi_squared"] > 100.) and (indices["j"] > 250):
+                light_curve.tags = ["variable star candidate"]
+                candidates.append(light_curve)
         
         ccd.close()
     
@@ -262,6 +266,10 @@ def save_candidates_to_mongodb(candidates, collection):
             continue
         
         light_curve_document = mongo.light_curve_to_document(candidate, indices=candidate.indices)
+        try:
+            light_curve_document["tags"] += candidate.tags
+        except AttributeError:
+            pass
         mongo.save_light_curve_document_to_collection(light_curve_document, collection)
     
     return True
@@ -287,7 +295,7 @@ if __name__ == "__main__":
                     help="Number of light curves to select from each CCD.")
     parser.add_argument("--num-simulations", dest="num_simulations", default=100, type=int,
                     help="Number of simulations per light curve.")
-    parser.add_argument("--min-observations", dest="min_num_observations", default=100, type=int,
+    parser.add_argument("--min-observations", dest="min_num_observations", default=50, type=int,
                     help="The minimum number of observations")
     
     args = parser.parse_args()
@@ -312,7 +320,7 @@ if __name__ == "__main__":
[…]

### 2012-10-15 paper_figures.py
massive update of bad git form

@@ -31,6 +31,8 @@ logger.addHandler(ch)
 # Project
 from ptf.simulation.simulatedlightcurve import SimulatedLightCurve
 from ptf.ptflightcurve import PTFLightCurve
+import ptf.globals as pg
+import ptf.variability_indices as vi
 
 try:
     import ptf.photometricdatabase as pdb
@@ -140,9 +142,9 @@ def microlensing_event_sim():
     fig.subplots_adjust(hspace=0.0, left=0.1, right=0.9)
     fig.savefig("plots/simulated_events.pdf", bbox_inches="tight", facecolor="white")
 
-def median_maximum_indices_plot(field_id, ccd_id=5):
-    """ Given a field ID, produce a figure with 10 panels: left column are light curves
-        with median values of the variability indices, right column are light curves with
+'''
+def maximum_outlier_indices_plot(field_id, ccd_id=5):
+    """ Given a field ID, produce a figure with light curves for the 
         maximum outlier values of the indices.
     """
     
@@ -151,18 +153,18 @@ def median_maximum_indices_plot(field_id, ccd_id=5):
     
     pickle_filename = "data/paper_figures/var_indices_field{}_ccd{}.pickle".format(field_id, ccd_id)
     
-    indices = ["eta", "delta_chi_squared", "sigma_mu", "j", "k", "con"]
+    indices = ["eta", "delta_chi_squared", "sigma_mu", "j", "k"]
     if not os.path.exists(pickle_filename):
         chip = ccd.read()
         
-        source_ids = chip.sources.readWhere("ngoodobs > 25")["matchedSourceID"]
+        source_ids = chip.sources.readWhere("ngoodobs > 600")["matchedSourceID"]
         logger.info("{} source ids selected".format(len(source_ids)))
         
         var_indices = []
         for source_id in source_ids:
             light_curve = ccd.light_curve(source_id, clean=True, barebones=True)
             
-            if len(light_curve.mjd) > 25:
+            if len(light_curve.mjd) > 500:
                 logger.debug("Source ID {} is good".format(source_id))
                 try:
                     lc_var_indices = (source_id,) + pa.compute_variability_indices(light_curve, indices, return_tuple=True)
@@ -181,106 +183,112 @@ def median_maximum_indices_plot(field_id, ccd_id=5):
     f = open(pickle_filename, "r")
     var_indices = pickle.load(f)
     f.close()
-        
[…]

### 2012-10-15 ptf/analyze/analyze.py

@@ -80,6 +80,7 @@ def fit_subtract_microlensing(light_curve):
     light_curve.t0 = params["t0"].value
     light_curve.u0 = params["u0"].value
     light_curve.m0 = params["m0"].value
+    light_curve.chisqr = result.chisqr
     
     return light_curve_new
 

### 2012-10-15 ptf/photometricdatabase.py

@@ -242,7 +242,7 @@ class Field(object):
         for ccdid,ccd in self.ccds.items():
             chip = ccd.read()
             exposures_per_ccd[ccdid] = chip.exposures[:]
-            ccd.close()
+            #ccd.close()
         
         return exposures_per_ccd
     
@@ -256,7 +256,7 @@ class Field(object):
             chip = ccd.read()
             mjds = np.sort(chip.exposures.col("obsMJD"))
             baseline_per_ccd[ccdid] = mjds[-1]-mjds[0]
-            ccd.close()
+            #ccd.close()
         
         return baseline_per_ccd
     

### 2012-10-15 ptf/ptflightcurve.py

@@ -76,7 +76,9 @@ class PTFLightCurve(object):
             plt.show()
             return
         
-        ax.errorbar(self.mjd, self.mag, self.error, ecolor='0.7', capsize=0, **kwargs)
+        # APW Hack
+        idx = np.where(self.error < 0.5)
+        ax.errorbar(self.mjd[idx], self.mag[idx], self.error[idx], ecolor='0.7', capsize=0, **kwargs)
         ax.set_ylim(ax.get_ylim()[::-1])
         return ax
     
@@ -106,6 +108,12 @@ class PTFLightCurve(object):
         else:
             raise ValueError("I don't know how to handle {} files!".format(ext))
 
+    def slice_mjd(self, min=None, max=None):
+        """ Select out a part of the light curve between two MJD values -- INCLUSIVE """
+        idx = (self.mjd >= min) & (self.mjd <= max)
+        
+        return PTFLightCurve(mjd=self.mjd[idx], mag=self.mag[idx], error=self.error[idx])
+        
 class PDBLightCurve(PTFLightCurve):
     """ Subclass of PTFLightCurve that requires a field_id, ccd_id, and source_id """
     

### 2012-10-15 ptf/variability_indices.py

@@ -86,11 +86,12 @@ def simulate_events_worker(sim_light_curve, tE, u0, reference_mag, indices):
     light_curve.reset()
     
     # APW HACK -- add microlensing events only on a data point, ignores the *survey* detection efficiency
-    t0 = light_curve.mjd[np.random.randint(len(light_curve.mjd))]
-    #t0 = None
+    #t0 = light_curve.mjd[np.random.randint(len(light_curve.mjd))]
+    t0 = None
     light_curve.addMicrolensingEvent(tE=tE, u0=u0, t0=t0)
     
     lc_var_indices = analyze.compute_variability_indices(light_curve, indices, return_tuple=True) + (light_curve.tE, light_curve.u0, reference_mag)
+    
     return lc_var_indices
 
 def simulate_events_compute_indices(light_curve, events_per_light_curve, indices, u0=None):
@@ -108,17 +109,19 @@ def simulate_events_compute_indices(light_curve, events_per_light_curve, indices
     def callback(result):
         var_indices_with_events.append(result)
     
+    # APW HACK
     pool = multiprocessing.Pool(processes=8)
     for event_id in range(events_per_light_curve):
         tE = 10**np.random.uniform(0.3, 3.)
         pool.apply_async(simulate_events_worker, args=(sim_light_curve, tE, u0, reference_mag, indices), callback=callback)
-        #callback(simulate_light_curves_worker(sim_light_curve, tE, u0, reference_mag, indices))
+        #callback(simulate_events_worker(sim_light_curve, tE, u0, reference_mag, indices))
     
     pool.close()
     pool.join()
     
     dtypes = [(index,float) for index in indices] + [("tE",float),("u0",float),("m",float)]
     var_indices_with_events = np.array(var_indices_with_events, dtype=dtypes)
+    
     return var_indices_with_events
 
 def simulate_light_curves_worker(sim_light_curve, indices):

### 2012-10-15 candidate_pipeline.py
clean up comments

@@ -224,11 +224,6 @@ def select_candidates(field, selection_criteria):
         for source in sources:
             light_curve = ccd.light_curve(source["matchedSourceID"], barebones=True, clean=True)
             
-            #if source["matchedSourceID"] == 2126:
-            #    srcd = chip.sourcedata.readWhere("matchedSourceID == 2126")
-            #    print light_curve.mag
-            #    sys.exit(0)
-            
             # If light curve doesn't have enough clean observations, skip it
             if len(light_curve.mjd) < min_number_of_good_observations: continue
             

### 2012-10-15 paper_figures.py

@@ -582,9 +582,7 @@ if __name__ == "__main__":
     np.random.seed(104)
     #make_survey_sampling_figure(10)
     #microlensing_event_sim()
-    #maximum_outlier_indices_plot(100049)
-    #maximum_outlier_indices_plot(100101) # APW TESTING
-    #maximum_outlier_indices_plot(100078) # APW TESTING
+    #maximum_outlier_indices_plot(100101)
     
     #variability_indices_distributions()
     #Don't use this! variability_indices_distributions_easy()

### 2013-01-24 scripts/event_fitter.py
long overdue

@@ -39,6 +39,7 @@ logger.addHandler(ch)
 # Third-party
 from apwlib.globals import greenText, yellowText, redText
 import numpy as np
+np.seterr(invalid="ignore")
 import emcee
 import matplotlib.gridspec as gridspec
 import matplotlib
@@ -221,9 +222,9 @@ def fit_model_to_light_curve(light_curve, nwalkers=200, nburn_in=100, nsamples=2
                                     args=[light_curve.mag, light_curve.mjd, light_curve.error],
                                     threads=4)
     pos, prob, state = sampler.run_mcmc(p0, nburn_in)
-    sampler.reset()
 
-    sampler.run_mcmc(pos, nsamples)
+    sampler.reset()
+    pos, prob, state = sampler.run_mcmc(pos, nsamples)
 
     logger.info("Mean acceptance fraction: {0:.3f}".format(np.mean(sampler.acceptance_fraction)))
 
@@ -234,10 +235,12 @@ def make_chain_distribution_figure(light_curve, sampler, filename=None):
 
     fig, axes = plt.subplots(2, 2, figsize=(10,14))
     param_to_label = {"m0" : r"$M_0$", "u0" : r"$u_0$", "t0" : r"$t_0$", "tE" : r"$t_E$"}
-    params = np.array([["m0", "u0"], ["t0", "tE"]])
+    params = ["m0", "u0", "t0", "tE"]
 
+    #params = np.array([["m0", "u0"], ["t0", "tE"]])
     max_idx = np.ravel(sampler.lnprobability).argmax()
 
+    """
     for ii in range(2):
         for jj in range(2):
             idx = 2*ii + jj
@@ -249,6 +252,53 @@ def make_chain_distribution_figure(light_curve, sampler, filename=None):
             axes[ii,jj].axvline(chain[:,idx][max_idx], color='r', linestyle='--')
             axes[ii,jj].set_title("{}".format(param_to_label[params[ii,jj]]), fontsize=26)
             axes[ii,jj].yaxis.set_ticks([])
+    """
+
+    # m0
+    idx = 0
+    mn = np.mean(chain,axis=0)[idx]
+    std = np.std(chain,axis=0)[idx]
+    bins = np.linspace(mn - 2*std, mn + 2*std, 100)
+
+    axes[0,0].hist(chain[:,idx], bins=bins, color="k", histtype="step")
[…]

### 2013-01-24 scripts/paper_figures.py

@@ -65,7 +65,7 @@ def make_survey_sampling_figure(N=10):
     field_ids = [100031, 2794, 4252, 100085, 110002, 3685]
     rand_fields = [pdb.Field(f, "R") for f in field_ids]
 
-    fig = plt.figure(figsize=(10,10))
+    fig = plt.figure(figsize=(12,8))
     ax = fig.add_subplot(111)
 
     min_mjd = 55250
@@ -80,7 +80,7 @@ def make_survey_sampling_figure(N=10):
         source_id = source_ids[0]
 
         logger.info("Reading source {} on {}, {}".format(source_id, field, ccd))
-        light_curve = ccd.light_curve(source_id, clean=False, barebones=True, where=["(mjd >= {})".format(min_mjd), "(mjd <= {})".format(max_mjd)])
+        light_curve = ccd.light_curve(source_id, clean=False, barebones=True)#, where=["(mjd >= {})".format(min_mjd), "(mjd <= {})".format(max_mjd)])
 
         ax.plot(light_curve.mjd, [ii]*len(light_curve.mjd), color="black", alpha=0.3, marker="o", markersize=7, linestyle="none", markeredgecolor="none")
         ax.text(55300, ii+0.1, "{} observations".format(len(light_curve.mjd)), size=14)
@@ -95,8 +95,8 @@ def make_survey_sampling_figure(N=10):
     ax.set_xlim(min_mjd, max_mjd)
     ax.set_ylim(-0.5, ii+0.5)
 
-    ax.set_xlabel("Time [day]", fontsize=20)
-    fig.savefig(os.path.join(pg.plots_path, "sampling_figure.pdf", facecolor="white", bbox_inches="tight"))
+    ax.set_xlabel("Time [days]", fontsize=20)
+    fig.savefig(os.path.join(pg.plots_path, "sampling_figure.pdf"), facecolor="white", bbox_inches="tight")
 
 def microlensing_event_sim():
     """ Create the multi-panel figure with simulated microlensing events for a single
@@ -146,6 +146,9 @@ def microlensing_event_sim():
 
         #for tick in axes[ii].yaxis.get_major_ticks():
         #    tick.label.set_fontsize(tick_font_size)
+
+        if ii == 0:
+            [tick.set_visible(False) for jj,tick in enumerate(axes[ii].get_yticklabels()) if jj % 2 != 0]
 
         if ii % 2 != 0:
             axes[ii].yaxis.tick_right()
[…]

### 2013-01-24 ptf/lightcurve/lightcurve.py
u0 to 1.5

@@ -277,7 +277,7 @@ class SimulatedLightCurve(PTFLightCurve):
 
         # If u0 is not specified, draw from u0 distribution
         #   - see for example Popowski & Alcock
-        if u0 == None: self.u0 = np.random.uniform(0., 1.34)
+        if u0 == None: self.u0 = np.random.uniform(0., 1.5)
         else: self.u0 = float(u0)
 
         # If t0 is not specified, draw from uniform distribution between days

### 2013-01-24 text/paper/ms_v3.tex

--- 2012-10-15 text/paper/ptf_microlensing.tex
@@
-%\documentclass{emulateapj}
-\documentclass[12pt,preprint]{aastex}
+\documentclass{emulateapj}
+%\documentclass[12pt,preprint]{aastex}
 \newcounter{address}
 \setcounter{address}{1}
-%\usepackage{lscape, longtable}
-\usepackage{lmodern}
-%\usepackage[T1]{fontenc}
-\usepackage{graphicx}
+
 \usepackage{amsmath}
@@
 \usepackage{amsfonts}
-\usepackage{subfigure}
 \usepackage{natbib}
-
-% for flowchart
-\usepackage{tikz}
-\usetikzlibrary{shapes,arrows}
+\usepackage{longtable, hyperref, graphicx, subfigure, epstopdf, multirow, setspace, verbatim, lmodern} 
+\usepackage{aas_macros}
 
@@
 \altaffiltext{\theaddress}{\stepcounter{address}\label{col} Department of Astronomy, Columbia University, 550 W 120th St., New York, NY 10027, USA}
-\altaffiltext{\theaddress}{\stepcounter{address}\label{nsf} NSF graduate fellow}
 \altaffiltext{\theaddress}{\stepcounter{address}\label{ucsb} Department of Physics, Broida Hall, University of California, Santa Barbara, CA 93106, USA}
@@
 %\altaffiltext{\theaddress}{\stepcounter{address}\label{dun} Dunlap Fellow}
+\altaffiltext{\theaddress}{\stepcounter{address}\label{nsf} NSF Graduate Fellow}
 \altaffiltext{\theaddress}{\stepcounter{address}\label{eins} Einstein Fellow}
@@
 In the case of the Palomar Transient Factory (PTF), several such sub-surveys are being conducted in parallel, leading to extremely non-uniform sampling over the survey's nearly $20,000~\mathrm{deg}^2$ footprint.
-While the median $7.26~\mathrm{deg}^2$ PTF field has been imaged 20 times in \textit{R}-band, \apwsim$2300~\mathrm{deg}^2$ have been observed more than 100 times.
-[We use the existing PTF data (\apwsim$6.4\times10^7$ light curves as of September 2012) to study the trade-off between searching for microlensing events in a survey footprint that is much larger than most microlensing searches, but with far-from-optimal sampling.] To examine the probability that microlensing events can be recovered in these data, we test previous statistics used on uniformly sampled data to identify variables and transients.
+While the median $7.26~\mathrm{deg}^2$ PTF field has been imaged \apwsim40 times in \textit{R}-band, \apwsim$2300~\mathrm{deg}^2$ have been observed more than 100 times.
+We use the existing PTF data to study the trade-off between searching for microlensing events in a survey footprint that is much larger than most microlensing searches, but with far-from-optimal sampling.
+To examine the probability that microlensing events can be recovered in these data, we test previous statistics used on uniformly sampled data to identify variables and transients.
 We find that one such statistic, the von Neumann ratio, performs best for identifying simulated microlensing events.
-[We develop a selection method using this statistic and apply it to data from all PTF fields with $>$100 observations to uncover a number of interesting candidate events.] This work can help constrain all-sky event rate predictions and tests microlensing signal recovery in large datasets, both of which will be useful to future wide-field, time-domain surveys such as the LSST.
+We develop a selection method using this statistic and apply it to data from all PTF fields with $>$10 observations (a total of $1.1\times10^9$ light curves) to uncover a number of interesting candidate events.
+This work can help constrain all-sky event rate predictions and tests microlensing signal recovery in large datasets, both of which will be useful to future wide-field, time-domain surveys such as the LSST.
 
@@
 \section{Introduction}
-Over the past 20 years, microlensing, in which gravitational lensing results in a transient increase in the flux from a background source, has been used to search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, to study Galactic structure and kinematics \citep{binney2000}, and, most recently.
-to identify extrasolar planets \citep{gaudi2011}.
-Such studies were once limited by the small number of detected events, but with advances in CCD technology and the development of dedicated microlensing surveys \citep{original_ogle, original_macho, eros_original}, $>$1000 events are now detected each year.
+Over the past 20 years, microlensing, in which gravitational lensing causes a transient increase in the flux from a background source, has been used to search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, to study Galactic structure and kinematics \citep{binney2000}, and, most recently, to identify extrasolar planets \citep[][and references therein]{gaudi2011}.
+Such studies were once limited by the small number of detected events, but with advances in CCD technology and the development of dedicated microlensing surveys \citep[e.g.,][]{original_ogle, original_macho, eros_original}, $>$1000 events are now detected each year.
 In the coming decade, this number is expected to rise as the next generation of photometric, time-domain surveys come online, providing an opportunity for the precise study of otherwise hard-to-characterize objects such as low-mass stars, substellar objects, and isolated neutron stars.
 
-The microlensing event rate in a given region of the sky depends on the space and velocity distributions of the sources and on the space, velocity, and mass distribution of the lenses.
+The microlensing event rate for a given region of the sky depends on the space and velocity distributions of the sources and on the space, velocity, and mass distribution of the lenses.
 To maximize the event rate, microlensing surveys have typically focused on high-density stellar regions such as the Galactic Bulge, M31, and the Magellanic Clouds \citep[e.g.][]{original_ogle, original_macho, eros_original, crotts1996}.
-Microlensing events are of course not limited to dense stellar fields, but only one event outside of such fields has been recorded: the so-called Tago event \citep{fukui2007, gaudi2008}.
-\cite{gaudi2008} were able to establish probabilistic limits on the mass, distance, proper motion, and magnitude of this lens using existing observations near the location of the event.
-With \textit{JHK} magnitudes of the lens itself or a direct measurement of the proper motion of the lens, as will be possible within the next year, these limits can be turned into precise measurements.
-Such a detailed analysis is not normally possible for events in the previously mentioned high-density regions where the lenses are much farther away, and blending makes precise photometry difficult.
+Microlensing events are of course not limited to dense stellar fields, and lensing events away from these fields are potentially very interesting.
+Such nearby events are expected to involve closer sources, closer lenses, and less crowded backgrounds compared to those found by modern microlensing surveys \citep{mesolensing}.
+Thus, the properties of the lenses are more likely to be well-constrained by observations, enabling precious mass measurements.
 
-Lensing events away from high density fields are generally expected to involve closer sources, closer lenses, and less crowded background fields compared to those found by modern microlensing surveys \citep{mesolensing}.
-Thus, the properties of the lenses are more likely to be well-constrained by direct observation, possibly enabling mass measurements of objects not bound in binary pairs.
-The event rate in low-density stellar regions will certainly be smaller than that near the Bulge \citep[e.g.,][]{wood_optical_depth, ogle_optical_depth, macho_optical_depth, eros_optical_depth}, but the opportunity for the precise study of even a handful of lenses would be extremely valuable.
+To date, however, only one event outside of a dense stellar field has been recorded: the Tago event \citep{fukui2007, gaudi2008}.
+\cite{gaudi2008} set probabilistic limits on the mass, distance, proper motion, and magnitude of this lens using existing observations near the location of the event.
+With \textit{JHK} magnitudes of the lens or a direct measurement of its proper motion, which may be possible today with high-resolution imaging, these limits can be turned into precise measurements.
+Such an analysis is not normally possible for events in high-density fields where the lenses are, on average, much farther away and blending makes photometry difficult.
+The event rate in low-density stellar regions will certainly be smaller than that near e.g., the Bulge \citep[][]{wood_optical_depth, ogle_optical_depth, macho_optical_depth, eros_optical_depth}, but uncovering even a handful of lenses that can be studied in detail would be extremely valuable.
 
-In this paper, we use the Palomar Transient Factory (PTF) survey data (described in \S2) as a test case for the frequency and detectability of microlensing events in all-sky, irregularly sampled time-domain surveys.
+\begin{figure*}[!ht]
+\centering\includegraphics[scale=0.4,trim=0 0 0 27, clip]{figures/R_coverage.pdf}
+\caption{PTF $R$-band survey footprint, in equatorial coordinates. The fields are color-coded by total number of observations.}\label{fig:survey_footprint}
+\end{figure*}
+
+In this paper, we use the Palomar Transient Factory (PTF) survey data (described in \S2) as a test case for studying the frequency and detectability of microlensing events in all-sky, irregularly-sampled, time-domain surveys.
 We apply previous selection methods for identifying microlensing event signatures in light curves to the data and compare these methods with an event identification procedure based on a set of variability statistics (\S\ref{sec:event_recovery}).
-[APW: We then present the detection efficiency for these methods as computed for a representative set of PTF fields.
-In \S4 we apply a set of criteria that use the results of these tests to all PTF light curves with $>$100 high-quality observations to identify candidate microlensing events and present three high confidence candidates.] We conclude in \S5.
+We then present the detection efficiency for these methods as computed for a representative set of PTF fields.
+In \S4 we apply a set of criteria that use the results of these tests to all PTF light curves with $>$10 high-quality observations to identify candidate microlensing events [APW: and present XXX high-confidence candidates.] We conclude in \S5.
 
-%General relativity tells us that mass alters the propagation of light.
-In particular, light from a background source passing some mass distribution will be deflected by an angle that depends on the distance between the source and mass distribution, as well as the structure and kinematics of the mass distribution itself.
-Microlensing refers to any unresolved gravitational lensing, where the images formed simply contribute to a transient increase in flux from the source.
-Since Einstein first mentioned gravitational microlensing \citep{einstein1936}, it has been applied to the search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, Galactic structure and kinematics \citep{binney2000}, and extrasolar planets \citep{gaudi2011}.
-Microlensing events were once relatively rare, but with advances in CCD technology and the help of several dedicated microlensing surveys \citep{original_ogle, original_macho, eros_original}, more than 1000 events are detected each year by several microlensing collaborations and surveys.
-Within the next decade, this number is expected to rise as the next generation of photometric, time-domain surveys come online.
-
-\section{PTF } % Better section header?
-The Palomar Transient Factory is a transient detection system incorporating three elements.
+\section{PTF: The Survey and The Data} % Better section header?
+PTF is a transient detection system incorporating three elements.
 The first is a wide-field survey camera mounted on the automated 48 inch Oschin Schmidt telescope (the P48) at Palomar Observatory, CA.
@@
 Under median seeing conditions (1.1$\arcsec$), observations in Mould $R$ or Sloan Digital Sky Survey \citep[SDSS;][]{york00} $g$ achieve 2.0$\arcsec$ full-width half-maximum images and reach 5$\sigma$ magnitudes of $R \approx 21.0$ and $g \approx 21.3$ mag in a standard 60~s exposure \citep{nick2010}.
-As of September 2012, the PTF footprint included $\apwsim$10,100 (2300) deg$^2$ imaged $>$25 ($>$100) times in $R$ and $\apwsim$2200 (170) deg$^2$ imaged that often in $g$ (\figurename~\ref{fig:survey_footprint})\footnote{For an interactive visualization of the PTF survey coverage, see: \url{http://bit.ly/PTFSky}}.
+As of November 2012, the PTF footprint included $\apwsim$15,224 (2766) deg$^2$ imaged $>$10 ($>$100) times in $R$ and $\apwsim$5430 (290) deg$^2$ imaged that often in $g$ (Figure~\ref{fig:survey_footprint})\footnote{For an interactive visualization of the PTF survey coverage, see: \url{http://bit.ly/PTFSky}.}.
 PTF is also a real-time data reduction pipeline that identifies transients of interest, passes these to a dedicated photometric follow-up telescope, and generates an archive of all detected sources \citep{nick2009,rau2009}.
-%[CHECK NUMBERS]
 
@@
 Each field has a unique sampling pattern determined by (1) which of the PTF sub-surveys it belongs to, (2) which time of the year it is visible from Palomar, and (3) how high a priority it is given by the scheduler.
-The survey design strategy was to allocate 92\% of the P48 time between three surveys: a 5-day cadence experiment covering 8000~deg$^2$ at $|b|>30^\circ$ to search for extragalactic transients, a dynamic cadence experiment to identify  short duration ($<$5~days) transients, and a high-cadence imaging campaign of a single field in Orion to find transiting Jupiter-mass planets around young stars \citep{nick2009}.
-As a result, PTF light curves often contain gaps, regions of high-cadence observations, and regions of low-cadence observations, leading to a massive data set of irregularly-sampled, time-domain photometry.
-\figurename~\ref{fig:sampling} shows six randomly-selected light curves and demonstrates the varying cadences and coverage that different fields may have over the same one-year baseline.
-We restrict our sample to the \apwsim$6.4\times10^7$ light curves with $>$25 \textit{R}-band observations.
+The survey design strategy was to allocate 92\% of the P48 time between three surveys: a 5-day cadence experiment covering 8000~deg$^2$ at $|b|>30^\circ$ to search for extragalactic transients, a dynamic cadence experiment to identify short duration ($<$5~days) transients, and a high-cadence imaging campaign of a field in Orion to find transiting planets around young stars \citep{nick2009}.
+PTF light curves often contain gaps, regions of high-cadence observations, and regions of low-cadence observations, leading to a massive data set of irregularly-sampled, time-domain photometry.
+Figure~\ref{fig:sampling} shows six randomly selected light curves and demonstrates the varying cadences and coverage that different fields may have over the same one-year baseline.
+Figure~\ref{fig:num_obs} shows the distributions of the total number of $R$-band observations and time baselines for all the PTF fields.
+Based on these plots, we restrict our sample to the \apwsim$1.1\times10^9$ light curves with $>$10 \textit{R}-band observations.
+% because there is a drop off in the number of fields with fewer than 10 observations and we don't expect such light curves to be useful for our search.
+%APW!
 
 What would a microlensing event look like in a typical PTF light curve?
-If the source is an un-blended point source and the lens is a foreground, dim object, a microlensing event is fully described by three parameters: the angular impact parameter $u_0$, the peak time of the event $t_0$, and the timescale of the event (Einstein crossing time) $t_E$ such that the flux is given by:
+If the source is an unblended point source and the lens is a foreground, dim object, a microlensing event is fully described by three parameters: the angular impact parameter $u_0$, the peak time of the event $t_0$, and the timescale of the event (Einstein crossing time) $t_E$.
+The amplification factor $A$ and flux $F$ can then be defined as a function of time:
 \begin{align}
-	F(t) &= A(t)\times F_{source} \\
-	A(u) &= \frac{u^2 + 2}{u\sqrt{u^2 + 4}}\\
-	u(t) &= \sqrt{u_0^2 + 2\Big(\frac{t-t_0}{t_E}\Big)}
+	u(t) &= \sqrt{u_0^2 + 2\Big(\frac{t-t_0}{t_E}\Big)},\\
+	A(t) &= \frac{u^2 + 2}{u\sqrt{u^2 + 4}},\\
+	F(t) &= A(t)\times F_{source}.
 \end{align}
-where $A$ is the amplification factor. The microlensing perturbation can also be expressed in terms of magnitudes as:
+The microlensing perturbation can also be expressed as:
 \begin{align}
@@
 
+\begin{figure}[!t]
+\begin{center}
+\includegraphics[width=0.45\textwidth, trim=2 2 5 5, clip]{figures/sampling_figure.pdf}
+\caption{Sampling of six different PTF fields over the course of a year.
+Darker points indicate that more exposures were taken that night; the total number of observations for each field is indicated.}\label{fig:sampling}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\centering\includegraphics[width=0.52\textwidth,trim=78 60 10 70, clip]{figures/num_observations_baseline.pdf}
+\caption{Distributions of the total number of $R$-band observations and time baselines for all the PTF fields. \\}
+\label{fig:num_obs}
+\end{figure}
+
 %However, if the event is \textit{too} bright, the pixels will saturate and there could be photometric errors flagging the most interesting data as bad.
-When $u_0$ is small ($u_0<<1$), the amplification is large.
-But even in cases of high amplification, a survey may miss or poorly sample an event if $t_E$ is short ($t_E \lesssim\mathrm{days}$), while if $t_E$ is long ($t_E \gtrsim 1~\mathrm{year}$) the event may be confused with other forms of long-duration variability.
-\figurename~\ref{fig:microlensing_sim} shows a well-sampled PTF light curve and illustrates what a microlensing event with a fixed $t_E=20~\mathrm{days}$ but different $u_0$ might look like in the data.
+Even in cases of high amplification ($u_0<<1$), a survey may miss or poorly sample an event if $t_E$ is short ($t_E \lesssim\mathrm{days}$), while if $t_E$ is long ($t_E \gtrsim 1~\mathrm{year}$) the event may be confused with other forms of long-duration variability.
+Figure~\ref{fig:microlensing_sim} illustrates the effect of simulated microlensing events with a fixed $t_E=20~\mathrm{days}$ but different $u_0$ on a well-sampled PTF light curve.
 
 \section{Microlensing event recovery} \label{sec:event_recovery}
-Microlensing surveys typically use difference difference image analysis \citep{alard1998} to identify transient events in raw imaging data.
+Microlensing surveys typically use difference image analysis \citep{alard1998} to identify transient events in raw imaging data.
 Light curves of any transient sources are then analyzed and vetted using a variety of selection methods to search for microlensing event candidates and distinguish them from e.g., variable stars, outbursting systems, and novae.
-Different surveys approach this problem differently (as described in \citealt{ogle_optical_depth, con_idx, alcock2000, macho_detection_efficiency, hamadache2009, wyrzykowski2009, sumi2011}), but the general idea is to require that:
-
+Surveys have approached this problem with various techniques (as described in \citealt{ogle_optical_depth, con_idx, alcock2000, macho_detection_efficiency, hamadache2009, wyrzykowski2009, sumi2011}), but the general idea is to require that:
 \begin{enumerate}
@@
 \end{enumerate}
-
-This process is ideal for studying crowded fields where photometry is difficult.
+Difference imaging is ideal for studying crowded fields where photometry is difficult.
 It also limits the number of light curves that must be fit with a model, thus saving on computation time and allowing for real-time detection of events, so that high-cadence follow-up observations of interesting sources can be triggered.
 
+\begin{figure}[!t]
+\centering\includegraphics[width=0.48\textwidth, trim=0 2 0 90, clip]{figures/simulated_events.pdf}
+\caption{Simulating the effect of a $t_E=20~\mathrm{days}$ microlensing event with differing impact parameters $u_0$ on a PTF light curve: a) the original light curve, b) the light curve with a $u_0=1.0$ event, c) with a $u_0=0.5$ event, d) with a $u_0=0.01$ event.
+The red dashed line shows the approximate saturation limit of the PTF camera in the $R$-band; such an event might therefore be missed by the survey.}\label{fig:microlensing_sim}
+\end{figure}
+
 Applying this prescription to the PTF data presents some obvious challenges.
-Microlensing surveys have the advantage of relatively uniform time sampling of their survey footprint over an observing season, but it is hard to apply a cut on consecutive points on survey data with significant (and irregular) gaps.
-In order to develop the most successful prescription for PTF data, we set out to examine the relative performance of a set of variability statistics in selecting simulated microlensing events.
+Microlensing surveys have the advantage of relatively uniform time sampling of their survey footprint over an observing season, but it is hard to justify a cut on consecutive points on survey data with significant (and irregular) gaps.
+In order to develop the most successful procedure for PTF data, we set out to examine the relative performance of a set of variability statistics in selecting simulated microlensing events.
 
@@
 We choose five statistical measures of variability compiled by \cite{shin2009} that have been previously applied to the classification and discovery of periodic variables, and evaluate their effectiveness in recovering simulated microlensing events in the PTF data.
-These variability indices are $\sigma/\mu$, $Con$, $\eta$, $J$, and $K$.\footnote{We do not to implement the sixth index described by \cite{shin2009}, $AoVM$, because it mainly helps with periodic sources.} $\sigma/\mu$ is the ratio of the root-variance to the sample mean,
+These variability indices are $\sigma/\mu$, $Con$, $\eta$, $J$, and $K$.\footnote{We do not to implement the sixth index described by \cite{shin2009}, $AoVM$, because it mainly helps identify periodic sources.} $\sigma/\mu$ is the ratio of the root-variance to the sample mean,
 \begin{align}
@@
 \end{align}
-where N is the total number of observations in the light curve.
+where N is the total number of observations.
 
-We modify the definition of $Con$ such that it is the number of clusters of three or more consecutive observations that are more than $3\sigma$ brighter than the reference magnitude of the source (e.g., for a single microlensing event in an otherwise flat light curve, $Con=1$).
+We modify the definition of $Con$ to be the number of clusters of three or more consecutive observations that are more than $3\sigma$ brighter than the reference magnitude of the source (e.g., for a single microlensing event in an otherwise flat light curve, $Con=1$).
 This change allows us to use the performance of $Con$ as a proxy for the consecutive point requirement described above.
@@
 \begin{align}
-	\delta_i &= \sqrt{\frac{N}{N-1}}\frac{x_i-\mu}{e_i}\\
-	J &= \sum^{N-1}_i sign(\delta_i \delta_{i+1})\sqrt{|\delta_i \delta_{i+1}|}\\
-	K &= \frac{1/N\sum^N_i |\delta_i|}{\sqrt{1/N\sum^N_i\delta_i^2}}
+	\delta_i &= \sqrt{\frac{N}{N-1}}\frac{x_i-\mu}{e_i},\\
+	J &= \sum^{N-1}_i sign(\delta_i \delta_{i+1})\sqrt{|\delta_i \delta_{i+1}|},\\
+	K &= \frac{1/N\sum^N_i |\delta_i|}{\sqrt{1/N\sum^N_i\delta_i^2}},
 \end{align}
@@
 
-We add one more variability index, $\Delta \chi^2$, defined as the difference in $\chi^2$ between fitting a Gaussian model and fitting a linear model to the light curve data.
+We add one more index, $\Delta \chi^2$, the difference in $\chi^2$ between fitting a Gaussian model and fitting a linear model to a light curve.
 This is the standard statistical test used by microlensing surveys, and allows us to compare the relative performance of the (slightly modified) \cite{shin2009} indices and of this approach.
 
-\figurename~\ref{fig:indices_examples} shows maximal outlier light curves for each variability statistic selected from \apwsim20,000 light curves for objects on a single CCD in PTF field 100101.
-Clearly, the indices are sensitive to different aspects of variability in the data, though for some fields a single object's variability is so large that multiple indices select out the same object.
-We would expect $\sigma/\mu$ to be more useful for discerning periodic or semi-periodic variability where the variance is large; however, the expectations are not so clear for other indices.
+Figure~\ref{fig:indices_examples} shows maximal outlier light curves for each variability statistic selected from \apwsim20,000 light curves for objects on a single CCD in PTF field 100049.
+Clearly, the indices are sensitive to different aspects of variability in the data.
+We expect $\sigma/\mu$ to be more useful for discerning periodic or semi-periodic variability where the variance is large; however, the expectations are not so clear for other indices.
 
-To determine the regions of parameter space where microlensing light curves fall, we conduct Monte Carlo simulations and inject artificial microlensing events into real PTF light curves and compare the distributions of variability indices for a set of light curves both with and without these simulated events.
-\figurename~\ref{fig:var_indices} shows projections of the distribution for various combinations of the variability indices.
+\begin{figure}
+\centering\includegraphics[width=.48\textwidth, trim=130 95 100 65, clip]{figures/max_outlier_light_curves.pdf}
+\caption{The maximum outlier light curves for each variability index applied to data from PTF field 100049.} \label{fig:indices_examples}
+\end{figure}
+
+We conduct Monte Carlo simulations and inject artificial microlensing events into real PTF light curves and compare the distributions of variability indices for a set of light curves both with and without these simulated events to determine the regions of parameter space where microlensing light curves fall.
+Figure~\ref{fig:var_indices} shows projections of the distribution for various combinations of the variability indices.
 These simulations do define regions to search for microlensing events, but there is still a large amount of overlap between the distributions with and without microlensing events.
@@
 % ^ discuss how the distributions change with injecting microlensing events
-%  Here discuss a figure that shows variance of the distribution of the indices for simulated light curves w/ varied noise sigma
+
+\begin{figure*}[t]
+\centering
+        \subfigure{
+	   \includegraphics[height=0.25\textheight] {figures/eta_vs_delta_chi_squared.pdf}
+	 }
+
+	 \subfigure{
+	   \includegraphics[height=0.25\textheight] {figures/eta_vs_j.pdf}
+	 }
+
+	 \subfigure{
+	   \includegraphics[height=0.25\textheight] {figures/delta_chi_squared_vs_j.pdf}
+	 }
+	\caption{Two-dimensional density histograms for projections of the six-dimensional variability statistic distribution for 10,000 light curves sampled from PTF field 100018 (right) and the equivalent with the same light curves with simulated microlensing events (left).}
+	\label{fig:var_indices}
+\end{figure*}
 
 \subsection{Detection Efficiency}
-The probability of detecting any given microlensing event depends on the parameters of the event, the survey properties (i.e., observing strategy, limiting magnitude, weather), and the nature of the stellar field (i.e., amount of crowding).
-In order to compute the detection efficiency, $\varepsilon$, for a given selection method, we must first define the selection criteria for each variability index.
-For a given PTF field, we randomly sample 1000 light curves with good observations \citep[these have no photometric flags; see description of processing pipeline in][]{nick2009}.
+The probability of detecting a microlensing event depends on the parameters of the event, the survey properties (i.e., observing strategy, limiting magnitude, weather), and the nature of the stellar field (i.e., amount of crowding).
+To compute the detection efficiency, $\varepsilon$, for a given selection method, we must first define the selection criteria for each variability index.
+For a given PTF field, we randomly sample 1000 light curves with good observations \citep[these have no ``bad'' photometric flags; see description of processing pipeline in][]{nick2009}.
 For each of these light curves, we then use the time and magnitude error information to simulate 100 light curves with purely Gaussian scatter.
-We find the value of each variability index such that the selection using that index to identify interesting light curves returns 1\% of these simulated data --- i.e., we set the limiting value of each variability index such that the false positive rate (FPR) is 1\%.
+We find the value of each variability index such that the selection using that index to identify interesting light curves returns 1\% of these simulated data --- i.e., we set the limiting value of each variability index such that the false positive recovery rate (FPR) is 1\%.
+We again randomly sample 1000 light curves from each CCD with $>$10 good observations, compute the set of variability statistics for these light curves, then add 100 different simulated microlensing events to each light curve and evaluate how often these events are recovered given the cuts defined above.
 
-We then run additional Monte Carlo simulations in which we artificially add microlensing events to real data from the same field and evaluate how often these events are recovered given the cuts defined above.
-For a given field, we randomly sample 1000 light curves from each CCD with $>$25 good observations.
-We compute the set of variability statistics for these light curves, then iterate to add 100 different simulated microlensing events to each light curve.
-
-The parameters for each simulated event are chosen as follows: $t_0$ is drawn from a uniform distribution between the minimum and maximum observation times, $u_0$ is drawn from a uniform distribution between 0 and 1.34 ([TODO:] cite?), and $t_E$ is drawn from a log-uniform distribution between 1 and 1000 days.
+The parameters for each simulated event are chosen as follows: $t_0$ is drawn from a uniform distribution between the minimum and maximum observation times, $u_0$ is drawn from a uniform distribution between 0 and 1 (the maximum impact parameter that causes a maximum deviation larger than the median photometric scatter for an $R\sim16$ source), and $t_E$ is drawn from a log-uniform distribution between 1 and 1000 days.
 For each iteration we recompute the variability statistics and store the event parameters.
-Figures~\ref{fig:detection_efficiency_4327}-\ref{fig:detection_efficiency_100152} show detection efficiency curves (bottom panels) computed using this simulation for three representative fields chosen for their different sampling patterns (top panels).
+Figures~\ref{fig:detection_efficiency_4327}-\ref{fig:detection_efficiency_100152} show the results for three representative fields chosen for their different sampling patterns (top panels) and detection efficiency curves computed using this simulation (bottom panels).
 We ignore the indices \textit{K} and $\sigma/\mu$ because their integrated detection efficiencies are each below 1\%.
@@
 \section{Searching for events in the full PTF data set}
-% FLUSH OUT AND UPDATE THIS SECTION
-As an initial test, we applied the 1\% FPR selection on the 325 PTF fields (\apwsim2400 deg$^2$) with $>$100 $R$-band observations using the variability index $\eta$. Our procedure is as follows:
+
+We applied the 1\% FPR selection on the 2097 PTF fields (\apwsim15,224 deg$^2$) with $>$10 $R$-band observations using the variability index $\eta$. Our procedure is as follows:
 \begin{enumerate}
-	\item select a field from the sample of fields with $>100$ good $R$-band observations, and run Monte Carlo simulations to compute the 1\% FPR limiting values for the variability indices;
+	\item select a field from the sample of fields with $>10$ good $R$-band observations, and run Monte Carlo simulations to compute the 1\% FPR limiting values for the variability indices;
 	\item compute $\eta$ for each light curve, and select all light curves that pass the cut determined from the previous simulations;
@@
 	\item we reject the light curve if the new value of $\eta$ still passes the cut (it is probably a periodic variable);
-	\item any remaining light curves are loaded into a local database to be vetted by hand.
+	\item any remaining light curves are loaded into a local database where we then evaluate the relative significance of a microlensing model compared to a variety of others;
+	\item the final sample are those for which the microlensing model best describes the data.
 \end{enumerate}
-% [THE ABOVE WOULD MAKE A VERY NICE FLOWCHART] --> for the actual paper..
+% Turn this into a table, with number of objects left over after each cut?
 
-From the \apwsim$10^7$ light curves in the 325 field footprint, this procedure selected 6306 candidates.
+From the \apwsim$10^9$ light curves in the 2097 field footprint, this procedure selected XXX candidates.
 For each candidate, we look at the SDSS colors (if the PTF field is in the SDSS footprint), search SIMBAD for any nearby sources within 10 arcseconds, and visually inspect the PTF images to vet the data.
 We require that the sources are not associated with SDSS galaxies so we may eliminate supernovae from our sample, but acknowledge the possibility of contamination from supernovae in faint galaxies.
-Of the 6306 candidates, \apwsim1/3 of the light curves had bad or poorly calibrated data that mimicked a transient increase in flux.
+[APW: update when pipeline finishes] Before the final model selection cut, \apwsim1/3 of the light curves had bad or poorly calibrated data that mimicked a transient increase in flux.
 Approximately 2/3 of the light curves were quasars, AGN, or objects with unknown long-term variability.
@@
 
-% In paper, explain MCMC fit in more detail...cit Emcee
+% In paper, explain MCMC fit in more detail...cite Emcee
 From this entire sample, we identified four microlensing event candidates.
@@
 Microlensing event rate predictions away from the Galactic Plane are not well constrained and only a single microlensing event has been observed in a sparse stellar field, despite the great scientific interest in observing such events.
-Using nearly all-sky, time-domain data from the Palomar Transient Factory, we (1) develop new methods for identifying interesting transients in massive light curve datasets, (2) search for sparse-field microlensing events in the data, and (3) will estimate the microlensing event rate over a large area of sky away from the Galactic plane.
+Using nearly all-sky, time-domain data from the Palomar Transient Factory, we develop new methods for identifying interesting transients in massive light curve datasets and search for sparse-field microlensing events.
 
@@
 Price-Whelan is supported by a National Science Foundation Graduate Research Fellowship under grant No.
-$<$Grant Number$>$.
+11-44155.
 
-\clearpage
 \setlength{\baselineskip}{0.6\baselineskip}
@@
 \centering
-	\caption{The PTF survey footprint shown in equatorial coordinates. Each field's color corresponds to some minimum number of observations of that field. }
-    \includegraphics[angle=270, scale=0.55]{figures/R_coverage.pdf}
-    \label{fig:survey_footprint}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\caption{A visualization of the different sampling regimes of six different PTF fields over the same one year baseline.
-Darker points mean more data (more exposures).
-Some fields have roughly uniform coverage, where others are very dense during some weeks and not others.
-}
-    \includegraphics[width=1.0\textwidth]{figures/sampling_figure.pdf}
-    \label{fig:sampling}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\caption{(a) the light curve plotted as is, (b) light curve with a simulated microlensing event with $u_0=1.0$, $t_E=20~\mathrm{days}$, (c) same as (b) with $u_0=0.5$, (d) same as (b) with $u_0=0.01$.
-The red dashed line shows the approximate saturation limit of the PTF camera in R-band.}
-    \includegraphics[width=1.0\textwidth]{figures/simulated_events.pdf}
-    \label{fig:microlensing_sim}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\caption{The maximum outlier light curves for each variability index on PTF field 100049.}
-    \includegraphics[width=1.0\textwidth]{figures/max_outlier_light_curves.pdf}
-    \label{fig:indices_examples}
-\end{figure}
-
-\begin{figure}
-        \centering
-        \caption{Two-dimensional density histograms for projections of the six-dimensional variability statistic distribution for 10,000 light curves sampled from PTF field 100018 (right) and the equivalent with the same light curves with simulated microlensing events (left).}
-        \subfigure{
-	   \includegraphics[height=0.3\textheight] {figures/eta_vs_delta_chi_squared.pdf}
-	 }
-
-	 \subfigure{
-	   \includegraphics[height=0.3\textheight] {figures/eta_vs_j.pdf}
-	 }
-
-	 \subfigure{
-	   \includegraphics[height=0.3\textheight] {figures/delta_chi_squared_vs_j.pdf}
-	 }
-          \label{fig:var_indices}
-\end{figure}
-
-\begin{figure}
-        \centering
-        \caption{Sample light curve (top) and detection efficiency curves (bottom) for PTF field 4327.}
 	\subfigure[A randomly selected light curve from this field. Note the sampling pattern.]{
@@
 	 }
-         \label{fig:detection_efficiency_4327}
+\caption{Sample light curve (top) and detection efficiency curves (bottom) for PTF field 4327.}  \label{fig:detection_efficiency_4327}
 \end{figure}
@@
 \centering
-        \caption{Same as \figurename~\ref{fig:detection_efficiency_4327} for PTF field 4588.}
         \subfigure{
@@
 	 }
-         \label{fig:detection_efficiency_4588}
+\caption{Same as \figurename~\ref{fig:detection_efficiency_4327} for PTF field 4588.}\label{fig:detection_efficiency_4588}
 \end{figure}
@@
 \centering
-        \caption{Same as \figurename~\ref{fig:detection_efficiency_4327} for PTF field 100152.}
         \subfigure{
@@
 	 }
-         \label{fig:detection_efficiency_100152}
+\caption{Same as \figurename~\ref{fig:detection_efficiency_4327} for PTF field 100152.}\label{fig:detection_efficiency_100152}
 \end{figure}
@@
 \centering
-        \caption{Zoomed-in (top) and full (bottom) light curves for four microlensing event candidates found in the PTF light curve database.
-Lines (red) show models with parameters sampled from the posterior probability distribution for the model fits.}
         \subfigure{
@@
 	 }
-         \label{fig:candidates}
+\caption{Zoomed-in (top) and full (bottom) light curves for four microlensing event candidates found in the PTF light curve database.
+Lines (red) show models with parameters sampled from the posterior probability distribution for the model fits.}\label{fig:candidates}
 \end{figure}
@@
 \centering
-        \caption{Posterior probability distributions for each microlensing event model parameter for each event candidate shown in \figurename~\ref{fig:candidates} }
         \subfigure{
@@
 	 }
-         \label{fig:candidate_probs}
+\caption{Posterior probability distributions for each microlensing event model parameter for each event candidate shown in \figurename~\ref{fig:candidates} }\label{fig:candidate_probs}
 \end{figure}

### 2013-01-24 text/paper/ptf_microlensing.tex

@@
-%\documentclass{emulateapj}
-\documentclass[12pt,preprint]{aastex}
+\documentclass{emulateapj}
+%\documentclass[12pt,preprint]{aastex}
 \newcounter{address}
@@
 Under median seeing conditions (1.1$\arcsec$), observations in Mould $R$ or Sloan Digital Sky Survey \citep[SDSS;][]{york00} $g$ achieve 2.0$\arcsec$ full-width half-maximum images and reach 5$\sigma$ magnitudes of $R \approx 21.0$ and $g \approx 21.3$ mag in a standard 60~s exposure \citep{nick2010}.
-As of September 2012, the PTF footprint included $\apwsim$10,100 (2300) deg$^2$ imaged $>$25 ($>$100) times in $R$ and $\apwsim$2200 (170) deg$^2$ imaged that often in $g$ (\figurename~\ref{fig:survey_footprint})\footnote{For an interactive visualization of the PTF survey coverage, see: \url{http://bit.ly/PTFSky}}.
+As of November 2012, the PTF footprint included $\apwsim$15,224 (2766) deg$^2$ imaged $>$10 ($>$100) times in $R$ and $\apwsim$5430 (290) deg$^2$ imaged that often in $g$ (\figurename~\ref{fig:survey_footprint})\footnote{For an interactive visualization of the PTF survey coverage, see: \url{http://bit.ly/PTFSky}}.
 PTF is also a real-time data reduction pipeline that identifies transients of interest, passes these to a dedicated photometric follow-up telescope, and generates an archive of all detected sources \citep{nick2009,rau2009}.
-%[CHECK NUMBERS]
 
@@
 \figurename~\ref{fig:sampling} shows six randomly-selected light curves and demonstrates the varying cadences and coverage that different fields may have over the same one-year baseline.
-We restrict our sample to the \apwsim$6.4\times10^7$ light curves with $>$25 \textit{R}-band observations.
+[APW:] Figure ??
+shows distributions of a) total number of $R$-band observations and b) field baseline for all fields in PTF.
+We restrict our sample to the \apwsim$1.1\times10^9$ light curves with $>$10 \textit{R}-band observations because there is a drop off in the number of fields with fewer than 10 observations and we don't expect such light curves to be useful for our search.
+%APW!
 
@@
 For each iteration we recompute the variability statistics and store the event parameters.
-Figures~\ref{fig:detection_efficiency_4327}-\ref{fig:detection_efficiency_100152} show detection efficiency curves (bottom panels) computed using this simulation for three representative fields chosen for their different sampling patterns (top panels).
+Figures~\ref{fig:detection_efficiency_4327}-\ref{fig:detection_efficiency_100152} show the results for three representative fields chosen for their different sampling patterns (top panels) and detection efficiency curves computed using this simulation (bottom panels).
 We ignore the indices \textit{K} and $\sigma/\mu$ because their integrated detection efficiencies are each below 1\%.
@@
 \section{Searching for events in the full PTF data set}
-% FLUSH OUT AND UPDATE THIS SECTION
+[APW:  FLUSH OUT AND UPDATE THIS SECTION when pipeline tests are done]
+
 As an initial test, we applied the 1\% FPR selection on the 325 PTF fields (\apwsim2400 deg$^2$) with $>$100 $R$-band observations using the variability index $\eta$. Our procedure is as follows:

### 2013-06-07 text/paper/ms.tex
derp

\documentclass{emulateapj}
%\documentclass[12pt,preprint]{aastex}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{natbib}
\usepackage{longtable, hyperref, graphicx, subfigure, epstopdf, multirow, setspace, verbatim, lmodern} 
\usepackage{aas_macros}

\newcommand{\Msun}{\ifmmode {M_{\odot}}\else${M_{\odot}}$\fi}
\newcommand{\Rsun}{\ifmmode {R_{\odot}}\else${R_{\odot}}$\fi}
\newcommand{\lapprox }{{\lower0.8ex\hbox{$\buildrel <\over\sim$}}}
\newcommand{\gapprox }{{\lower0.8ex\hbox{$\buildrel >\over\sim$}}}
\def\amin{\ifmmode^{\prime}\else$^{\prime}$\fi}
\def\asec{\ifmmode^{\prime\prime}\else$^{\prime\prime}$\fi}

\newcommand{\apwsim}{\raisebox{0.2ex}{\scriptsize$\sim$\normalsize}} 
\newcommand{\inlinecode}{\texttt}

\slugcomment{DRAFT \today}
\shorttitle{Microlensing \& PTF}
\shortauthors{Price-Whelan et al.}

\bibliographystyle{apj}

\begin{document}

\title{Identifying Microlensing Events in Large, Non-Uniformly Sampled Surveys: The Case
 of the Palomar Transient Factory}
\author{Adrian~M.~Price-Whelan\altaffilmark{1,10} Marcel~A.~Ag\"ueros\altaffilmark{1}, Amanda Fournier\altaffilmark{2}, Rachel Street\altaffilmark{3}, Eran O. Ofek\altaffilmark{4}, Kevin~R.~Covey\altaffilmark{5}, David Levitan\altaffilmark{6}, Russ R.~Laher\altaffilmark{7}, Branimir Sesar\altaffilmark{8}, Jason Surace\altaffilmark{9}}
%, Joshua S.\ Bloom\altaffilmark{\ref{cal}}, S.\ Bradley Cenko\altaffilmark{\ref{cal}}, Mansi M.\ Kasliwal\altaffilmark{\ref{calt}}, Shrinivas R.\ Kulkarni\altaffilmark{\ref{calt}}, Nicholas~M.~Law\altaffilmark{\ref{to},\ref{dun}}, Peter Nugent\altaffilmark{\ref{lbnl}}, Dovi Poznanski\altaffilmark{\ref{cal},\ref{calt},\ref{eins}}, Robert M.\ Quimby\altaffilmark{\ref{calt}}}

\altaffiltext{1}{Department of Astronomy, Columbia University, 550 W 120th St., New York, NY 10027, USA}
\altaffiltext{2}{Department of Physics, Broida Hall, University of California, Santa Barbara, CA 93106, USA}
\altaffiltext{3}{Las Cumbres Observatory Global Telescope Network, Inc., 6740 Cortona Dr.\ Suite 102, Santa Barbara, CA 93117, USA}
\altaffiltext{4}{Benoziyo Center for Astrophysics, Weizmann Institute of Science, 76100 Rehovot, Israel}
\altaffiltext{5}{Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
\altaffiltext{6}{Lowell Observatory, 1400 West Mars Hill Rd, Flagstaff, AZ 86001, USA}
\altaffiltext{7}{Spitzer Science Center, California Institute of Technology, Mail Stop 314-6, Pasadena, CA 91125, USA}
\altaffiltext{8}{Division of Physics, Mathematics, and Astronomy, California Institute of Technology, Pasadena, CA 91125, USA}
\altaffiltext{9}{Infrared Processing and Analysis Center, California Institute of Technology, Mail Stop 100-22, Pasadena, CA 91125, USA}
%\altaffiltext{\theaddress}{\stepcounter{address}\label{cal} Department of Astronomy, University of California, Berkeley, CA  94720, USA}
%\altaffiltext{\theaddress}{\stepcounter{address}\label{to} Dunlap Institute for Astronomy and Astrophysics, University of Toronto, 50 St.\ George Street, Toronto M5S 3H4, Ontario, Canada}
%\altaffiltext{\theaddress}{\stepcounter{address}\label{lbnl} Computational Cosmology Center, Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA}
%\altaffiltext{\theaddress}{\stepcounter{address}\label{dun} Dunlap Fellow}
\altaffiltext{10}{NSF Graduate Fellow}

\begin{abstract}
Many current photometric, time-domain surveys are driven by specific goals such as searches for supernovae or transiting exoplanets, which set the cadence with which individual fields are re-imaged. In the case of the Palomar Transient Factory (PTF), several such sub-surveys are being conducted in parallel, leading to extremely non-uniform sampling over the survey's nearly $20,000~\mathrm{deg}^2$ footprint. While the median $7.26~\mathrm{deg}^2$ PTF field has been imaged \apwsim40 times in \textit{R}-band, \apwsim$2300~\mathrm{deg}^2$ have been observed more than 100 times. We use the existing PTF data to study the trade-off between searching for microlensing events in a survey whose footprint is much larger than that of typical microlensing searches, but that has far-from-optimal sampling. To examine the probability that microlensing events can be recovered in these data, we test previous statistics used on uniformly sampled data to identify variables and transients. We find that one such statistic, the von Neumann ratio, performs best for identifying simulated microlensing events. We develop a selection method using this statistic and apply it to data from all PTF fields with $>$10 $R$-band observations, a total of $1.1\times10^9$ light curves, to uncover three candidate sparse-field microlensing events. This work can help constrain all-sky event rate predictions and tests microlensing signal recovery in large datasets, both of which will be useful to future wide-field, time-domain surveys such as that planned with the Large Synoptic Survey Telescope.

% Second to last sentence will change when I run on all fields with >25 observations
\end{abstract}

\keywords{
  survey science
  ---
  gravitational microlensing
  ---
  time-domain astrophysics
}

\section{Introduction}
Over the past 20 years, microlensing, in which gravitational lensing causes a transient increase in the flux from a background point source, has been used to search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, to study Galactic structure and kinematics \citep{binney2000}, to determine the shape of stars \citep{rattenbury2005} and, most recently, to identify extrasolar planets \citep[][and references therein]{gaudi2011}. These studies were once limited by the small number of detected events, but thanks to advances in CCD technology and the development of dedicated microlensing surveys, a few thousand events are now observed each year. In the coming decade, this number is expected to increase as the next generation of photometric, time-domain surveys comes online, providing an opportunity for the precise study of otherwise hard-to-characterize objects such as low-mass stars, sub-stellar objects, and isolated neutron stars.

%The microlensing event rate for a given region of the sky depends on the space and velocity distributions of the sources and on the space, velocity, and mass distribution of the lenses.
To maximize the event rate, microlensing surveys typically focus on high-density stellar regions such as the Galactic Bulge, M31, or the Magellanic Clouds \citep[e.g.][]{original_ogle, original_macho, eros_original, crotts1996}. Of course, microlensing events are not limited to dense stellar fields. Indeed, lensing events away from these fields are potentially very interesting. While the event rate in low-density stellar regions will certainly be lower than that near e.g., the Bulge \citep[][]{wood_optical_depth, ogle_optical_depth, macho_optical_depth, eros_optical_depth}, these events are expected to involve closer sources and lenses and less crowded backgrounds compared to those found by modern microlensing surveys \citep{mesolensing}. Thus, such lenses offer more opportunities to constrain the properties of the event.

\begin{figure*}[!t]
\centering\includegraphics[scale=0.4,trim=0 0 0 27, clip]{figures/R_coverage_crop.pdf}
\caption{PTF $R$-band survey footprint, in equatorial coordinates. The fields are color-coded by the total number of observations. The field size corresponds to the actual area covered by a single PTF exposure.}\label{fig:survey_footprint}
\end{figure*}

To date, however, only one event outside of a dense stellar field has been recorded: the Tago event \citep{fukui2007, gaudi2008}, a serendipitous detection of a microlensing event of a bright source. \cite{gaudi2008} set probabilistic limits on the mass, distance, proper motion, and magnitude of this lens using existing observations near the event's location, and further observations may turn these limits into precise measurements. Usually, such an analysis is impossible for events in high-density fields where the lenses are, on average, much farther away and blending makes photometry --- and therefore mass measurements --- challenging. By contrast, high Galactic latitude events probe lenses with distances $\lesssim 1$~kpc, thus the Einstein radii tend to be larger (a few milliarcseconds) and may cause detectable astrometric signatures. A measurement of the astrometric shift of a source due to microlensing will break the distance-mass degeneracy in the lensing parameters and enable a direct lens mass measurement. Finding even a handful of nearby lenses that can be studied in the same kind of detail as the Tago event would therefore be very valuable.

In this paper, we use the Palomar Transient Factory (PTF) survey, described in Section 2, as a test case for studying the frequency and detectability of microlensing events in all-sky, irregularly sampled, time-domain surveys. We examine the selection methods used by microlensing surveys to identify event signatures in light curves and an event identification procedure based on a set of variability statistics. We then present the detection efficiency for these methods as computed for a representative set of PTF fields (Section~\ref{sec:event_recovery}). In Section~4 we apply a set of criteria that use the results of these tests to all PTF light curves with $>$10 high-quality observations to identify candidate microlensing events. We discuss eight interesting transients and three candidate events identified in this manner before concluding in Section~5.

\section{PTF: Survey Design and Consequences for Microlensing Searches} 
PTF is an automated transient detection system \citep{nick2009,rau2009}. Data are collected using the former Canada-France-Hawaii Telescope 12K$\times$8K mosaic camera, which has 11 working chips, $0.92\times10^8$ pixels, and a 7.26 deg$^2$ field-of-view \citep{rahmer2008}, mounted on the 48 inch Oschin Schmidt telescope (the P48) at Palomar Observatory, CA. Under median seeing conditions (1.1$\arcsec$), observations in Mould $R$ or Sloan Digital Sky Survey $g$ \citep[SDSS;][]{york00} achieve 2.0$\arcsec$ full-width half-maximum images and reach 5$\sigma$ magnitudes of $R \approx 21.0$ and $g \approx 21.3$ mag in each 60~s exposure \citep{nick2010}. The data are processed in real time by a pipeline that identifies transients of interest and passes these to a dedicated photometric follow-up telescope; this has proved an excellent tool for e.g., early detection of supernovae \citep{nugent2011}. In addition, all detected sources are archived\footnote{The first photometric catalog of non-variable sources drawn from this archive is described in \citet[][]{ofek2012}.}; this enables the study of periodic forms of variability \citep[e.g., stellar or asteroid rotation;][]{agueros11,polishook2012}.

Images are processed using standard reduction procedures including de-biasing, flat-fielding, and astrometric calibration (Laher et al. in prep). Catalogs are generated using SExtractor for source identification \citep{bertin96}. The absolute photometric calibration of such images is good to a systematic limit of $\sim2\%$ for photometric nights, as described in \cite{ofek2012}. The final light curves are produced using relative photometric calibration which refines the calibration to $<1\%$ for photometric nights and improves the calibration for bad nights \citep[Levitan et al., in prep; for algorithm details see][]{ofek2012, levitan2011}.

As of 2012 Dec, the PTF footprint includes $\apwsim$15,224 (2766) deg$^2$ imaged $>$10 ($>$100) times in $R$-band and $\apwsim$5430 (290) deg$^2$ imaged that often in $g$-band (see Figure~\ref{fig:survey_footprint})\footnote{For an interactive visualization of the PTF survey footprint, see \url{http://bit.ly/PTFSky}.}. The PTF survey footprint is not uniformly sampled either spatially or temporally. Each field has a unique sampling pattern determined by which of the PTF sub-surveys it belongs to, which time of the year it is visible from Palomar, and how high a priority it is given by the scheduler. The survey cadences mostly range from 1-5 days but include a high-cadence imaging campaign of a field in Orion to find transiting planets around young stars \citep{nick2009}. The result is that PTF light curves often contain gaps and regions of high-cadence observations and/or of low-cadence observations, so that the archive is a massive dataset of irregularly-sampled, time-domain photometry. Figure~\ref{fig:sampling} shows six randomly selected light curves and illustrates the varying cadences and coverage that different fields may have over the same one-year period.

In Figure~\ref{fig:num_obs}, we plot the number of $R$-band exposures for each field against the observational baseline (the number of days between the first and last exposure). Given the exposure distribution, which peaks at 30-40 but includes a long tail to larger numbers, we restrict our sample to the light curves with $>$10 \textit{R}-band observations. % because there is a drop off in the number of fields with fewer than 10 observations and we don't expect such light curves to be useful for our search. %APW!

What would a microlensing event look like in a typical PTF light curve? If the source is an unblended point source and the lens is a foreground, dim object, a microlensing event is fully described by three parameters: the angular impact parameter $u_0$, the peak time of the event $t_0$, and the timescale of the event (Einstein crossing time) $t_E$. The amplification factor $A$ and flux $F$ can then be defined as a function of time: 
\begin{align}
	u(t) &= \sqrt{u_0^2 + 2\Big(\frac{t-t_0}{t_E}\Big)},\\
	A(t) &= \frac{u^2 + 2}{u\sqrt{u^2 + 4}},\\
	F(t) &= A(t)\times F_{source}.
\end{align}
The microlensing perturbation can also be expressed as:
\begin{align}\label{eq:ml_model}
	m(t) &= m_0(t) - 2.5\log A(t)
\end{align}
where $m_0(t)$ is the unperturbed, but possibly time-variable, magnitude of the source.

\begin{figure}[!t]
\begin{center}
\includegraphics[width=0.45\textwidth, trim=2 2 5 5, clip]{figures/sampling_figure.pdf}
\caption{Sampling of six different PTF fields over the course of a year. Darker points indicate that more exposures were taken that night; the total number of observations for each field is indicated.}\label{fig:sampling}
\end{center}
\end{figure}

\begin{figure}[h]
\centering\includegraphics[width=0.5\textwidth, trim=21 20 25 70, clip]{figures/num_observations_baseline_90deg.pdf}
\caption{Total number of $R$-band exposures and baselines (number of days between first and last exposure) for all PTF fields. The histogram to the right shows that the exposure distribution is peaked at 30-40; the histogram at the top shows that the baseline distribution has three peaks, corresponding to fields observed over $\lapprox$weeks, $\sim$1 year, and $\sim$3 years (the full length of the survey).}
\label{fig:num_obs}
\end{figure}

%However, if the event is \textit{too} bright, the pixels will saturate and there could be photometric errors flagging the most interesting data as bad.
Even in cases of high amplification ($u_0<<1$), a survey may miss or poorly sample an event if $t_E$ is short ($t_E \lesssim\mathrm{days}$), while if $t_E$ is long ($t_E \gtrsim 1~\mathrm{year}$) the event may be confused with other forms of long-duration variability. Figure~\ref{fig:microlensing_sim} illustrates the effect of simulated microlensing events with a fixed $t_E=20~\mathrm{days}$ but different $u_0$ on a random well-sampled PTF light curve.

\begin{figure}[!h]
\centering\includegraphics[width=0.48\textwidth, trim=5 5 5 7, clip]{figures/simulated_events.pdf}
\caption{Simulating the effect of a $20~\mathrm{day}$ microlensing event with differing impact parameters $u_0$ on a PTF light curve: a) the original light curve, b) the light curve with a $u_0=1.0$ event, c) with a $u_0=0.5$ event, d) with a $u_0=0.01$ event. The dashed line in the bottom panel shows the approximate saturation limit of the PTF camera in the $R$-band; such an event with a brighter source might therefore be missed by the survey.}\label{fig:microlensing_sim}
\end{figure}

\section{Microlensing event recovery} \label{sec:event_recovery}
Microlensing surveys typically use difference image analysis \citep{alard1998} to identify transient events in raw imaging data. Light curves of any transient sources are then analyzed and vetted using a variety of selection methods to search for microlensing event candidates and distinguish them from e.g., variable stars, outbursting systems, and novae. Surveys have approached this process differently (as described in \citealt{ogle_optical_depth, con_idx, alcock2000, macho_detection_efficiency, udalski03,  hamadache2009, wyrzykowski2009, sumi2011}), but the general idea is to require that: 
\begin{enumerate}
	\item any selected light curve has some number of consecutive datapoints brighter than some threshold,
	\item a microlensing model best describe the data (compared to flat or linear light-curve models), as determined by using a $\Delta\chi^2$ test, and
	\item the model event parameters have physically reasonable values.
\end{enumerate}
%Difference imaging is ideal for studying crowded fields where photometric measurements are difficult. It also limits the number of light curves that must be fit with a model, thus saving on computation time and allowing for real-time detection of events, so that high-cadence follow-up observations of interesting sources can be triggered.

Applying this prescription to the PTF data presents obvious challenges. For example, microlensing surveys have relatively uniform time sampling of their survey footprint over an observing season, justifying the first requirement above. But it is harder to motivate such a cut on data with significant and irregular gaps. In order to develop the most successful procedure for PTF data, we therefore examine the relative performance of a set of variability indices and of the traditional prescription in selecting simulated microlensing events.

\subsection{The Variability Indices} %\label{sec:event_selection}
We choose five statistical measures of variability compiled by \cite{shin2009} that have been previously applied to the classification and discovery of periodic variables. These variability indices are $\sigma/\mu$, $Con$, $\eta$, $J$, and $K$\footnote{We do not to implement the sixth index described by \cite{shin2009}, $AoVM$, because it mainly helps identify periodic sources.}. $\sigma/\mu$ is the ratio of the root-variance to the sample mean, 
\begin{align}
	\frac{\sigma}{\mu} = \frac{\sqrt{\sum^N_i (x_i - \mu)^2 / (N-1)}}{\sum^N_i x_i/N},
\end{align}
where N is the total number of observations.

We modify the definition of $Con$ to be the number of clusters of three or more consecutive observations that are more than $3\sigma$ brighter than the reference magnitude of the source (e.g., for a single microlensing event in an otherwise flat light curve, $Con=1$). This change allows us to use the performance of $Con$ as a proxy for the consecutive-point requirement described above.

$\eta$, the von Neumann ratio \citep{von_neumann1941}, is the mean square successive difference divided by the sample variance:
\begin{align}
	\eta = \frac{\delta^2}{\sigma^2} = \frac{\sum^{N-1}_i(x_{i+1} - x_i)^2/(N-1)}{\sigma^2}.
\end{align}
$\eta$ is small when there is strong positive serial correlation between successive data points.

$J$ and $K$ were suggested by \cite{stetson1996}:
\begin{align}
	\delta_i &= \sqrt{\frac{N}{N-1}}\frac{x_i-\mu}{e_i},\\
	J &= \sum^{N-1}_i sign(\delta_i \delta_{i+1})\sqrt{|\delta_i \delta_{i+1}|},\\
	K &= \frac{1/N\sum^N_i |\delta_i|}{\sqrt{1/N\sum^N_i\delta_i^2}},
\end{align}
where $e_i$ is the photometric error of each data point, and the $sign$ function returns $\pm$1, depending on the sign of the argument. $J$ tends to 0 for non-variable stars, but is large when there are significant differences between successive data points in a light curve. $K$ is a measure of the kurtosis of the distribution of data points.

We add one more index, $\Delta \chi^2$, the difference in $\chi^2$ between fitting a Gaussian model and fitting a linear model to a light curve. This is the standard statistical test used by microlensing surveys, and allows us to compare the relative performance of the (slightly modified) \cite{shin2009} indices and of this approach.

Figure~\ref{fig:indices_examples} shows maximal outlier light curves for each variability statistic selected from \apwsim20,000 light curves for objects on a single CCD in PTF field 100101. %, centered at (J2000) 17:21:31.44 $+$61:07:46.6. 
Clearly, the indices are sensitive to different aspects of variability in the data. We expect $\sigma/\mu$ to be most useful for discerning periodic or semi-periodic variability where the variance is large; the expectations are not as clear for other indices.

\begin{figure}
\centering\includegraphics[width=.48\textwidth, trim=130 85 90 65, clip]{figures/max_outlier_light_curves.pdf}
\caption{The maximum outlier light curves for each variability index applied to data from PTF field 100101. For $\eta$ and $K$, this corresponds to the light curve with the \emph{minimum} value of the index over the entire field. For the other indices -- $\Delta\chi^2$, $J$, $\sigma/\mu$, $Con$ -- these are the light curves with the \emph{maximum} value of the relevant index over the field. } \label{fig:indices_examples}
\end{figure}

We conduct Monte Carlo simulations and inject artificial microlensing events into real PTF light curves. We then compare the distributions of variability indices for a set of light curves with and without these simulated events to determine the regions of parameter space where microlensing light curves fall. Figure~\ref{fig:var_indices} shows projections of the distribution for various combinations of the variability indices for another PTF field. These simulations do define regions to search for microlensing events, but there is still a large amount of overlap between the distributions with and without microlensing events. Below we describe our method for defining selection boundaries for each variability statistic to determine which is most efficient for detecting microlensing events.

% ^ discuss how the distributions change with injecting microlensing events

\begin{figure}[t]
\centering
        \subfigure{
	   \includegraphics[width=.467\textwidth, trim=5 9 5 3, clip] {figures/eta_vs_delta_chi_squared.pdf}
	 }

	 \subfigure{
	   \includegraphics[width=.47\textwidth, trim=10 9 0 7, clip] {figures/eta_vs_j.pdf}
	 }

	 \subfigure{
	   \includegraphics[width=.47\textwidth, trim=10 6 0 7, clip] {figures/delta_chi_squared_vs_j.pdf}
	 }
	\caption{Two-dimensional density histograms for projections of the six-dimensional variability statistic distribution for 10,000 light curves from PTF field 100018 (left) and with the addition of simulated microlensing events (right). Red (dashed) lines are the 1\% false positive recovery selection boundaries for each index.} 
	\label{fig:var_indices}
\end{figure}

\subsection{Event Selection and Detection Efficiency} \label{sec:detection_eff}
%The probability of detecting a microlensing event depends on the parameters of the event, the survey properties (i.e., observing strategy, limiting magnitude, weather), and the nature of the stellar field (i.e., amount of crowding). Before computing the detection efficiency, $\varepsilon$, for a given selection method,
We begin by randomly sampling 1000 light curves from each chip in a given PTF field. At least half of the individual observations in these light curves have to be defined as ``good'' \citep[see description of processing pipeline in][]{nick2009}. The light curves we reject are for objects that are either faint or near bright stars whose scattered light and diffraction spikes cause large photometric errors.

For each of these light curves, we use the date and magnitude error information to simulate 100 light curves with purely Gaussian scatter. We find the value of each variability index such that the selection using that index to identify interesting light curves returns 1\% of these scrambled light curves --- i.e., we set the limiting value of each index such that the false positive recovery (FPR) rate is 1\% per trial. We again randomly sample 1000 light curves from each CCD with $>$10 good observations and compute the set of indices for these light curves. Finally, we add 100 different simulated microlensing events to each light curve and evaluate how often these events are recovered given the 1\% FPR cut defined above.

The parameters for each simulated event are chosen as follows: $t_0$ is drawn from a uniform distribution between the minimum and maximum observation times, $u_0$ is drawn from a uniform distribution between 0 and \apwsim1.5 \citep[the impact parameter that causes a maximum deviation larger than \apwsim$5\sigma$ for an $R\sim17.5$ mag source; e.g.,][]{macho_detection_efficiency}, and $t_E$ is drawn from a log-uniform distribution between 1 and 1000 days. For each iteration we recompute the variability indices and store the event parameters.

Figures~\ref{fig:detection_efficiency_4327}-\ref{fig:detection_efficiency_100152} show the results for three representative fields with different sampling patterns (top panels) and detection efficiency curves computed using this simulation (bottom panels). We ignore the indices \textit{K}, $\sigma/\mu$ and $Con$ because their integrated detection efficiencies are below 1\%. At a fixed FPR of 1\%, $Con$ performs poorly, but it may be useful as an initial cut by allowing a higher FPR. In this work, we only consider a single-index selection method and thus reject $Con$. For each of these fields we find that $\eta$ consistently performs better than the other indices at recovering microlensing events. This is especially interesting because computing $\eta$ is $\apwsim100\times$ faster per light curve than computing $\Delta\chi^2$, and that identifying candidate microlensing events in the full PTF dataset using this statistic is therefore computationally plausible.

%\figurename~\ref{fig:detection_efficiency} shows the detection efficiency for each variability statistic as functions of microlensing impact parameter and source limiting magnitude [APW: TODO, create this figure].

%The parameters for the random microlensing events are chosen as follows: 
%\begin{align}
%	t_0&\sim\mathcal{U}[t_{min}, t_{max}] \\
%	u_0&\sim\mathcal{U}[0, 1] \\
%	\log_{10}t_E&\sim\mathcal{U}[0, 3] 
%\end{align}
%where $\mathcal{U}$ is a uniform distribution.

\begin{figure}
\centering
	\subfigure{
	   \includegraphics[width=.45\textwidth, trim=0 8 6 3, clip] {figures/example_light_curve_f4327_ccd2.pdf}
	   \label{fig:subfig1}
	 }

	 \subfigure{
	   \includegraphics[width=.46\textwidth, trim=8 8 5 5, clip] {figures/field004327_Nperccd1000_Nevents100.pdf}
	   \label{fig:subfig2}
	 }
\caption{{\it Top:} A randomly selected light curve from PTF field 4327. Note the sampling pattern. {\it Bottom:} Detection efficiency $\varepsilon$ for $\eta$, $\Delta\chi^2$, and $J$ as functions of the simulated microlensing event parameters $t_E$, $u_0$, and $m_0$, the microlensing event timescale, angular impact parameter, and `quiescent' source magnitude, respectively. Black (dashed) lines show the normalized input distributions, red (solid) lines show the recovered distributions.}  \label{fig:detection_efficiency_4327}
\end{figure}

\begin{figure}
\centering
        \subfigure{
	   \includegraphics[width=.45\textwidth, trim=0 8 6 3, clip] {figures/example_light_curve_f4588_ccd2.pdf}
	   \label{fig:subfig1}
	 }

	 \subfigure{
	   \includegraphics[width=.46\textwidth, trim=8 8 5 5, clip] {figures/field004588_Nperccd1000_Nevents100.pdf}
	   \label{fig:subfig2}
	 }
\caption{Same as \figurename~\ref{fig:detection_efficiency_4327}, for PTF field 4588.}\label{fig:detection_efficiency_4588}
\end{figure}

\begin{figure}
\centering
        \subfigure{
	   \includegraphics[width=.45\textwidth, trim=0 8 6 5, clip] {figures/example_light_curve_f100152_ccd2.pdf}
	   \label{fig:subfig1}
	 }

	 \subfigure{
	   \includegraphics[width=.46\textwidth, trim=8 8 5 5, clip] {figures/field100152_Nperccd1000_Nevents100.pdf}
	   \label{fig:subfig2}
	 }
\caption{Same as \figurename~\ref{fig:detection_efficiency_4327}, for PTF field 100152.}\label{fig:detection_efficiency_100152}
\end{figure}

\section{Searching for events in the full PTF dataset}
To identify candidate microlensing events in the $1.1\times10^9$ light curves with $>$10 $R$-band observations, we:
%We apply the 1\% FPR cut to the 2097 PTF fields (\apwsim15,224 deg$^2$) with $>$10 $R$-band observations using the variability index $\eta$. Our procedure is as follows:
\begin{enumerate}
	\item select a field from the 2097 fields with $>$10 good $R$-band observations and run Monte Carlo simulations to compute the 1\% FPR limiting values for $\eta$, $\Delta \chi^2$, and $J$ for that field;
	\item compute $\eta$ for all the light curves in the field, and select the light curves that pass the quality cut described in Section~\ref{sec:detection_eff} and the FPR cut determined from the Monte Carlo simulations;
	\item fit a microlensing event model and subtract the model, then recompute $\eta$ for all the selected light curves;
	\item reject the light curves for which the new value of $\eta$ still passes the cut\footnote{These are probably periodic variables.};
	\item retain any remaining light curves for further inspection.
\end{enumerate}
% Turn this into a table, with number of objects left over after each cut?

This procedure identifies 2377 candidates from among the initial sample of $1.1\times10^9$ light curves. We search SIMBAD\footnote{This research has made use of the SIMBAD database, operated at CDS, Strasbourg, France.} and remove any candidate with a known extragalactic counterpart within 10\asec. For those in the SDSS footprint, we use Data Release 8 photometry \citep{DR8paper} to identify candidate quasars \citep{richards02} and examine the SDSS images to remove extended objects \citep{sdss_images}. Removing extragalactic objects from our candidate list reduces contamination by e.g., supernovae.

We visually inspect each of the remaining \apwsim2000 light curves. We identify \apwsim1100 quasars that did not meet the \cite{richards02} color cut, or objects with unknown long-term variability (e.g., Mira-type variables). An additional \apwsim600 light curves have bad or poorly calibrated data that mimics a transient increase in flux: diffraction spikes, ghosts, and scattered light can cause non-Gaussian, seeing-dependent variability. Where appropriate, as an additional test, we examine the PTF images to verify the data quality. A small fraction of the selected light curves are variable stars that made it past our periodic variability cut.

We are left with $\sim$300 unclassified transients. Through visual inspection, we classify the bulk of these as novae, supernovae, flares, or outbursting systems, and identify 11 microlensing event candidates. We then search the literature for additional photometric data for these 11 objects \citep[2MASS,WISE;][]{cutri03,WISE}; see Table~\ref{table:photometry}. We also obtained spectroscopy for several of these candidates. Below we detail what can be learned from these data. Generally, additional deeper optical/infrared imaging and/or optical spectroscopy is required to draw firm conclusions about the nature of the object.

\subsubsection*{PTF1J033545.80$-$041849.1} %w
%PTFS1203am
%This object was flagged as a transient by the PTF marshal but not followed up. 
PTF1J0335 has a counterpart in 2MASS, and its $(J-H)$ and $(H-K)$ colors indicate that it is unlikely to be a star \citep{kev07}. Because it is only detected in the $W1$ and $W2$ bands, the WISE color does not help much: its $(W1-W2)$ color is consistent with that of stars and of compact galaxies \citep[cf.\ Figure 14 in][]{yan2013}. We tentatively conclude that this not a good microlensing candidate.

\subsubsection*{PTF1J061800.25$+$203142.5} % (a)}
%PTFS1206i
% - Star
% - 2MASS photometry believable
% - Detected in WISE -- appears blended with nearby object, photometry claims to have successfully deblended. 
PTF1J0618 is detected in 2MASS and in all four WISE filters. The corresponding colors are difficult to interpret, however: the object's $(J-H)$ color is consistent with that of a mid-M dwarf \citep{kev07}, but its $(H-K)$ is significantly redder than expected for such a star. Meanwhile, PTF1J0618's WISE colors suggest it is an extragalactic object \citep[cf.\ Figure 14 in][]{yan2013}.

We observed PTF1J0618 with the Double Spectrograph \citep[DBSP;][]{oke_gunn} on the Hale 5-m telescope at Palomar Observatory, CA, on 2013 Feb 19; see Figure~\ref{fig:1206i_spectrum}. The spectrum was obtained with the D55 dichroic; from the atmospheric cutoff to 5500 \AA, the grating had 600 line~mm$^{-1}$ and was blazed at 4000 \AA, giving a resolution of 1.1 \AA. From 6300-8800 \AA, the grating was 158 line~mm$^{-1}$, was blazed at 7500\ \AA, and gave a resolution was 2.5 \AA. The data reduction was performed by S.~Tang, and included flat fielding, extraction, and wavelength calibration using standard IRAF routines\footnote{IRAF is distributed by the National Optical Astronomy Observatories, which are operated by the Association of Universities for Research in Astronomy, Inc., under cooperative agreement with the National Science Foundation.}.

We analyzed this spectrum with the HAMMER IDL package \citep{kev07}. The HAMMER measures a suite of spectral features and provides automated spectral types by comparison to templates. The typical uncertainty in HAMMER spectral typing is $\lesssim$1 spectral subclass. PTF1J0618 is clearly an early M star, with a spectral type of M2-M3.

M stars are well-known sources of stellar flares, and in sparse data these could easily be confused for a microlensing event. However, typical flares on M stars last minutes to hours, with the longest being of order eight hours \citep{kowal2010}. By contrast, the candidate event detected by PTF had a duration $\sim$30 days and is therefore highly unlikely to have been a flare.

Could PTF1J0618 be an outbursting symbiotic star? To test this hypothesis, we compare the DBSP spectrum to a set of M1-M2 dwarf and giant star spectra. In particular, we inspect the depth of the Ca II triplet lines; we also calculate several of the spectral indices collected by \citet{kev06} to distinguish between dwarfs and giants (e.g., the fluxes in the Na D 5900 and CN 7900 bands). These all indicate that PTF1J0618 is a dwarf star, rendering it unlikely that the PTF event was a symbiotic-type outburst. We further note that these typically have much longer durations, e.g., the systems discussed in \citet{tomov2013}, for which outbursts last a few years.

%Moderate-resolution optical spectra of PTF1J033545.80$-$041849.1 and PTF1J061800.25$+$203142.5 were obtained on 2013-01-10 and 2013-02-19, respectively, with the Double Spectrograph \citep[DBSP;][]{oke_gunn} on the Hale 5 m telescope at Palomar Observatory (Figures~\ref{fig:1206i_spectrum}\&\ref{fig:1203am_spectrum}). The observation on 2013-01-10 was made by Eric Bellm and the observation on 2013-02-19 was made by B. Sesar. Observations with the DBSP used the D55 dichroic to obtain spectra with a 600 $line~mm^{-1}$ grating blazed at 4000\AA~from the atmospheric cutoff to 5500\AA~at 1.1 A resolution, and with a 158 $line~mm^{-1}$ grating blazed at 7500\AA~over the range 6300-8800\AA~at 2.5\AA~dispersion. DBSP data reduction was performed by B. Sesar and Sumin Tang, respectively, including flat fielding, extraction, and wavelength calibration, using standard routines in the IRAF environment.

\begin{figure}[!t]
	\centering
	\includegraphics[width=.48\textwidth, trim=17 15 15 40, clip] {figures/PTFS1206i_spectrum515.pdf}
\caption{\label{fig:1206i_spectrum} DBSP spectrum of PTF1J0618 smoothed with a Gaussian filter ($\sigma = 1$). This object is most likely a M2 dwarf star. The timescale of the event detected by PTF is inconsistent with those generally observed in flares, which are the mostly likely contaminants in identifying microlensing events involving M stars.}
\end{figure}

\subsubsection*{PTF1J120642.60$-$192016.4} % (b)}
%PTFS1212ak
PTF1J1206 has no counterpart within 10\asec\ in 2MASS. While a counterpart is listed in the WISE data 3\asec\ from the PTF position, a visual comparison of the co-added PTF $R$-band images and of the WISE images indicates that the PTF object may be blended with other objects in the WISE data or not detected at all. However, the WISE photometric flags for this nearby source suggest it is clean and unblended. The available information is insufficient to draw even tentative conclusions about the nature of this object.

\subsubsection*{PTF1J131500.09$+$715032.5} % (x)}
%PTFS1213ap
PTF1J1315 has no counterpart within 10\asec\ in 2MASS, but is detected in all four WISE bands. %The $W1$ and $W2$ images look slightly extended, but the photometry flags do not suggest that this object is blended or an extended source.
The $W3$ image appears to have an offset nucleus compared to other filters; this could be a background source, and we therefore do not use the $W3$ magnitude because of this possible contamination. The $(W1-W2)$ color places the object well away from the stellar locus \citep[cf.\ Figure 14 in][]{yan2013}. We tentatively conclude that this is a likely extragalactic source and therefore not a good microlensing candidate.

\subsubsection*{PTF1J153202.91$+$674825.1} % (y)}
%PTFS1215aj
PTF1J1532 has no counterpart in 2MASS, but is detected in the WISE $W1$ and $W2$ bands. The corresponding $(W1-W2) = -0.43\pm0.2$ color is somewhat unusual, as most objects appear to have $(W1-W2)\ \gapprox\ 0.0$, with compact galaxies being the rare exceptions \citep[cf.\ Figure 14 in][]{yan2013}. Here again, we cannot draw any conclusions about the nature of this object.

\subsubsection*{PTF1J161502.39$+$540053.8} % (c)}
%PTFS1216bt
PTF1J1615 is one of two candidates that fall in the SDSS footprint; its counterpart, SDSS J161502.43$+$540053.6, has a proper motion of 2.66 mas/year, consistent with its automated SDSS classification as a star. The star has $u = 22.9$ mag\footnote{These are SDSS PSF magnitudes. Typically, PSF fitting provides better estimates of isolated star magnitudes; see \citet{stoughton02}.}, beyond the $u = 22.0$ 95\% completeness limit for SDSS \citep{stoughton02}, and we therefore focus on its $griz$ photometry. With $(g-r) = 1.10\pm0.05$, $(r-i) = 1.00\pm0.03$, and $(i-z) = 0.56\pm0.04$, the star's colors are consistent with those of a late K/early M star \citep{kev07}. PTF1J1615 has no counterpart in 2MASS, and is only detected in the WISE $W1$ and $W2$ bands, making it hard to draw any conclusions as to the candidate's nature based on these data. Still, its $(W1-W2)$ is consistent with what is seen for stars \citep[cf.\ Figure 14 in][]{yan2013}.

Here again, the most likely source of contamination is flaring. However, the event detected by PTF had a duration $\sim$20 days, and so is highly unlikely to have been a flare.

\renewcommand{\thefootnote}{\alph{footnote}}
\begin{table*}
\small
\centering
\caption{Photometric Properties of the Microlensing Candidates} 
\begin{tabular}{| c | c | c | c | c | c | c | c | c | c |}
\hline
PTF1 & $R$ & $(g'-R)$ & $J$ & $(J-H)$ & $(H-K)$ & $W1$ & $(W1-W2)$ & $(W2-W3)$ & $(W3-W4)$  \\
\hline
\hline
J0335\tablenotemark{a} & 18.28$\pm$0.12 & \nodata & 16.59$\pm$0.13 & 0.73$\pm$0.18 & 0.79$\pm$0.18 & 15.06$\pm$0.04 & 0.18$\pm$0.09 & \nodata & \nodata \\\hline
{\bf J0618} & 20.13$\pm$0.10 & 1.13$\pm$0.22 & 16.23$\pm$0.09 & 0.77$\pm$0.15 & 0.70$\pm$0.15 & 14.08$\pm$0.03 & 0.49$\pm$0.05 & 3.22$\pm$0.09 & 2.65$\pm$0.19 \\\hline
J1206 & 19.24$\pm$0.27 & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata \\\hline
J1315\tablenotemark{a} & 19.69$\pm$0.28 & \nodata & \nodata & \nodata & \nodata & 15.37$\pm$0.04 & 0.81$\pm$0.06 & 2.46$\pm$0.17 & 2.96$\pm$0.37 \\\hline
J1532 & 19.25$\pm$0.21 & \nodata & \nodata & \nodata & \nodata & 16.20$\pm$0.05 & $-0.43$$\pm$0.21 & \nodata & \nodata \\\hline
{\bf J1615} & 20.01$\pm$0.24 & 1.10$\pm$0.29 & \nodata & \nodata & \nodata & 16.36$\pm$0.04 & 0.09$\pm$0.11 & \nodata & \nodata \\\hline
J1716\tablenotemark{a} & 20.27$\pm$0.20 & 0.17$\pm$0.28 & \nodata & \nodata & \nodata & 15.56$\pm$0.04 & 0.85$\pm$0.07 & 2.93$\pm$0.15 & \nodata \\\hline
{\bf J1728} & 16.20$\pm$0.10 & 0.82$\pm$0.10 & 14.70$\pm$0.04 & 0.58$\pm$0.06 & 0.16$\pm$0.09 & 13.92$\pm$0.02 & 0.01$\pm$0.03 & $-$0.28$\pm$0.50 & \nodata \\\hline
J1733 & 20.15$\pm$0.25 & 0.00$\pm$0.34 & 17.18$\pm$0.21 & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata \\\hline
J1747\tablenotemark{a} & 17.37$\pm$0.10 & 0.20$\pm$0.12 & 16.12$\pm$0.07 & 0.49$\pm$0.13 & 0.92$\pm$0.13 & 13.10$\pm$0.03 & 1.01$\pm$0.03 & 2.96$\pm$0.04 & 2.68$\pm$0.06 \\\hline
J1933\tablenotemark{a} & 17.53$\pm$0.10 & 1.13$\pm$0.27 & 15.77$\pm$0.07 & 0.74$\pm$0.11 & 0.59$\pm$0.13 & 14.16$\pm$0.03 & 0.15$\pm$0.04 & 0.94$\pm$0.44 & \nodata \\
\hline
\end{tabular}
\footnotetext{We tentatively rule out these candidates based on their photometry and/or visual inspection of the PTF/WISE images.}
\tablecomments{The bolded objects are plausible microlensing candidates.}
\label{table:photometry}
\end{table*}
\renewcommand{\thefootnote}{\arabic{footnote}}

\subsubsection*{PTF1J171606.82$+$474423.7} % (d)}
%PTFS1217ce
%PTF1J1716 was flagged as a variable star by the PTF pipeline, but only exhibits one peak in the full baseline of PTF data. The object 
PTF1J1716 does not have a counterpart in 2MASS. Its $(W1-W2)$ and $(W2-W3)$ colors are consistent with this object being a QSO \citep[cf.\ Figure 14 in][]{yan2013}. We tentatively conclude that this not a good microlensing candidate.

\subsubsection*{PTF1J172826.08$+$692501.1} % (e)}
%PTFS1217cv
% http://skyserver.sdss3.org/dr9/en/tools/explore/obj.asp?id=1237655748311122026
PTF1J1728 is the other candidate in the SDSS footprint. Its SDSS match, SDSS J172826.07$+$692501.3, has a proper motion of 1.81 mas/year, consistent with its automated classification as a star. PTF1J1728 also has a counterpart in 2MASS, and its SDSS/2MASS colors --- $(u-g) = 1.81\pm0.04$, $(g-r) = 0.79\pm 0.03$, $(r-i) = 0.35\pm0.03$, $(i-z) = 0.21\pm0.03$, $(z-J) = 0.97 \pm0.05$, $(J-H) = 0.58\pm0.06$, and $(H-K) = 0.16\pm 0.09$ --- are consistent with those of a K star \citep{kev07}. Its $(W1-W2)$ and $(W2-W3)$ colors place PTF1J1728 in the stellar locus in Figure 14 of \citet[][]{yan2013}.

E.~Bowsher and S.~Douglas obtained a spectrum of PTF1J1728 with the MDM Observatory Modular Spectrograph (MODspec) on the Hiltner 2.4-m telescope at MDM Observatory, Kitt Peak, AZ, on 2012 Nov 14\footnote{The MDM Observatory is operated by Dartmouth College, Columbia University, Ohio State University, Ohio University, and the University of Michigan.}. MODspec was configured to provide coverage from 4500 to 7500 \AA~with $\sim$1.8 \AA~sampling and a spectral resolution of $\sim$3300. The spectrum was trimmed, overscan- and bias-corrected, cleaned of cosmic rays, flat-fielded, extracted, dispersion-corrected, and flux-calibrated by S.~Douglas using standard IRAF tasks. All the spectra from this night suffered from significant noise blueward of 5000 \AA~and redward of 7000 \AA; in  Figure~\ref{fig:1217cv_spectrum}, we therefore present the spectrum over that 2000 \AA~range. Although the spectrum is noisy, PTF1J1728 appears to be a mid-K star, with a HAMMER spectral type of K3.
%The spectrum was reduced with a script written in PyRAF, the Python-based command language for IRAF\footnote{PyRAF is a product of the Space Telescope Science Institute, which is operated by AURA for NASA.}, by S.~Douglas.

\begin{figure}[!ht]
	\centering
	\includegraphics[width=.48\textwidth, trim=17 15 15 40, clip] {figures/PTFS1217cv.pdf}
\caption{\label{fig:1217cv_spectrum} MODspec spectrum of PTF1J172826.08$+$692501.1.}
\end{figure}

\subsubsection*{PTF1J173301.07$+$374311.9} % (f)}
%PTFS1217df
%PTF1J1733 was flagged as a transient by the PTF pipeline but not followed up with high-cadence imaging or spectroscopy. 
PTF1J1733 is detected in the 2MASS $J$ and $K$ bands (but not $H$). There is no object in the WISE catalog within 10\asec\ of PTF1J1733. Although the $(J-K$) color appears inconsistent with PTF1J1733 being a star \citep{kev07}, we cannot draw any firm conclusions about the nature of this object.

\subsubsection*{PTF1J174736.51$+$300506.9} % (g)}
%PTFS1217dk
PTF1J1747 has counterparts in both 2MASS and WISE. The 2MASS colors suggest that this is not a star \citep{kev07}, while the WISE colors are consistent with those of a QSO \citep[cf.\ Figure 14 in][]{yan2013}. %Visual comparison of the PTF and WISE images suggests that there may be some blending of objects in the WISE data, however the WISE photometric quality flags are excellent for each band and suggest that the source is not blended. 
We tentatively conclude that this not a good microlensing candidate.

% The WISE colors are consistent with a Seyfert galaxy, . Comparing WISE images to PTF, there may be some blending in the WISE data. %The 2MASS ...? $J= 16.119\pm0.074$ $H= 15.633\pm0.106$ $K= 14.710\pm0.071$

\subsubsection*{PTF1J193330.11$+$451501.1} % (z)}
%PTFS1219ad
PTF1J1933 is detected in both 2MASS and WISE. The 2MASS colors suggest that this is not a star \citep{kev07}, while the WISE colors are consistent with PTF1J1933 being an elliptical galaxy. This interpretation is strengthened by a visual inspection of the PTF and WISE images, which show that the object is elongated in the Northeast-Southwest direction. We tentatively conclude that this not a good microlensing candidate. \linebreak

To summarize: we rule out five of the 11 candidates as being likely extragalactic transients based on photometry from 2MASS and WISE. Three of the remaining objects are ambiguous because they are not detected in WISE and/or 2MASS; follow-up imaging or spectroscopy is needed to draw any conclusions about these objects. The remaining three objects are likely stars and plausible microlensing event candidates, but lacking simultaneous, multi-color photometry during the events detected by PTF, we cannot claim these as microlensing events with any confidence.

% If they were ... [todo: once I re-run the fitting code on these three, some comment about the derived parameters.]

%\subsection*{}
%Moderate-resolution optical spectra of PTF1J033545.80$-$041849.1 and PTF1J061800.25$+$203142.5 were obtained on 2013-01-10 and 2013-02-19, respectively, with the Double Spectrograph \citep[DBSP;][]{oke_gunn} on the Hale 5 m telescope at Palomar Observatory (Figures~\ref{fig:1206i_spectrum}\&\ref{fig:1203am_spectrum}). The observation on 2013-01-10 was made by Eric Bellm and the observation on 2013-02-19 was made by B. Sesar. Observations with the DBSP used the D55 dichroic to obtain spectra with a 600 $line~mm^{-1}$ grating blazed at 4000\AA~from the atmospheric cutoff to 5500\AA~at 1.1 A resolution, and with a 158 $line~mm^{-1}$ grating blazed at 7500\AA~over the range 6300-8800\AA~at 2.5\AA~dispersion. DBSP data reduction was performed by B. Sesar and Sumin Tang, respectively, including flat fielding, extraction, and wavelength calibration, using standard routines in the IRAF environment.

% ------------------------------------

\section{Conclusions and Future Work}
Microlensing event rate predictions away from the Galactic Plane are not well constrained and only a single microlensing event has been observed in a sparse stellar field, despite the great scientific interest in observing such events. Using nearly all-sky, time-domain data from the Palomar Transient Factory, we have developed new methods for identifying interesting transients in massive light curve datasets and searching for sparse-field microlensing events.

We examine the detection efficiency of recovering simulated microlensing events using a set of variability statistics. We first determine selection criteria for each statistic using Monte Carlo simulations to simulate flat light curves with Gaussian noise and choose the selection boundaries such that our cuts achieve a 1\% false positive rate. We then simulate microlensing events in real PTF data and compute the detection efficiency for each selection method. We find that the von Neumann ratio, $\eta$, performs better than previously used statistics in recovering injected microlensing events in non-uniformly sampled data. We use $\eta$ to develop a selection procedure for extracting microlensing event candidates from the PTF light curve database and identify 11 interesting, plausible microlensing events, along with numerous unclassified variable objects. Of these 11 candidates, eight are tentatively ruled out  by examining 2MASS or WISE colors, or due to lack of data. We tentatively label the three remaining objects as candidate microlensing events, but we cannot now confirm whether these candidates are true microlensing events without having simultaneous multi-color imaging during the event.

Figures~\ref{fig:not_candidates1}-\ref{fig:candidates} show the light curves for each of the 11 candidates. For the three candidate events, we use a Markov-chain Monte-Carlo algorithm \citep{goodman, dfm} to derive posterior probability distributions over each parameter in the point-source, point-lens microlensing model (eq.~\ref{eq:ml_model}). Overplotted on Figure~\ref{fig:candidates} are samples from these posterior distributions along with the maximum a posteriori model.

A large fraction of the contaminants in our selected sample of candidates are AGN or quasars exhibiting long-term, peaked variability. Many more of these objects are filtered out using a quasar color cut where the SDSS footprint overlaps with that of PTF \citep{richards02} or spectroscopy from the SDSS. We also recovered a large number of transient events such as novae, outbursts, and flares. We have cross-referenced this list of objects with the PTF follow-up marshal and found that most of these objects are known sources, but our pipeline discovered at least two supernovae that the PTF pipeline had missed.

Our algorithm identified \apwsim 2000 transient candidates from a database of over one billion light curves, which we then visually inspected to approximately classify this sample. Though it is not necessarily scalable for future surveys to look at all light curves selected by this method, these 2000 classified light curves could serve as a training dataset for a more sophisticated machine learning approach to classifying the remaining data. Many of the light curves we reject through inspection present interesting forms of variability. Rigorously classifying and cataloging such sources is outside of the scope of this article, but this problem is being approached by e.g., \cite{bloomPasp}.

% TODO: Some reference to these needs a home: PTF data web interfaces\footnote{For example: \url{http://bit.ly/PTFSky}, \url{http://bit.ly/PTFDSky}, \url{http://bit.ly/PTFCandidates}}.

\acknowledgments
We thank E.~Bowsher, S.~Douglas, and S.~Tang for their assistance in obtaining spectra for our candidates, and M.~Modjaz for her help in interpreting several of these spectra.

This paper is based upon work supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155 and on observations obtained with the Samuel Oschin Telescope as part of the Palomar Transient Factory project, a scientific collaboration between the California Institute of Technology, Columbia University, Las Cumbres Observatory, the Lawrence Berkeley National Laboratory, the National Energy Research Scientific Computing Center, the University of Oxford and the Weizmann Institute of Science.

Funding for SDSS-III has been provided by the Alfred P. Sloan Foundation, the Participating Institutions, the National Science Foundation, and the U.S. Department of Energy Office of Science. The SDSS-III web site is http://www.sdss3.org/.

SDSS-III is managed by the Astrophysical Research Consortium for the Participating Institutions of the SDSS-III Collaboration including the University of Arizona, the Brazilian Participation Group, Brookhaven National Laboratory, University of Cambridge, Carnegie Mellon University, University of Florida, the French Participation Group, the German Participation Group, Harvard University, the Instituto de Astrofisica de Canarias, the Michigan State/Notre Dame/JINA Participation Group, Johns Hopkins University, Lawrence Berkeley National Laboratory, Max Planck Institute for Astrophysics, Max Planck Institute for Extraterrestrial Physics, New Mexico State University, New York University, Ohio State University, Pennsylvania State University, University of Portsmouth, Princeton University, the Spanish Participation Group, University of Tokyo, University of Utah, Vanderbilt University, University of Virginia, University of Washington, and Yale University.

This publication makes use of data products from the Wide-field Infrared Survey Explorer, which is a joint project of the University of California, Los Angeles, and the Jet Propulsion Laboratory/California Institute of Technology, funded by the National Aeronautics and Space Administration.

This research made use of Astropy (\url{http://www.astropy.org}), a community-developed core Python package for Astronomy.

%\appendix
%\label{appendix}

%\subsubsection*{}
%Follow-up spectra were obtained for two candidates (PTF1J033545.80$-$041849.1 and PTF1J153202.91$+$674825.1) which classified these sources as galaxies, thus excluding them from the candidate sample as probable supernovae. A third object (PTF1J193330.11$+$451501.1) appears elongated in the PTF coadded image and is also rejected.

\bigskip
\setlength{\baselineskip}{0.6\baselineskip}
\bibliography{references_3}
\setlength{\baselineskip}{1.667\baselineskip}

\clearpage

\begin{figure}
\centering
\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1203am_lc.png}}
\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1212ak_lc.png}}
\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1213ap_lc.png}}
\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1215aj_lc.png}}
\caption{Full light curves (bottom) and zooms around the transient maximum (top) for rejected microlensing event candidates found in PTF. Bad (flagged) data points have been removed.}\label{fig:not_candidates1}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1217ce_lc.png}}
\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1217df_lc.png}}
\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1217dk_lc.png}}
\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1219ad_lc.png}}
\caption{Same as Figure~\ref{fig:not_candidates1}. }\label{fig:not_candidates2}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1206i_lc.png}}
\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1216bt_lc.png}}
\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1217cv_lc.png}}
\caption{Same as Figure~\ref{fig:not_candidates1} for three microlensing event candidates. Lines (grey, thin) show models with parameters sampled from the posterior probability distribution over the four parameters in the point-lens, point-source microlensing event model. Red (thick) line shows the maximum a posteriori model. }\label{fig:candidates}
\end{figure}

\renewcommand*\arraystretch{1.8}
\begin{table}
\scriptsize
\centering
\caption{Candidates} 
\begin{tabular}{| c | c | c | c | c | c | c | c | c | c |}
\hline
ID & $u_0$ & $t_E$ [day] & $t_0$ [MJD] & $m_0$ [R] \\
\hline
J0618 & 0.160 $\pm$ 0.015 & 22.9$\pm$1.6 & 55879.975$\pm$0.248 & 20.216 $\pm$ 0.032\\\hline
J1615 & 0.414$\pm$0.025 & 31.8$\pm$2.3  & 55895.797$\pm$1.956 & 20.016 $\pm$ 0.005\\\hline
J1728 & 1.338$\pm$0.001 & 158.3$\pm$2.1 & 55194.313$\pm$0.702 & 16.2594$\pm$0.0006 \\\hline
\end{tabular}
\label{table:candidates1}
\end{table}

\end{document}

### 2013-06-07 text/paper/ms.tex
v6

@@
 \begin{abstract}
-Many current photometric, time-domain surveys are driven by specific goals such as searches for supernovae or transiting exoplanets, which set the cadence with which individual fields are re-imaged.
+Many current photometric, time-domain surveys are driven by specific goals, such as searches for supernovae or transiting exoplanets, which set the cadence with which individual fields are re-imaged.
 In the case of the Palomar Transient Factory (PTF), several such sub-surveys are being conducted in parallel, leading to extremely non-uniform sampling over the survey's nearly $20,000~\mathrm{deg}^2$ footprint.
-While the median $7.26~\mathrm{deg}^2$ PTF field has been imaged \apwsim40 times in \textit{R}-band, \apwsim$2300~\mathrm{deg}^2$ have been observed more than 100 times.
+While the median $7.26~\mathrm{deg}^2$ PTF field has been imaged $\sim$40 times in \textit{R}-band, $\sim$$2300~\mathrm{deg}^2$ have been observed more than 100 times.
 We use the existing PTF data to study the trade-off between searching for microlensing events in a survey whose footprint is much larger than that of typical microlensing searches, but that has far-from-optimal sampling.
@@
 \section{Introduction}
-Over the past 20 years, microlensing, in which gravitational lensing causes a transient increase in the flux from a background point source, has been used to search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, to study Galactic structure and kinematics \citep{binney2000}, to determine the shape of stars \citep{rattenbury2005} and, most recently, to identify extrasolar planets \citep[][and references therein]{gaudi2011}.
-These studies were once limited by the small number of detected events, but thanks to advances in CCD technology and the development of dedicated microlensing surveys, a few thousand events are now observed each year.
+Over the past 20 years, microlensing, in which gravitational lensing causes a transient increase in the flux from a background point source, has been used to search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, to study Galactic structure and kinematics \citep{binney2000}, to determine the shape of stars \citep{rattenbury2005}, and, most recently, to identify extrasolar planets \citep[][and references therein]{gaudi2011}.
+Such studies were once limited by the small number of detected events, but thanks to advances in CCD technology and the development of dedicated microlensing surveys, a few thousand events are now observed each year.
 In the coming decade, this number is expected to increase as the next generation of photometric, time-domain surveys comes online, providing an opportunity for the precise study of otherwise hard-to-characterize objects such as low-mass stars, sub-stellar objects, and isolated neutron stars.
@@
 Indeed, lensing events away from these fields are potentially very interesting.
-While the event rate in low-density stellar regions will certainly be lower than that near e.g., the Bulge \citep[][]{wood_optical_depth, ogle_optical_depth, macho_optical_depth, eros_optical_depth}, these events are expected to involve closer sources and lenses and less crowded backgrounds compared to those found by modern microlensing surveys \citep{mesolensing}.
-Thus, such lenses offer more opportunities to constrain the properties of the event.
+While the event rate in low-density stellar regions will certainly be lower than that near e.g., the Bulge \citep[][]{ogle_optical_depth, macho_optical_depth, wood_optical_depth, eros_optical_depth}, these events are expected to involve closer sources and lenses, and less crowded backgrounds, compared to those found by modern microlensing surveys \citep{mesolensing}.
+Thus, such lenses offer more opportunities to constrain the properties of the events.
 
@@
 
-To date, however, only one event outside of a dense stellar field has been recorded: the Tago event \citep{fukui2007, gaudi2008}, a serendipitous detection of a microlensing event of a bright source.
+Only one event outside of a dense stellar field has been recorded to date: the Tago event \citep{fukui2007, gaudi2008}, a serendipitous detection of the microlensing of a bright source.
 \cite{gaudi2008} set probabilistic limits on the mass, distance, proper motion, and magnitude of this lens using existing observations near the event's location, and further observations may turn these limits into precise measurements.
-Usually, such an analysis is impossible for events in high-density fields where the lenses are, on average, much farther away and blending makes photometry --- and therefore mass measurements --- challenging.
-By contrast, high Galactic latitude events probe lenses with distances $\lesssim 1$~kpc, thus the Einstein radii tend to be larger (a few milliarcseconds) and may cause detectable astrometric signatures.
-A measurement of the astrometric shift of a source due to microlensing will break the distance-mass degeneracy in the lensing parameters and enable a direct lens mass measurement.
+Such an analysis is usually not possible for events in high-density fields where the lenses are, on average, much farther away and blending makes photometry --- and therefore mass measurements --- challenging.
+By contrast, high-Galactic-latitude events probe lenses with distances $\lesssim$1~kpc; the Einstein radii therefore tend to be larger (a few milliarcseconds) and may cause detectable astrometric signatures.
+A measurement of the astrometric shift due to microlensing will break the distance-mass degeneracy in the lensing parameters and enable a direct measurement of the lens mass.
 Finding even a handful of nearby lenses that can be studied in the same kind of detail as the Tago event would therefore be very valuable.
 
-In this paper, we use the Palomar Transient Factory (PTF) survey, described in Section 2, as a test case for studying the frequency and detectability of microlensing events in all-sky, irregularly sampled, time-domain surveys.
+In this paper, we use the Palomar Transient Factory (PTF) survey, described in Section~\ref{sec:ptf}, as a test case for studying the frequency and detectability of microlensing events in all-sky, irregularly sampled, time-domain surveys.
 We examine the selection methods used by microlensing surveys to identify event signatures in light curves and an event identification procedure based on a set of variability statistics.
 We then present the detection efficiency for these methods as computed for a representative set of PTF fields (Section~\ref{sec:event_recovery}).
-In Section~4 we apply a set of criteria that use the results of these tests to all PTF light curves with $>$10 high-quality observations to identify candidate microlensing events.
-We discuss eight interesting transients and three candidate events identified in this manner before concluding in Section~5.
+In Section~\ref{sec:search} we apply a set of criteria that use the results of these tests to all PTF light curves with $>$10 high-quality observations to identify candidate microlensing events.
+We discuss eight interesting transients and three candidate events identified in this manner before concluding in Section~\ref{sec:conc}.
 
-\section{PTF: Survey Design and Consequences for Microlensing Searches} 
+\section{PTF: Survey Design and Consequences for Microlensing Searches}\label{sec:ptf} 
 PTF is an automated transient detection system \citep{nick2009,rau2009}.
 Data are collected using the former Canada-France-Hawaii Telescope 12K$\times$8K mosaic camera, which has 11 working chips, $0.92\times10^8$ pixels, and a 7.26 deg$^2$ field-of-view \citep{rahmer2008}, mounted on the 48 inch Oschin Schmidt telescope (the P48) at Palomar Observatory, CA.
-Under median seeing conditions (1.1$\arcsec$), observations in Mould $R$ or Sloan Digital Sky Survey $g$ \citep[SDSS;][]{york00} achieve 2.0$\arcsec$ full-width half-maximum images and reach 5$\sigma$ magnitudes of $R \approx 21.0$ and $g \approx 21.3$ mag in each 60~s exposure \citep{nick2010}.
+Under median seeing conditions (1.1$\arcsec$), observations in Mould $R$ or Sloan Digital Sky Survey \citep[SDSS;][]{york00} $g$ achieve 2.0$\arcsec$ full-width half-maximum images and reach 5$\sigma$ magnitudes of $R \approx 21.0$ and $g \approx 21.3$ mag in each 60~s exposure \citep{nick2010}.
 The data are processed in real time by a pipeline that identifies transients of interest and passes these to a dedicated photometric follow-up telescope; this has proved an excellent tool for e.g., early detection of supernovae \citep{nugent2011}.
-In addition, all detected sources are archived\footnote{The first photometric catalog of non-variable sources drawn from this archive is described in \citet[][]{ofek2012}.}; this enables the study of periodic forms of variability \citep[e.g., stellar or asteroid rotation;][]{agueros11,polishook2012}.
 
-Images are processed using standard reduction procedures including de-biasing, flat-fielding, and astrometric calibration (Laher et al.
-in prep).
-Catalogs are generated using SExtractor for source identification \citep{bertin96}.
-The absolute photometric calibration of such images is good to a systematic limit of $\sim2\%$ for photometric nights, as described in \cite{ofek2012}.
-The final light curves are produced using relative photometric calibration which refines the calibration to $<1\%$ for photometric nights and improves the calibration for bad nights \citep[Levitan et al., in prep; for algorithm details see][]{ofek2012, levitan2011}.
+Images are processed using standard reduction procedures including de-biasing, flat-fielding, and astrometric calibration (Laher et al., in prep).
+Catalogs are then generated using SExtractor for source identification \citep{bertin96}.
+The absolute photometric calibration of the images is good to a systematic limit of $\sim2\%$ for photometric nights, as described in \cite{ofek2012}.
+The final light curves are produced using relative photometric calibration, which refines the calibration to $<$1\% for photometric nights and improves the calibration for bad nights \citep[Levitan et al., in prep; for algorithm details see][]{levitan2011, ofek2012}.
+The resulting archive is an excellent resource for the study of periodic forms of variability \citep[e.g., stellar or asteroid rotation;][]{agueros11,polishook2012}.
 
-As of 2012 Dec, the PTF footprint includes $\apwsim$15,224 (2766) deg$^2$ imaged $>$10 ($>$100) times in $R$-band and $\apwsim$5430 (290) deg$^2$ imaged that often in $g$-band (see Figure~\ref{fig:survey_footprint})\footnote{For an interactive visualization of the PTF survey footprint, see \url{http://bit.ly/PTFSky}.}.
+%In addition, all detected sources are archived\footnote{The first photometric catalog of non-variable sources drawn from this archive is described in \citet[][]{ofek2012}.};
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.47\textwidth, trim=2 2 5 5, clip]{figures/sampling_figure.pdf}
+\caption{Sampling of six different PTF fields over the course of a year.
+Darker points indicate that more exposures were taken that night; the total number of observations for each field is indicated.}\label{fig:sampling}
+\end{center}
+\end{figure}
+
+\begin{figure}[t]
+\centering\includegraphics[width=0.5\textwidth, trim=21 20 25 70, clip]{figures/num_observations_baseline_90deg.pdf}
+\caption{Total number of $R$-band exposures and baselines (number of days between first and last exposure) for all PTF fields.
+The histogram to the right shows that the exposure distribution is peaked at 30-40; the histogram at the top shows that the baseline distribution has three peaks, corresponding to fields observed over $\lapprox$weeks, $\sim$1 year, and $\sim$3 years (the full length of the survey).}
+\label{fig:num_obs}
+\end{figure}
+
+\begin{figure}[!h]
+\centering\includegraphics[width=0.48\textwidth, trim=5 5 5 7, clip]{figures/simulated_events.pdf}
+\caption{Simulating the effect of $20$-day microlensing events with impact parameters $u_0 = 1.0$, 0.5, and 0.01, on a random PTF $R$-band light curve.
+The original light curve is shown at the top.
+The dashed line in the bottom panel shows the approximate saturation limit of the PTF camera in the $R$-band; such an event with a brighter source might therefore be missed by the survey.}\label{fig:microlensing_sim}
+\end{figure}
+
+As of 2012 Dec, the PTF footprint includes $\sim$15,224 (2766) deg$^2$ imaged $>$10 ($>$100) times in $R$-band and $\sim$5430 (290) deg$^2$ imaged that often in $g$-band (see Figure~\ref{fig:survey_footprint})\footnote{For an interactive visualization of the PTF survey footprint, see \url{http://bit.ly/PTFSky}.}.
 The PTF survey footprint is not uniformly sampled either spatially or temporally.
-Each field has a unique sampling pattern determined by which of the PTF sub-surveys it belongs to, which time of the year it is visible from Palomar, and how high a priority it is given by the scheduler.
-The survey cadences mostly range from 1-5 days but include a high-cadence imaging campaign of a field in Orion to find transiting planets around young stars \citep{nick2009}.
+Each field has a unique sampling pattern determined by which of the PTF sub-surveys it belongs to, what time of the year it is visible, and how high a priority it is given by the scheduler.
+The imaging cadences range from 1-5 days, but there is a higher-cadence campaign of a field in Orion to find transiting planets around young stars \citep{nick2009}.
 The result is that PTF light curves often contain gaps and regions of high-cadence observations and/or of low-cadence observations, so that the archive is a massive dataset of irregularly-sampled, time-domain photometry.
@@
 \begin{align}\label{eq:ml_model}
-	m(t) &= m_0(t) - 2.5\log A(t)
+	m(t) &= m_0(t) - 2.5\log A(t),
 \end{align}
@@
 
-\begin{figure}[!t]
-\begin{center}
-\includegraphics[width=0.45\textwidth, trim=2 2 5 5, clip]{figures/sampling_figure.pdf}
-\caption{Sampling of six different PTF fields over the course of a year.
-Darker points indicate that more exposures were taken that night; the total number of observations for each field is indicated.}\label{fig:sampling}
-\end{center}
-\end{figure}
-
-\begin{figure}[h]
-\centering\includegraphics[width=0.5\textwidth, trim=21 20 25 70, clip]{figures/num_observations_baseline_90deg.pdf}
-\caption{Total number of $R$-band exposures and baselines (number of days between first and last exposure) for all PTF fields.
-The histogram to the right shows that the exposure distribution is peaked at 30-40; the histogram at the top shows that the baseline distribution has three peaks, corresponding to fields observed over $\lapprox$weeks, $\sim$1 year, and $\sim$3 years (the full length of the survey).}
-\label{fig:num_obs}
-\end{figure}
-
 %However, if the event is \textit{too} bright, the pixels will saturate and there could be photometric errors flagging the most interesting data as bad.
 Even in cases of high amplification ($u_0<<1$), a survey may miss or poorly sample an event if $t_E$ is short ($t_E \lesssim\mathrm{days}$), while if $t_E$ is long ($t_E \gtrsim 1~\mathrm{year}$) the event may be confused with other forms of long-duration variability.
-Figure~\ref{fig:microlensing_sim} illustrates the effect of simulated microlensing events with a fixed $t_E=20~\mathrm{days}$ but different $u_0$ on a random well-sampled PTF light curve.
-
-\begin{figure}[!h]
-\centering\includegraphics[width=0.48\textwidth, trim=5 5 5 7, clip]{figures/simulated_events.pdf}
-\caption{Simulating the effect of a $20~\mathrm{day}$ microlensing event with differing impact parameters $u_0$ on a PTF light curve: a) the original light curve, b) the light curve with a $u_0=1.0$ event, c) with a $u_0=0.5$ event, d) with a $u_0=0.01$ event.
-The dashed line in the bottom panel shows the approximate saturation limit of the PTF camera in the $R$-band; such an event with a brighter source might therefore be missed by the survey.}\label{fig:microlensing_sim}
-\end{figure}
+Figure~\ref{fig:microlensing_sim} illustrates the effect of simulated microlensing events with a fixed $t_E=20~\mathrm{days}$, but different $u_0$, on a random well-sampled PTF light curve.
 
@@
 	\item any selected light curve has some number of consecutive datapoints brighter than some threshold,
-	\item a microlensing model best describe the data (compared to flat or linear light-curve models), as determined by using a $\Delta\chi^2$ test, and
-	\item the model event parameters have physically reasonable values.
+	\item compared to flat or linear light-curve models, a microlensing model best describe the data, as determined by using a $\Delta\chi^2$ test, and
+	\item the microlensing model event parameters have physically reasonable values.
 \end{enumerate}
@@
 Applying this prescription to the PTF data presents obvious challenges.
-For example, microlensing surveys have relatively uniform time sampling of their survey footprint over an observing season, justifying the first requirement above.
+For example, microlensing surveys have relatively uniform time sampling of their survey footprint over an observing season, justifying the first requirement.
 But it is harder to motivate such a cut on data with significant and irregular gaps.
@@
 
-Figure~\ref{fig:indices_examples} shows maximal outlier light curves for each variability statistic selected from \apwsim20,000 light curves for objects on a single CCD in PTF field 100101.
+Figure~\ref{fig:indices_examples} shows maximally outlying light curves for each variability statistic selected from $\sim$20,000 light curves for objects on a single CCD in PTF field 100101.
 %, centered at (J2000) 17:21:31.44 $+$61:07:46.6.
@@
 \centering\includegraphics[width=.48\textwidth, trim=130 85 90 65, clip]{figures/max_outlier_light_curves.pdf}
-\caption{The maximum outlier light curves for each variability index applied to data from PTF field 100101.
+\caption{The maximally outlying light curves for each variability index applied to data from PTF field 100101.
 For $\eta$ and $K$, this corresponds to the light curve with the \emph{minimum} value of the index over the entire field.
-For the other indices -- $\Delta\chi^2$, $J$, $\sigma/\mu$, $Con$ -- these are the light curves with the \emph{maximum} value of the relevant index over the field.
+For the other indices, these are the light curves with the \emph{maximum} value of the relevant index over the field.
 } \label{fig:indices_examples}
@@
 We conduct Monte Carlo simulations and inject artificial microlensing events into real PTF light curves.
-We then compare the distributions of variability indices for a set of light curves with and without these simulated events to determine the regions of parameter space where microlensing light curves fall.
+We then compare the distributions of variability indices for a set of light curves with and without these simulated events to determine the regions of parameter space where microlensed light curves fall.
 Figure~\ref{fig:var_indices} shows projections of the distribution for various combinations of the variability indices for another PTF field.
@@
 Below we describe our method for defining selection boundaries for each variability statistic to determine which is most efficient for detecting microlensing events.
-
-% ^ discuss how the distributions change with injecting microlensing events
 
@@
 We begin by randomly sampling 1000 light curves from each chip in a given PTF field.
-At least half of the individual observations in these light curves have to be defined as ``good'' \citep[see description of processing pipeline in][]{nick2009}.
-The light curves we reject are for objects that are either faint or near bright stars whose scattered light and diffraction spikes cause large photometric errors.
+At least half of the individual observations in these light curves must be defined as ``good'' \citep[see description of processing pipeline in][]{nick2009}.
+The rejected light curves are for objects that are either faint or near bright stars whose scattered light and diffraction spikes cause large photometric errors.
 
@@
 
-The parameters for each simulated event are chosen as follows: $t_0$ is drawn from a uniform distribution between the minimum and maximum observation times, $u_0$ is drawn from a uniform distribution between 0 and \apwsim1.5 \citep[the impact parameter that causes a maximum deviation larger than \apwsim$5\sigma$ for an $R\sim17.5$ mag source; e.g.,][]{macho_detection_efficiency}, and $t_E$ is drawn from a log-uniform distribution between 1 and 1000 days.
+The parameters for each simulated event are chosen as follows: $t_0$ is drawn from a uniform distribution between the first and last observation date, $u_0$ is drawn from a uniform distribution between 0 and $\sim$1.5 \citep[the impact parameter that causes a maximum deviation larger than $\sim$$5\sigma$ for a $R\sim17.5$ mag source; e.g.,][]{macho_detection_efficiency}, and $t_E$ is drawn from a log-uniform distribution between 1 and 1000 days.
 For each iteration we recompute the variability indices and store the event parameters.
@@
 Figures~\ref{fig:detection_efficiency_4327}-\ref{fig:detection_efficiency_100152} show the results for three representative fields with different sampling patterns (top panels) and detection efficiency curves computed using this simulation (bottom panels).
-We ignore the indices \textit{K}, $\sigma/\mu$ and $Con$ because their integrated detection efficiencies are below 1\%.
-At a fixed FPR of 1\%, $Con$ performs poorly, but it may be useful as an initial cut by allowing a higher FPR.
+We ignore the indices \textit{K} and $\sigma/\mu$ because their integrated detection efficiencies are below 1\%.
+At a fixed FPR of 1\%, $Con$ performs poorly, but it may be useful as an initial cut if a higher FPR is used.
 In this work, we only consider a single-index selection method and thus reject $Con$.
+
 For each of these fields we find that $\eta$ consistently performs better than the other indices at recovering microlensing events.
-This is especially interesting because computing $\eta$ is $\apwsim100\times$ faster per light curve than computing $\Delta\chi^2$, and that identifying candidate microlensing events in the full PTF dataset using this statistic is therefore computationally plausible.
+This is especially interesting because computing $\eta$ is $\sim$100$\times$ faster per light curve than computing $\Delta\chi^2$.
+Identifying candidate microlensing events in the full PTF dataset using this statistic is therefore computationally plausible.
 
@@
 Note the sampling pattern.
-{\it Bottom:} Detection efficiency $\varepsilon$ for $\eta$, $\Delta\chi^2$, and $J$ as functions of the simulated microlensing event parameters $t_E$, $u_0$, and $m_0$, the microlensing event timescale, angular impact parameter, and `quiescent' source magnitude, respectively.
+{\it Bottom:} Detection efficiency $\varepsilon$ for $\eta$, $\Delta\chi^2$, and $J$, as a function of the simulated microlensing event parameters $t_E$, $u_0$, and $m_0$ (the event timescale, impact parameter, and `quiescent' source magnitude, respectively).
 Black (dashed) lines show the normalized input distributions, red (solid) lines show the recovered distributions.}  \label{fig:detection_efficiency_4327}
@@
 
-\section{Searching for events in the full PTF dataset}
+\section{Searching for events in the full PTF dataset}\label{sec:search}
 To identify candidate microlensing events in the $1.1\times10^9$ light curves with $>$10 $R$-band observations, we:
-%We apply the 1\% FPR cut to the 2097 PTF fields (\apwsim15,224 deg$^2$) with $>$10 $R$-band observations using the variability index $\eta$. Our procedure is as follows:
+%We apply the 1\% FPR cut to the 2097 PTF fields ($\sim$15,224 deg$^2$) with $>$10 $R$-band observations using the variability index $\eta$. Our procedure is as follows:
 \begin{enumerate}
@@
 This procedure identifies 2377 candidates from among the initial sample of $1.1\times10^9$ light curves.
-We search SIMBAD\footnote{This research has made use of the SIMBAD database, operated at CDS, Strasbourg, France.} and remove any candidate with a known extragalactic counterpart within 10\asec.
-For those in the SDSS footprint, we use Data Release 8 photometry \citep{DR8paper} to identify candidate quasars \citep{richards02} and examine the SDSS images to remove extended objects \citep{sdss_images}.
-Removing extragalactic objects from our candidate list reduces contamination by e.g., supernovae.
+We search SIMBAD\footnote{This research has made use of the SIMBAD database, operated at CDS, Strasbourg, France.} and remove any candidate among these with a known extragalactic counterpart within 10\asec.
+For those in the SDSS footprint, we use Data Release 8 photometry \citep{DR8paper} to identify candidate quasars \citep[using the cuts described in][]{richards02} and examine the SDSS images \citep{sdss_images} to remove extended objects.
+Removing extragalactic objects in this manner from our candidate list reduces contamination by e.g., supernovae.
 
-We visually inspect each of the remaining \apwsim2000 light curves.
-We identify \apwsim1100 quasars that did not meet the \cite{richards02} color cut, or objects with unknown long-term variability (e.g., Mira-type variables).
-An additional \apwsim600 light curves have bad or poorly calibrated data that mimics a transient increase in flux: diffraction spikes, ghosts, and scattered light can cause non-Gaussian, seeing-dependent variability.
+We visually inspect each of the remaining $\sim$2000 light curves.
+We identify $\sim$1100 quasars that did not meet the \cite{richards02} color cut, or objects with unknown long-term variability (e.g., Mira-type variables).
+An additional $\sim$600 light curves have bad or poorly calibrated data that mimics a transient increase in flux: diffraction spikes, ghosts, and scattered light can cause non-Gaussian, seeing-dependent variability.
 Where appropriate, as an additional test, we examine the PTF images to verify the data quality.
@@
 Through visual inspection, we classify the bulk of these as novae, supernovae, flares, or outbursting systems, and identify 11 microlensing event candidates.
-We then search the literature for additional photometric data for these 11 objects \citep[2MASS,WISE;][]{cutri03,WISE}; see Table~\ref{table:photometry}.
+We then search the literature for additional photometric data for these 11 objects.
+Most have a counterpart in WISE \citep{WISE}, while a few have a counterpart in 2MASS \citep{cutri03}; see Table~\ref{table:photometry} for a summary.
 We also obtained spectroscopy for several of these candidates.
@@
 Generally, additional deeper optical/infrared imaging and/or optical spectroscopy is required to draw firm conclusions about the nature of the object.
+
+To summarize: we rule out five of the 11 candidates as being likely extragalactic transients based on photometry from WISE and/or 2MASS.
+Three of the remaining objects are ambiguous because they are not detected in WISE and/or 2MASS; follow-up imaging or spectroscopy is needed to draw any conclusions about these candidates.
+The remaining three objects (PTF1J061800.25$+$203142.5,
+J161502.39$+$540053.8, and J172826.08$+$692501.1) are likely stars and the PTF-detected variability plausible microlensing events.
+We emphasize, however, that lacking simultaneous, multi-color photometry during the events, we cannot claim these as microlensing events with any confidence.
+%The reader may wish to focus on the description of the available data for these three candidates before moving on to the next section.
 
@@
 PTF1J0335 has a counterpart in 2MASS, and its $(J-H)$ and $(H-K)$ colors indicate that it is unlikely to be a star \citep{kev07}.
-Because it is only detected in the $W1$ and $W2$ bands, the WISE color does not help much: its $(W1-W2)$ color is consistent with that of stars and of compact galaxies \citep[cf.\ Figure 14 in][]{yan2013}.
+Because it is only detected in the $W1$ and $W2$ bands, the WISE color does not help much: it is consistent with that of stars and of compact galaxies \citep[cf.\ Figure 14 in][]{yan2013}.
 We tentatively conclude that this not a good microlensing candidate.
@@
 From 6300-8800 \AA, the grating was 158 line~mm$^{-1}$, was blazed at 7500\ \AA, and gave a resolution was 2.5 \AA.
-The data reduction was performed by S.~Tang, and included flat fielding, extraction, and wavelength calibration using standard IRAF routines\footnote{IRAF is distributed by the National Optical Astronomy Observatories, which are operated by the Association of Universities for Research in Astronomy, Inc., under cooperative agreement with the National Science Foundation.}.
+The data reduction was performed by S.~Tang using standard IRAF routines\footnote{IRAF is distributed by the National Optical Astronomy Observatories, which are operated by the Association of Universities for Research in Astronomy, Inc., under cooperative agreement with the National Science Foundation.}.
 
@@
 M stars are well-known sources of stellar flares, and in sparse data these could easily be confused for a microlensing event.
-However, typical flares on M stars last minutes to hours, with the longest being of order eight hours \citep{kowal2010}.
+However, typical flares on M stars last minutes to hours, with the longest seen being of order eight hours \citep{kowal2010}.
 By contrast, the candidate event detected by PTF had a duration $\sim$30 days and is therefore highly unlikely to have been a flare.
-
-Could PTF1J0618 be an outbursting symbiotic star?
-To test this hypothesis, we compare the DBSP spectrum to a set of M1-M2 dwarf and giant star spectra.
-In particular, we inspect the depth of the Ca II triplet lines; we also calculate several of the spectral indices collected by \citet{kev06} to distinguish between dwarfs and giants (e.g., the fluxes in the Na D 5900 and CN 7900 bands).
-These all indicate that PTF1J0618 is a dwarf star, rendering it unlikely that the PTF event was a symbiotic-type outburst.
-We further note that these typically have much longer durations, e.g., the systems discussed in \citet{tomov2013}, for which outbursts last a few years.
 
@@
 
-\begin{figure}[!t]
+\begin{figure}[h]
 	\centering
@@
 \end{figure}
+
+Could PTF1J0618 be an outbursting symbiotic star?
+To test this hypothesis, we compare the DBSP spectrum to a set of M1-M2 dwarf and giant star spectra.
+In particular, we inspect the depth of the Ca II triplet lines; we also calculate several of the spectral indices collected by \citet{kev06} to distinguish between dwarfs and giants (e.g., the fluxes in the Na D 5900 and CN 7900 bands).
+These all indicate that PTF1J0618 is a dwarf star, rendering it unlikely that the PTF event was a symbiotic-type outburst.
+We further note that these typically have much longer durations, e.g., the systems discussed in \citet{tomov2013}, for which outbursts last a few years.
 
@@
 Still, its $(W1-W2)$ is consistent with what is seen for stars \citep[cf.\ Figure 14 in][]{yan2013}.
+Here again, the most likely source of contamination is flaring.
+However, the event detected by PTF had a duration $\sim$20 days, and so is highly unlikely to have been a flare.
 
-Here again, the most likely source of contamination is flaring. However, the event detected by PTF had a duration $\sim$20 days, and so is highly unlikely to have been a flare.
-
-\renewcommand{\thefootnote}{\alph{footnote}}
+%\renewcommand{\thefootnote}{\alph{footnote}}
 \begin{table*}
@@
 \end{tabular}
-\footnotetext{We tentatively rule out these candidates based on their photometry and/or visual inspection of the PTF/WISE images.}
-\tablecomments{The bolded objects are plausible microlensing candidates.}
+%\footnotetext{We tentatively rule out these candidates based on their photometry and/or visual inspection of the PTF/WISE images.}
+\tablecomments{The bolded objects are plausible microlensing candidates.
+Candidates flagged with an ``a'' are tentatively ruled out based on their photometry and/or visual inspection of the PTF/WISE images.}
 \label{table:photometry}
 \end{table*}
-\renewcommand{\thefootnote}{\arabic{footnote}}
+%\renewcommand{\thefootnote}{\arabic{footnote}}
 
@@
 MODspec was configured to provide coverage from 4500 to 7500 \AA~with $\sim$1.8 \AA~sampling and a spectral resolution of $\sim$3300.
-The spectrum was trimmed, overscan- and bias-corrected, cleaned of cosmic rays, flat-fielded, extracted, dispersion-corrected, and flux-calibrated by S.~Douglas using standard IRAF tasks.
+The data reduction was performed by S.~Douglas using standard IRAF routines.
+% The spectrum was trimmed, overscan- and bias-corrected, cleaned of cosmic rays, flat-fielded, extracted, dispersion-corrected, and flux-calibrated by S.~Douglas using standard IRAF tasks.
 All the spectra from this night suffered from significant noise blueward of 5000 \AA~and redward of 7000 \AA; in  Figure~\ref{fig:1217cv_spectrum}, we therefore present the spectrum over that 2000 \AA~range.
 Although the spectrum is noisy, PTF1J1728 appears to be a mid-K star, with a HAMMER spectral type of K3.
-%The spectrum was reduced with a script written in PyRAF, the Python-based command language for IRAF\footnote{PyRAF is a product of the Space Telescope Science Institute, which is operated by AURA for NASA.}, by S.~Douglas.
 
-\begin{figure}[!ht]
+\begin{figure}[h]
 	\centering
@@
 PTF1J1933 is detected in both 2MASS and WISE.
-The 2MASS colors suggest that this is not a star \citep{kev07}, while the WISE colors are consistent with PTF1J1933 being an elliptical galaxy.
+The 2MASS colors suggest that this is not a star \citep{kev07}, while the WISE colors are consistent with PTF1J1933 being an elliptical galaxy \citep[cf.\ Figure 14 in][]{yan2013}.
 This interpretation is strengthened by a visual inspection of the PTF and WISE images, which show that the object is elongated in the Northeast-Southwest direction.
 We tentatively conclude that this not a good microlensing candidate.
-\linebreak
 
-To summarize: we rule out five of the 11 candidates as being likely extragalactic transients based on photometry from 2MASS and WISE.
-Three of the remaining objects are ambiguous because they are not detected in WISE and/or 2MASS; follow-up imaging or spectroscopy is needed to draw any conclusions about these objects.
-The remaining three objects are likely stars and plausible microlensing event candidates, but lacking simultaneous, multi-color photometry during the events detected by PTF, we cannot claim these as microlensing events with any confidence.
+\section{Conclusions and Future Work}\label{sec:conc}
+Microlensing event rate predictions away from the Galactic Plane are not well constrained.
+Only a single event has been observed in a sparse stellar field, despite the great scientific interest in observing such events.
+Using time-domain data from the Palomar Transient Factory, we have developed new methods for identifying interesting transients in massive light curve datasets and for searching for sparse-field microlensing events.
 
-% If they were ... [todo: once I re-run the fitting code on these three, some comment about the derived parameters.]
+We have examined the detection efficiency of recovering simulated microlensing events using a set of variability statistics adapted from \citet{shin2009}.
+We determined selection criteria for each statistic using Monte Carlo simulations to simulate flat light curves with Gaussian noise, and by choosing the selection boundaries such that our cuts achieved a 1\% false positive recovery rate.
+We then simulated microlensing events in real PTF data and computed the detection efficiency for each statistic.
 
-%\subsection*{}
-%Moderate-resolution optical spectra of PTF1J033545.80$-$041849.1 and PTF1J061800.25$+$203142.5 were obtained on 2013-01-10 and 2013-02-19, respectively, with the Double Spectrograph \citep[DBSP;][]{oke_gunn} on the Hale 5 m telescope at Palomar Observatory (Figures~\ref{fig:1206i_spectrum}\&\ref{fig:1203am_spectrum}).
-The observation on 2013-01-10 was made by Eric Bellm and the observation on 2013-02-19 was made by B.
-Sesar.
-Observations with the DBSP used the D55 dichroic to obtain spectra with a 600 $line~mm^{-1}$ grating blazed at 4000\AA~from the atmospheric cutoff to 5500\AA~at 1.1 A resolution, and with a 158 $line~mm^{-1}$ grating blazed at 7500\AA~over the range 6300-8800\AA~at 2.5\AA~dispersion.
-DBSP data reduction was performed by B.
-Sesar and Sumin Tang, respectively, including flat fielding, extraction, and wavelength calibration, using standard routines in the IRAF environment.
+We found that the von Neumann ratio, $\eta$ \citep{von_neumann1941}, performs better than previously used statistics in recovering injected microlensing events in non-uniformly sampled data.
+We used $\eta$ to develop a selection procedure for extracting microlensing event candidates from the PTF light-curve archive.
+Among the $1.1 \times 10^9$ light curves with $>$10 $R$-band observations we first identified $\sim$2000 interesting candidate transients, from which we selected 11 candidate microlensing events (see Figures~\ref{fig:not_candidates1}-\ref{fig:candidates}).
 
-% ------------------------------------
-
-\section{Conclusions and Future Work}
-Microlensing event rate predictions away from the Galactic Plane are not well constrained and only a single microlensing event has been observed in a sparse stellar field, despite the great scientific interest in observing such events.
-Using nearly all-sky, time-domain data from the Palomar Transient Factory, we have developed new methods for identifying interesting transients in massive light curve datasets and searching for sparse-field microlensing events.
-
-We examine the detection efficiency of recovering simulated microlensing events using a set of variability statistics.
-We first determine selection criteria for each statistic using Monte Carlo simulations to simulate flat light curves with Gaussian noise and choose the selection boundaries such that our cuts achieve a 1\% false positive rate.
-We then simulate microlensing events in real PTF data and compute the detection efficiency for each selection method.
-We find that the von Neumann ratio, $\eta$, performs better than previously used statistics in recovering injected microlensing events in non-uniformly sampled data.
-We use $\eta$ to develop a selection procedure for extracting microlensing event candidates from the PTF light curve database and identify 11 interesting, plausible microlensing events, along with numerous unclassified variable objects.
-Of these 11 candidates, eight are tentatively ruled out  by examining 2MASS or WISE colors, or due to lack of data.
-We tentatively label the three remaining objects as candidate microlensing events, but we cannot now confirm whether these candidates are true microlensing events without having simultaneous multi-color imaging during the event.
-
-Figures~\ref{fig:not_candidates1}-\ref{fig:candidates} show the light curves for each of the 11 candidates.
-For the three candidate events, we use a Markov-chain Monte-Carlo algorithm \citep{goodman, dfm} to derive posterior probability distributions over each parameter in the point-source, point-lens microlensing model (eq.~\ref{eq:ml_model}).
-Overplotted on Figure~\ref{fig:candidates} are samples from these posterior distributions along with the maximum a posteriori model.
-
-A large fraction of the contaminants in our selected sample of candidates are AGN or quasars exhibiting long-term, peaked variability.
-Many more of these objects are filtered out using a quasar color cut where the SDSS footprint overlaps with that of PTF \citep{richards02} or spectroscopy from the SDSS.
+A large fraction of the contaminants among the $\sim$2000 interesting candidates were AGN or quasars exhibiting long-term, peaked variability.
 We also recovered a large number of transient events such as novae, outbursts, and flares.
@@
 
-Our algorithm identified \apwsim 2000 transient candidates from a database of over one billion light curves, which we then visually inspected to approximately classify this sample.
-Though it is not necessarily scalable for future surveys to look at all light curves selected by this method, these 2000 classified light curves could serve as a training dataset for a more sophisticated machine learning approach to classifying the remaining data.
-Many of the light curves we reject through inspection present interesting forms of variability.
-Rigorously classifying and cataloging such sources is outside of the scope of this article, but this problem is being approached by e.g., \cite{bloomPasp}.
+Of the 11 candidate microlensing events, we tentatively ruled out eight by examining their 2MASS or WISE photometry, or because they lacked such data.
+We labeled the three remaining objects as plausible microlensing events, but lacking simultaneous multi-color imaging during the events, we cannot confirm these candidates as true microlensing events.
+
+For the three plausible events, we use a Markov-chain Monte-Carlo algorithm \citep{goodman, dfm} to derive posterior probability distributions over each parameter in the point-source, point-lens microlensing model (Eq.~\ref{eq:ml_model}).
+Overplotted on Figure~\ref{fig:candidates} are samples from these posterior distributions along with the maximum {\it a posteriori} model; the corresponding event parameters are listed in Table~\ref{table:candidates1}.
+
+%\renewcommand*\arraystretch{1.8}
+\begin{table*}
+\small
+\centering
+\caption{Microlensing Event Parameters Derived for Plausible Candidates} 
+\begin{tabular}{| c | c | c | c | c | c | c | c | c | c |}
+\hline
+ID & $u_0$ & $t_E$ [day] & $t_0$ [MJD] & $m_0$ [R] \\
+\hline
+J0618 & $0.160\pm0.015$ & $22.9\pm1.6$ & $55879.975\pm0.248$ & $20.216\pm0.032$\\\hline
+J1615 & $0.414\pm0.025$ & $31.8\pm2.3$  & $55895.797\pm1.956$ & $20.016\pm0.005$\\\hline
+J1728 & $1.338\pm0.001$ & $158.3\pm2.1$ & $55194.313\pm0.702$ & $16.259\pm0.001$ \\\hline
+\end{tabular}
+\label{table:candidates1}
+\end{table*}
+
+Our algorithm identified $\sim$2000 transient candidates from a database of over $10^9$ light curves, each of which we then visually inspected.
+Rigorously classifying and cataloging these sources was outside of the scope of this paper, but many  presented interesting forms of variability.
+Though it is not necessarily scalable for future surveys to look at all light curves selected by using our algorithm, these $\sim$2000 light curves, once carefully classified, could serve as a training set for a more sophisticated machine-learning approach \cite[as discussed by e.g.,][]{bloomPasp} to classifying the light curves that survive our selection criteria.
+Approaches that combine statistical cuts such as those described in this paper and some type of machine learning will be essential to making the most of the future time-domain data such as that collected by the Large Synoptic Survey Telescope.
 
@@
 \acknowledgments
-We thank E.~Bowsher, S.~Douglas, and S.~Tang for their assistance in obtaining spectra for our candidates, and M.~Modjaz for her help in interpreting several of these spectra.
+We thank E.~Bowsher, S.~Douglas, and S.~Tang for their assistance in obtaining spectra for our candidates, and M.~Modjaz for help in interpreting several of these spectra.
 
@@
 
-This research made use of Astropy (\url{http://www.astropy.org}), a community-developed core Python package for Astronomy.
+This research made use of Astropy (\url{http://www.astropy.org}), a community-developed core Python package for astronomy.
 
@@
 \setlength{\baselineskip}{0.6\baselineskip}
-\bibliography{references_3}
+\bibliography{references_5}
 \setlength{\baselineskip}{1.667\baselineskip}
@@
 \subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1215aj_lc.png}}
-\caption{Full light curves (bottom) and zooms around the transient maximum (top) for rejected microlensing event candidates found in PTF.
+\caption{Full light curves (top) and zooms around the transient maximum (bottom) for rejected microlensing event candidates found in PTF.
 Bad (flagged) data points have been removed.}\label{fig:not_candidates1}
@@
 \caption{Same as Figure~\ref{fig:not_candidates1} for three microlensing event candidates.
-Lines (grey, thin) show models with parameters sampled from the posterior probability distribution over the four parameters in the point-lens, point-source microlensing event model.
-Red (thick) line shows the maximum a posteriori model.
+Lines (gray, thin) show models with parameters sampled from the posterior probability distribution over the four parameters in the point-lens, point-source microlensing event model.
+Red (thick) line shows the maximum {\it a posteriori} model.
 }\label{fig:candidates}
@@
 
-\renewcommand*\arraystretch{1.8}
-\begin{table}
-\scriptsize
-\centering
-\caption{Candidates} 
-\begin{tabular}{| c | c | c | c | c | c | c | c | c | c |}
-\hline
-ID & $u_0$ & $t_E$ [day] & $t_0$ [MJD] & $m_0$ [R] \\
-\hline
-J0618 & 0.160 $\pm$ 0.015 & 22.9$\pm$1.6 & 55879.975$\pm$0.248 & 20.216 $\pm$ 0.032\\\hline
-J1615 & 0.414$\pm$0.025 & 31.8$\pm$2.3  & 55895.797$\pm$1.956 & 20.016 $\pm$ 0.005\\\hline
-J1728 & 1.338$\pm$0.001 & 158.3$\pm$2.1 & 55194.313$\pm$0.702 & 16.2594$\pm$0.0006 \\\hline
-\end{tabular}
-\label{table:candidates1}
-\end{table}
-
 \end{document}

### 2013-06-20 text/paper/ms.tex
paper update

@@
 
-\usepackage{amsmath}
-\usepackage{amssymb}
+\usepackage{amssymb,amsmath}
 \usepackage{amsfonts}
 \usepackage{natbib}
-\usepackage{longtable, hyperref, graphicx, subfigure, epstopdf, multirow, setspace, verbatim, lmodern} 
+\usepackage{longtable, graphicx, subfigure, epstopdf, multirow, setspace, verbatim, lmodern} 
 \usepackage{aas_macros}
+\newcounter{address}
+\setcounter{address}{1}
 
@@
 
-\title{Identifying Microlensing Events in Large, Non-Uniformly Sampled Surveys: The Case
- of the Palomar Transient Factory}
-\author{Adrian~M.~Price-Whelan\altaffilmark{1,10} Marcel~A.~Ag\"ueros\altaffilmark{1}, Amanda Fournier\altaffilmark{2}, Rachel Street\altaffilmark{3}, Eran O.
-Ofek\altaffilmark{4}, Kevin~R.~Covey\altaffilmark{5}, David Levitan\altaffilmark{6}, Russ R.~Laher\altaffilmark{7}, Branimir Sesar\altaffilmark{8}, Jason Surace\altaffilmark{9}}
+\title{Identifying Microlensing Events in Large, Non-Uniformly Sampled Surveys: A Statistical Search Through Palomar Transient Factory Data}
+\author{Adrian~M.~Price-Whelan\altaffilmark{1,8} Marcel~A.~Ag\"ueros\altaffilmark{1}, Amanda P.
+Fournier\altaffilmark{2,3}, Rachel Street\altaffilmark{3}, Eran O.
+Ofek\altaffilmark{4}, Kevin~R.~Covey\altaffilmark{5}, David Levitan\altaffilmark{6}, Russ R.~Laher\altaffilmark{7}, Branimir Sesar\altaffilmark{6}, Jason Surace\altaffilmark{7}}
 %, Joshua S.\ Bloom\altaffilmark{\ref{cal}}, S.\ Bradley Cenko\altaffilmark{\ref{cal}}, Mansi M.\ Kasliwal\altaffilmark{\ref{calt}}, Shrinivas R.\ Kulkarni\altaffilmark{\ref{calt}}, Nicholas~M.~Law\altaffilmark{\ref{to},\ref{dun}}, Peter Nugent\altaffilmark{\ref{lbnl}}, Dovi Poznanski\altaffilmark{\ref{cal},\ref{calt},\ref{eins}}, Robert M.\ Quimby\altaffilmark{\ref{calt}}}
@@
 \altaffiltext{4}{Benoziyo Center for Astrophysics, Weizmann Institute of Science, 76100 Rehovot, Israel}
-\altaffiltext{5}{Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
-\altaffiltext{6}{Lowell Observatory, 1400 West Mars Hill Rd, Flagstaff, AZ 86001, USA}
+\altaffiltext{5}{Lowell Observatory, 1400 West Mars Hill Rd, Flagstaff, AZ 86001, USA}
+\altaffiltext{6}{Division of Physics, Mathematics, and Astronomy, California Institute of Technology, Pasadena, CA 91125, USA}
 \altaffiltext{7}{Spitzer Science Center, California Institute of Technology, Mail Stop 314-6, Pasadena, CA 91125, USA}
-\altaffiltext{8}{Division of Physics, Mathematics, and Astronomy, California Institute of Technology, Pasadena, CA 91125, USA}
-\altaffiltext{9}{Infrared Processing and Analysis Center, California Institute of Technology, Mail Stop 100-22, Pasadena, CA 91125, USA}
+%\altaffiltext{9}{Infrared Processing and Analysis Center, California Institute of Technology, Mail Stop 100-22, Pasadena, CA 91125, USA}
 %\altaffiltext{\theaddress}{\stepcounter{address}\label{cal} Department of Astronomy, University of California, Berkeley, CA  94720, USA}
@@
 %\altaffiltext{\theaddress}{\stepcounter{address}\label{dun} Dunlap Fellow}
-\altaffiltext{10}{NSF Graduate Fellow}
+\altaffiltext{8}{NSF Graduate Fellow}
 
@@
 We develop a selection method using this statistic and apply it to data from all PTF fields with $>$10 $R$-band observations, a total of $1.1\times10^9$ light curves, to uncover three candidate sparse-field microlensing events.
-This work can help constrain all-sky event rate predictions and tests microlensing signal recovery in large datasets, both of which will be useful to future wide-field, time-domain surveys such as that planned with the Large Synoptic Survey Telescope.
+This work can help constrain all-sky event rate predictions and tests microlensing signal recovery in large data sets, both of which will be useful to future wide-field, time-domain surveys, such as that planned with the Large Synoptic Survey Telescope.
 
@@
 Such studies were once limited by the small number of detected events, but thanks to advances in CCD technology and the development of dedicated microlensing surveys, a few thousand events are now observed each year.
-In the coming decade, this number is expected to increase as the next generation of photometric, time-domain surveys comes online, providing an opportunity for the precise study of otherwise hard-to-characterize objects such as low-mass stars, sub-stellar objects, and isolated neutron stars.
+In the coming decade, this number is expected to increase as the next generation of photometric, time-domain surveys comes online, providing an opportunity for the precise study of otherwise hard-to-characterize objects, such as low-mass stars, sub-stellar objects, and isolated neutron stars.
 
 %The microlensing event rate for a given region of the sky depends on the space and velocity distributions of the sources and on the space, velocity, and mass distribution of the lenses.
-To maximize the event rate, microlensing surveys typically focus on high-density stellar regions such as the Galactic Bulge, M31, or the Magellanic Clouds \citep[e.g.][]{original_ogle, original_macho, eros_original, crotts1996}.
+To maximize the event rate, microlensing surveys typically focus on high-density stellar regions, such as the Galactic Bulge, M31, or the Magellanic Clouds \citep[e.g.][]{original_ogle, original_macho, eros_original, crotts1996}.
 Of course, microlensing events are not limited to dense stellar fields.
 Indeed, lensing events away from these fields are potentially very interesting.
-While the event rate in low-density stellar regions will certainly be lower than that near e.g., the Bulge \citep[][]{ogle_optical_depth, macho_optical_depth, wood_optical_depth, eros_optical_depth}, these events are expected to involve closer sources and lenses, and less crowded backgrounds, compared to those found by modern microlensing surveys \citep{mesolensing}.
+While the event rate in low-density stellar regions will certainly be lower than that near, e.g., the Bulge \citep[][]{ogle_optical_depth, macho_optical_depth, wood_optical_depth, eros_optical_depth}, these events are expected to involve closer sources and lenses, and less crowded backgrounds, compared to those found by modern microlensing surveys \citep{mesolensing}.
 Thus, such lenses offer more opportunities to constrain the properties of the events.
@@
 \cite{gaudi2008} set probabilistic limits on the mass, distance, proper motion, and magnitude of this lens using existing observations near the event's location, and further observations may turn these limits into precise measurements.
-Such an analysis is usually not possible for events in high-density fields where the lenses are, on average, much farther away and blending makes photometry --- and therefore mass measurements --- challenging.
-By contrast, high-Galactic-latitude events probe lenses with distances $\lesssim$1~kpc; the Einstein radii therefore tend to be larger (a few milliarcseconds) and may cause detectable astrometric signatures.
+Such an analysis is usually not possible for events in high-density fields where the lenses are, on average, much farther away and blending makes photometry --- and, therefore, mass measurements --- challenging.
+By contrast, high-Galactic-latitude events probe lenses with distances $\lesssim$1~kpc; the Einstein radii, therefore, tend to be larger (a few milliarcseconds) and may cause detectable astrometric signatures.
 A measurement of the astrometric shift due to microlensing will break the distance-mass degeneracy in the lensing parameters and enable a direct measurement of the lens mass.
-Finding even a handful of nearby lenses that can be studied in the same kind of detail as the Tago event would therefore be very valuable.
+Finding even a handful of nearby lenses that can be studied in the same kind of detail as the Tago event would, therefore, be very valuable.
 
-In this paper, we use the Palomar Transient Factory (PTF) survey, described in Section~\ref{sec:ptf}, as a test case for studying the frequency and detectability of microlensing events in all-sky, irregularly sampled, time-domain surveys.
+In this paper, we use the Palomar Transient Factory (PTF) survey, which is described in Section~\ref{sec:ptf}, as a test case for studying the frequency and detectability of microlensing events in all-sky, irregularly sampled, time-domain surveys.
+A simple extrapolation from previous all-sky rate estimates suggests $\sim$1--10 detectable microlensing events within the PTF footprint over the three year survey \citep{han2008, gaudi2008}.
 We examine the selection methods used by microlensing surveys to identify event signatures in light curves and an event identification procedure based on a set of variability statistics.
 We then present the detection efficiency for these methods as computed for a representative set of PTF fields (Section~\ref{sec:event_recovery}).
-In Section~\ref{sec:search} we apply a set of criteria that use the results of these tests to all PTF light curves with $>$10 high-quality observations to identify candidate microlensing events.
+In Section~\ref{sec:search}, we apply a set of criteria that use the results of these tests to all PTF light curves with $>$10 high-quality observations to identify candidate microlensing events.
 We discuss eight interesting transients and three candidate events identified in this manner before concluding in Section~\ref{sec:conc}.
+A detailed investigation of the all-sky microlensing event rate,  generated from a model that includes realistic mass and velocity distributions, will be presented in Fournier et al.
+(in prep).
 
 \section{PTF: Survey Design and Consequences for Microlensing Searches}\label{sec:ptf} 
-PTF is an automated transient detection system \citep{nick2009,rau2009}.
-Data are collected using the former Canada-France-Hawaii Telescope 12K$\times$8K mosaic camera, which has 11 working chips, $0.92\times10^8$ pixels, and a 7.26 deg$^2$ field-of-view \citep{rahmer2008}, mounted on the 48 inch Oschin Schmidt telescope (the P48) at Palomar Observatory, CA.
-Under median seeing conditions (1.1$\arcsec$), observations in Mould $R$ or Sloan Digital Sky Survey \citep[SDSS;][]{york00} $g$ achieve 2.0$\arcsec$ full-width half-maximum images and reach 5$\sigma$ magnitudes of $R \approx 21.0$ and $g \approx 21.3$ mag in each 60~s exposure \citep{nick2010}.
-The data are processed in real time by a pipeline that identifies transients of interest and passes these to a dedicated photometric follow-up telescope; this has proved an excellent tool for e.g., early detection of supernovae \citep{nugent2011}.
+PTF data are collected using the former Canada-France-Hawaii Telescope 12K$\times$8K mosaic camera, which has 11 working chips, $0.92\times10^8$ pixels, and a 7.26 deg$^2$ field-of-view \citep{rahmer2008}, mounted on the 48-nch Oschin Schmidt Telescope (the P48) at Palomar Observatory, CA.
+Under median seeing conditions (2.2$\arcsec$), observations in Mould $R$ or Sloan Digital Sky Survey \citep[SDSS;][]{york00} $g$ achieve 2.0$\arcsec$ full-width half-maximum images and reach 5$\sigma$ magnitudes of $R \approx 21.0$ and $g \approx 21.3$ mag in each 60~s exposure \citep{nick2009,rau2009,nick2010}.
+%The data are processed in real time by a pipeline that i
 
+While a real-time image-differencing pipeline identifies transients of interest and passes these to a dedicated photometric follow-up telescope, a separate pipeline generates PTF light curves.
 Images are processed using standard reduction procedures including de-biasing, flat-fielding, and astrometric calibration (Laher et al., in prep).
-Catalogs are then generated using SExtractor for source identification \citep{bertin96}.
-The absolute photometric calibration of the images is good to a systematic limit of $\sim2\%$ for photometric nights, as described in \cite{ofek2012}.
+Catalogs are produced using SExtractor for source identification \citep{bertin96}.
+The absolute photometric calibration of the images is good to a systematic limit of $\sim$2\% for photometric nights, as described in \cite{ofek2012}.
 The final light curves are produced using relative photometric calibration, which refines the calibration to $<$1\% for photometric nights and improves the calibration for bad nights \citep[Levitan et al., in prep; for algorithm details see][]{levitan2011, ofek2012}.
 The resulting archive is an excellent resource for the study of periodic forms of variability \citep[e.g., stellar or asteroid rotation;][]{agueros11,polishook2012}.
+%A separate pipeline identifies transients of interest and passes these to a dedicated photometric follow-up telescope; this has proved an excellent tool for e.g., early detection of supernovae \citep{nugent2011}.
 
@@
 
-As of 2012 Dec, the PTF footprint includes $\sim$15,224 (2766) deg$^2$ imaged $>$10 ($>$100) times in $R$-band and $\sim$5430 (290) deg$^2$ imaged that often in $g$-band (see Figure~\ref{fig:survey_footprint})\footnote{For an interactive visualization of the PTF survey footprint, see \url{http://bit.ly/PTFSky}.}.
-The PTF survey footprint is not uniformly sampled either spatially or temporally.
-Each field has a unique sampling pattern determined by which of the PTF sub-surveys it belongs to, what time of the year it is visible, and how high a priority it is given by the scheduler.
+As of 2012 Dec, the PTF footprint includes $\sim$15,224 (2766) deg$^2$ imaged $>$10 ($>$100) times in $R$-band and $\sim$5430 (290) deg$^2$ imaged that often in $g$-band (see Figure~\ref{fig:survey_footprint}).
+The PTF survey footprint is not uniformly sampled either spatially or temporally.\footnote{For an interactive visualization of the PTF survey footprint, see \url{http://bit.ly/PTFSky}.} Each field has a unique sampling pattern determined by which of the PTF sub-surveys it belongs to, what time of the year it is visible, and how high a priority it is given by the scheduler.
 The imaging cadences range from 1-5 days, but there is a higher-cadence campaign of a field in Orion to find transiting planets around young stars \citep{nick2009}.
@@
 If the source is an unblended point source and the lens is a foreground, dim object, a microlensing event is fully described by three parameters: the angular impact parameter $u_0$, the peak time of the event $t_0$, and the timescale of the event (Einstein crossing time) $t_E$.
-The amplification factor $A$ and flux $F$ can then be defined as a function of time:
+In terms of a dimensionless projected distance between the source and lens, $u$, the amplification factor $A$ and flux $F$ as a function of time can be defined as:
 \begin{align}
@@
 \end{align}
-The microlensing perturbation can also be expressed as:
+\citep{paczynski1986}. The microlensing perturbation can also be expressed as:
 \begin{align}\label{eq:ml_model}
@@
 Even in cases of high amplification ($u_0<<1$), a survey may miss or poorly sample an event if $t_E$ is short ($t_E \lesssim\mathrm{days}$), while if $t_E$ is long ($t_E \gtrsim 1~\mathrm{year}$) the event may be confused with other forms of long-duration variability.
-Figure~\ref{fig:microlensing_sim} illustrates the effect of simulated microlensing events with a fixed $t_E=20~\mathrm{days}$, but different $u_0$, on a random well-sampled PTF light curve.
+Figure~\ref{fig:microlensing_sim} illustrates the effect of simulated microlensing events with a fixed $t_E=20~\mathrm{days}$, but different angular impact parameter, $u_0$, on a random well-sampled PTF light curve.
 
@@
 Microlensing surveys typically use difference image analysis \citep{alard1998} to identify transient events in raw imaging data.
-Light curves of any transient sources are then analyzed and vetted using a variety of selection methods to search for microlensing event candidates and distinguish them from e.g., variable stars, outbursting systems, and novae.
+The light curves of transient sources are then analyzed and vetted using a variety of selection methods to search for microlensing event candidates and distinguish them from, e.g., variable stars, outbursting systems, and novae.
 Surveys have approached this process differently (as described in \citealt{ogle_optical_depth, con_idx, alcock2000, macho_detection_efficiency, udalski03,  hamadache2009, wyrzykowski2009, sumi2011}), but the general idea is to require that:
 \begin{enumerate}
-	\item any selected light curve has some number of consecutive datapoints brighter than some threshold,
+	\item any selected light curve has some number of consecutive data points brighter than some threshold,
 	\item compared to flat or linear light-curve models, a microlensing model best describe the data, as determined by using a $\Delta\chi^2$ test, and
@@
 We choose five statistical measures of variability compiled by \cite{shin2009} that have been previously applied to the classification and discovery of periodic variables.
-These variability indices are $\sigma/\mu$, $Con$, $\eta$, $J$, and $K$\footnote{We do not to implement the sixth index described by \cite{shin2009}, $AoVM$, because it mainly helps identify periodic sources.}.
-$\sigma/\mu$ is the ratio of the root-variance to the sample mean,
+These variability indices are $\sigma/\mu$, $Con$, $\eta$, $J$, and $K$\footnote{We do not implement the sixth index described by \cite{shin2009}, $AoVM$, because it mainly helps identify periodic sources.}.
+$\sigma/\mu$ is the ratio of the sample standard deviation to the sample mean,
 \begin{align}
@@
 We search SIMBAD\footnote{This research has made use of the SIMBAD database, operated at CDS, Strasbourg, France.} and remove any candidate among these with a known extragalactic counterpart within 10\asec.
-For those in the SDSS footprint, we use Data Release 8 photometry \citep{DR8paper} to identify candidate quasars \citep[using the cuts described in][]{richards02} and examine the SDSS images \citep{sdss_images} to remove extended objects.
+For those in the SDSS footprint, we use Data Release 7 photometry \citep{DR7paper} to identify candidate quasars \citep[using the cuts described in][]{richards02} and examine the SDSS images \citep{sdss_images} to remove extended objects.
 Removing extragalactic objects in this manner from our candidate list reduces contamination by e.g., supernovae.
@@
 We visually inspect each of the remaining $\sim$2000 light curves.
-We identify $\sim$1100 quasars that did not meet the \cite{richards02} color cut, or objects with unknown long-term variability (e.g., Mira-type variables).
+We identify an additional $\sim$1100 objects with unknown long-term variability (e.g, Mira-type variables, quasars not in the SDSS footprint).
 An additional $\sim$600 light curves have bad or poorly calibrated data that mimics a transient increase in flux: diffraction spikes, ghosts, and scattered light can cause non-Gaussian, seeing-dependent variability.
@@
 We are left with $\sim$300 unclassified transients.
-Through visual inspection, we classify the bulk of these as novae, supernovae, flares, or outbursting systems, and identify 11 microlensing event candidates.
+Through visual inspection of these light curves, we classify the bulk of these as novae, supernovae, flares, or outbursting systems, and identify 11 microlensing event candidates.
 We then search the literature for additional photometric data for these 11 objects.
@@
 The remaining three objects (PTF1J061800.25$+$203142.5,
-J161502.39$+$540053.8, and J172826.08$+$692501.1) are likely stars and the PTF-detected variability plausible microlensing events.
+J161502.39$+$540053.8, and J172826.08$+$692501.1) are likely stars and, given the observed PTF variability, plausible microlensing events.
 We emphasize, however, that lacking simultaneous, multi-color photometry during the events, we cannot claim these as microlensing events with any confidence.
@@
 The spectrum was obtained with the D55 dichroic; from the atmospheric cutoff to 5500 \AA, the grating had 600 line~mm$^{-1}$ and was blazed at 4000 \AA, giving a resolution of 1.1 \AA.
-From 6300-8800 \AA, the grating was 158 line~mm$^{-1}$, was blazed at 7500\ \AA, and gave a resolution was 2.5 \AA.
-The data reduction was performed by S.~Tang using standard IRAF routines\footnote{IRAF is distributed by the National Optical Astronomy Observatories, which are operated by the Association of Universities for Research in Astronomy, Inc., under cooperative agreement with the National Science Foundation.}.
+From 6300-8800 \AA, the grating was 158 line~mm$^{-1}$, was blazed at 7500\ \AA, and gave a resolution of 2.5 \AA.
+The data were reduced using standard IRAF routines\footnote{IRAF is distributed by the National Optical Astronomy Observatories, which are operated by the Association of Universities for Research in Astronomy, Inc., under cooperative agreement with the National Science Foundation.}.
 
@@
 	\centering
-	\includegraphics[width=.48\textwidth, trim=17 15 15 40, clip] {figures/PTFS1206i_spectrum515.pdf}
+	\includegraphics[width=.48\textwidth, trim=17 15 15 40, clip] {figures/PTFS1206i_new.pdf}
 \caption{\label{fig:1206i_spectrum} DBSP spectrum of PTF1J0618 smoothed with a Gaussian filter ($\sigma = 1$).
@@
 We further note that these typically have much longer durations, e.g., the systems discussed in \citet{tomov2013}, for which outbursts last a few years.
+We conclude that this is a good candidate microlensing event.
 
@@
 However, the event detected by PTF had a duration $\sim$20 days, and so is highly unlikely to have been a flare.
+We conclude that this is a good candidate microlensing event.
 
@@
 Although the spectrum is noisy, PTF1J1728 appears to be a mid-K star, with a HAMMER spectral type of K3.
+We conclude that this is a good candidate microlensing event.
 
@@
 \section{Conclusions and Future Work}\label{sec:conc}
-Microlensing event rate predictions away from the Galactic Plane are not well constrained.
+Microlensing event rate predictions away from the Galactic Plane are not well constrained but are being modeled in detail by Fournier et al.
+(in prep).
 Only a single event has been observed in a sparse stellar field, despite the great scientific interest in observing such events.
@@
 We also recovered a large number of transient events such as novae, outbursts, and flares.
-We have cross-referenced this list of objects with the PTF follow-up marshal and found that most of these objects are known sources, but our pipeline discovered at least two supernovae that the PTF pipeline had missed.
+We have cross-referenced this list of objects with the PTF transient detection system and found that most of these objects are known sources, but our pipeline discovered at least two supernovae that the PTF pipeline had missed.
 
@@
 Overplotted on Figure~\ref{fig:candidates} are samples from these posterior distributions along with the maximum {\it a posteriori} model; the corresponding event parameters are listed in Table~\ref{table:candidates1}.
+The event durations for J0618 and J1615 are reasonable, given that a ${\sim}0.1M_\odot$ lens at ${\sim}500~\mathrm{pc}$ in the thick disk ($v_{tan}\sim 50~\mathrm{km}~\mathrm{s}^{-1}$) has a typical event duration $t_E \sim 20~\mathrm{days}$, but the significantly longer duration of J1728 ($>150~\mathrm{days}$) could be a sign of a greater mass lens or larger distance.
 
@@
 \hline
-ID & $u_0$ & $t_E$ [day] & $t_0$ [MJD] & $m_0$ [R] \\
+ID & $u_0$ & $t_E$ [days] & $t_0$ [MJD] & $m_0$ [R] \\
 \hline
@@
 Rigorously classifying and cataloging these sources was outside of the scope of this paper, but many  presented interesting forms of variability.
-Though it is not necessarily scalable for future surveys to look at all light curves selected by using our algorithm, these $\sim$2000 light curves, once carefully classified, could serve as a training set for a more sophisticated machine-learning approach \cite[as discussed by e.g.,][]{bloomPasp} to classifying the light curves that survive our selection criteria.
+Though it is not necessarily scalable for future surveys to look at all light curves selected by using our algorithm, these $\sim$2000 light curves, once carefully classified, could serve as a training set for a more sophisticated machine-learning approach \cite[as discussed by e.g.,][]{bloomPasp} to classify the light curves that survive our selection criteria.
 Approaches that combine statistical cuts such as those described in this paper and some type of machine learning will be essential to making the most of the future time-domain data such as that collected by the Large Synoptic Survey Telescope.
@@
 \setlength{\baselineskip}{0.6\baselineskip}
-\bibliography{references_5}
+%\bibliography{references_5}
+\bibliography{references_3}
 \setlength{\baselineskip}{1.667\baselineskip}

### 2013-09-30 text/final_paper/ms.tex
final paper

--- 2013-06-20 text/paper/ms.tex
@@
 
-\title{Identifying Microlensing Events in Large, Non-Uniformly Sampled Surveys: A Statistical Search Through Palomar Transient Factory Data}
+\title{Statistical Searches for Microlensing Events in Large, Non-Uniformly Sampled Time-Domain Surveys: A Test Using Palomar Transient Factory Data}
 \author{Adrian~M.~Price-Whelan\altaffilmark{1,8} Marcel~A.~Ag\"ueros\altaffilmark{1}, Amanda P.
@@
 Ofek\altaffilmark{4}, Kevin~R.~Covey\altaffilmark{5}, David Levitan\altaffilmark{6}, Russ R.~Laher\altaffilmark{7}, Branimir Sesar\altaffilmark{6}, Jason Surace\altaffilmark{7}}
-%, Joshua S.\ Bloom\altaffilmark{\ref{cal}}, S.\ Bradley Cenko\altaffilmark{\ref{cal}}, Mansi M.\ Kasliwal\altaffilmark{\ref{calt}}, Shrinivas R.\ Kulkarni\altaffilmark{\ref{calt}}, Nicholas~M.~Law\altaffilmark{\ref{to},\ref{dun}}, Peter Nugent\altaffilmark{\ref{lbnl}}, Dovi Poznanski\altaffilmark{\ref{cal},\ref{calt},\ref{eins}}, Robert M.\ Quimby\altaffilmark{\ref{calt}}}
 
@@
 \altaffiltext{7}{Spitzer Science Center, California Institute of Technology, Mail Stop 314-6, Pasadena, CA 91125, USA}
-%\altaffiltext{9}{Infrared Processing and Analysis Center, California Institute of Technology, Mail Stop 100-22, Pasadena, CA 91125, USA}
-%\altaffiltext{\theaddress}{\stepcounter{address}\label{cal} Department of Astronomy, University of California, Berkeley, CA  94720, USA}
-%\altaffiltext{\theaddress}{\stepcounter{address}\label{to} Dunlap Institute for Astronomy and Astrophysics, University of Toronto, 50 St.\ George Street, Toronto M5S 3H4, Ontario, Canada}
-%\altaffiltext{\theaddress}{\stepcounter{address}\label{lbnl} Computational Cosmology Center, Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA}
-%\altaffiltext{\theaddress}{\stepcounter{address}\label{dun} Dunlap Fellow}
-\altaffiltext{8}{NSF Graduate Fellow}
 
@@
 While the median $7.26~\mathrm{deg}^2$ PTF field has been imaged $\sim$40 times in \textit{R}-band, $\sim$$2300~\mathrm{deg}^2$ have been observed more than 100 times.
-We use the existing PTF data to study the trade-off between searching for microlensing events in a survey whose footprint is much larger than that of typical microlensing searches, but that has far-from-optimal sampling.
+We use the existing PTF data to study the trade-off between searching for microlensing events in a survey whose footprint is much larger than that of typical microlensing searches, but that has far-from-optimal time sampling.
 To examine the probability that microlensing events can be recovered in these data, we test previous statistics used on uniformly sampled data to identify variables and transients.
-We find that one such statistic, the von Neumann ratio, performs best for identifying simulated microlensing events.
-We develop a selection method using this statistic and apply it to data from all PTF fields with $>$10 $R$-band observations, a total of $1.1\times10^9$ light curves, to uncover three candidate sparse-field microlensing events.
+We find that one such statistic, the von Neumann ratio, performs best for identifying simulated microlensing events in our data.
+We develop a selection method using this statistic and apply it to data from all PTF fields with $>$10 $R$-band observations --- a total of $1.1\times10^9$ light curves --- to uncover three candidate sparse-field microlensing events.
+While these candidates cannot be confirmed, as we lack simultaneous, multi-color photometry of the events, this number is consistent with simple predictions for the microlensing event rate in the PTF footprint over the survey's three years of operations.
 This work can help constrain all-sky event rate predictions and tests microlensing signal recovery in large data sets, both of which will be useful to future wide-field, time-domain surveys, such as that planned with the Large Synoptic Survey Telescope.
-
-% Second to last sentence will change when I run on all fields with >25 observations
 \end{abstract}
@@
 \keywords{
-  survey science
+  gravitational lensing: micro
   ---
-  gravitational microlensing
+  surveys
   ---
-  time-domain astrophysics
+  methods: statistical
 }
@@
 \section{Introduction}
-Over the past 20 years, microlensing, in which gravitational lensing causes a transient increase in the flux from a background point source, has been used to search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, to study Galactic structure and kinematics \citep{binney2000}, to determine the shape of stars \citep{rattenbury2005}, and, most recently, to identify extrasolar planets \citep[][and references therein]{gaudi2011}.
+Over the past 20 years, microlensing, in which gravitational lensing causes a transient increase in the flux from a background point source, has been used to search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, to study Galactic structure and kinematics \citep{binney2000}, to determine the shape of stars \citep{rattenbury2005}, and to identify extrasolar planets \citep[][and references therein]{gaudi2011}.
 Such studies were once limited by the small number of detected events, but thanks to advances in CCD technology and the development of dedicated microlensing surveys, a few thousand events are now observed each year.
@@
 
-%The microlensing event rate for a given region of the sky depends on the space and velocity distributions of the sources and on the space, velocity, and mass distribution of the lenses.
-To maximize the event rate, microlensing surveys typically focus on high-density stellar regions, such as the Galactic Bulge, M31, or the Magellanic Clouds \citep[e.g.][]{original_ogle, original_macho, eros_original, crotts1996}.
-Of course, microlensing events are not limited to dense stellar fields.
+Microlensing surveys typically focus on high-density stellar regions, such as the Galactic Bulge, M31, or the Magellanic Clouds \citep[e.g.,][]{original_ogle, original_macho, eros_original, crotts1996}, but microlensing events are not limited to dense stellar fields.
 Indeed, lensing events away from these fields are potentially very interesting.
-While the event rate in low-density stellar regions will certainly be lower than that near, e.g., the Bulge \citep[][]{ogle_optical_depth, macho_optical_depth, wood_optical_depth, eros_optical_depth}, these events are expected to involve closer sources and lenses, and less crowded backgrounds, compared to those found by modern microlensing surveys \citep{mesolensing}.
+While the microlensing event rate in low-density stellar regions will be smaller, compared to events found by modern microlensing surveys, these events are expected to involve closer sources and lenses, and less crowded backgrounds \citep{mesolensing}.
 Thus, such lenses offer more opportunities to constrain the properties of the events.
@@
 \begin{figure*}[!t]
-\centering\includegraphics[scale=0.4,trim=0 0 0 27, clip]{figures/R_coverage_crop.pdf}
+\centering\includegraphics[scale=0.4,trim=0 0 0 27, clip]{fig1.eps}
 \caption{PTF $R$-band survey footprint, in equatorial coordinates.
@@
 Such an analysis is usually not possible for events in high-density fields where the lenses are, on average, much farther away and blending makes photometry --- and, therefore, mass measurements --- challenging.
-By contrast, high-Galactic-latitude events probe lenses with distances $\lesssim$1~kpc; the Einstein radii, therefore, tend to be larger (a few milliarcseconds) and may cause detectable astrometric signatures.
+By contrast, high-Galactic-latitude events probe lenses with distances $\lesssim$1~kpc; the Einstein radii, therefore, tend to be larger ($\sim$milliarcseconds) and may cause detectable astrometric signatures.
 A measurement of the astrometric shift due to microlensing will break the distance-mass degeneracy in the lensing parameters and enable a direct measurement of the lens mass.
@@
 
-In this paper, we use the Palomar Transient Factory (PTF) survey, which is described in Section~\ref{sec:ptf}, as a test case for studying the frequency and detectability of microlensing events in all-sky, irregularly sampled, time-domain surveys.
-A simple extrapolation from previous all-sky rate estimates suggests $\sim$1--10 detectable microlensing events within the PTF footprint over the three year survey \citep{han2008, gaudi2008}.
-We examine the selection methods used by microlensing surveys to identify event signatures in light curves and an event identification procedure based on a set of variability statistics.
+In this paper, we use the Palomar Transient Factory (PTF) survey, described in Section~\ref{sec:ptf}, as a test case for studying the frequency and detectability of microlensing events in all-sky, irregularly sampled, time-domain surveys.
+Although to date no microlensing event has been reported using PTF data, a simple extrapolation from previous all-sky rate estimates suggests $\sim$1--10 detectable microlensing events within the PTF footprint over the survey's three years of operation \citep{han2008, gaudi2008}.
+We examine the selection methods used by microlensing surveys to identify event signatures in light curves as well as an event identification procedure based on a set of variability statistics.
 We then present the detection efficiency for these methods as computed for a representative set of PTF fields (Section~\ref{sec:event_recovery}).
@@
 Under median seeing conditions (2.2$\arcsec$), observations in Mould $R$ or Sloan Digital Sky Survey \citep[SDSS;][]{york00} $g$ achieve 2.0$\arcsec$ full-width half-maximum images and reach 5$\sigma$ magnitudes of $R \approx 21.0$ and $g \approx 21.3$ mag in each 60~s exposure \citep{nick2009,rau2009,nick2010}.
-%The data are processed in real time by a pipeline that i
 
@@
 Images are processed using standard reduction procedures including de-biasing, flat-fielding, and astrometric calibration (Laher et al., in prep).
-Catalogs are produced using SExtractor for source identification \citep{bertin96}.
+SExtractor is used for source identification \citep{bertin96}.
 The absolute photometric calibration of the images is good to a systematic limit of $\sim$2\% for photometric nights, as described in \cite{ofek2012}.
@@
 The resulting archive is an excellent resource for the study of periodic forms of variability \citep[e.g., stellar or asteroid rotation;][]{agueros11,polishook2012}.
-%A separate pipeline identifies transients of interest and passes these to a dedicated photometric follow-up telescope; this has proved an excellent tool for e.g., early detection of supernovae \citep{nugent2011}.
-
-%In addition, all detected sources are archived\footnote{The first photometric catalog of non-variable sources drawn from this archive is described in \citet[][]{ofek2012}.};
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=0.47\textwidth, trim=2 2 5 5, clip]{figures/sampling_figure.pdf}
-\caption{Sampling of six different PTF fields over the course of a year.
-Darker points indicate that more exposures were taken that night; the total number of observations for each field is indicated.}\label{fig:sampling}
-\end{center}
-\end{figure}
-
-\begin{figure}[t]
-\centering\includegraphics[width=0.5\textwidth, trim=21 20 25 70, clip]{figures/num_observations_baseline_90deg.pdf}
-\caption{Total number of $R$-band exposures and baselines (number of days between first and last exposure) for all PTF fields.
-The histogram to the right shows that the exposure distribution is peaked at 30-40; the histogram at the top shows that the baseline distribution has three peaks, corresponding to fields observed over $\lapprox$weeks, $\sim$1 year, and $\sim$3 years (the full length of the survey).}
-\label{fig:num_obs}
-\end{figure}
-
-\begin{figure}[!h]
-\centering\includegraphics[width=0.48\textwidth, trim=5 5 5 7, clip]{figures/simulated_events.pdf}
-\caption{Simulating the effect of $20$-day microlensing events with impact parameters $u_0 = 1.0$, 0.5, and 0.01, on a random PTF $R$-band light curve.
-The original light curve is shown at the top.
-The dashed line in the bottom panel shows the approximate saturation limit of the PTF camera in the $R$-band; such an event with a brighter source might therefore be missed by the survey.}\label{fig:microlensing_sim}
-\end{figure}
 
@@
 
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.47\textwidth, trim=2 2 5 5, clip]{fig2.eps}
+\caption{Sampling of six different PTF fields over the course of a year.
+Darker points indicate that more exposures were taken that night; the total number of observations for each field is indicated.}\label{fig:sampling}
+\end{center}
+\end{figure}
+
+\begin{figure}[t]
+\centering\includegraphics[width=0.5\textwidth, trim=21 20 25 70, clip]{fig3.eps}
+\caption{Total number of $R$-band exposures and baselines (number of days between first and last exposure) for all PTF fields.
+The histogram to the right shows that the exposure distribution is peaked at 30-40; the histogram at the top shows that the baseline distribution has three peaks, corresponding to fields observed over $\lapprox$weeks, $\sim$1 year, and $\sim$3 years (the full length of the survey).}
+\label{fig:num_obs}
+\end{figure}
+
+\begin{figure}[!h]
+\centering\includegraphics[width=0.48\textwidth, trim=5 0 0 7, clip]{fig4.eps}
+\caption{Simulating the effect of $20$-day microlensing events with impact parameters $u_0 = 1.0$, 0.5, and 0.01, on a random PTF $R$-band light curve.
+The original light curve is shown at the top.
+The dashed line in the bottom panel shows the approximate saturation limit of the PTF camera in the $R$-band; such an event with a brighter source might therefore be missed by the survey.}\label{fig:microlensing_sim}
+\end{figure}
+
 In Figure~\ref{fig:num_obs}, we plot the number of $R$-band exposures for each field against the observational baseline (the number of days between the first and last exposure).
 Given the exposure distribution, which peaks at 30-40 but includes a long tail to larger numbers, we restrict our sample to the light curves with $>$10 \textit{R}-band observations.
-% because there is a drop off in the number of fields with fewer than 10 observations and we don't expect such light curves to be useful for our search.
-%APW!
 
@@
 
-%However, if the event is \textit{too} bright, the pixels will saturate and there could be photometric errors flagging the most interesting data as bad.
 Even in cases of high amplification ($u_0<<1$), a survey may miss or poorly sample an event if $t_E$ is short ($t_E \lesssim\mathrm{days}$), while if $t_E$ is long ($t_E \gtrsim 1~\mathrm{year}$) the event may be confused with other forms of long-duration variability.
@@
 \end{enumerate}
-%Difference imaging is ideal for studying crowded fields where photometric measurements are difficult.
-It also limits the number of light curves that must be fit with a model, thus saving on computation time and allowing for real-time detection of events, so that high-cadence follow-up observations of interesting sources can be triggered.
 
@@
 
-\subsection{The Variability Indices} %\label{sec:event_selection}
+\begin{figure}[h]
+\centering\includegraphics[width=.48\textwidth, trim=130 85 90 65, clip]{fig5.eps}
+\caption{The maximally outlying light curves for each variability index applied to data from PTF field 100101.
+For $\eta$ and $K$, this corresponds to the light curve with the \emph{minimum} value of the index over the entire field.
+For the other indices, these are the light curves with the \emph{maximum} value of the relevant index over the field.
+} \label{fig:indices_examples}
+\end{figure}
+
+\subsection{The Variability Indices} 
 We choose five statistical measures of variability compiled by \cite{shin2009} that have been previously applied to the classification and discovery of periodic variables.
@@
 
-\begin{figure}
-\centering\includegraphics[width=.48\textwidth, trim=130 85 90 65, clip]{figures/max_outlier_light_curves.pdf}
-\caption{The maximally outlying light curves for each variability index applied to data from PTF field 100101.
-For $\eta$ and $K$, this corresponds to the light curve with the \emph{minimum} value of the index over the entire field.
-For the other indices, these are the light curves with the \emph{maximum} value of the relevant index over the field.
-} \label{fig:indices_examples}
+\begin{figure}[h]
+\centering
+        \subfigure{
+	   \includegraphics[width=.467\textwidth, trim=5 9 5 3, clip] {fig6_a.eps}
+	 }
+
+	 \subfigure{
+	   \includegraphics[width=.47\textwidth, trim=10 9 0 7, clip] {fig6_b.eps}
+	 }
+
+	 \subfigure{
+	   \includegraphics[width=.47\textwidth, trim=10 6 0 7, clip] {fig6_c.eps}
+	 }
+	\caption{Two-dimensional density histograms for projections of the six-dimensional variability statistic distribution for 10,000 light curves from PTF field 100018 (left) and with the addition of simulated microlensing events (right).
+Red (dashed) lines are the 1\% false positive recovery selection boundaries for each index.
+Shaded (red) regions show the subspaces where microlensing events are expected to fall, as defined by the two statistics in question.
+}
+	\label{fig:var_indices}
 \end{figure}
@@
 Below we describe our method for defining selection boundaries for each variability statistic to determine which is most efficient for detecting microlensing events.
-
-\begin{figure}[t]
-\centering
-        \subfigure{
-	   \includegraphics[width=.467\textwidth, trim=5 9 5 3, clip] {figures/eta_vs_delta_chi_squared.pdf}
-	 }
-
-	 \subfigure{
-	   \includegraphics[width=.47\textwidth, trim=10 9 0 7, clip] {figures/eta_vs_j.pdf}
-	 }
-
-	 \subfigure{
-	   \includegraphics[width=.47\textwidth, trim=10 6 0 7, clip] {figures/delta_chi_squared_vs_j.pdf}
-	 }
-	\caption{Two-dimensional density histograms for projections of the six-dimensional variability statistic distribution for 10,000 light curves from PTF field 100018 (left) and with the addition of simulated microlensing events (right).
-Red (dashed) lines are the 1\% false positive recovery selection boundaries for each index.}
-	\label{fig:var_indices}
-\end{figure}
 
@@
 Finally, we add 100 different simulated microlensing events to each light curve and evaluate how often these events are recovered given the 1\% FPR cut defined above.
+
+\begin{figure}
+\centering
+	\subfigure{
+	   \includegraphics[width=.45\textwidth, trim=0 8 6 3, clip] {fig7_a.eps}
+	   \label{fig:subfig1}
+	 }
+
+	 \subfigure{
+	   \includegraphics[width=.46\textwidth, trim=8 8 5 5, clip] {fig7_b.eps}
+	   \label{fig:subfig2}
+	 }
+\caption{{\it Top:} A randomly selected light curve from PTF field 4327.
+Note the sampling pattern.
+{\it Bottom:} Detection efficiency $\varepsilon$ for $\eta$, $\Delta\chi^2$, and $J$, as a function of the simulated microlensing event parameters $t_E$, $u_0$, and $m_0$ (the event timescale, impact parameter, and `quiescent' source magnitude, respectively).
+Black (dashed) lines show the normalized input distributions, red (solid) lines show the recovered distributions.}  \label{fig:detection_efficiency_4327}
+\end{figure}
 
@@
         \subfigure{
-	   \includegraphics[width=.45\textwidth, trim=0 8 6 3, clip] {figures/example_light_curve_f4327_ccd2.pdf}
+	   \includegraphics[width=.45\textwidth, trim=0 8 6 3, clip] {fig8_a.eps}
 	   \label{fig:subfig1}
@@
 	 \subfigure{
-	   \includegraphics[width=.46\textwidth, trim=8 8 5 5, clip] {figures/field004327_Nperccd1000_Nevents100.pdf}
-	   \label{fig:subfig2}
-	 }
-\caption{{\it Top:} A randomly selected light curve from PTF field 4327.
-Note the sampling pattern.
-{\it Bottom:} Detection efficiency $\varepsilon$ for $\eta$, $\Delta\chi^2$, and $J$, as a function of the simulated microlensing event parameters $t_E$, $u_0$, and $m_0$ (the event timescale, impact parameter, and `quiescent' source magnitude, respectively).
-Black (dashed) lines show the normalized input distributions, red (solid) lines show the recovered distributions.}  \label{fig:detection_efficiency_4327}
-\end{figure}
-
-\begin{figure}
-\centering
-        \subfigure{
-	   \includegraphics[width=.45\textwidth, trim=0 8 6 3, clip] {figures/example_light_curve_f4588_ccd2.pdf}
-	   \label{fig:subfig1}
-	 }
-
-	 \subfigure{
-	   \includegraphics[width=.46\textwidth, trim=8 8 5 5, clip] {figures/field004588_Nperccd1000_Nevents100.pdf}
+	   \includegraphics[width=.46\textwidth, trim=8 8 5 5, clip] {fig8_b.eps}
 	   \label{fig:subfig2}
@@
         \subfigure{
-	   \includegraphics[width=.45\textwidth, trim=0 8 6 5, clip] {figures/example_light_curve_f100152_ccd2.pdf}
+	   \includegraphics[width=.45\textwidth, trim=0 8 6 5, clip] {fig9_a.eps}
 	   \label{fig:subfig1}
@@
 	 \subfigure{
-	   \includegraphics[width=.46\textwidth, trim=8 8 5 5, clip] {figures/field100152_Nperccd1000_Nevents100.pdf}
+	   \includegraphics[width=.46\textwidth, trim=8 8 5 5, clip] {fig9_b.eps}
 	   \label{fig:subfig2}
@@
 For those in the SDSS footprint, we use Data Release 7 photometry \citep{DR7paper} to identify candidate quasars \citep[using the cuts described in][]{richards02} and examine the SDSS images \citep{sdss_images} to remove extended objects.
-Removing extragalactic objects in this manner from our candidate list reduces contamination by e.g., supernovae.
+Removing extragalactic objects in this manner from our candidate list reduces contamination by, e.g., supernovae.
+
+\begin{figure*}[htb]
+\centering
+%\includegraphics[width=2in] {PTFS1206i_lc.png}%}
+%\hspace{.2cm}
+%\includegraphics[width=2in, trim = 0 0 -30 0, clip] {PTFS1216bt_lc.png}%}
+%\hspace{.2cm}
+%\includegraphics[width=2in] {PTFS1217cv_lc.png}%}
+\includegraphics[width=2.1\columnwidth]{fig10.eps}
+\caption{Full light curves (top) and zooms around the transient maximum (bottom) for the three plausible microlensing event candidates.
+Bad (flagged) data points have been removed.
+Lines (gray, thin) show models with parameters sampled from the posterior probability distribution over the four parameters in the point-lens, point-source microlensing event model.
+Black (thick) line shows the maximum {\it a posteriori} model.
+}\label{fig:candidates}
+\end{figure*}
+
+\begin{deluxetable*}{cccccccccc}[!th]
+%\tablewidth{0pt}
+%\tabletypesize{\scriptsize}
+\tablecaption{Photometric Properties of the Microlensing Candidates \label{table:photometry}}
+\tablehead{
+\colhead{PTF1} & \colhead{$R$}  & \colhead{$(g'-R)$}   & \colhead{$J$}   & \colhead{$(J-H)$} & \colhead{$(H-K)$} &  \colhead{$W1$}  & \colhead{$(W1-W2)$}  & \colhead{$(W2-W3)$} & \colhead{$(W3-W4)$}
+}
+\startdata
+J0335\tablenotemark{a} & 18.28$\pm$0.12 & \nodata & 16.59$\pm$0.13 & 0.73$\pm$0.18 & 0.79$\pm$0.18 & 15.06$\pm$0.04 & 0.18$\pm$0.09 & \nodata & \nodata \\
+{\bf J0618} & 20.13$\pm$0.10 & 1.13$\pm$0.22 & 16.23$\pm$0.09 & 0.77$\pm$0.15 & 0.70$\pm$0.15 & 14.08$\pm$0.03 & 0.49$\pm$0.05 & 3.22$\pm$0.09 & 2.65$\pm$0.19 \\
+J1206\tablenotemark{b} & 19.24$\pm$0.27 & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata \\
+J1315\tablenotemark{a} & 19.69$\pm$0.28 & \nodata & \nodata & \nodata & \nodata & 15.37$\pm$0.04 & 0.81$\pm$0.06 & 2.46$\pm$0.17 & 2.96$\pm$0.37 \\
+J1532\tablenotemark{b} & 19.25$\pm$0.21 & \nodata & \nodata & \nodata & \nodata & 16.20$\pm$0.05 & $-0.43$$\pm$0.21 & \nodata & \nodata \\
+{\bf J1615} & 20.01$\pm$0.24 & 1.10$\pm$0.29 & \nodata & \nodata & \nodata & 16.36$\pm$0.04 & 0.09$\pm$0.11 & \nodata & \nodata \\
+J1716\tablenotemark{a} & 20.27$\pm$0.20 & 0.17$\pm$0.28 & \nodata & \nodata & \nodata & 15.56$\pm$0.04 & 0.85$\pm$0.07 & 2.93$\pm$0.15 & \nodata \\
+{\bf J1728} & 16.20$\pm$0.10 & 0.82$\pm$0.10 & 14.70$\pm$0.04 & 0.58$\pm$0.06 & 0.16$\pm$0.09 & 13.92$\pm$0.02 & 0.01$\pm$0.03 & $-$0.28$\pm$0.50 & \nodata \\
+J1733\tablenotemark{b} & 20.15$\pm$0.25 & 0.00$\pm$0.34 & 17.18$\pm$0.21 & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata \\
+J1747\tablenotemark{a} & 17.37$\pm$0.10 & 0.20$\pm$0.12 & 16.12$\pm$0.07 & 0.49$\pm$0.13 & 0.92$\pm$0.13 & 13.10$\pm$0.03 & 1.01$\pm$0.03 & 2.96$\pm$0.04 & 2.68$\pm$0.06 \\
+J1933\tablenotemark{a} & 17.53$\pm$0.10 & 1.13$\pm$0.27 & 15.77$\pm$0.07 & 0.74$\pm$0.11 & 0.59$\pm$0.13 & 14.16$\pm$0.03 & 0.15$\pm$0.04 & 0.94$\pm$0.44 & \nodata
+\enddata
+\tablenotetext{a}{Photometry and/or visual inspection of the PTF/WISE images rule these out as plausible candidate events.}
+\tablenotetext{b}{Available data are insufficient to determine the likelihood that these are plausible candidate events.}
+\tablecomments{Objects in bold are plausible microlensing candidates.}
+\end{deluxetable*}
 
@@
 Where appropriate, as an additional test, we examine the PTF images to verify the data quality.
-A small fraction of the selected light curves are variable stars that made it past our periodic variability cut.
+A small fraction of the light curves are variable stars that survived our periodic variability cut.
 
@@
 
-To summarize: we rule out five of the 11 candidates as being likely extragalactic transients based on photometry from WISE and/or 2MASS.
-Three of the remaining objects are ambiguous because they are not detected in WISE and/or 2MASS; follow-up imaging or spectroscopy is needed to draw any conclusions about these candidates.
-The remaining three objects (PTF1J061800.25$+$203142.5,
-J161502.39$+$540053.8, and J172826.08$+$692501.1) are likely stars and, given the observed PTF variability, plausible microlensing events.
+To summarize what follows: we rule out five of the 11 candidates as being likely extragalactic transients based on photometry and/or imaging from WISE, 2MASS, and/or PTF.
+Three of the candidates are not detected in WISE and/or 2MASS; follow-up imaging or spectroscopy is needed to draw any conclusions about these objects.
+The remaining three objects (PTF1J061800.25$+$203142.5, J161502.39$+$540053.8, and J172826.08$+$692501.1) are likely stars and, given the observed PTF variability, plausible microlensing events.
 We emphasize, however, that lacking simultaneous, multi-color photometry during the events, we cannot claim these as microlensing events with any confidence.
-%The reader may wish to focus on the description of the available data for these three candidates before moving on to the next section.
+These three candidates are shown in Figure~\ref{fig:candidates}, and PTF cutouts of the associated sources at quiescence and near peak brightness are shown in Figure~\ref{fig:postage_stamps}.
+Light curves for the other eight events are shown in Figure~\ref{fig:not_candidates}.
 
-\subsubsection*{PTF1J033545.80$-$041849.1} %w
-%PTFS1203am
-%This object was flagged as a transient by the PTF marshal but not followed up. 
-PTF1J0335 has a counterpart in 2MASS, and its $(J-H)$ and $(H-K)$ colors indicate that it is unlikely to be a star \citep{kev07}.
-Because it is only detected in the $W1$ and $W2$ bands, the WISE color does not help much: it is consistent with that of stars and of compact galaxies \citep[cf.\ Figure 14 in][]{yan2013}.
-We tentatively conclude that this not a good microlensing candidate.
+For these three plausible events, we use a Markov-chain Monte-Carlo algorithm \citep{goodman, dfm} to derive posterior probability distributions over each parameter in the point-source, point-lens microlensing model (Eq.~\ref{eq:ml_model}).
+Overplotted on Figure~\ref{fig:candidates} are samples from these posterior distributions along with the maximum {\it a posteriori} (MAP) model; the corresponding MAP event parameters are listed in Table~\ref{table:candidates1}.
+The event durations for PTF1J0618 and PTF1J1615 are reasonable, given that a ${\sim}0.1\ M_\odot$ lens at ${\sim}500~\mathrm{pc}$ in the thick disk ($v_{tan}\sim 50~\mathrm{km}~\mathrm{s}^{-1}$) has a typical event duration $t_E \sim 20~\mathrm{days}$, but the significantly longer duration of PTF1J1728 ($>$150 days) could be a sign of a greater mass lens or larger distance.
 
+% postage stamps
+\begin{figure}
+\centering
+\includegraphics[width=1.0\columnwidth] {fig11_0.eps}
+\includegraphics[width=1.0\columnwidth] {fig11_1.eps}
+\includegraphics[width=1.0\columnwidth] {fig11_2.eps}
+\caption{$2'$ by $2'$ PTF $R$-band images of the three sources in Figure~\ref{fig:candidates}.
+{\it Left} ---  Sources at quiescence.
+{\it Right} --- PTF image closest in time to the peak event time $t_0$ obtained by our model fits.
+Images are shown North-up, East-left.}\label{fig:postage_stamps}
+\end{figure}
+
+PTF1J0618 and PTF1J1728 have counterparts in 2MASS, and we can therefore use the \citet{kev07} $J$-band absolute magnitudes for stars of the relevant spectral types to estimate rough distances to these stars, assuming they have solar metallicity (see below for discussion of how we obtain spectral types for these stars).
+For PTF1J0618, this results in a distance that ranges from 1.1 kpc (if it is a M1 star) to 0.8 kpc (if it is a M3 star), corresponding to heights above the Galactic Plane 43 and 31 pc, respectively.
+For PTF1J1728, assuming it is a K0-K5 star results in distances ranging from 1.3 to 0.7 kpc, and heights of 0.9 to 0.5~kpc.
+
+We lack a spectrum for PTF1J1615, so the distance estimate to this star is even more uncertain.
+Still, we compare the PTF $R$ magnitude to the stellar library compiled by \citet{Pickles1998} to estimate a distance, assuming (from its photometry) that it is a K5-M2 star\footnote{This ignores differences between Mould and Cousins $R$ filters.}.
+This results in distances ranging from 4.2 to 1.9 kpc and heights from 2.3 to 1.0~kpc.
+The estimated heights above the Galactic Plane suggest that PTF1J0618 is a thin disk star, while the other two, PTF1J1615 and PT1J1728, are plausible thick disk members \citep{Bochanski2010}.
+
+\begin{deluxetable*}{ccccc}[!h]
+%\tablewidth{0pt}
+\tabletypesize{\scriptsize}
+\tablecaption{Derived Microlensing Event Parameters for Plausible Candidates \label{table:candidates1}}
+\tablehead{
+\colhead{PTF1} & \colhead{$u_0$}  & \colhead{$t_E$ [days]}   & \colhead{$t_0$ [MJD]}   & \colhead{$m_0$ [R]}     
+}
+\startdata
+J0618 & $0.178\pm0.019$ & $22.8\pm1.4$ & $55879.975\pm0.248$ & $20.226\pm0.026$\\
+J1615 & $0.388\pm0.015$ & $32.5\pm2.3$  & $55895.797\pm1.956$ & $20.016\pm0.005$\\
+J1728 & $1.338\pm0.001$ & $156.6\pm1.2$ & $55194.313\pm0.702$ & $16.259\pm0.001$ 
+\enddata
+\end{deluxetable*}
+
+\subsection{Plausible Microlensing Candidates}
 \subsubsection*{PTF1J061800.25$+$203142.5} % (a)}
@@
 Meanwhile, PTF1J0618's WISE colors suggest it is an extragalactic object \citep[cf.\ Figure 14 in][]{yan2013}.
-
-We observed PTF1J0618 with the Double Spectrograph \citep[DBSP;][]{oke_gunn} on the Hale 5-m telescope at Palomar Observatory, CA, on 2013 Feb 19; see Figure~\ref{fig:1206i_spectrum}.
-The spectrum was obtained with the D55 dichroic; from the atmospheric cutoff to 5500 \AA, the grating had 600 line~mm$^{-1}$ and was blazed at 4000 \AA, giving a resolution of 1.1 \AA.
-From 6300-8800 \AA, the grating was 158 line~mm$^{-1}$, was blazed at 7500\ \AA, and gave a resolution of 2.5 \AA.
+We observed PTF1J0618 for 1200~s with the Double Spectrograph \citep[DBSP;][]{oke_gunn} on the Hale 5-m telescope at Palomar Observatory, CA, on 2013 Feb 19; see Figure~\ref{fig:1206i_spectrum}.
+The spectrum was obtained with the D55 dichroic; from the atmospheric cutoff to 5500~\AA, the grating had 600 line~mm$^{-1}$ and was blazed at 4000~\AA, giving a resolution of 1.1~\AA.
+From 6300-8800~\AA, the grating was 158 line~mm$^{-1}$, was blazed at 7500~\AA, and gave a resolution of 2.5~\AA.
 The data were reduced using standard IRAF routines\footnote{IRAF is distributed by the National Optical Astronomy Observatories, which are operated by the Association of Universities for Research in Astronomy, Inc., under cooperative agreement with the National Science Foundation.}.
@@
 
+\begin{figure}[!h]
+	\centering
+	\includegraphics[width=.48\textwidth, trim=0 0 15 40, clip] {fig12.eps}
+\caption{\label{fig:1206i_spectrum} DBSP spectrum of PTF1J0618, an early M dwarf star, smoothed with a Gaussian filter ($\sigma = 1$).
+The timescale of the event detected by PTF is inconsistent with those generally observed in flares, which are the mostly likely contaminants in identifying microlensing events involving M stars.}
+\end{figure}
+
+Could PTF1J0618 be an outbursting symbiotic star?
+To test this hypothesis, we compare the DBSP spectrum to a set of M1-M2 dwarf and giant star spectra.
+We inspect the depth of the Ca II triplet lines and also calculate several of the spectral indices collected by \citet{kev06} to distinguish between dwarfs and giants (e.g., the fluxes in the Na D 5900 and CN 7900 bands).
+These all indicate that PTF1J0618 is a dwarf star, rendering it unlikely that the PTF event was a symbiotic-type outburst.
+We further note that those outbursts typically last $\sim$years \citep[e.g.,][]{tomov2013}.
+We conclude that this is a good candidate microlensing event.
+
 %Moderate-resolution optical spectra of PTF1J033545.80$-$041849.1 and PTF1J061800.25$+$203142.5 were obtained on 2013-01-10 and 2013-02-19, respectively, with the Double Spectrograph \citep[DBSP;][]{oke_gunn} on the Hale 5 m telescope at Palomar Observatory (Figures~\ref{fig:1206i_spectrum}\&\ref{fig:1203am_spectrum}).
@@
 
-\begin{figure}[h]
+\subsubsection*{PTF1J161502.39$+$540053.8} % (c)}
+%PTFS1216bt
+PTF1J1615 is one of two candidates that fall in the SDSS footprint; its counterpart, SDSS J161502.43$+$540053.6, has a proper motion of 2.66 mas/year, consistent with its automated SDSS classification as a star.
+The star has $u = 22.9$ mag\footnote{These are SDSS PSF magnitudes.
+Typically, PSF fitting provides better estimates of isolated star magnitudes; see \citet{stoughton02}.}, beyond the $u = 22.0$ 95\% completeness limit for SDSS \citep{stoughton02}, and we therefore focus on its $griz$ photometry.
+With $(g-r) = 1.10\pm0.05$, $(r-i) = 1.00\pm0.03$, and $(i-z) = 0.56\pm0.04$, the star's colors are consistent with those of a late K/early M star \citep{kev07}.
+PTF1J1615 has no counterpart in 2MASS, and is only detected in the WISE $W1$ and $W2$ bands, making it hard to draw any conclusions based on these data.
+Still, its $(W1-W2)$ is consistent with what is seen for stars \citep[cf.\ Figure 14 in][]{yan2013}.
+Here again, the most likely source of contamination is flaring.
+However, the event detected by PTF had a duration $\sim$20 days, and so is highly unlikely to have been a flare.
+We conclude that this is a good candidate microlensing event.
+
+\subsubsection*{PTF1J172826.08$+$692501.1} % (e)}
+%PTFS1217cv
+% http://skyserver.sdss3.org/dr9/en/tools/explore/obj.asp?id=1237655748311122026
+PTF1J1728 is the other candidate in the SDSS footprint.
+Its SDSS match, SDSS J172826.07$+$692501.3, has a proper motion of 1.81 mas/year, consistent with its automated classification as a star.
+PTF1J1728 also has a counterpart in 2MASS, and its SDSS/2MASS colors --- $(u-g) = 1.81\pm0.04$, $(g-r) = 0.79\pm 0.03$, $(r-i) = 0.35\pm0.03$, $(i-z) = 0.21\pm0.03$, $(z-J) = 0.97 \pm0.05$, $(J-H) = 0.58\pm0.06$, and $(H-K) = 0.16\pm 0.09$ --- are consistent with those of a K star \citep{kev07}.
+Its $(W1-W2)$ and $(W2-W3)$ colors place PTF1J1728 in the stellar locus in Figure 14 of \citet[][]{yan2013}.
+
+\begin{figure}[!h]
 	\centering
-	\includegraphics[width=.48\textwidth, trim=17 15 15 40, clip] {figures/PTFS1206i_new.pdf}
-\caption{\label{fig:1206i_spectrum} DBSP spectrum of PTF1J0618 smoothed with a Gaussian filter ($\sigma = 1$).
-This object is most likely a M2 dwarf star.
-The timescale of the event detected by PTF is inconsistent with those generally observed in flares, which are the mostly likely contaminants in identifying microlensing events involving M stars.}
+	\includegraphics[width=.48\textwidth, trim=0 0 15 40, clip] {fig13.eps}
+\caption{\label{fig:1217cv_spectrum} MODspec spectrum of PTF1J172826.08$+$692501.1, a mid-K star.}
 \end{figure}
 
-Could PTF1J0618 be an outbursting symbiotic star?
-To test this hypothesis, we compare the DBSP spectrum to a set of M1-M2 dwarf and giant star spectra.
-In particular, we inspect the depth of the Ca II triplet lines; we also calculate several of the spectral indices collected by \citet{kev06} to distinguish between dwarfs and giants (e.g., the fluxes in the Na D 5900 and CN 7900 bands).
-These all indicate that PTF1J0618 is a dwarf star, rendering it unlikely that the PTF event was a symbiotic-type outburst.
-We further note that these typically have much longer durations, e.g., the systems discussed in \citet{tomov2013}, for which outbursts last a few years.
+E.~Bowsher and S.~Douglas obtained two 1200-s spectra of PTF1J1728 with the MDM Observatory Modular Spectrograph (MODspec) on the Hiltner 2.4-m telescope at MDM Observatory, Kitt Peak, AZ, on 2012 Nov 14\footnote{The MDM Observatory is operated by Dartmouth College, Columbia University, Ohio State University, Ohio University, and the University of Michigan.}.
+MODspec was configured to provide coverage from 4500 to 7500 \AA~with $\sim$1.8 \AA~sampling and a spectral resolution of $\sim$3300.
+The data reduction was performed by S.~Douglas using standard IRAF routines.
+% The spectrum was trimmed, overscan- and bias-corrected, cleaned of cosmic rays, flat-fielded, extracted, dispersion-corrected, and flux-calibrated by S.~Douglas using standard IRAF tasks.
+All the spectra from this night suffered from significant noise blueward of 5000 \AA~and redward of 7000 \AA; in  Figure~\ref{fig:1217cv_spectrum}, we therefore present the spectrum over that 2000 \AA~range.
+While the spectrum is noisy, PTF1J1728 appears to be a mid-K star; its HAMMER spectral type is K3.
 We conclude that this is a good candidate microlensing event.
 
-\subsubsection*{PTF1J120642.60$-$192016.4} % (b)}
-%PTFS1212ak
-PTF1J1206 has no counterpart within 10\asec\ in 2MASS.
-While a counterpart is listed in the WISE data 3\asec\ from the PTF position, a visual comparison of the co-added PTF $R$-band images and of the WISE images indicates that the PTF object may be blended with other objects in the WISE data or not detected at all.
-However, the WISE photometric flags for this nearby source suggest it is clean and unblended.
-The available information is insufficient to draw even tentative conclusions about the nature of this object.
+\subsection{Candidates Ruled out by Photometry and/or Imaging}
+\subsubsection*{PTF1J033545.80$-$041849.1} %w
+%PTFS1203am
+%This object was flagged as a transient by the PTF marshal but not followed up. 
+PTF1J0335 has a counterpart in 2MASS, and its $(J-H)$ and $(H-K)$ colors indicate that it is unlikely to be a star \citep{kev07}.
+Because it is only detected in the $W1$ and $W2$ bands, the WISE color does not help much: it is consistent with that of stars and of compact galaxies \citep[cf.\ Figure 14 in][]{yan2013}.
+We tentatively conclude that this not a good microlensing candidate.
 
@@
 
-\subsubsection*{PTF1J153202.91$+$674825.1} % (y)}
-%PTFS1215aj
-PTF1J1532 has no counterpart in 2MASS, but is detected in the WISE $W1$ and $W2$ bands.
-The corresponding $(W1-W2) = -0.43\pm0.2$ color is somewhat unusual, as most objects appear to have $(W1-W2)\ \gapprox\ 0.0$, with compact galaxies being the rare exceptions \citep[cf.\ Figure 14 in][]{yan2013}.
-Here again, we cannot draw any conclusions about the nature of this object.
-
-\subsubsection*{PTF1J161502.39$+$540053.8} % (c)}
-%PTFS1216bt
-PTF1J1615 is one of two candidates that fall in the SDSS footprint; its counterpart, SDSS J161502.43$+$540053.6, has a proper motion of 2.66 mas/year, consistent with its automated SDSS classification as a star.
-The star has $u = 22.9$ mag\footnote{These are SDSS PSF magnitudes.
-Typically, PSF fitting provides better estimates of isolated star magnitudes; see \citet{stoughton02}.}, beyond the $u = 22.0$ 95\% completeness limit for SDSS \citep{stoughton02}, and we therefore focus on its $griz$ photometry.
-With $(g-r) = 1.10\pm0.05$, $(r-i) = 1.00\pm0.03$, and $(i-z) = 0.56\pm0.04$, the star's colors are consistent with those of a late K/early M star \citep{kev07}.
-PTF1J1615 has no counterpart in 2MASS, and is only detected in the WISE $W1$ and $W2$ bands, making it hard to draw any conclusions as to the candidate's nature based on these data.
-Still, its $(W1-W2)$ is consistent with what is seen for stars \citep[cf.\ Figure 14 in][]{yan2013}.
-Here again, the most likely source of contamination is flaring.
-However, the event detected by PTF had a duration $\sim$20 days, and so is highly unlikely to have been a flare.
-We conclude that this is a good candidate microlensing event.
-
-%\renewcommand{\thefootnote}{\alph{footnote}}
-\begin{table*}
-\small
-\centering
-\caption{Photometric Properties of the Microlensing Candidates} 
-\begin{tabular}{| c | c | c | c | c | c | c | c | c | c |}
-\hline
-PTF1 & $R$ & $(g'-R)$ & $J$ & $(J-H)$ & $(H-K)$ & $W1$ & $(W1-W2)$ & $(W2-W3)$ & $(W3-W4)$  \\
-\hline
-\hline
-J0335\tablenotemark{a} & 18.28$\pm$0.12 & \nodata & 16.59$\pm$0.13 & 0.73$\pm$0.18 & 0.79$\pm$0.18 & 15.06$\pm$0.04 & 0.18$\pm$0.09 & \nodata & \nodata \\\hline
-{\bf J0618} & 20.13$\pm$0.10 & 1.13$\pm$0.22 & 16.23$\pm$0.09 & 0.77$\pm$0.15 & 0.70$\pm$0.15 & 14.08$\pm$0.03 & 0.49$\pm$0.05 & 3.22$\pm$0.09 & 2.65$\pm$0.19 \\\hline
-J1206 & 19.24$\pm$0.27 & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata \\\hline
-J1315\tablenotemark{a} & 19.69$\pm$0.28 & \nodata & \nodata & \nodata & \nodata & 15.37$\pm$0.04 & 0.81$\pm$0.06 & 2.46$\pm$0.17 & 2.96$\pm$0.37 \\\hline
-J1532 & 19.25$\pm$0.21 & \nodata & \nodata & \nodata & \nodata & 16.20$\pm$0.05 & $-0.43$$\pm$0.21 & \nodata & \nodata \\\hline
-{\bf J1615} & 20.01$\pm$0.24 & 1.10$\pm$0.29 & \nodata & \nodata & \nodata & 16.36$\pm$0.04 & 0.09$\pm$0.11 & \nodata & \nodata \\\hline
-J1716\tablenotemark{a} & 20.27$\pm$0.20 & 0.17$\pm$0.28 & \nodata & \nodata & \nodata & 15.56$\pm$0.04 & 0.85$\pm$0.07 & 2.93$\pm$0.15 & \nodata \\\hline
-{\bf J1728} & 16.20$\pm$0.10 & 0.82$\pm$0.10 & 14.70$\pm$0.04 & 0.58$\pm$0.06 & 0.16$\pm$0.09 & 13.92$\pm$0.02 & 0.01$\pm$0.03 & $-$0.28$\pm$0.50 & \nodata \\\hline
-J1733 & 20.15$\pm$0.25 & 0.00$\pm$0.34 & 17.18$\pm$0.21 & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata \\\hline
-J1747\tablenotemark{a} & 17.37$\pm$0.10 & 0.20$\pm$0.12 & 16.12$\pm$0.07 & 0.49$\pm$0.13 & 0.92$\pm$0.13 & 13.10$\pm$0.03 & 1.01$\pm$0.03 & 2.96$\pm$0.04 & 2.68$\pm$0.06 \\\hline
-J1933\tablenotemark{a} & 17.53$\pm$0.10 & 1.13$\pm$0.27 & 15.77$\pm$0.07 & 0.74$\pm$0.11 & 0.59$\pm$0.13 & 14.16$\pm$0.03 & 0.15$\pm$0.04 & 0.94$\pm$0.44 & \nodata \\
-\hline
-\end{tabular}
-%\footnotetext{We tentatively rule out these candidates based on their photometry and/or visual inspection of the PTF/WISE images.}
-\tablecomments{The bolded objects are plausible microlensing candidates.
-Candidates flagged with an ``a'' are tentatively ruled out based on their photometry and/or visual inspection of the PTF/WISE images.}
-\label{table:photometry}
-\end{table*}
-%\renewcommand{\thefootnote}{\arabic{footnote}}
-
 \subsubsection*{PTF1J171606.82$+$474423.7} % (d)}
@@
 We tentatively conclude that this not a good microlensing candidate.
-
-\subsubsection*{PTF1J172826.08$+$692501.1} % (e)}
-%PTFS1217cv
-% http://skyserver.sdss3.org/dr9/en/tools/explore/obj.asp?id=1237655748311122026
-PTF1J1728 is the other candidate in the SDSS footprint.
-Its SDSS match, SDSS J172826.07$+$692501.3, has a proper motion of 1.81 mas/year, consistent with its automated classification as a star.
-PTF1J1728 also has a counterpart in 2MASS, and its SDSS/2MASS colors --- $(u-g) = 1.81\pm0.04$, $(g-r) = 0.79\pm 0.03$, $(r-i) = 0.35\pm0.03$, $(i-z) = 0.21\pm0.03$, $(z-J) = 0.97 \pm0.05$, $(J-H) = 0.58\pm0.06$, and $(H-K) = 0.16\pm 0.09$ --- are consistent with those of a K star \citep{kev07}.
-Its $(W1-W2)$ and $(W2-W3)$ colors place PTF1J1728 in the stellar locus in Figure 14 of \citet[][]{yan2013}.
-
-E.~Bowsher and S.~Douglas obtained a spectrum of PTF1J1728 with the MDM Observatory Modular Spectrograph (MODspec) on the Hiltner 2.4-m telescope at MDM Observatory, Kitt Peak, AZ, on 2012 Nov 14\footnote{The MDM Observatory is operated by Dartmouth College, Columbia University, Ohio State University, Ohio University, and the University of Michigan.}.
-MODspec was configured to provide coverage from 4500 to 7500 \AA~with $\sim$1.8 \AA~sampling and a spectral resolution of $\sim$3300.
-The data reduction was performed by S.~Douglas using standard IRAF routines.
-% The spectrum was trimmed, overscan- and bias-corrected, cleaned of cosmic rays, flat-fielded, extracted, dispersion-corrected, and flux-calibrated by S.~Douglas using standard IRAF tasks.
-All the spectra from this night suffered from significant noise blueward of 5000 \AA~and redward of 7000 \AA; in  Figure~\ref{fig:1217cv_spectrum}, we therefore present the spectrum over that 2000 \AA~range.
-Although the spectrum is noisy, PTF1J1728 appears to be a mid-K star, with a HAMMER spectral type of K3.
-We conclude that this is a good candidate microlensing event.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=.48\textwidth, trim=17 15 15 40, clip] {figures/PTFS1217cv.pdf}
-\caption{\label{fig:1217cv_spectrum} MODspec spectrum of PTF1J172826.08$+$692501.1.}
-\end{figure}
-
-\subsubsection*{PTF1J173301.07$+$374311.9} % (f)}
-%PTFS1217df
-%PTF1J1733 was flagged as a transient by the PTF pipeline but not followed up with high-cadence imaging or spectroscopy. 
-PTF1J1733 is detected in the 2MASS $J$ and $K$ bands (but not $H$).
-There is no object in the WISE catalog within 10\asec\ of PTF1J1733.
-Although the $(J-K$) color appears inconsistent with PTF1J1733 being a star \citep{kev07}, we cannot draw any firm conclusions about the nature of this object.
-
 \subsubsection*{PTF1J174736.51$+$300506.9} % (g)}
@@
 
+\subsection{Candidates Lacking Sufficient Data to Evaluate Likelihood}
+\subsubsection*{PTF1J120642.60$-$192016.4} % (b)}
+%PTFS1212ak
+PTF1J1206 has no counterpart within 10\asec\ in 2MASS.
+While a counterpart is listed in the WISE data 3\asec\ from the PTF position, a visual comparison of the co-added PTF $R$-band images and of the WISE images indicates that the PTF object may be blended with other objects in the WISE data or not detected at all.
+However, the WISE photometric flags for this nearby source suggest it is clean and unblended.
+The available information is insufficient to draw even tentative conclusions about the nature of this object.
+
+\subsubsection*{PTF1J153202.91$+$674825.1} % (y)}
+%PTFS1215aj
+PTF1J1532 has no counterpart in 2MASS, but is detected in the WISE $W1$ and $W2$ bands.
+The corresponding $(W1-W2) = -0.43\pm0.2$ color is somewhat unusual, as most objects appear to have $(W1-W2)\ \gapprox\ 0.0$, with compact galaxies being the rare exceptions \citep[cf.\ Figure 14 in][]{yan2013}.
+Here again, we cannot draw any conclusions about the nature of this object.
+
+\subsubsection*{PTF1J173301.07$+$374311.9} % (f)}
+%PTFS1217df
+%PTF1J1733 was flagged as a transient by the PTF pipeline but not followed up with high-cadence imaging or spectroscopy. 
+PTF1J1733 is detected in the 2MASS $J$ and $K$ bands (but not $H$).
+There is no object in the WISE catalog within 10\asec\ of PTF1J1733.
+Although the $(J-K$) color appears inconsistent with PTF1J1733 being a star \citep{kev07}, we cannot draw any firm conclusions about the nature of this object.
+
 \section{Conclusions and Future Work}\label{sec:conc}
-Microlensing event rate predictions away from the Galactic Plane are not well constrained but are being modeled in detail by Fournier et al.
-(in prep).
-Only a single event has been observed in a sparse stellar field, despite the great scientific interest in observing such events.
-Using time-domain data from the Palomar Transient Factory, we have developed new methods for identifying interesting transients in massive light curve datasets and for searching for sparse-field microlensing events.
-
+Using time-domain data from the Palomar Transient Factory, we have developed a new method for identifying interesting transients in massive light-curve data sets and for searching for sparse-field microlensing events.
 We have examined the detection efficiency of recovering simulated microlensing events using a set of variability statistics adapted from \citet{shin2009}.
@@
 We used $\eta$ to develop a selection procedure for extracting microlensing event candidates from the PTF light-curve archive.
-Among the $1.1 \times 10^9$ light curves with $>$10 $R$-band observations we first identified $\sim$2000 interesting candidate transients, from which we selected 11 candidate microlensing events (see Figures~\ref{fig:not_candidates1}-\ref{fig:candidates}).
+Among the $1.1 \times 10^9$ light curves with $>$10 $R$-band observations, we first identified $\sim$2000 interesting candidate transients, from which we selected 11 candidate microlensing events.
 
@@
 We also recovered a large number of transient events such as novae, outbursts, and flares.
-We have cross-referenced this list of objects with the PTF transient detection system and found that most of these objects are known sources, but our pipeline discovered at least two supernovae that the PTF pipeline had missed.
+We have cross-referenced this list of objects with the PTF transient detection system and found that most of these objects are known sources, but we discovered at least two supernovae that the PTF pipeline had missed.
 
-Of the 11 candidate microlensing events, we tentatively ruled out eight by examining their 2MASS or WISE photometry, or because they lacked such data.
+Of the 11 candidate microlensing events, we tentatively ruled out eight by examining the available photometry and images, or because they lacked such data.
 We labeled the three remaining objects as plausible microlensing events, but lacking simultaneous multi-color imaging during the events, we cannot confirm these candidates as true microlensing events.
 
-For the three plausible events, we use a Markov-chain Monte-Carlo algorithm \citep{goodman, dfm} to derive posterior probability distributions over each parameter in the point-source, point-lens microlensing model (Eq.~\ref{eq:ml_model}).
-Overplotted on Figure~\ref{fig:candidates} are samples from these posterior distributions along with the maximum {\it a posteriori} model; the corresponding event parameters are listed in Table~\ref{table:candidates1}.
-The event durations for J0618 and J1615 are reasonable, given that a ${\sim}0.1M_\odot$ lens at ${\sim}500~\mathrm{pc}$ in the thick disk ($v_{tan}\sim 50~\mathrm{km}~\mathrm{s}^{-1}$) has a typical event duration $t_E \sim 20~\mathrm{days}$, but the significantly longer duration of J1728 ($>150~\mathrm{days}$) could be a sign of a greater mass lens or larger distance.
+\begin{figure*}
+\centering
+\includegraphics[width=2.1\columnwidth] {fig14_0.eps}
+\includegraphics[width=2.1\columnwidth] {fig14_1.eps}
+\includegraphics[width=2.1\columnwidth] {fig14_2.eps}
+\caption{Full light curves (top) and zooms around the transient maximum (bottom) for the rejected microlensing event candidates. Bad (flagged) data points have been removed.}\label{fig:not_candidates}
+\end{figure*}
 
-%\renewcommand*\arraystretch{1.8}
-\begin{table*}
-\small
-\centering
-\caption{Microlensing Event Parameters Derived for Plausible Candidates} 
-\begin{tabular}{| c | c | c | c | c | c | c | c | c | c |}
-\hline
-ID & $u_0$ & $t_E$ [days] & $t_0$ [MJD] & $m_0$ [R] \\
-\hline
-J0618 & $0.160\pm0.015$ & $22.9\pm1.6$ & $55879.975\pm0.248$ & $20.216\pm0.032$\\\hline
-J1615 & $0.414\pm0.025$ & $31.8\pm2.3$  & $55895.797\pm1.956$ & $20.016\pm0.005$\\\hline
-J1728 & $1.338\pm0.001$ & $158.3\pm2.1$ & $55194.313\pm0.702$ & $16.259\pm0.001$ \\\hline
-\end{tabular}
-\label{table:candidates1}
-\end{table*}
+For these three plausible events, we use a Markov-chain Monte-Carlo algorithm to derive posterior probability distributions over each parameter in the point-source, point-lens microlensing model.
+The event durations for two of the candidates are reasonable, but the significantly longer duration of the third candidate could be a sign of a greater mass lens or larger distance.
+
+While this number of candidate events is consistent with simple predictions, microlensing event rate predictions away from the Galactic Plane are generally not well constrained.
+In a forthcoming paper, we will present a detailed investigation of the all-sky microlensing event rate,  generated from a model that includes realistic mass and velocity distributions
+(Fournier et al., in prep).
 
@@
 Rigorously classifying and cataloging these sources was outside of the scope of this paper, but many  presented interesting forms of variability.
-Though it is not necessarily scalable for future surveys to look at all light curves selected by using our algorithm, these $\sim$2000 light curves, once carefully classified, could serve as a training set for a more sophisticated machine-learning approach \cite[as discussed by e.g.,][]{bloomPasp} to classify the light curves that survive our selection criteria.
+Though it is not necessarily scalable for future surveys to look at all light curves selected by using our algorithm, these $\sim$2000 light curves, once carefully classified, could serve as a training set for a more sophisticated machine-learning approach \cite[as discussed by, e.g.,][]{bloomPasp} to classify the light curves that survive our selection criteria.
 Approaches that combine statistical cuts such as those described in this paper and some type of machine learning will be essential to making the most of the future time-domain data such as that collected by the Large Synoptic Survey Telescope.
-
-% TODO: Some reference to these needs a home: PTF data web interfaces\footnote{For example: \url{http://bit.ly/PTFSky}, \url{http://bit.ly/PTFDSky}, \url{http://bit.ly/PTFCandidates}}.
 
@@
 
-%\appendix
-%\label{appendix}
+%\bigskip
+%\setlength{\baselineskip}{0.6\baselineskip}
+%\bibliography{references_5}
+%\bibliography{references_3}
+%\setlength{\baselineskip}{1.667\baselineskip}
 
-%\subsubsection*{}
-%Follow-up spectra were obtained for two candidates (PTF1J033545.80$-$041849.1 and PTF1J153202.91$+$674825.1) which classified these sources as galaxies, thus excluding them from the candidate sample as probable supernovae.
-A third object (PTF1J193330.11$+$451501.1) appears elongated in the PTF coadded image and is also rejected.
+\begin{thebibliography}{53}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
 
-\bigskip
-\setlength{\baselineskip}{0.6\baselineskip}
-%\bibliography{references_5}
-\bibliography{references_3}
-\setlength{\baselineskip}{1.667\baselineskip}
+\bibitem[{{Abazajian} {et~al.}(2009)}]{DR7paper}
+{Abazajian}, K.~N., {et~al.} 2009, \apjs, 182, 543
 
-\clearpage
+\bibitem[{{Ag{\"u}eros} {et~al.}(2011){Ag{\"u}eros}, {Covey}, {Lemonias},
+  {Law}, {Kraus}, {Batalha}, {Bloom}, {Cenko}, {Kasliwal}, {Kulkarni},
+  {Nugent}, {Ofek}, {Poznanski}, \& {Quimby}}]{agueros11}
+{Ag{\"u}eros}, M.~A., {Covey}, K.~R., {Lemonias}, J.~J., {et~al.} 2011, \apj,
+  740, 110
 
-\begin{figure}
-\centering
-\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1203am_lc.png}}
-\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1212ak_lc.png}}
-\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1213ap_lc.png}}
-\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1215aj_lc.png}}
-\caption{Full light curves (top) and zooms around the transient maximum (bottom) for rejected microlensing event candidates found in PTF.
-Bad (flagged) data points have been removed.}\label{fig:not_candidates1}
-\end{figure}
+\bibitem[{{Alard} \& {Lupton}(1998)}]{alard1998}
+{Alard}, C., \& {Lupton}, R.~H. 1998, \apj, 503, 325
 
-\begin{figure}
-\centering
-\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1217ce_lc.png}}
-\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1217df_lc.png}}
-\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1217dk_lc.png}}
-\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1219ad_lc.png}}
-\caption{Same as Figure~\ref{fig:not_candidates1}. }\label{fig:not_candidates2}
-\end{figure}
+\bibitem[{{Alcock} {et~al.}(1993){Alcock}, {Allsman}, {Axelrod}, {Bennett},
+  {Cook}, {Park}, {Marshall}, {Stubbs}, {Griest}, {Perlmutter}, {Sutherland},
+  {Freeman}, {Peterson}, {Quinn}, \& {Rodgers}}]{original_macho}
+{Alcock}, C., {Allsman}, R.~A., {Axelrod}, T.~S., {et~al.} 1993, in
+  Astronomical Society of the Pacific Conference Series, Vol.~43, Sky Surveys.
+  Protostars to Protogalaxies, ed. {B.~T.~Soifer}, 291
 
-\begin{figure}
-\centering
-\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1206i_lc.png}}
-\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1216bt_lc.png}}
-\subfigure{\includegraphics[width=0.48\textwidth, trim=-5 25 30 15, clip] {figures/PTFS1217cv_lc.png}}
-\caption{Same as Figure~\ref{fig:not_candidates1} for three microlensing event candidates.
-Lines (gray, thin) show models with parameters sampled from the posterior probability distribution over the four parameters in the point-lens, point-source microlensing event model.
-Red (thick) line shows the maximum {\it a posteriori} model.
-}\label{fig:candidates}
-\end{figure}
+\bibitem[{{Alcock} {et~al.}(2000){Alcock}, {Allsman}, {Alves}, {Axelrod},
+  {Becker}, {Bennett}, {Cook}, {Drake}, {Freeman}, {Geha}, {Griest}, {Lehner},
+  {Marshall}, {Minniti}, {Nelson}, {Peterson}, {Popowski}, {Pratt}, {Quinn},
+  {Stubbs}, {Sutherland}, {Tomaney}, {Vandehei}, {Welch}, \& {MACHO
+  Collaboration}}]{alcock2000}
+{Alcock}, C., {Allsman}, R.~A., {Alves}, D.~R., {et~al.} 2000, \apj, 541, 734
+
+\bibitem[{{Alcock} {et~al.}(2001){Alcock}, {Allsman}, {Alves}, {Axelrod},
+  {Becker}, {Bennett}, {Cook}, {Drake}, {Freeman}, {Geha}, {Griest}, {Lehner},
+  {Marshall}, {Minniti}, {Nelson}, {Peterson}, {Popowski}, {Pratt}, {Quinn},
+  {Stubbs}, {Sutherland}, {Tomaney}, {Vandehei}, \&
+  {Welch}}]{macho_detection_efficiency}
+{Alcock}, C., {Allsman}, R.~A., {Alves}, D., {et~al.} 2001, \apjs, 136, 439
+
+\bibitem[{{Aubourg} {et~al.}(1993){Aubourg}, {Bareyre}, {Brehin}, {Gros},
+  {Lachieze-Rey}, {Laurent}, {Lesquoy}, {Magneville}, {Milsztajn}, {Moscoso},
+  {Queinnec}, {Rich}, {Spiro}, {Vigroux}, {Zylberajch}, {Ansari}, {Cavalier},
+  {Moniez}, {Beaulieu}, {Ferlet}, {Grison}, {Vidal-Madjar}, {Guibert},
+  {Moreau}, {Tajahmady}, {Maurice}, {Prevot}, \& {Gry}}]{eros_original}
+{Aubourg}, E., {Bareyre}, P., {Brehin}, S., {et~al.} 1993, The Messenger, 72,
+  20
+
+\bibitem[{{Bertin} \& {Arnouts}(1996)}]{bertin96}
+{Bertin}, E., \& {Arnouts}, S. 1996, \aaps, 117, 393
+
+\bibitem[{{Binney} {et~al.}(2000){Binney}, {Bissantz}, \&
+  {Gerhard}}]{binney2000}
+{Binney}, J., {Bissantz}, N., \& {Gerhard}, O. 2000, \apjl, 537, L99
+
+\bibitem[{{Bloom} {et~al.}(2012){Bloom}, {Richards}, {Nugent}, {Quimby},
+  {Kasliwal}, {Starr}, {Poznanski}, {Ofek}, {Cenko}, {Butler}, {Kulkarni},
+  {Gal-Yam}, \& {Law}}]{bloomPasp}
+{Bloom}, J.~S., {Richards}, J.~W., {Nugent}, P.~E., {et~al.} 2012, \pasp, 124,
+  1175
+
+\bibitem[{{Bochanski} {et~al.}(2010){Bochanski}, {Hawley}, {Covey}, {West},
+  {Reid}, {Golimowski}, \& {Ivezi{\'c}}}]{Bochanski2010}
+{Bochanski}, J.~J., {Hawley}, S.~L., {Covey}, K.~R., {et~al.} 2010, \aj, 139,
+  2679
+
+\bibitem[{{Covey}(2006)}]{kev06}
+{Covey}, K.~R. 2006, PhD thesis, University of Washington, Washington, USA
+
+\bibitem[{{Covey} {et~al.}(2007){Covey}, {Ivezi{\'c}}, {Schlegel},
+  {Finkbeiner}, {Padmanabhan}, {Lupton}, {Ag{\"u}eros}, {Bochanski}, {Hawley},
+  {West}, {Seth}, {Kimball}, {Gogarten}, {Claire}, {Haggard}, {Kaib},
+  {Schneider}, \& {Sesar}}]{kev07}
+{Covey}, K.~R., {Ivezi{\'c}}, {\v Z}., {Schlegel}, D., {et~al.} 2007, \aj, 134,
+  2398
+
+\bibitem[{{Crotts} \& {Tomaney}(1996)}]{crotts1996}
+{Crotts}, A.~P.~S., \& {Tomaney}, A.~B. 1996, \apjl, 473, L87
+
+\bibitem[{{Cutri} {et~al.}(2003)}]{cutri03}
+{Cutri}, R.~M., {et~al.} 2003, {2MASS All Sky Catalog of point sources.} (The
+  IRSA 2MASS All-Sky Point Source Catalog, NASA/IPAC Infrared Science
+  Archive.~http://irsa.ipac.caltech.edu/applications/Gator/)
+
+\bibitem[{{Di Stefano}(2008)}]{mesolensing}
+{Di Stefano}, R. 2008, \apj, 684, 46
+
+\bibitem[{{Foreman-Mackey} {et~al.}(2012){Foreman-Mackey}, {Hogg}, {Lang}, \&
+  {Goodman}}]{dfm}
+{Foreman-Mackey}, D., {Hogg}, D.~W., {Lang}, D., \& {Goodman}, J. 2012, ArXiv
+  e-prints
+
+\bibitem[{{Fukui} {et~al.}(2007){Fukui}, {Abe}, {Ayani}, {Fujii}, {Iizuka},
+  {Itow}, {Kabumoto}, {Kamiya}, {Kawabata}, {Kawanomoto}, {Kinugasa}, {Koff},
+  {Krajci}, {Naito}, {Nogami}, {Narusawa}, {Ohishi}, {Ohnishi}, {Sumi}, \&
+  {Tsumuraya}}]{fukui2007}
+{Fukui}, A., {Abe}, F., {Ayani}, K., {et~al.} 2007, \apj, 670, 423
+
+\bibitem[{{Gaudi}(2011)}]{gaudi2011}
+{Gaudi}, B.~S. 2011, {Microlensing by Exoplanets}, ed. S.~{Piper}, 79--110
+
+\bibitem[{{Gaudi} {et~al.}(2008){Gaudi}, {Patterson}, {Spiegel}, {Krajci},
+  {Koff}, {Pojma{\'n}ski}, {Dong}, {Gould}, {Prieto}, {Blake}, {Roming},
+  {Bennett}, {Bloom}, {Boyd}, {Eyler}, {de Ponthi{\`e}re}, {Mirabal}, {Morgan},
+  {Remillard}, {Vanmunster}, {Wagner}, \& {Watson}}]{gaudi2008}
+{Gaudi}, B.~S., {Patterson}, J., {Spiegel}, D.~S., {et~al.} 2008, \apj, 677,
+  1268
+
+\bibitem[{{Goodman} \& {Weare}(2010)}]{goodman}
+{Goodman}, J., \& {Weare}, J. 2010, Comm. App. Math. Comp. Sci., 5
+
+\bibitem[{{Hamadache} {et~al.}(2006){Hamadache}, {Le Guillou}, {Tisserand},
+  {Afonso}, {Albert}, {Andersen}, {Ansari}, {Aubourg}, {Bareyre}, {Beaulieu},
+  {Charlot}, {Coutures}, {Ferlet}, {Fouqu{\'e}}, {Glicenstein}, {Goldman},
+  {Gould}, {Graff}, {Gros}, {Haissinski}, {de Kat}, {Lesquoy}, {Loup},
+  {Magneville}, {Marquette}, {Maurice}, {Maury}, {Milsztajn}, {Moniez},
+  {Palanque-Delabrouille}, {Perdereau}, {Rahal}, {Rich}, {Spiro},
+  {Vidal-Madjar}, {Vigroux}, \& {Zylberajch}}]{hamadache2009}
+{Hamadache}, C., {Le Guillou}, L., {Tisserand}, P., {et~al.} 2006, \aap, 454,
+  185
+
+\bibitem[{{Han}(2008)}]{han2008}
+{Han}, C. 2008, \apj, 681, 806
+
+\bibitem[{{Kowalski} {et~al.}(2010){Kowalski}, {Hawley}, {Holtzman},
+  {Wisniewski}, \& {Hilton}}]{kowal2010}
+{Kowalski}, A.~F., {Hawley}, S.~L., {Holtzman}, J.~A., {Wisniewski}, J.~P., \&
+  {Hilton}, E.~J. 2010, \apjl, 714, L98
+
+\bibitem[{{Law} {et~al.}(2009){Law}, {Kulkarni}, {Dekany}, {Ofek}, {Quimby},
+  {Nugent}, {Surace}, {Grillmair}, {Bloom}, {Kasliwal}, {Bildsten}, {Brown},
+  {Cenko}, {Ciardi}, {Croner}, {Djorgovski}, {van Eyken}, {Filippenko}, {Fox},
+  {Gal-Yam}, {Hale}, {Hamam}, {Helou}, {Henning}, {Howell}, {Jacobsen},
+  {Laher}, {Mattingly}, {McKenna}, {Pickles}, {Poznanski}, {Rahmer}, {Rau},
+  {Rosing}, {Shara}, {Smith}, {Starr}, {Sullivan}, {Velur}, {Walters}, \&
+  {Zolkower}}]{nick2009}
+{Law}, N.~M., {Kulkarni}, S.~R., {Dekany}, R.~G., {et~al.} 2009, \pasp, 121,
+  1395
+
+\bibitem[{{Law} {et~al.}(2010){Law}, {Dekany}, {Rahmer}, {Hale}, {Smith},
+  {Quimby}, {Ofek}, {Kasliwal}, {Zolkower}, {Velur}, {Henning}, {Bui},
+  {McKenna}, {Nugent}, {Jacobsen}, {Walters}, {Bloom}, {Surace}, {Grillmair},
+  {Laher}, {Mattingly}, \& {Kulkarni}}]{nick2010}
+{Law}, N.~M., {Dekany}, R.~G., {Rahmer}, G., {et~al.} 2010, in Society of
+  Photo-Optical Instrumentation Engineers (SPIE) Conference Series, Vol. 7735
+
+\bibitem[{{Levitan} {et~al.}(2011){Levitan}, {Fulton}, {Groot}, {Kulkarni},
+  {Ofek}, {Prince}, {Shporer}, {Bloom}, {Cenko}, {Kasliwal}, {Law}, {Nugent},
+  {Poznanski}, {Quimby}, {Horesh}, {Sesar}, \& {Sternberg}}]{levitan2011}
+{Levitan}, D., {Fulton}, B.~J., {Groot}, P.~J., {et~al.} 2011, \apj, 739, 68
+
+\bibitem[{{Nieto--Santisteban} {et~al.}(2004)}]{sdss_images}
+{Nieto--Santisteban}, M.~A., {et~al.} 2004, in ASP Conf. Ser. 314: Astronomical
+  Data Analysis Software and Systems (ADASS) XIII, ed. F.~{Ochsenbein}, M.~G.
+  {Allen}, \& D.~{Egret}, 666--+
+
+\bibitem[{{Ofek} {et~al.}(2012){Ofek}, {Laher}, {Surace}, {Levitan}, {Sesar},
+  {Horesh}, {Law}, {van Eyken}, {Kulkarni}, {Prince}, {Nugent}, {Sullivan},
+  {Yaron}, {Pickles}, {Ag{\"u}eros}, {Arcavi}, {Bildsten}, {Bloom}, {Cenko},
+  {Gal-Yam}, {Grillmair}, {Helou}, {Kasliwal}, {Poznanski}, \&
+  {Quimby}}]{ofek2012}
+{Ofek}, E.~O., {Laher}, R., {Surace}, J., {et~al.} 2012, \pasp, 124, 854
+
+\bibitem[{{Oke} \& {Gunn}(1982)}]{oke_gunn}
+{Oke}, J.~B., \& {Gunn}, J.~E. 1982, \pasp, 94, 586
+
+\bibitem[{{Os{\l}owski} {et~al.}(2008){Os{\l}owski}, {Moderski}, {Bulik}, \&
+  {Belczynski}}]{oslowski2008}
+{Os{\l}owski}, S., {Moderski}, R., {Bulik}, T., \& {Belczynski}, K. 2008, \aap,
+  478, 429
+
+\bibitem[{{Paczynski}(1986)}]{paczynski1986}
+{Paczynski}, B. 1986, \apj, 304, 1
+
+\bibitem[{{Pickles}(1998)}]{Pickles1998}
+{Pickles}, A.~J. 1998, \pasp, 110, 863
+
+\bibitem[{{Polishook} {et~al.}(2012){Polishook}, {Ofek}, {Waszczak},
+  {Kulkarni}, {Gal-Yam}, {Aharonson}, {Laher}, {Surace}, {Klein}, {Bloom},
+  {Brosch}, {Prialnik}, {Grillmair}, {Cenko}, {Kasliwal}, {Law}, {Levitan},
+  {Nugent}, {Poznanski}, \& {Quimby}}]{polishook2012}
+{Polishook}, D., {Ofek}, E.~O., {Waszczak}, A., {et~al.} 2012, \mnras, 421,
+  2094
+
+\bibitem[{{Rahmer} {et~al.}(2008){Rahmer}, {Smith}, {Velur}, {Hale}, {Law},
+  {Bui}, {Petrie}, \& {Dekany}}]{rahmer2008}
+{Rahmer}, G., {Smith}, R., {Velur}, V., {et~al.} 2008, in Society of
+  Photo-Optical Instrumentation Engineers (SPIE) Conference Series, Vol. 7014
+
+\bibitem[{{Rattenbury} {et~al.}(2005){Rattenbury}, {Abe}, {Bennett}, {Bond},
+  {Calitz}, {Claret}, {Cook}, {Furuta}, {Gal-Yam}, {Glicenstein}, {Hearnshaw},
+  {Hauschildt}, {Kilmartin}, {Kurata}, {Masuda}, {Maoz}, {Matsubara},
+  {Meintjes}, {Moniez}, {Muraki}, {Noda}, {Ofek}, {Okajima}, {Philpott},
+  {Rhie}, {Sako}, {Sullivan}, {Sumi}, {Terndrup}, {Tristram}, {Wood},
+  {Yanagisawa}, \& {Yock}}]{rattenbury2005}
+{Rattenbury}, N.~J., {Abe}, F., {Bennett}, D.~P., {et~al.} 2005, \aap, 439, 645
+
+\bibitem[{{Rau} {et~al.}(2009){Rau}, {Kulkarni}, {Law}, {Bloom}, {Ciardi},
+  {Djorgovski}, {Fox}, {Gal-Yam}, {Grillmair}, {Kasliwal}, {Nugent}, {Ofek},
+  {Quimby}, {Reach}, {Shara}, {Bildsten}, {Cenko}, {Drake}, {Filippenko},
+  {Helfand}, {Helou}, {Howell}, {Poznanski}, \& {Sullivan}}]{rau2009}
+{Rau}, A., {Kulkarni}, S.~R., {Law}, N.~M., {et~al.} 2009, \pasp, 121, 1334
+
+\bibitem[{{Richards} {et~al.}(2002)}]{richards02}
+{Richards}, G.~T., {et~al.} 2002, \aj, 123, 2945
+
+\bibitem[{{Sartore} \& {Treves}(2010)}]{sartore2010}
+{Sartore}, N., \& {Treves}, A. 2010, \aap, 523, A33
+
+\bibitem[{{Shin} {et~al.}(2009){Shin}, {Sekora}, \& {Byun}}]{shin2009}
+{Shin}, M.-S., {Sekora}, M., \& {Byun}, Y.-I. 2009, \mnras, 400, 1897
+
+\bibitem[{{Stetson}(1996)}]{stetson1996}
+{Stetson}, P.~B. 1996, \pasp, 108, 851
+
+\bibitem[{{Stoughton} {et~al.}(2002){Stoughton}, {Lupton}, {Bernardi},
+  {Blanton}, {Burles}, {Castander}, {Connolly}, {Eisenstein}, {Frieman},
+  {Hennessy}, {Hindsley}, {Ivezi{\'c}}, {Kent}, {Kunszt}, {Lee}, {Meiksin},
+  {Munn}, {Newberg}, {Nichol}, {Nicinski}, {Pier}, {Richards}, {Richmond},
+  {Schlegel}, {Smith}, {Strauss}, {SubbaRao}, {Szalay}, {Thakar}, {Tucker},
+  {Vanden Berk}, {Yanny}, {Adelman}, {Anderson}, {Anderson}, {Annis},
+  {Bahcall}, {Bakken}, {Bartelmann}, {Bastian}, {Bauer}, {Berman},
+  {B{\"o}hringer}, {Boroski}, {Bracker}, {Briegel}, {Briggs}, {Brinkmann},
+  {Brunner}, {Carey}, {Carr}, {Chen}, {Christian}, {Colestock}, {Crocker},
+  {Csabai}, {Czarapata}, {Dalcanton}, {Davidsen}, {Davis}, {Dehnen},
+  {Dodelson}, {Doi}, {Dombeck}, {Donahue}, {Ellman}, {Elms}, {Evans}, {Eyer},
+  {Fan}, {Federwitz}, {Friedman}, {Fukugita}, {Gal}, {Gillespie}, {Glazebrook},
+  {Gray}, {Grebel}, {Greenawalt}, {Greene}, {Gunn}, {de Haas}, {Haiman},
+  {Haldeman}, {Hall}, {Hamabe}, {Hansen}, {Harris}, {Harris}, {Harvanek},
+  {Hawley}, {Hayes}, {Heckman}, {Helmi}, {Henden}, {Hogan}, {Hogg}, {Holmgren},
+  {Holtzman}, {Huang}, {Hull}, {Ichikawa}, {Ichikawa}, {Johnston}, {Kauffmann},
+  {Kim}, {Kimball}, {Kinney}, {Klaene}, {Kleinman}, {Klypin}, {Knapp},
+  {Korienek}, {Krolik}, {Kron}, {Krzesi{\'n}ski}, {Lamb}, {Leger},
+  {Limmongkol}, {Lindenmeyer}, {Long}, {Loomis}, {Loveday}, {MacKinnon},
+  {Mannery}, {Mantsch}, {Margon}, {McGehee}, {McKay}, {McLean}, {Menou},
+  {Merelli}, {Mo}, {Monet}, {Nakamura}, {Narayanan}, {Nash}, {Neilsen},
+  {Newman}, {Nitta}, {Odenkirchen}, {Okada}, {Okamura}, {Ostriker}, {Owen},
+  {Pauls}, {Peoples}, {Peterson}, {Petravick}, {Pope}, {Pordes}, {Postman},
+  {Prosapio}, {Quinn}, {Rechenmacher}, {Rivetta}, {Rix}, {Rockosi}, {Rosner},
+  {Ruthmansdorfer}, {Sandford}, {Schneider}, {Scranton}, {Sekiguchi}, {Sergey},
+  {Sheth}, {Shimasaku}, {Smee}, {Snedden}, {Stebbins}, {Stubbs}, {Szapudi},
+  {Szkody}, {Szokoly}, {Tabachnik}, {Tsvetanov}, {Uomoto}, {Vogeley}, {Voges},
+  {Waddell}, {Walterbos}, {Wang}, {Watanabe}, {Weinberg}, {White}, {White},
+  {Wilhite}, {Wolfe}, {Yasuda}, {York}, {Zehavi}, \& {Zheng}}]{stoughton02}
+{Stoughton}, C., {Lupton}, R.~H., {Bernardi}, M., {et~al.} 2002, \aj, 123, 485
+
+\bibitem[{{Sumi} {et~al.}(2011){Sumi}, {Kamiya}, {Bennett}, {Bond}, {Abe},
+  {Botzler}, {Fukui}, {Furusawa}, {Hearnshaw}, {Itow}, {Kilmartin}, {Korpela},
+  {Lin}, {Ling}, {Masuda}, {Matsubara}, {Miyake}, {Motomura}, {Muraki},
+  {Nagaya}, {Nakamura}, {Ohnishi}, {Okumura}, {Perrott}, {Rattenbury}, {Saito},
+  {Sako}, {Sullivan}, {Sweatman}, {Tristram}, {Udalski}, {Szyma{\'n}ski},
+  {Kubiak}, {Pietrzy{\'n}ski}, {Poleski}, {Soszy{\'n}ski}, {Wyrzykowski},
+  {Ulaczyk}, \& {Microlensing Observations in Astrophysics (MOA)
+  Collaboration}}]{sumi2011}
+{Sumi}, T., {Kamiya}, K., {Bennett}, D.~P., {et~al.} 2011, \nat, 473, 349
+
+\bibitem[{{Tomov} {et~al.}(2013){Tomov}, {Tomova}, \& {Bisikalo}}]{tomov2013}
+{Tomov}, N.~A., {Tomova}, M.~T., \& {Bisikalo}, D.~V. 2013, ArXiv e-prints
+
+\bibitem[{{Udalski}(2003)}]{udalski03}
+{Udalski}, A. 2003, \actaa, 53, 291
+
+\bibitem[{{Udalski} {et~al.}(1992){Udalski}, {Szymanski}, {Kaluzny}, {Kubiak},
+  \& {Mateo}}]{original_ogle}
+{Udalski}, A., {Szymanski}, M., {Kaluzny}, J., {Kubiak}, M., \& {Mateo}, M.
+  1992, \actaa, 42, 253
+
+\bibitem[{{Udalski} {et~al.}(1994){Udalski}, {Szymanski}, {Stanek}, {Kaluzny},
+  {Kubiak}, {Mateo}, {Krzeminski}, {Paczynski}, \&
+  {Venkat}}]{ogle_optical_depth}
+{Udalski}, A., {Szymanski}, M., {Stanek}, K.~Z., {et~al.} 1994, \actaa, 44, 165
+
+\bibitem[{{von Neumann} {et~al.}(1941){von Neumann}, {Kent}, {Bellinson}, \&
+  {Hart}}]{von_neumann1941}
+{von Neumann}, J., {Kent}, R.~H., {Bellinson}, H.~R., \& {Hart}, B.~I. 1941,
+  The Annals of Mathematical Statistics, 12, 153
+
+\bibitem[{{Wozniak}(2000)}]{con_idx}
+{Wozniak}, P.~R. 2000, \actaa, 50, 421
+
+\bibitem[{{Wright} {et~al.}(2010){Wright}, {Eisenhardt}, {Mainzer}, {Ressler},
+  {Cutri}, {Jarrett}, {Kirkpatrick}, {Padgett}, {McMillan}, {Skrutskie},
+  {Stanford}, {Cohen}, {Walker}, {Mather}, {Leisawitz}, {Gautier}, {McLean},
+  {Benford}, {Lonsdale}, {Blain}, {Mendez}, {Irace}, {Duval}, {Liu}, {Royer},
+  {Heinrichsen}, {Howard}, {Shannon}, {Kendall}, {Walsh}, {Larsen}, {Cardon},
+  {Schick}, {Schwalm}, {Abid}, {Fabinsky}, {Naes}, \& {Tsai}}]{WISE}
+{Wright}, E.~L., {Eisenhardt}, P.~R.~M., {Mainzer}, A.~K., {et~al.} 2010, \aj,
+  140, 1868
+
+\bibitem[{{Wyrzykowski} {et~al.}(2009){Wyrzykowski}, {Koz{\l}owski}, {Skowron},
+  {Belokurov}, {Smith}, {Udalski}, {Szyma{\'n}ski}, {Kubiak},
+  {Pietrzy{\'n}ski}, {Soszy{\'n}ski}, {Szewczyk}, \&
+  {{\.Z}ebru{\'n}}}]{wyrzykowski2009}
+{Wyrzykowski}, {\L}., {Koz{\l}owski}, S., {Skowron}, J., {et~al.} 2009, \mnras,
+  397, 1228
+
+\bibitem[{Yan {et~al.}(2013)Yan, Donoso, Tsai, Stern, Assef, Eisenhardt, Blain,
+  Cutri, Jarrett, Stanford, Wright, Bridge, \& Riechers}]{yan2013}
+Yan, L., Donoso, E., Tsai, C.-W., {et~al.} 2013, \aj, 145, 55
+
+\bibitem[{{York} {et~al.}(2000)}]{york00}
+{York}, D.~G., {et~al.} 2000, \aj, 120, 1579
+
+\end{thebibliography}
 

### 2013-10-01 text/final_paper/ms.tex
added reference to abstract, added url footnote

@@
 We develop a selection method using this statistic and apply it to data from all PTF fields with $>$10 $R$-band observations --- a total of $1.1\times10^9$ light curves --- to uncover three candidate sparse-field microlensing events.
-While these candidates cannot be confirmed, as we lack simultaneous, multi-color photometry of the events, this number is consistent with simple predictions for the microlensing event rate in the PTF footprint over the survey's three years of operations.
+While these candidates cannot be confirmed, as we lack simultaneous, multi-color photometry of the events, this number is consistent with predictions for the microlensing event rate in the PTF footprint over the survey's three years of operations, as estimated from the model proposed in \citet{han2008}.
 This work can help constrain all-sky event rate predictions and tests microlensing signal recovery in large data sets, both of which will be useful to future wide-field, time-domain surveys, such as that planned with the Large Synoptic Survey Telescope.
@@
 Over the past 20 years, microlensing, in which gravitational lensing causes a transient increase in the flux from a background point source, has been used to search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, to study Galactic structure and kinematics \citep{binney2000}, to determine the shape of stars \citep{rattenbury2005}, and to identify extrasolar planets \citep[][and references therein]{gaudi2011}.
-Such studies were once limited by the small number of detected events, but thanks to advances in CCD technology and the development of dedicated microlensing surveys, a few thousand events are now observed each year.
-In the coming decade, this number is expected to increase as the next generation of photometric, time-domain surveys comes online, providing an opportunity for the precise study of otherwise hard-to-characterize objects, such as low-mass stars, sub-stellar objects, and isolated neutron stars.
+Such studies were once limited by the small number of detected events, but thanks to advances in CCD technology and the development of dedicated microlensing surveys, a few thousand events are now observed each year.\footnote{http://ogle.astrouw.edu.pl/ogle4/ews/ews.html} In the coming decade, this number is expected to increase as the next generation of photometric, time-domain surveys comes online, providing an opportunity for the precise study of otherwise hard-to-characterize objects, such as low-mass stars, sub-stellar objects, and isolated neutron stars.
 

### 2013-10-01 text/final_paper/ms.tex
add reference to han 2008

@@
 Such an analysis is usually not possible for events in high-density fields where the lenses are, on average, much farther away and blending makes photometry --- and, therefore, mass measurements --- challenging.
-By contrast, high-Galactic-latitude events probe lenses with distances $\lesssim$1~kpc; the Einstein radii, therefore, tend to be larger ($\sim$milliarcseconds) and may cause detectable astrometric signatures.
+By contrast, high-Galactic-latitude events probe lenses with distances $\lesssim$1~kpc; the Einstein radii, therefore, tend to be larger ($\sim$milliarcseconds) and may cause detectable astrometric signatures \citep[][]{han2008}.
 A measurement of the astrometric shift due to microlensing will break the distance-mass degeneracy in the lensing parameters and enable a direct measurement of the lens mass.

### 2013-10-01 text/final_paper/ms.tex
removed reference to in prep work
removed footnote to bitly

@@
 We discuss eight interesting transients and three candidate events identified in this manner before concluding in Section~\ref{sec:conc}.
-A detailed investigation of the all-sky microlensing event rate,  generated from a model that includes realistic mass and velocity distributions, will be presented in Fournier et al.
+% A detailed investigation of the all-sky microlensing event rate,  generated from a model that includes realistic mass and velocity distributions, will be presented in Fournier et al.
 (in prep).
@@
 As of 2012 Dec, the PTF footprint includes $\sim$15,224 (2766) deg$^2$ imaged $>$10 ($>$100) times in $R$-band and $\sim$5430 (290) deg$^2$ imaged that often in $g$-band (see Figure~\ref{fig:survey_footprint}).
-The PTF survey footprint is not uniformly sampled either spatially or temporally.\footnote{For an interactive visualization of the PTF survey footprint, see \url{http://bit.ly/PTFSky}.} Each field has a unique sampling pattern determined by which of the PTF sub-surveys it belongs to, what time of the year it is visible, and how high a priority it is given by the scheduler.
+The PTF survey footprint is not uniformly sampled either spatially or temporally.
+Each field has a unique sampling pattern determined by which of the PTF sub-surveys it belongs to, what time of the year it is visible, and how high a priority it is given by the scheduler.
 The imaging cadences range from 1-5 days, but there is a higher-cadence campaign of a field in Orion to find transiting planets around young stars \citep{nick2009}.

### 2013-10-01 text/final_paper/ms.tex
added note about units of u0, moved discussion of PTF light curves to sec 3

@@
 Figure~\ref{fig:sampling} shows six randomly selected light curves and illustrates the varying cadences and coverage that different fields may have over the same one-year period.
+The PTF data is not publicly available, but ...
+TODO
 
@@
 
+\section{Microlensing event recovery} \label{sec:event_recovery}
+Microlensing surveys typically use difference image analysis \citep{alard1998} to identify transient events in raw imaging data.
+The light curves of transient sources are then analyzed and vetted using a variety of selection methods to search for microlensing event candidates and distinguish them from, e.g., variable stars, outbursting systems, and novae.
+Surveys have approached this process differently (as described in \citealt{ogle_optical_depth, con_idx, alcock2000, macho_detection_efficiency, udalski03,  hamadache2009, wyrzykowski2009, sumi2011}), but the general idea is to require that:
+\begin{enumerate}
+	\item any selected light curve has some number of consecutive data points brighter than some threshold,
+	\item compared to flat or linear light-curve models, a microlensing model best describe the data, as determined by using a $\Delta\chi^2$ test, and
+	\item the microlensing model event parameters have physically reasonable values.
+\end{enumerate}
+
 What would a microlensing event look like in a typical PTF light curve?
 If the source is an unblended point source and the lens is a foreground, dim object, a microlensing event is fully described by three parameters: the angular impact parameter $u_0$, the peak time of the event $t_0$, and the timescale of the event (Einstein crossing time) $t_E$.
-In terms of a dimensionless projected distance between the source and lens, $u$, the amplification factor $A$ and flux $F$ as a function of time can be defined as:
+In terms of a dimensionless projected distance between the source and lens (in units of Einstein radius), $u$, the amplification factor $A$ and flux $F$ as a function of time can be defined as:
 \begin{align}
@@
 
-\section{Microlensing event recovery} \label{sec:event_recovery}
-Microlensing surveys typically use difference image analysis \citep{alard1998} to identify transient events in raw imaging data.
-The light curves of transient sources are then analyzed and vetted using a variety of selection methods to search for microlensing event candidates and distinguish them from, e.g., variable stars, outbursting systems, and novae.
-Surveys have approached this process differently (as described in \citealt{ogle_optical_depth, con_idx, alcock2000, macho_detection_efficiency, udalski03,  hamadache2009, wyrzykowski2009, sumi2011}), but the general idea is to require that:
-\begin{enumerate}
-	\item any selected light curve has some number of consecutive data points brighter than some threshold,
-	\item compared to flat or linear light-curve models, a microlensing model best describe the data, as determined by using a $\Delta\chi^2$ test, and
-	\item the microlensing model event parameters have physically reasonable values.
-\end{enumerate}
-
-Applying this prescription to the PTF data presents obvious challenges.
+Applying the above prescription to the PTF data presents obvious challenges.
 For example, microlensing surveys have relatively uniform time sampling of their survey footprint over an observing season, justifying the first requirement.

### 2013-10-01 text/final_paper/ms.tex
added durbin-watson thing
removed mention of delta chi squared before defining it

@@
 	\item any selected light curve has some number of consecutive data points brighter than some threshold,
-	\item compared to flat or linear light-curve models, a microlensing model best describe the data, as determined by using a $\Delta\chi^2$ test, and
+	\item compared to flat or linear light-curve models, a microlensing model best describe the data, and
 	\item the microlensing model event parameters have physically reasonable values.
@@
 
-$\eta$, the von Neumann ratio \citep{von_neumann1941}, is the mean square successive difference divided by the sample variance:
+$\eta$, the von Neumann ratio \citep[][also known as the Durbin-Watson statistic]{von_neumann1941}, is the mean square successive difference divided by the sample variance:
 \begin{align}

### 2013-10-01 text/final_paper/ms.tex
added detail about delta chi squared and AIC BIC etc

@@
 We add one more index, $\Delta \chi^2$, the difference in $\chi^2$ between fitting a Gaussian model and fitting a linear model to a light curve.
+In detail, we use a Levenberg-Marquardt optimizer to perform a least-squares fit with each of these models to the light curves and then compute $\Delta\chi^2=\chi_{linear} - \chi_{gaussian}$.
+Out tests with $\Delta \chi^2$ below compare the distribution of values over light curves on the same chip --- because of this, we neglect including an AIC or BIC correction to make this a proper likelihood test as the number of data points and number of model parameters are constant.
+For the Gaussian fit, the optimizer is initialized with a standard-deviation of 10 days, centered on the brightest data point.
 This is the standard statistical test used by microlensing surveys, and allows us to compare the relative performance of the (slightly modified) \cite{shin2009} indices and of this approach.

### 2013-10-01 scripts/event_fitter.py
seed the mcmc so produces same links each time

@@ -121,6 +121,8 @@ def ln_p_t0(t0, mjd, mag, t0_sigma=25.):
 
 def ln_p_tE(tE, tE_sigma=1.3):
     """ Prior on duration of event tE """
+    if tE < 0:
+        return -np.inf
     return -0.5 * (np.log(2.*np.pi*tE_sigma) + (np.log(tE) - np.log(30.))**2 / tE_sigma**2) # Log-normal centered at 30 days
 
 def ln_p_m0(m0, median_mag, mag_rms):
@@ -179,6 +181,7 @@ def plot_priors(light_curve, ln=False):
     plt.savefig("plots/test.png")
 
 def ln_posterior(p, mag, mjd, sigma):
+    #print p, ln_prior(p, mag, mjd), ln_likelihood(p, mag, mjd, sigma)
     return ln_prior(p, mag, mjd) + ln_likelihood(p, mag, mjd, sigma)
 
 def test_ln_prior():
@@ -203,7 +206,7 @@ def test_ln_prior():
     print "Bad tE:", ln_prior([15, 0.3, 50., 0.2], light_curve.mag, light_curve.mjd)
     print "Bad tE:", ln_prior([15, 0.3, 50., 2000.], light_curve.mag, light_curve.mjd)
 
-def fit_model_to_light_curve(light_curve, nwalkers=200, nburn_in=100, nsamples=250):
+def fit_model_to_light_curve(light_curve, nwalkers=200, nburn_in=100, nsamples=250, seed=None):
     """ Fit a microlensing model to a given light curve using Emcee
 
     """
@@ -221,7 +224,7 @@ def fit_model_to_light_curve(light_curve, nwalkers=200, nburn_in=100, nsamples=2
     sampler = emcee.EnsembleSampler(nwalkers, ndim, ln_posterior,
                                     args=[light_curve.mag, light_curve.mjd, light_curve.error],
                                     threads=4)
-    pos, prob, state = sampler.run_mcmc(p0, nburn_in)
+    pos, prob, state = sampler.run_mcmc(p0, nburn_in, rstate0=seed)
 
     sampler.reset()
     pos, prob, state = sampler.run_mcmc(pos, nsamples)
@@ -233,7 +236,7 @@ def fit_model_to_light_curve(light_curve, nwalkers=200, nburn_in=100, nsamples=2
[…]

### 2013-10-01 scripts/paper_figures.py
new figures

@@ -16,7 +16,7 @@ import time
 
 # Third-party
 import matplotlib
-matplotlib.use("Agg")
+matplotlib.use("TkAgg")
 import matplotlib.cm as cm
 import matplotlib.colors as mc
 import matplotlib.pyplot as plt
@@ -47,6 +47,18 @@ except ImportError:
     logger.warning("photometric database modules failed to load! If this is on Navtara, you made a boo-boo.")
 
 tick_font_size = 14
+
+ptfid_to_longname = {"PTFS1203am" : "PTF1J033545.80-041849.1",
+                     "PTFS1206i" : "PTF1J061800.25+203142.5",
+                     "PTFS1212ak" : "PTF1J120642.60-192016.4", 
+                     "PTFS1213ap" : "PTF1J131500.09+715032.5",
+                     "PTFS1215aj" : "PTF1J153202.91+674825.1",
+                     "PTFS1216bt" : "PTF1J161502.39+540053.8",
+                     "PTFS1217ce" : "PTF1J171606.82+474423.7",
+                     "PTFS1217cv" : "PTF1J172826.08+692501.1",
+                     "PTFS1217df" : "PTF1J173301.07+374311.9",
+                     "PTFS1217dk" : "PTF1J174736.51+300506.9",
+                     "PTFS1219ad" : "PTF1J193330.11+451501.1"}
 
 def make_survey_sampling_figure(N=10):
     """ Generate the survey sampling figure that shows how different the
@@ -694,24 +706,307 @@ def ml_parameter_distributions(overwrite=False):
 
     fig.savefig("plots/fit_events/all_parameters.png")
 
-def fit_candidates():
-    candidates = np.genfromtxt("data/candidate_list.txt", names=True, dtype=int)
+def lightcurve_from_filename(filename):
+    import astropy.io.ascii as ascii
+    
+    with open(filename) as f:
+        data_start = 0
+        for line in f:
+            if line.startswith("#"):
+                data_start += 1
+    
+    candidate = ascii.read(filename, header_start=data_start-1)#, data_start=data_start)
+    c = candidate[(candidate['filterID'] == 2) & (candidate['mag'] > 0.)]
+    flag_idx = c['relPhotFlags']==0
+    
+    light_curve = PTFLightCurve(mjd=c['epoch'], mag=c['mag'], error=c['magerr'])
+    
+    return light_curve
[…]

### 2013-10-01 text/paper/ms.tex
final version of paper in dir it should be in

@@
 While the median $7.26~\mathrm{deg}^2$ PTF field has been imaged $\sim$40 times in \textit{R}-band, $\sim$$2300~\mathrm{deg}^2$ have been observed more than 100 times.
-We use the existing PTF data to study the trade-off between searching for microlensing events in a survey whose footprint is much larger than that of typical microlensing searches, but that has far-from-optimal sampling.
+We use the existing PTF data to study the trade-off between searching for microlensing events in a survey whose footprint is much larger than that of typical microlensing searches, but that has far-from-optimal time sampling.
 To examine the probability that microlensing events can be recovered in these data, we test previous statistics used on uniformly sampled data to identify variables and transients.
-We find that one such statistic, the von Neumann ratio, performs best for identifying simulated microlensing events.
-We develop a selection method using this statistic and apply it to data from all PTF fields with $>$10 $R$-band observations, a total of $1.1\times10^9$ light curves, to uncover three candidate sparse-field microlensing events.
+We find that one such statistic, the von Neumann ratio, performs best for identifying simulated microlensing events in our data.
+We develop a selection method using this statistic and apply it to data from all PTF fields with $>$10 $R$-band observations --- a total of $1.1\times10^9$ light curves --- to uncover three candidate sparse-field microlensing events.
+While these candidates cannot be confirmed, as we lack simultaneous, multi-color photometry of the events, this number is consistent with predictions for the microlensing event rate in the PTF footprint over the survey's three years of operations, as estimated from the model proposed in \citet{han2008}.
 This work can help constrain all-sky event rate predictions and tests microlensing signal recovery in large data sets, both of which will be useful to future wide-field, time-domain surveys, such as that planned with the Large Synoptic Survey Telescope.
-
-% Second to last sentence will change when I run on all fields with >25 observations
 \end{abstract}
@@
 \section{Introduction}
-Over the past 20 years, microlensing, in which gravitational lensing causes a transient increase in the flux from a background point source, has been used to search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, to study Galactic structure and kinematics \citep{binney2000}, to determine the shape of stars \citep{rattenbury2005}, and, most recently, to identify extrasolar planets \citep[][and references therein]{gaudi2011}.
-Such studies were once limited by the small number of detected events, but thanks to advances in CCD technology and the development of dedicated microlensing surveys, a few thousand events are now observed each year.
-In the coming decade, this number is expected to increase as the next generation of photometric, time-domain surveys comes online, providing an opportunity for the precise study of otherwise hard-to-characterize objects, such as low-mass stars, sub-stellar objects, and isolated neutron stars.
+Over the past 20 years, microlensing, in which gravitational lensing causes a transient increase in the flux from a background point source, has been used to search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, to study Galactic structure and kinematics \citep{binney2000}, to determine the shape of stars \citep{rattenbury2005}, and to identify extrasolar planets \citep[][and references therein]{gaudi2011}.
+Such studies were once limited by the small number of detected events, but thanks to advances in CCD technology and the development of dedicated microlensing surveys, a few thousand events are now observed each year.\footnote{http://ogle.astrouw.edu.pl/ogle4/ews/ews.html} In the coming decade, this number is expected to increase as the next generation of photometric, time-domain surveys comes online, providing an opportunity for the precise study of otherwise hard-to-characterize objects, such as low-mass stars, sub-stellar objects, and isolated neutron stars.
 
-%The microlensing event rate for a given region of the sky depends on the space and velocity distributions of the sources and on the space, velocity, and mass distribution of the lenses.
-To maximize the event rate, microlensing surveys typically focus on high-density stellar regions, such as the Galactic Bulge, M31, or the Magellanic Clouds \citep[e.g.][]{original_ogle, original_macho, eros_original, crotts1996}.
-Of course, microlensing events are not limited to dense stellar fields.
+Microlensing surveys typically focus on high-density stellar regions, such as the Galactic Bulge, M31, or the Magellanic Clouds \citep[e.g.,][]{original_ogle, original_macho, eros_original, crotts1996}, but microlensing events are not limited to dense stellar fields.
 Indeed, lensing events away from these fields are potentially very interesting.
-While the event rate in low-density stellar regions will certainly be lower than that near, e.g., the Bulge \citep[][]{ogle_optical_depth, macho_optical_depth, wood_optical_depth, eros_optical_depth}, these events are expected to involve closer sources and lenses, and less crowded backgrounds, compared to those found by modern microlensing surveys \citep{mesolensing}.
+While the microlensing event rate in low-density stellar regions will be smaller, compared to events found by modern microlensing surveys, these events are expected to involve closer sources and lenses, and less crowded backgrounds \citep{mesolensing}.
 Thus, such lenses offer more opportunities to constrain the properties of the events.
@@
 Such an analysis is usually not possible for events in high-density fields where the lenses are, on average, much farther away and blending makes photometry --- and, therefore, mass measurements --- challenging.
-By contrast, high-Galactic-latitude events probe lenses with distances $\lesssim$1~kpc; the Einstein radii, therefore, tend to be larger (a few milliarcseconds) and may cause detectable astrometric signatures.
+By contrast, high-Galactic-latitude events probe lenses with distances $\lesssim$1~kpc; the Einstein radii, therefore, tend to be larger ($\sim$milliarcseconds) and may cause detectable astrometric signatures \citep[][]{han2008}.
 A measurement of the astrometric shift due to microlensing will break the distance-mass degeneracy in the lensing parameters and enable a direct measurement of the lens mass.
@@
 The resulting archive is an excellent resource for the study of periodic forms of variability \citep[e.g., stellar or asteroid rotation;][]{agueros11,polishook2012}.
-%A separate pipeline identifies transients of interest and passes these to a dedicated photometric follow-up telescope; this has proved an excellent tool for e.g., early detection of supernovae \citep{nugent2011}.
 
-%In addition, all detected sources are archived\footnote{The first photometric catalog of non-variable sources drawn from this archive is described in \citet[][]{ofek2012}.};
+As of 2012 Dec, the PTF footprint includes $\sim$15,224 (2766) deg$^2$ imaged $>$10 ($>$100) times in $R$-band and $\sim$5430 (290) deg$^2$ imaged that often in $g$-band (see Figure~\ref{fig:survey_footprint}).
+The PTF survey footprint is not uniformly sampled either spatially or temporally.
+Each field has a unique sampling pattern determined by which of the PTF sub-surveys it belongs to, what time of the year it is visible, and how high a priority it is given by the scheduler.
+The imaging cadences range from 1-5 days, but there is a higher-cadence campaign of a field in Orion to find transiting planets around young stars \citep{nick2009}.
+The result is that PTF light curves often contain gaps and regions of high-cadence observations and/or of low-cadence observations, so that the archive is a massive dataset of irregularly-sampled, time-domain photometry.
+Figure~\ref{fig:sampling} shows six randomly selected light curves and illustrates the varying cadences and coverage that different fields may have over the same one-year period.
+The PTF data is not publicly available, but ...
+TODO
 
@@
 \begin{center}
-\includegraphics[width=0.47\textwidth, trim=2 2 5 5, clip]{figures/sampling_figure.pdf}
+\includegraphics[width=0.47\textwidth, trim=2 2 5 5, clip]{fig2.eps}
 \caption{Sampling of six different PTF fields over the course of a year.
@@
 \begin{figure}[t]
-\centering\includegraphics[width=0.5\textwidth, trim=21 20 25 70, clip]{figures/num_observations_baseline_90deg.pdf}
+\centering\includegraphics[width=0.5\textwidth, trim=21 20 25 70, clip]{fig3.eps}
 \caption{Total number of $R$-band exposures and baselines (number of days between first and last exposure) for all PTF fields.
@@
 \begin{figure}[!h]
-\centering\includegraphics[width=0.48\textwidth, trim=5 5 5 7, clip]{figures/simulated_events.pdf}
+\centering\includegraphics[width=0.48\textwidth, trim=5 0 0 7, clip]{fig4.eps}
 \caption{Simulating the effect of $20$-day microlensing events with impact parameters $u_0 = 1.0$, 0.5, and 0.01, on a random PTF $R$-band light curve.
@@
 
-As of 2012 Dec, the PTF footprint includes $\sim$15,224 (2766) deg$^2$ imaged $>$10 ($>$100) times in $R$-band and $\sim$5430 (290) deg$^2$ imaged that often in $g$-band (see Figure~\ref{fig:survey_footprint}).
-The PTF survey footprint is not uniformly sampled either spatially or temporally.\footnote{For an interactive visualization of the PTF survey footprint, see \url{http://bit.ly/PTFSky}.} Each field has a unique sampling pattern determined by which of the PTF sub-surveys it belongs to, what time of the year it is visible, and how high a priority it is given by the scheduler.
-The imaging cadences range from 1-5 days, but there is a higher-cadence campaign of a field in Orion to find transiting planets around young stars \citep{nick2009}.
-The result is that PTF light curves often contain gaps and regions of high-cadence observations and/or of low-cadence observations, so that the archive is a massive dataset of irregularly-sampled, time-domain photometry.
-Figure~\ref{fig:sampling} shows six randomly selected light curves and illustrates the varying cadences and coverage that different fields may have over the same one-year period.
-
 In Figure~\ref{fig:num_obs}, we plot the number of $R$-band exposures for each field against the observational baseline (the number of days between the first and last exposure).
 Given the exposure distribution, which peaks at 30-40 but includes a long tail to larger numbers, we restrict our sample to the light curves with $>$10 \textit{R}-band observations.
-% because there is a drop off in the number of fields with fewer than 10 observations and we don't expect such light curves to be useful for our search.
-%APW!
+
+\section{Microlensing event recovery} \label{sec:event_recovery}
+Microlensing surveys typically use difference image analysis \citep{alard1998} to identify transient events in raw imaging data.
+The light curves of transient sources are then analyzed and vetted using a variety of selection methods to search for microlensing event candidates and distinguish them from, e.g., variable stars, outbursting systems, and novae.
+Surveys have approached this process differently (as described in \citealt{ogle_optical_depth, con_idx, alcock2000, macho_detection_efficiency, udalski03,  hamadache2009, wyrzykowski2009, sumi2011}), but the general idea is to require that:
+\begin{enumerate}
+	\item any selected light curve has some number of consecutive data points brighter than some threshold,
+	\item compared to flat or linear light-curve models, a microlensing model best describe the data, and
+	\item the microlensing model event parameters have physically reasonable values.
+\end{enumerate}
 
@@
 If the source is an unblended point source and the lens is a foreground, dim object, a microlensing event is fully described by three parameters: the angular impact parameter $u_0$, the peak time of the event $t_0$, and the timescale of the event (Einstein crossing time) $t_E$.
-In terms of a dimensionless projected distance between the source and lens, $u$, the amplification factor $A$ and flux $F$ as a function of time can be defined as:
+In terms of a dimensionless projected distance between the source and lens (in units of Einstein radius), $u$, the amplification factor $A$ and flux $F$ as a function of time can be defined as:
 \begin{align}
@@
 
-\section{Microlensing event recovery} \label{sec:event_recovery}
-Microlensing surveys typically use difference image analysis \citep{alard1998} to identify transient events in raw imaging data.
-The light curves of transient sources are then analyzed and vetted using a variety of selection methods to search for microlensing event candidates and distinguish them from, e.g., variable stars, outbursting systems, and novae.
-Surveys have approached this process differently (as described in \citealt{ogle_optical_depth, con_idx, alcock2000, macho_detection_efficiency, udalski03,  hamadache2009, wyrzykowski2009, sumi2011}), but the general idea is to require that:
-\begin{enumerate}
-	\item any selected light curve has some number of consecutive data points brighter than some threshold,
-	\item compared to flat or linear light-curve models, a microlensing model best describe the data, as determined by using a $\Delta\chi^2$ test, and
-	\item the microlensing model event parameters have physically reasonable values.
-\end{enumerate}
-%Difference imaging is ideal for studying crowded fields where photometric measurements are difficult.
-It also limits the number of light curves that must be fit with a model, thus saving on computation time and allowing for real-time detection of events, so that high-cadence follow-up observations of interesting sources can be triggered.
-
-Applying this prescription to the PTF data presents obvious challenges.
+Applying the above prescription to the PTF data presents obvious challenges.
 For example, microlensing surveys have relatively uniform time sampling of their survey footprint over an observing season, justifying the first requirement.
@@
 Finally, we add 100 different simulated microlensing events to each light curve and evaluate how often these events are recovered given the 1\% FPR cut defined above.
+
+\begin{figure}
+\centering
+	\subfigure{
+	   \includegraphics[width=.45\textwidth, trim=0 8 6 3, clip] {fig7_a.eps}
+	   \label{fig:subfig1}
+	 }
+
+	 \subfigure{
+	   \includegraphics[width=.46\textwidth, trim=8 8 5 5, clip] {fig7_b.eps}
+	   \label{fig:subfig2}
+	 }
+\caption{{\it Top:} A randomly selected light curve from PTF field 4327.
+Note the sampling pattern.
+{\it Bottom:} Detection efficiency $\varepsilon$ for $\eta$, $\Delta\chi^2$, and $J$, as a function of the simulated microlensing event parameters $t_E$, $u_0$, and $m_0$ (the event timescale, impact parameter, and `quiescent' source magnitude, respectively).
+Black (dashed) lines show the distributions for all light curves (normalized, so scale of y-axis is arbitrary), red (solid) lines show the recovered distributions.}  \label{fig:detection_efficiency_4327}
+\end{figure}
 

### 2013-10-01 scripts/candidate_info.py
adding stuff that should be in repo

@@ -1,0 +1,47 @@
+import sys
+import numpy as np
+from astropy.io import ascii
+
+candidates = ascii.read("data/candidate_list_ipac.tbl.txt")
+
+ptf_name = sys.argv[1]
+
+row = np.array(candidates[candidates["object"] == ptf_name])
+
+latex_row = ""
+
+try:
+    print "2MASS"
+    print "-----"
+    print "J = {0} +/- {1}, Quality: {2}".format(row["j_m_2mass"][0], row["j_msig_2mass"][0], row["ph_qual_2mass"][0][0])
+    print "H = {0} +/- {1}, Quality: {2}".format(row["h_m_2mass"][0], row["h_msig_2mass"][0], row["ph_qual_2mass"][0][1])
+    print "K = {0} +/- {1}, Quality: {2}".format(row["k_m_2mass"][0], row["k_msig_2mass"][0], row["ph_qual_2mass"][0][2])
+    print "J-K = {0} +/- {1}".format(row["j_m_2mass"][0]-row["k_m_2mass"][0], np.sqrt(row["j_msig_2mass"][0]**2+row["k_msig_2mass"][0]**2))
+    
+    latex_row += "{0:.3f}$\pm${1:.3f} & {2:.3f}$\pm${3:.3f} & {4:.3f}$\pm${5:.3f}".format(row["j_m_2mass"][0], row["j_msig_2mass"][0],
+                                                                                    row["h_m_2mass"][0], row["h_msig_2mass"][0],
+                                                                                    row["k_m_2mass"][0], row["k_msig_2mass"][0])
+except IndexError:
+    pass
+
+try:
+    print "-----"
+    print "WISE"
+    print "-----"
+    print "W1 = {0} +/- {1}, Quality: {2[0]}, Detection? {3[0]}".format(row["w1mpro"][0], row["w1sigmpro"][0], row["ph_qual"][0], "{0:0<4d}".format(int(str(bin(row["det_bit"][0]))[2:])))
+    print "W2 = {0} +/- {1}, Quality: {2[1]}, Detection? {3[1]}".format(row["w2mpro"][0], row["w2sigmpro"][0], row["ph_qual"][0], "{0:0<4d}".format(int(str(bin(row["det_bit"][0]))[2:])))
[…]

### 2013-10-01 ptf/analyze/analyze.py
added project root, word wrapping stuff

@@ -20,7 +20,8 @@ from .statistics import *
 from ..util import get_logger
 logger = get_logger(__name__)
 
-__all__ = ["fit_subtract_microlensing", "fit_microlensing_event", "fit_constant_line", "iscandidate"]
+__all__ = ["fit_subtract_microlensing", "fit_microlensing_event", 
+           "fit_constant_line", "iscandidate"]
 
 def _parameters_to_dict(parameters):
     """ Convert an lmfit Parameters object to a Python dictionary """

### 2013-10-01 ptf/db/photometric_database.py

@@ -21,17 +21,26 @@ logger = get_logger(__name__)
 try:
     import tables
 except ImportError:
-    logger.warning("PyTables not found! Some functionality won't work.\n (on navtara, try running with: /scr4/dlevitan/sw/epd-7.1-1-rh5-x86_64/bin/python instead.)")
+    logger.warning("PyTables not found! Some functionality won't work."
+                   "\n (on navtara, try running with: /scr4/dlevitan/sw/"
+                    "epd-7.1-1-rh5-x86_64/bin/python instead.)")
 
 try:
     import galacticutils
 except ImportError:
-    logger.warning("galacticutils not found! SDSS search functionality won't work.")
+    logger.warning("galacticutils not found! SDSS search functionality "
+                   "won't work.")
 
-#all_fields = np.load(os.path.join(ptf_params.config["PROJECTPATH"], "data", "all_fields.npy"))
-all_fields = np.load(os.path.join(os.path.split(globals._base_path)[0], "data", "all_fields.npy"))
+# Path to npy file containing all field IDs
+#_all_fields = os.path.join(ptf_params.config["PROJECTPATH"], 
+#                           "data", "all_fields.npy")
+_all_fields = os.path.join(os.path.split(globals._base_path)[0], 
+                           "data", "all_fields.npy")
+all_fields = np.load(_all_fields)
+
 match_path = "/scr4/dlevitan/matches"
-pytable_base_string = os.path.join(match_path, "match_{filter.id:02d}_{field.id:06d}_{ccd.id:02d}.pytable")
+pytable_base_string = os.path.join(match_path, "match_{filter.id:02d}"
+                                   "_{field.id:06d}_{ccd.id:02d}.pytable")
 filter_map = {"R" : 2, "g" : 1}
 inv_filter_map = {2 : "R", 1 : "g"}
 

### 2013-10-01 ptf/util.py

@@ -3,7 +3,11 @@
 
 # Standard library
 import os, sys
+import re
 import logging
+
+# Third-party
+import numpy as np
 
 # Convert my names for variability indices to the PDB names
 pdb_index_name = dict(eta="vonNeumannRatio", \
@@ -11,6 +15,18 @@ pdb_index_name = dict(eta="vonNeumannRatio", \
                       k="stetsonK", \
                       delta_chi_squared="chiSQ", \
                       sigma_mu=["magRMS","referenceMag"])
+
+# This code will find the root directory of the project
+_pattr = re.compile("(.*)\/ptf")
+try:
+    matched_path = _pattr.search(os.getcwd()).groups()[0]
+except AttributeError: # match not found, try __file__ instead
+    matched_path = _pattr.search(__file__).groups()[0]
+
+if os.path.basename(matched_path) == "ptf":
+    project_root = matched_path
+else:
+    project_root = os.path.join(matched_path, "ptf")
 
 def source_index_name_to_pdb_index(source, index_name):
     """ Given a source (a row from chip.sources) and an index name (e.g. eta),
@@ -39,26 +55,41 @@ ch = logging.StreamHandler()
 formatter = logging.Formatter("%(name)s / %(levelname)s / %(message)s")
 ch.setFormatter(formatter)
 logger.addHandler(ch)
+
 def get_logger(name):
     return logger
 
-def convert_all_fields_txt(txt_filename="ptf_allfields.txt", npy_filename="all_fields.npy"):
-    """ Convert the text file containing all PTF Field ID's and positons to a
-        numpy .npy save file.
+def cache_all_fields(txt_filename="PTF_AllFields_ID.txt", 
+                     npy_filename="all_fields.npy",
+                     path="data/cache", overwrite=False):
+    """ Convert the text file containing all PTF Field ID's and 
+        positons to a numpy .npy save file.
 
         It is much faster to load the .npy file than to generate an array on
         the fly from the text file. The text file can be found here:
             http://astro.caltech.edu/~eran/PTF_AllFields_ID.txt
     """
-    all_fields = np.genfromtxt(txt_filename, \
-                           skiprows=4, \
-                           usecols=[0,1,2,5,6,7], \
-                           dtype=[("id", int), ("ra", float), ("dec", float), ("gal_lon", float), ("gal_lat", float), ("Eb_v", float)]).\
-                           view(np.recarray)
+    txt_filename = os.path.join(project_root, path, txt_filename)
[…]

### 2013-10-01 setup.py

@@ -19,3 +19,7 @@ setup(name='PTF',
       description="adrn's PTF project",
       cmdclass={'build_ext': build_ext},
       ext_modules=[error_func, aov])
+
+# Now generate various data files:
+from ptf.util import cache_all_fields
+cache_all_fields()

### 2013-10-01 ptf/globals.py
change all_fields path

@@ -24,7 +24,7 @@ with open(os.path.join(_base_path, "config"), "r") as f:
         except ValueError:
             config[key] = val
 
-all_fields = np.load(os.path.join(os.path.split(_base_path)[0], "data", "all_fields.npy"))
+all_fields = np.load(os.path.join(os.path.split(_base_path)[0], "data", "cache", "all_fields.npy"))
 data_path = os.path.join(os.path.split(_base_path)[0], "data")
 plots_path = os.path.join(os.path.split(_base_path)[0], "plots")
 

### 2013-10-01 scripts/try_svm.py
script for trying an SVM classifier

@@ -1,0 +1,42 @@
+# coding: utf-8
+
+""" ...explain... """
+
+from __future__ import division, print_function
+
+__author__ = "adrn <adrn@astro.columbia.edu>"
+
+# Standard library
+import os, sys
+import logging
+
+# Third-party
+import tables
+import numpy as np
+import astropy.units as u
+
+# Project
+from ptf.globals import all_fields
+
+# Create logger
+logger = logging.getLogger(__name__)
+
+lcBaseDir = '/scr2/ptf/variable/matches/'
+indexFile = os.path.join(lcBaseDir, 'index.npy')
+index = np.load(indexFile)
+
+np.random.seed(142)
+idx = np.random.randint(len(all_fields))
+print(all_fields[idx])
+
+sys.exit(0)
+#filterID = 
+
+matchFile = os.path.join(lcBaseDir, 'match_{:02d}_{:06d}_{:02d}.pytable'\
+                          .format(filterID, fieldID, chip))
+chip = tables.openFile(matchFile)
+
+sourceTable = chip.getNode('/filter{:02d}/field{:06d}/chip{:02d}/sources'\
+                           .format(filterID, fieldID, chipID))
+
+print(len(sourceTable))

### 2013-10-01 scripts/try_svm.py
read actual file

@@ -27,13 +27,14 @@ index = np.load(indexFile)
 
 np.random.seed(142)
 idx = np.random.randint(len(all_fields))
-print(all_fields[idx])
 
 sys.exit(0)
-#filterID = 
+filterID = 2
+fieldID = all_fields[idx][0]
+chipID = 1
 
 matchFile = os.path.join(lcBaseDir, 'match_{:02d}_{:06d}_{:02d}.pytable'\
-                          .format(filterID, fieldID, chip))
+                          .format(filterID, fieldID, chipID))
 chip = tables.openFile(matchFile)
 
 sourceTable = chip.getNode('/filter{:02d}/field{:06d}/chip{:02d}/sources'\

### 2013-10-06 ptf/db/photometric_database.py
import fu

@@ -34,7 +34,7 @@ except ImportError:
 # Path to npy file containing all field IDs
 #_all_fields = os.path.join(ptf_params.config["PROJECTPATH"], 
 #                           "data", "all_fields.npy")
-_all_fields = os.path.join(os.path.split(globals._base_path)[0], 
+_all_fields = os.path.join(os.path.split(_base_path)[0], 
                            "data", "all_fields.npy")
 all_fields = np.load(_all_fields)
 

### 2013-10-06 ptf/db/photometric_database.py
remove redefinition of all_fields

@@ -31,13 +31,6 @@ except ImportError:
     logger.warning("galacticutils not found! SDSS search functionality "
                    "won't work.")
 
-# Path to npy file containing all field IDs
-#_all_fields = os.path.join(ptf_params.config["PROJECTPATH"], 
-#                           "data", "all_fields.npy")
-_all_fields = os.path.join(os.path.split(_base_path)[0], 
-                           "data", "all_fields.npy")
-all_fields = np.load(_all_fields)
-
 match_path = "/scr4/dlevitan/matches"
 pytable_base_string = os.path.join(match_path, "match_{filter.id:02d}"
                                    "_{field.id:06d}_{ccd.id:02d}.pytable")

### 2013-10-06 ptf/coverage/__init__.py
add coverage code

@@ -1,0 +1,2 @@
+from .core import *
+from .js import *

### 2013-10-06 ptf/coverage/core.py

@@ -1,0 +1,120 @@
+# coding: utf-8
+from __future__ import division
+
+""" 
+    This module contains utilities for computing PTF's survey coverage, both in time and position. 
+"""
+
+# Standard library
+import sys, os
+import datetime
+import time
+
+# Third-party
+import numpy as np
+
+# PTF
+from ..globals import ccd_size, all_fields, _base_path
+from ..lightcurve import PTFLightCurve, PDBLightCurve
+from ..analyze import compute_variability_indices
+from ..util import get_logger
+logger = get_logger(__name__)
+
+__all__ = ["SurveyInfo"]
+
+def get_fields_exposures(filter, filename=None, overwrite=False):
+    """ Given a filter, go to the PTF photometric database and get information about all PTF
+        fields for that filter, and the number of good exposures per field.
+        
+        Parameters
+        ----------
+        filter : ptf.photometricdatabase.Filter
+            Must be a Filter object (see the above module)
+        filename : str (optional)
+            The filename to store this data to.
+        overwrite : bool (optional)
+            Overwrite 'filename' if it already exists.
+        
+    """
+    
+    #if not isinstance(filter, pdb.Filter):
+    #    raise ValueError("Filter must be a valid Filter() object!")
+    
+    if filename == None:
+        filename = os.path.join(_base_path, "data", "survey_coverage", "fields_observations_{}.npy".format(str(filter)))
+    
+    if os.path.exists(filename) and overwrite:
+        logger.debug("Data file already exists, but you want to overwrite it")
+        os.remove(filename)
+        logger.debug("File {} deleted".format(filename))
+    elif os.path.exists(filename) and not overwrite:
+        logger.info("Data file already exists: {}".format(filename))
+    
+    if not os.path.exists(filename):
+        logger.info("Data file doesn't exist -- it could take some time to create it!")
+        
+        fields = []
+        exposures = []
+        
+        pattr = re.compile(".*match_(\d+)_(\d+)_(\d+)")
+        for match_filename in glob.glob("/scr4/dlevitan/matches/match_{:02d}_*.pytable".format(filter.id)):
+            logger.debug("Reading file: {}".format(match_filename))
+            
+            filter_id, field_id, ccd_id = map(int, pattr.search(match_filename).groups())
+            
+            if field_id in fields:
+                continue
+    
+            try:
+                file = tables.openFile(match_filename)
[…]

### 2013-10-06 ptf/coverage/js.py

@@ -1,0 +1,135 @@
+# coding: utf-8
+from __future__ import division
+
+""" This module contains utilities for computing PTF's survey coverage for JavaScript visualizations. """
+
+__author__ = "adrn <adrn@astro.columbia.edu>"
+
+# Standard library
+import json
+import math
+
+# Third-party
+import numpy as np
+import apwlib.geometry as g
+
+# Project
+from ..util import get_logger
+logger = get_logger(__name__)
+
+__all__ = ["field_to_feature", "field_list_to_json"]
+
+def field_to_feature(field):
+    """ Converts a PTF Field object into a 'feature' object to be stuffed into JSON for the 
+        interactive survey coverage viewer.
+        
+        Parameters
+        ----------
+        field : ptf.photometricdatabase.Field
+            Must be a PTF Field object. See above module for details
+    """
+    feature = dict(type="Feature", id=str(field.id))
+    properties = dict(name=str(field.id))
+    geometry = dict(type="Polygon")
+    
+    ra = field.ra.degrees-180.0
+    dec = field.dec.degrees
+    
+    ra_offset = camera_size_degrees[0]/math.cos(math.radians(dec))/2.
+    dec_offset = camera_size_degrees[1]/2.
+    
+    min_ra = ra - ra_offset
+    max_ra = ra + ra_offset
+    
+    min_dec = dec - dec_offset
+    max_dec = dec + dec_offset
+    
+    coordinates = [[ [min_ra, min_dec], [max_ra, min_dec], [max_ra, max_dec], [min_ra, max_dec], [min_ra, min_dec] ]]
+    geometry["coordinates"] = coordinates
+    feature["geometry"] = geometry
+    feature["properties"] = properties
+
+    return feature
+
+def field_list_to_json(fields, filename=None):
+    """ Given a list of fields, create a FeatureCollection JSON file to use with the 
+        interactive survey coverage viewer.
+        
+        The final structure should look like this, where "name" is the field id,
+        and coordinates are a list of the coordinates of the 4 corners of the field.
+        
+            {"type" : "FeatureCollection",
+                "features": [
+                    {"type" : "Feature",
+                     "properties" : { "name" : "2471"},
+                                      "geometry": { "type" : "Polygon",
+                                                    "coordinates" : [[ [100.0, 0.0], [101.0, 0.0], [101.0, 1.0], [100.0, 1.0], [100.0, 0.0] ]]
+                                                  },
+                     "id":"2471"}, etc.
+                            ]
+            }      
+        
[…]

### 2013-10-06 ptf/globals.py
read in config

@@ -16,7 +16,6 @@ config = dict()
 _base_path = os.path.split(__file__)[0]
 
 # Need this for mongodb config
-'''
 with open(os.path.join(_base_path, "config"), "r") as f:
     for line in f.readlines():
         key,val = line.split()
@@ -24,7 +23,6 @@ with open(os.path.join(_base_path, "config"), "r") as f:
             config[key] = int(val)
         except ValueError:
             config[key] = val
-'''
 
 all_fields = np.load(os.path.join(os.path.split(_base_path)[0], "data", "cache", "all_fields.npy"))
 data_path = os.path.join(os.path.split(_base_path)[0], "data")

### 2013-10-07 scripts/candidate_pipeline.py
mostly line wrapping

@@ -219,7 +219,8 @@ if __name__ == "__main__":
         field = pdb.Field(field_id, "R")
         logger.info("Field: {}".format(field.id))
 
-        # There is some strangeness to getting the coordinates for a Field -- this just makes it stupidproof
+        # There is some strangeness to getting the coordinates for a Field 
+        # -- this just makes it stupidproof
         try:
             if field.ra == None or field.dec == None:
                 raise AttributeError()
@@ -240,7 +241,8 @@ if __name__ == "__main__":
                     continue
 
         if args.overwrite_lcs:
-            field_collection.update({"_id" : field.id}, {"$set" : {"already_searched" : False}})
+            field_collection.update({"_id" : field.id}, 
+                                    {"$set" : {"already_searched" : False}})
             #light_curve_collection.remove({"field_id" : field.id})
 
             # Only remove light curves that haven't been looked at
@@ -264,7 +266,9 @@ if __name__ == "__main__":
             continue
 
         # Check to see if the selection criteria for this field is already loaded into the database
-        selection_criteria = field_collection.find_one({"_id" : field.id}, fields=["selection_criteria"])["selection_criteria"]
+        selection_criteria = field_collection.find_one({"_id" : field.id}, 
+                                                       fields=["selection_criteria"])
+        selection_criteria = selection_criteria["selection_criteria"]
 
         if args.overwrite or selection_criteria == None:
             logger.info("Selection criteria not available for field.")
@@ -295,8 +299,11 @@ if __name__ == "__main__":
             field_collection.update({"_id" : field.id}, {"$set" : {"selection_criteria" : selection_criteria_document}})
 
         logger.debug("Selection criteria loaded")
-        selection_criteria_document = field_collection.find_one({"_id" : field.id}, fields=["selection_criteria"])["selection_criteria"]["eta"]
-
+        selection_criteria_document = field_collection.find_one({"_id" : field.id}, 
+                                                                fields=["selection_criteria"])
+        selection_criteria_document = selection_criteria_document["selection_criteria"]["eta"]
+        
+        # APW: ok so what is select_candidates doing?
         selected_light_curves = select_candidates(field, selection_criteria_document)
         logger.info("Selected {} light curves.".format(len(selected_light_curves)))
 

### 2013-10-07 scripts/candidate_pipeline.py
word wrapping

@@ -26,7 +26,9 @@ import scipy.optimize as so
 try:
     from apwlib.globals import greenText, yellowText, redText
 except ImportError:
-    raise ImportError("apwlib not found! \nDo: 'git clone git@github.com:adrn/apwlib.git' and run 'python setup.py install' to install.")
+    raise ImportError("apwlib not found! \nDo: 'git clone "
+        "git@github.com:adrn/apwlib.git' and run 'python setup.py "
+        "install' to install.")
 
 # Project
 import ptf.db.photometric_database as pdb
@@ -40,8 +42,9 @@ logger = get_logger(__name__)
 def select_candidates(field, selection_criteria, num_fit_attempts=10):
     """ Select candidates from a field given the log10(selection criteria) from mongodb.
 
-        The current selection scheme is to first select on eta, then to sanity check with
-        delta chi-squared by making sure it's positive and >10.
+        The current selection scheme is to first select on eta, then to
+        sanity check with delta chi-squared by making sure it's positive 
+        and >10.
 
     """
 
@@ -52,29 +55,36 @@ def select_candidates(field, selection_criteria, num_fit_attempts=10):
         logger.info(greenText("Starting with CCD {}".format(ccd.id)))
         chip = ccd.read()
 
-        source_ids = chip.sources.readWhere("(ngoodobs > {}) & \
-                                          (vonNeumannRatio > 0.0) & \
-                                          (vonNeumannRatio < {}) & \
-                                          ((ngoodobs/nobs) > 0.5)".format(min_number_of_good_observations, eta_cut), \
-                                          field="matchedSourceID")
+        cdtn = ("(ngoodobs > {}) & (vonNeumannRatio > 0.0) & "
+                "(vonNeumannRatio < {}) & ((ngoodobs/nobs) > 0.5)")
+        cdtn = cdtn.format(min_number_of_good_observations, eta_cut)
+        source_ids = chip.sources.readWhere(cdtn, field="matchedSourceID")
 
-        logger.info("\tSelected {} pre-candidates from PDB".format(len(source_ids)))
+        logger.info("\tSelected {} pre-candidates from PDB"\
+                    .format(len(source_ids)))
 
         for source_id in source_ids:
-            # APW: TODO -- this is still the biggest time hog!!! It turns out it's still faster than reading the whole thing into memory, though!
-            light_curve = ccd.light_curve(source_id, barebones=True, clean=True)
+            # APW: TODO -- this is still the biggest time hog!!! It turns 
[…]

### 2013-10-07 scripts/candidate_pipeline.py
wrapping

@@ -181,27 +181,28 @@ if __name__ == "__main__":
     from argparse import ArgumentParser
 
     parser = ArgumentParser(description="")
-    parser.add_argument("-v", "--verbose", action="store_true", dest="verbose", default=False,
-                    help="Be chatty! (default = False)")
-    parser.add_argument("-q", "--quiet", action="store_true", dest="quiet", default=False,
-                    help="Be quiet! (default = False)")
-
-    parser.add_argument("-o", "--overwrite", action="store_true", dest="overwrite", default=False,
-                    help="Overwrite any existing / cached files")
-    parser.add_argument("--overwrite-light-curves", action="store_true", dest="overwrite_lcs", default=False,
-                    help="Keep the selection criteria, just redo the search.")
-
-    parser.add_argument("-a", "--all", dest="all", default=False, action="store_true",
-                    help="Run on all fields.")
-    parser.add_argument("-f", "--field-id", dest="field_id", default=[], nargs="+", type=int,
-                    help="The PTF field IDs to run on")
+    parser.add_argument("-v","--verbose", action="store_true", dest="verbose",
+                        default=False, help="Be chatty! (default = False)")
+    parser.add_argument("-q", "--quiet", action="store_true", dest="quiet", 
+                        default=False, help="Be quiet! (default = False)")
+    parser.add_argument("-o", "--overwrite", action="store_true", 
+                        dest="overwrite", default=False, 
+                        help="Overwrite any existing / cached files")
+    parser.add_argument("--overwrite-light-curves", action="store_true", 
+                        dest="overwrite_lcs", default=False,
+                        help="Keep the selection criteria, just redo the search.")
+    parser.add_argument("-a", "--all", dest="all", default=False, 
+                        action="store_true", help="Run on all fields.")
+    parser.add_argument("-f", "--field-id", dest="field_id", default=[], 
+                        nargs="+",type=int,help="The PTF field IDs to run on")
     parser.add_argument("-r", "--field-range", dest="field_range", default=[],
-                    help="A range of PTF field IDs to run on")
-
-    parser.add_argument("--num-light-curves", dest="num_light_curves", default=100, type=int,
-                    help="Number of light curves to select from each CCD.")
-    parser.add_argument("--num-simulations", dest="num_simulations", default=100, type=int,
-                    help="Number of simulations per light curve.")
+                        help="A range of PTF field IDs to run on")
[…]

### 2013-10-07 text/paper/ms.tex
add percentages for each step, how many light curves removed

@@
 	\item select a field from the 2097 fields with $>$10 good $R$-band observations and run Monte Carlo simulations to compute the 1\% FPR limiting values for $\eta$, $\Delta \chi^2$, and $J$ for that field;
-	\item compute $\eta$ for all the light curves in the field, and select the light curves that pass the quality cut described in Section~\ref{sec:detection_eff} and the FPR cut determined from the Monte Carlo simulations;
+	\item compute $\eta$ for all the light curves in the field, and select the light curves that pass the quality cut described in Section~\ref{sec:detection_eff} ($\sim$10\%) and the FPR cut determined from the Monte Carlo simulations ($\sim$0.1-0.5\%);
+	\item for fields with SDSS coverage, we use Data Release 9 photometry \citep{dr9paper} to remove sources typed as Galaxy or QSO (using \texttt{objc\_type}\footnote{http://www.sdss3.org/dr9/algorithms/classify.php}) and flag candidate quasars \citep[using the cuts described in][]{richards02} ($\sim$10\%);
 	\item fit a microlensing event model and subtract the model, then recompute $\eta$ for all the selected light curves;
-	\item reject the light curves for which the new value of $\eta$ still passes the cut\footnote{These are probably periodic variables.};
+	\item reject the light curves for which the new value of $\eta$ still passes the cut\footnote{These are probably periodic variables.} ($\sim$5-10\%);
 	\item retain any remaining light curves for further inspection.
@@
 We search SIMBAD\footnote{This research has made use of the SIMBAD database, operated at CDS, Strasbourg, France.} and remove any candidate among these with a known extragalactic counterpart within 10\asec.
-For those in the SDSS footprint, we use Data Release 7 photometry \citep{DR7paper} to identify candidate quasars \citep[using the cuts described in][]{richards02} and examine the SDSS images \citep{sdss_images} to remove extended objects.
+For those in the SDSS footprint, we examine the SDSS images \citep{sdss_images} to remove extended objects.
 Removing extragalactic objects in this manner from our candidate list reduces contamination by, e.g., supernovae.
@@
 %\setlength{\baselineskip}{0.6\baselineskip}
-%\bibliography{references_5}
 %\bibliography{references_3}
@@
 
-\bibitem[{{Abazajian} {et~al.}(2009)}]{DR7paper}
-{Abazajian}, K.~N., {et~al.} 2009, \apjs, 182, 543
-
 \bibitem[{{Ag{\"u}eros} {et~al.}(2011){Ag{\"u}eros}, {Covey}, {Lemonias},
@@
   740, 110
+
+\bibitem[{{Ahn} {et~al.}(2012){Ahn}, {Alexandroff}, {Allende Prieto},
+  {Anderson}, {Anderton}, {Andrews}, {Aubourg}, {Bailey}, {Balbinot}, {Barnes},
+  \& et~al.}]{dr9paper}
+{Ahn}, C.~P., {Alexandroff}, R., {Allende Prieto}, C., {et~al.} 2012, \apjs,
+  203, 21
 

### 2013-10-07 scripts/paper_figures.py
rmag

@@ -262,12 +262,13 @@ def maximum_outlier_indices_plot(field_id):
         axes[ii].set_title(pu.index_to_label(index), fontsize=28)
         #axes[ii].set_xlim(best_outlier_lightcurve.mjd.min()-2, best_outlier_lightcurve.mjd.max()+2)
         #axes[ii, 1].set_xlim(55350, 55600)
+        axes[ii].set_ylabel("$R$ [mag]")
 
     for ax in fig.axes[:-1]:
         ax.xaxis.set_visible(False)
-        ax.yaxis.set_visible(False)
+        #ax.yaxis.set_visible(False)
 
-    fig.axes[-1].yaxis.set_visible(False)
+    #fig.axes[-1].yaxis.set_visible(False)
     for ticklabel in fig.axes[-1].get_xticklabels():
         ticklabel.set_fontsize(22)
 

### 2013-10-07 scripts/paper_figures.py
make fig5 nicer

@@ -196,7 +196,7 @@ def maximum_outlier_indices_plot(field_id):
     pdb_indices = ["bestVonNeumannRatio", "bestChiSQ", "bestStetsonJ", "bestStetsonK", ["bestMagRMS", "bestMedianMag"], "bestCon"]
     field = pdb.Field(field_id, "R")
 
-    fig, axes = plt.subplots(len(indices), 1, sharex=True, figsize=(18,24))
+    fig, axes = plt.subplots(len(indices), 1, sharex=True, figsize=(15,20))
 
     for ii, (index, pdb_index) in enumerate(zip(indices, pdb_indices)):
         all_outlier_sources = []
@@ -262,18 +262,24 @@ def maximum_outlier_indices_plot(field_id):
         axes[ii].set_title(pu.index_to_label(index), fontsize=28)
         #axes[ii].set_xlim(best_outlier_lightcurve.mjd.min()-2, best_outlier_lightcurve.mjd.max()+2)
         #axes[ii, 1].set_xlim(55350, 55600)
-        axes[ii].set_ylabel("$R$ [mag]")
+        axes[ii].set_ylabel("$R$ [mag]", fontsize=26)
+        axes[ii].set_yticks(axes[ii].get_yticks()[1:-1])
+        for ticklabel in axes[ii].get_yticklabels():
+            ticklabel.set_fontsize(22)
 
     for ax in fig.axes[:-1]:
         ax.xaxis.set_visible(False)
         #ax.yaxis.set_visible(False)
 
     #fig.axes[-1].yaxis.set_visible(False)
+    ticklbls = []
     for ticklabel in fig.axes[-1].get_xticklabels():
         ticklabel.set_fontsize(22)
+        ticklbls.append(int(ticklabel.get_text()))
+    fig.axes[-1].set_xticklabels([str(x-min(ticklbls)) for x in ticklbls])
 
-    axes[-1].set_xlabel("MJD", fontsize=26)
-    fig.subplots_adjust(hspace=0.2, top=0.95, bottom=0.08)
+    axes[-1].set_xlabel("time [days]", fontsize=26)
+    fig.subplots_adjust(hspace=0.0, top=0.95, bottom=0.08)
     fig.savefig(os.path.join(pg.plots_path, "paper_figures", "max_outlier_light_curves.pdf".format(field_id)))#, bbox_inches="tight")
 
 def intersect_light_curves(light_curve1, light_curve2):

### 2013-10-07 scripts/paper_figures.py
plot dammit

@@ -16,7 +16,7 @@ import time
 
 # Third-party
 import matplotlib
-matplotlib.use("TkAgg")
+matplotlib.use("Agg")
 import matplotlib.cm as cm
 import matplotlib.colors as mc
 import matplotlib.pyplot as plt
@@ -1035,14 +1035,14 @@ if __name__ == "__main__":
 
     #make_survey_sampling_figure(10)
     #microlensing_event_sim()
-    #maximum_outlier_indices_plot(100101)
+    maximum_outlier_indices_plot(100101)
     #variability_indices_distributions()
     #num_observations_distribution()
     #num_observations_distribution_90deg()
     #after_eta_cut_num_observations()
     
     #np.random.seed(400)
-    fit_candidates(Nsamples=1000, Nburn=500, Nwalkers=100)
+    #fit_candidates(Nsamples=1000, Nburn=500, Nwalkers=100)
     
     #np.random.seed(104)
     #fit_non_candidates(Nsamples=500, Nburn=500, Nwalkers=100)

### 2013-10-07 scripts/paper_figures.py
updated figure 5

@@ -259,24 +259,26 @@ def maximum_outlier_indices_plot(field_id):
                 elif min_max[index] == "max":
                     idx_vals[idx_vals.argmax()] = -1E8
 
-        axes[ii].set_title(pu.index_to_label(index), fontsize=28)
         #axes[ii].set_xlim(best_outlier_lightcurve.mjd.min()-2, best_outlier_lightcurve.mjd.max()+2)
         #axes[ii, 1].set_xlim(55350, 55600)
         axes[ii].set_ylabel("$R$ [mag]", fontsize=26)
-        axes[ii].set_yticks(axes[ii].get_yticks()[1:-1])
+        yticks = np.linspace(min(axes[ii].get_yticks()), max(axes[ii].get_yticks()), 5)[1:-1]
+        axes[ii].set_yticks([round(yt,2) for yt in yticks])
         for ticklabel in axes[ii].get_yticklabels():
             ticklabel.set_fontsize(22)
+        
+        axes[ii].text(55900, yticks[0], "{0}({1})".format(min_max[index], pu.index_to_label(index)), fontsize=30)
 
     for ax in fig.axes[:-1]:
         ax.xaxis.set_visible(False)
         #ax.yaxis.set_visible(False)
 
     #fig.axes[-1].yaxis.set_visible(False)
-    ticklbls = []
     for ticklabel in fig.axes[-1].get_xticklabels():
         ticklabel.set_fontsize(22)
-        ticklbls.append(int(ticklabel.get_text()))
-    fig.axes[-1].set_xticklabels([str(x-min(ticklbls)) for x in ticklbls])
+
+    xticks = fig.axes[-1].get_xticks()
+    axes[-1].set_xticklabels(["{0:d}".format(int(x-min(xticks))) for x in xticks])
 
     axes[-1].set_xlabel("time [days]", fontsize=26)
     fig.subplots_adjust(hspace=0.0, top=0.95, bottom=0.08)

### 2013-10-07 text/paper/ms.tex
fixed fig 5 caption
updated some minor grammar

@@
 Figure~\ref{fig:sampling} shows six randomly selected light curves and illustrates the varying cadences and coverage that different fields may have over the same one-year period.
-The PTF data is not publicly available, but ...
+The PTF data are not publicly available, but ...
 TODO
@@
 \begin{figure}[h]
-\centering\includegraphics[width=.48\textwidth, trim=130 85 90 65, clip]{fig5.eps}
-\caption{The maximally outlying light curves for each variability index applied to data from PTF field 100101.
+\centering\includegraphics[width=.48\textwidth, trim=10 50 50 50, clip]{fig5.eps}
+\caption{Light curves selected from PTF field 100101 with maximally outlying values for each variability index.
 For $\eta$ and $K$, this corresponds to the light curve with the \emph{minimum} value of the index over the entire field.
 For the other indices, these are the light curves with the \emph{maximum} value of the relevant index over the field.
+These light curves represent the type of variability that each index is most sensitive to.
 } \label{fig:indices_examples}
@@
 For the Gaussian fit, the optimizer is initialized with a standard-deviation of 10 days, centered on the brightest data point.
-This is the standard statistical test used by microlensing surveys, and allows us to compare the relative performance of the (slightly modified) \cite{shin2009} indices and of this approach.
+This is a standard statistical test used by microlensing surveys, and allows us to compare the relative performance of the (slightly modified) \cite{shin2009} indices and of this approach.
 

### 2013-10-07 scripts/paper_dataset.py
find 100000 random field, ccd pairs

@@ -1,1 +1,55 @@
+# coding: utf-8
+from __future__ import division
 
+""" Select 100000 random light curves from PTF to publish. """
+
+__author__ = "adrn <adrn@astro.columbia.edu>"
+
+# Standard library
+import os, sys
+import logging
+import random
+
+# Third-party
+import numpy as np
+
+# Project
+import ptf.db.photometric_database as pdb
+from ptf.globals import all_fields
+
+np.random.seed(42)
+Nfields = len(all_fields)
+Nlightcurves = 100000
+Nselected = 0
+
+ccd_ids = range(13)
+ccd_ids.pop(3)
+
+field_ccd = dict()
+while Nselected < Nlightcurves:
+    field_row = all_fields[np.random.randint(Nfields)]
+    field_id = field_row['id']
+    ccd_id = random.choice(ccd_ids)
+
+    print(field_id, ccd_id)
+
+    try:
+        ccd = pdb.CCD(ccd_id, field_id, 'R')
+    except ValueError:
+        continue
+
+    try:
+        field_ccd[field_id]
+    except KeyError:
+        field_ccd[field_id] = dict()
+
+    field = field_ccd[field_id]
+
+    try:
+        field[ccd_id]
+    except KeyError:
+        field[ccd_id] = 0
+
+    field_ccd[field_id][ccd_id] += 1
+    Nlightcurves += 1
+    sys.exit(0)    

### 2013-10-16 scripts/paper_figures.py
remove outliers from fig 5 plot

@@ -250,6 +250,11 @@ def maximum_outlier_indices_plot(field_id):
                 best_outlier_source = all_outlier_sources[idx_vals.argmax()]
                 best_outlier_lightcurve = all_ccds[idx_vals.argmax()].light_curve(best_outlier_source["matchedSourceID"], clean=clean, barebones=True)
 
+            sig,mn = np.std(best_outlier_lightcurve.mag), np.median(best_outlier_lightcurve.mag)
+            ix = (best_outlier_lightcurve.mag < (md+8*sig)) & (best_outlier_lightcurve.mag > (md-8*sig))
+            best_outlier_lightcurve = PTFLightCurve(best_outlier_lightcurve.mjd, 
+                                                    best_outlier_lightcurve.mag,
+                                                    best_outlier_lightcurve.error)
             try:
                 best_outlier_lightcurve.plot(axes[ii], ms=4)
                 break

### 2013-10-16 text/paper/ms.tex
Added durbin-watson reference

@@
 
-$\eta$, the von Neumann ratio \citep[][also known as the Durbin-Watson statistic]{von_neumann1941}, is the mean square successive difference divided by the sample variance:
+$\eta$, the von Neumann ratio \citep[][also known as the Durbin-Watson statistic]{von_neumann1941, durbin50}, is the mean square successive difference divided by the sample variance:
 \begin{align}
@@
 
-\begin{thebibliography}{53}
+\begin{thebibliography}{54}
 \expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
@@
 {Di Stefano}, R. 2008, \apj, 684, 46
+
+\bibitem[Durbin \& Watson(1950)]{durbin50}
+J.~Durbin and G.~S. Watson.
+\newblock Testing for serial correlation in least squares regression: I.
+\newblock {\em Biometrika}, 37(3/4):pp. 409--428, 1950.
 

### 2013-10-16 scripts/paper_figures.py
changed field for outliers

@@ -250,11 +250,11 @@ def maximum_outlier_indices_plot(field_id):
                 best_outlier_source = all_outlier_sources[idx_vals.argmax()]
                 best_outlier_lightcurve = all_ccds[idx_vals.argmax()].light_curve(best_outlier_source["matchedSourceID"], clean=clean, barebones=True)
 
-            sig,mn = np.std(best_outlier_lightcurve.mag), np.median(best_outlier_lightcurve.mag)
+            sig,md = np.std(best_outlier_lightcurve.mag), np.median(best_outlier_lightcurve.mag)
             ix = (best_outlier_lightcurve.mag < (md+8*sig)) & (best_outlier_lightcurve.mag > (md-8*sig))
-            best_outlier_lightcurve = PTFLightCurve(best_outlier_lightcurve.mjd, 
-                                                    best_outlier_lightcurve.mag,
-                                                    best_outlier_lightcurve.error)
+            best_outlier_lightcurve = PTFLightCurve(best_outlier_lightcurve.mjd[ix], 
+                                                    best_outlier_lightcurve.mag[ix],
+                                                    best_outlier_lightcurve.error[ix])
             try:
                 best_outlier_lightcurve.plot(axes[ii], ms=4)
                 break
@@ -272,7 +272,10 @@ def maximum_outlier_indices_plot(field_id):
         for ticklabel in axes[ii].get_yticklabels():
             ticklabel.set_fontsize(22)
         
-        axes[ii].text(55900, yticks[0], "{0}({1})".format(min_max[index], pu.index_to_label(index)), fontsize=30)
+        bbox_props = dict(boxstyle="round,pad=0.3", fc="white", ec="k", lw=1)
+        axes[ii].text(55975, yticks[2], "{0}({1})".format(min_max[index], pu.index_to_label(index)), 
+                      fontsize=30,
+                      bbox=bbox_props)
 
     for ax in fig.axes[:-1]:
         ax.xaxis.set_visible(False)
@@ -287,7 +290,7 @@ def maximum_outlier_indices_plot(field_id):
 
     axes[-1].set_xlabel("time [days]", fontsize=26)
     fig.subplots_adjust(hspace=0.0, top=0.95, bottom=0.08)
-    fig.savefig(os.path.join(pg.plots_path, "paper_figures", "max_outlier_light_curves.pdf".format(field_id)))#, bbox_inches="tight")
[…]

### 2013-10-16 ptf/db/photometric_database.py
comment out ximage yimage stuff cause missing from db now...

@@ -31,7 +31,7 @@ except ImportError:
     logger.warning("galacticutils not found! SDSS search functionality "
                    "won't work.")
 
-match_path = "/scr4/dlevitan/matches"
+match_path = "/scr2/ptf/variable/matches/"
 pytable_base_string = os.path.join(match_path, "match_{filter.id:02d}"
                                    "_{field.id:06d}_{ccd.id:02d}.pytable")
 filter_map = {"R" : 2, "g" : 1}
@@ -89,7 +89,6 @@ def quality_cut(sourcedata, source_id=None):
     y_cut1, y_cut2 = ccd_edge_cutoff, ccd_size[1] - ccd_edge_cutoff
 
     # TODO: Do a robust test to figure out what is faster: many constraints in 'where', or doing it with numpy
-
     if source_id == None:
         where_string = ""
     else:
@@ -105,9 +104,10 @@ def quality_cut(sourcedata, source_id=None):
 
     # Saturation limit, 14.3, based on email conversation with David Levitan
     # UPDATE: I NO LONGER CUT ON SATURATION LIMIT (data["mag"] > 14.3) &
-    cut_data = data[(data["x_image"] > 15) & (data["x_image"] < 2033) & \
-                    (data["y_image"] > 15) & (data["y_image"] < 4081) & \
-                    (data["relPhotFlags"] < 4) & \
+    #(data["x_image"] > 15) & (data["x_image"] < 2033) & \
+    #(data["y_image"] > 15) & (data["y_image"] < 4081) & \
+                    
+    cut_data = data[(data["relPhotFlags"] < 4) & \
                     (data["mag"] < 21) & \
                     ((data["sextractorFlags"] & 251) == 0) & \
                     ((data["ipacFlags"] & 6077) == 0) & \

### 2013-10-16 scripts/paper_dataset.py
generate list of 10^5 light curves to download

@@ -9,13 +9,14 @@ __author__ = "adrn <adrn@astro.columbia.edu>"
 import os, sys
 import logging
 import random
+import cPickle as pickle
 
 # Third-party
 import numpy as np
 
 # Project
 import ptf.db.photometric_database as pdb
-from ptf.globals import all_fields
+from ptf.globals import all_fields, data_path
 
 np.random.seed(42)
 Nfields = len(all_fields)
@@ -29,27 +30,25 @@ field_ccd = dict()
 while Nselected < Nlightcurves:
     field_row = all_fields[np.random.randint(Nfields)]
     field_id = field_row['id']
-    ccd_id = random.choice(ccd_ids)
-
-    print(field_id, ccd_id)
-
-    try:
-        ccd = pdb.CCD(ccd_id, field_id, 'R')
-    except ValueError:
+    
+    field = pdb.Field(field_id, 'R')
+    if len(field.ccds) == 0:
         continue
 
-    try:
-        field_ccd[field_id]
-    except KeyError:
-        field_ccd[field_id] = dict()
-
-    field = field_ccd[field_id]
+    ccd = random.choice(field.ccds.values())
 
     try:
-        field[ccd_id]
+        field_ccd[field.id]
     except KeyError:
-        field[ccd_id] = 0
+        field_ccd[field.id] = dict()
 
-    field_ccd[field_id][ccd_id] += 1
-    Nlightcurves += 1
-    sys.exit(0)    
+    try:
+        field_ccd[field_id][ccd.id]
+    except KeyError:
+        field_ccd[field_id][ccd.id] = 0
+
+    field_ccd[field.id][ccd.id] += 1
+    Nselected += 1
+
+with open(os.path.join(data_path, "publish_field_ccds.pickle"), 'w') as f:
+    pickle.dump(field_ccd, f)

### 2013-10-16 scripts/paper_dataset.py
make into function

@@ -19,36 +19,40 @@ import ptf.db.photometric_database as pdb
 from ptf.globals import all_fields, data_path
 
 np.random.seed(42)
-Nfields = len(all_fields)
-Nlightcurves = 100000
-Nselected = 0
 
-ccd_ids = range(13)
-ccd_ids.pop(3)
+def make_lookup_dict(Nlightcurves=100000):
+    Nfields = len(all_fields)
+    Nlightcurves = 100000
+    Nselected = 0
 
-field_ccd = dict()
-while Nselected < Nlightcurves:
-    field_row = all_fields[np.random.randint(Nfields)]
-    field_id = field_row['id']
+    ccd_ids = range(13)
+    ccd_ids.pop(3)
     
-    field = pdb.Field(field_id, 'R')
-    if len(field.ccds) == 0:
-        continue
+    field_ccd = dict()
+    while Nselected < Nlightcurves:
+        field_row = all_fields[np.random.randint(Nfields)]
+        field_id = field_row['id']
+    
+        field = pdb.Field(field_id, 'R')
+        if len(field.ccds) == 0:
+            continue
 
-    ccd = random.choice(field.ccds.values())
+        ccd = random.choice(field.ccds.values())
 
-    try:
-        field_ccd[field.id]
-    except KeyError:
-        field_ccd[field.id] = dict()
+        try:
+            field_ccd[field.id]
+        except KeyError:
+            field_ccd[field.id] = dict()
 
-    try:
-        field_ccd[field_id][ccd.id]
-    except KeyError:
-        field_ccd[field_id][ccd.id] = 0
+        try:
+            field_ccd[field_id][ccd.id]
+        except KeyError:
+            field_ccd[field_id][ccd.id] = 0
 
-    field_ccd[field.id][ccd.id] += 1
-    Nselected += 1
+            field_ccd[field.id][ccd.id] += 1
+            Nselected += 1
 
-with open(os.path.join(data_path, "publish_field_ccds.pickle"), 'w') as f:
-    pickle.dump(field_ccd, f)
+    with open(os.path.join(data_path, "publish_field_ccds.pickle"), 'w') as f:
+        pickle.dump(field_ccd, f)
+   
+def 

### 2013-10-18 scripts/paper_dataset.py
wrote function to store light curves

@@ -13,12 +13,14 @@ import cPickle as pickle
 
 # Third-party
 import numpy as np
+from astropy.io.misc import fnunpickle
 
 # Project
 import ptf.db.photometric_database as pdb
 from ptf.globals import all_fields, data_path
 
 np.random.seed(42)
+dict_file = os.path.join(data_path, "publish_field_ccds.pickle")
 
 def make_lookup_dict(Nlightcurves=100000):
     Nfields = len(all_fields)
@@ -52,7 +54,64 @@ def make_lookup_dict(Nlightcurves=100000):
             field_ccd[field.id][ccd.id] += 1
             Nselected += 1
 
-    with open(os.path.join(data_path, "publish_field_ccds.pickle"), 'w') as f:
+    with open(dict_file, 'w') as f:
         pickle.dump(field_ccd, f)
    
-def 
+def random_light_curves(ccd, N):
+    chip = ccd.read()
+    sources = chip.sources.readWhere("ngoodobs > 10")
+    np.random.shuffle(sources)
+
+    lcs = []
+    for source in sources:
+        lc = ccd.light_curve(source["matchedSourceID"], clean=True, barebones=True)
+        if len(lc) > 10:
+            lcs.append(lc)
+
+        if len(lcs) >= N:
+            return lcs
+
+def store_light_curves():
+    field_ccd = fnunpickle(dict_file)
+
+    all_lightcurves = []
+    for ii,field_id in enumerate(sorted(field_ccd.keys())):
+        this_file = os.path.join(data_path, "publish_light_curves", "{0}.fits".format(ii % 100))
+        if os.path.exists(this_file):
+            continue
+
+        field = pdb.Field(field_id, 'R')
+
+        for ccd_id in sorted(field_ccd[field_id].keys()):
+            num = field_ccd[field_id][ccd_id]
+            ccd = field.ccds[ccd_id]
+
+            lcs = random_light_curves(ccd, num)
+            all_lightcurves += lcs
+
+        if ii > 0 and (ii % 100) == 0:
+            #write to file
+
+            data = None
+            for lc in all_lightcurves:
+                n = len(lc.mjd)
+                f = [lc.field_id]*n
+                c = [lc.ccd_id]*n
+                s = [lc.source_id]*n
+                
+                if data is None:
+                    data = np.array([f, c, s, lc.mjd, lc.mag, lc.error])
+                else:
[…]

### 2013-10-18 text/paper/ms.tex
added brink citation

--- 2013-10-07 text/paper/ms.tex
@@
 
-\slugcomment{DRAFT \today}
+%\slugcomment{DRAFT \today}
 \shorttitle{Microlensing \& PTF}
@@
 \title{Statistical Searches for Microlensing Events in Large, Non-Uniformly Sampled Time-Domain Surveys: A Test Using Palomar Transient Factory Data}
-\author{Adrian~M.~Price-Whelan\altaffilmark{1,8} Marcel~A.~Ag\"ueros\altaffilmark{1}, Amanda P.
+\author{Adrian~M.~Price-Whelan\altaffilmark{1} Marcel~A.~Ag\"ueros\altaffilmark{1}, Amanda P.
 Fournier\altaffilmark{2,3}, Rachel Street\altaffilmark{3}, Eran O.
@@
 
-\altaffiltext{1}{Department of Astronomy, Columbia University, 550 W 120th St., New York, NY 10027, USA}
+\altaffiltext{1}{Department of Astronomy, Columbia University, 550 W 120th St., New York, NY 10027, USA; adrn@astro.columbia.edu}
 \altaffiltext{2}{Department of Physics, Broida Hall, University of California, Santa Barbara, CA 93106, USA}
@@
 To examine the probability that microlensing events can be recovered in these data, we test previous statistics used on uniformly sampled data to identify variables and transients.
-We find that one such statistic, the von Neumann ratio, performs best for identifying simulated microlensing events in our data.
+We find that one such statistic, the von Neumann ratio, also known as the Durbin-Watson statistic, performs best for identifying simulated microlensing events in our data.
 We develop a selection method using this statistic and apply it to data from all PTF fields with $>$10 $R$-band observations --- a total of $1.1\times10^9$ light curves --- to uncover three candidate sparse-field microlensing events.
-While these candidates cannot be confirmed, as we lack simultaneous, multi-color photometry of the events, this number is consistent with predictions for the microlensing event rate in the PTF footprint over the survey's three years of operations, as estimated from the model proposed in \citet{han2008}.
+These candidates cannot be confirmed, as we lack simultaneous, multi-color photometry of the events.
+However, their number is consistent with predictions for the microlensing event rate in the PTF footprint over the survey's three years of operations, as estimated from near-field microlensing models.
 This work can help constrain all-sky event rate predictions and tests microlensing signal recovery in large data sets, both of which will be useful to future wide-field, time-domain surveys, such as that planned with the Large Synoptic Survey Telescope.
@@
 Over the past 20 years, microlensing, in which gravitational lensing causes a transient increase in the flux from a background point source, has been used to search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, to study Galactic structure and kinematics \citep{binney2000}, to determine the shape of stars \citep{rattenbury2005}, and to identify extrasolar planets \citep[][and references therein]{gaudi2011}.
-Such studies were once limited by the small number of detected events, but thanks to advances in CCD technology and the development of dedicated microlensing surveys, a few thousand events are now observed each year.\footnote{http://ogle.astrouw.edu.pl/ogle4/ews/ews.html} In the coming decade, this number is expected to increase as the next generation of photometric, time-domain surveys comes online, providing an opportunity for the precise study of otherwise hard-to-characterize objects, such as low-mass stars, sub-stellar objects, and isolated neutron stars.
+Such studies were once limited by the small number of detected events, but thanks to advances in CCD technology and the development of dedicated microlensing surveys, a few thousand events are now observed each year.\footnote{e.g., {\tt http://ogle.astrouw.edu.pl/ogle4/ews/ews.html}} In the coming decade, this number is expected to increase as the next generation of photometric, time-domain surveys comes online, providing an opportunity for the precise study of otherwise hard-to-characterize objects, such as low-mass stars, sub-stellar objects, and isolated neutron stars.
 
@@
 Indeed, lensing events away from these fields are potentially very interesting.
-While the microlensing event rate in low-density stellar regions will be smaller, compared to events found by modern microlensing surveys, these events are expected to involve closer sources and lenses, and less crowded backgrounds \citep{mesolensing}.
+While the microlensing event rate in low-density stellar regions will be smaller, compared to events found by modern microlensing surveys, these are expected to involve closer sources and lenses, and less crowded backgrounds \citep{mesolensing}.
 Thus, such lenses offer more opportunities to constrain the properties of the events.
@@
 
+% TODO: Han citation above, also mentioned in one of Rosanne's papers
+
 In this paper, we use the Palomar Transient Factory (PTF) survey, described in Section~\ref{sec:ptf}, as a test case for studying the frequency and detectability of microlensing events in all-sky, irregularly sampled, time-domain surveys.
-Although to date no microlensing event has been reported using PTF data, a simple extrapolation from previous all-sky rate estimates suggests $\sim$1--10 detectable microlensing events within the PTF footprint over the survey's three years of operation \citep{han2008, gaudi2008}.
+Although to date no microlensing event has been reported using PTF data, a simple extrapolation from previous all-sky rate estimates suggests $\sim$1--10 detectable microlensing events within the PTF footprint over the survey's three years of operation \citep{gaudi2008, han2008}.
 We examine the selection methods used by microlensing surveys to identify event signatures in light curves as well as an event identification procedure based on a set of variability statistics.
@@
 Figure~\ref{fig:sampling} shows six randomly selected light curves and illustrates the varying cadences and coverage that different fields may have over the same one-year period.
-The PTF data are not publicly available, but ...
-TODO
+While the PTF database is not yet public, we have made $10^5$ light curves available at {\tt }.
+These light curves meet the quality criteria described in Section~\ref{sec:detection_eff} and can be used to test the statistical methods described here (or others!).
 
@@
 
-Applying the above prescription to the PTF data presents obvious challenges.
+Applying the standard microlensing-search prescription to the PTF data presents obvious challenges.
 For example, microlensing surveys have relatively uniform time sampling of their survey footprint over an observing season, justifying the first requirement.
@@
 \centering\includegraphics[width=.48\textwidth, trim=10 50 50 50, clip]{fig5.eps}
-\caption{Light curves selected from PTF field 100101 with maximally outlying values for each variability index.
+\caption{Light curves selected from PTF field 3756 with maximally outlying values for each variability index.
 For $\eta$ and $K$, this corresponds to the light curve with the \emph{minimum} value of the index over the entire field.
 For the other indices, these are the light curves with the \emph{maximum} value of the relevant index over the field.
-These light curves represent the type of variability that each index is most sensitive to.
+These light curves illustrate the type of variability that each index is most sensitive to.
 } \label{fig:indices_examples}
@@
 We choose five statistical measures of variability compiled by \cite{shin2009} that have been previously applied to the classification and discovery of periodic variables.
-These variability indices are $\sigma/\mu$, $Con$, $\eta$, $J$, and $K$\footnote{We do not implement the sixth index described by \cite{shin2009}, $AoVM$, because it mainly helps identify periodic sources.}.
-$\sigma/\mu$ is the ratio of the sample standard deviation to the sample mean,
+These variability indices are $\sigma/\mu$, $Con$, $\eta$, $J$, and $K$.\footnote{We do not implement the sixth index described by \cite{shin2009}, $AoVM$, because it mainly helps identify periodic sources.} $\sigma/\mu$ is the ratio of the sample standard deviation ($\sigma$) to the sample mean,
 \begin{align}
@@
 
-$\eta$, the von Neumann ratio \citep[][also known as the Durbin-Watson statistic]{von_neumann1941}, is the mean square successive difference divided by the sample variance:
+$\eta$, the von Neumann ratio \citep[also known as the Durbin-Watson statistic;][]{von_neumann1941, durbin50}, is the mean square successive difference divided by the sample variance:
 \begin{align}
@@
 We add one more index, $\Delta \chi^2$, the difference in $\chi^2$ between fitting a Gaussian model and fitting a linear model to a light curve.
-In detail, we use a Levenberg-Marquardt optimizer to perform a least-squares fit with each of these models to the light curves and then compute $\Delta\chi^2=\chi_{linear} - \chi_{gaussian}$.
-Out tests with $\Delta \chi^2$ below compare the distribution of values over light curves on the same chip --- because of this, we neglect including an AIC or BIC correction to make this a proper likelihood test as the number of data points and number of model parameters are constant.
-For the Gaussian fit, the optimizer is initialized with a standard-deviation of 10 days, centered on the brightest data point.
+Specifically, we use a Levenberg-Marquardt optimizer to perform a least-squares fit with each of these models to the light curves and then compute $\Delta\chi^2=\chi_{linear} - \chi_{gaussian}$.
+Our tests with $\Delta \chi^2$ below compare the distribution of values over light curves on the same chip.
+As a result, the number of data points and number of model parameters are constant, and we therefore do not include an Akaike or Bayesian information criterion term.
+
+For the Gaussian fit, the optimizer is initialized with a $\sigma$ of 10 days, centered on the brightest data point.
 This is a standard statistical test used by microlensing surveys, and allows us to compare the relative performance of the (slightly modified) \cite{shin2009} indices and of this approach.
@@
 Note the sampling pattern.
-{\it Bottom:} Detection efficiency $\varepsilon$ for $\eta$, $\Delta\chi^2$, and $J$, as a function of the simulated microlensing event parameters $t_E$, $u_0$, and $m_0$ (the event timescale, impact parameter, and `quiescent' source magnitude, respectively).
-Black (dashed) lines show the distributions for all light curves (normalized, so scale of y-axis is arbitrary), red (solid) lines show the recovered distributions.}  \label{fig:detection_efficiency_4327}
+{\it Bottom:} Detection efficiency $\varepsilon$ for $\eta$, $\Delta\chi^2$, and $J$, as a function of the simulated microlensing event parameters $t_E$, $u_0$, and $m_0$ (the event timescale, impact parameter, and unmagnified source magnitude, respectively).
+Black (dashed) lines show the distributions for all light curves (normalized, so y-axis scale is arbitrary), red (solid) lines show the recovered distributions.}  \label{fig:detection_efficiency_4327}
 \end{figure}
@@
 	\item select a field from the 2097 fields with $>$10 good $R$-band observations and run Monte Carlo simulations to compute the 1\% FPR limiting values for $\eta$, $\Delta \chi^2$, and $J$ for that field;
-	\item compute $\eta$ for all the light curves in the field, and select the light curves that pass the quality cut described in Section~\ref{sec:detection_eff} ($\sim$10\%) and the FPR cut determined from the Monte Carlo simulations ($\sim$0.1-0.5\%);
-	\item for fields with SDSS coverage, we use Data Release 9 photometry \citep{dr9paper} to remove sources typed as Galaxy or QSO (using \texttt{objc\_type}\footnote{http://www.sdss3.org/dr9/algorithms/classify.php}) and flag candidate quasars \citep[using the cuts described in][]{richards02} ($\sim$10\%);
+	\item compute $\eta$ for all the light curves in the field.
+While $\sim$10\% of the light curves pass the quality cut described in Section~\ref{sec:detection_eff}, only $\sim$0.1-0.5\% of these pass the FPR cut determined from the Monte Carlo simulations;
+	\item for fields with SDSS coverage, we use Data Release 9 photometry \citep{dr9paper} to remove sources typed as Galaxy or QSO (using \texttt{objc\_type}\footnote{{\tt http://www.sdss3.org/dr9/algorithms/classify.php}}) and flag candidate quasars \citep[using the cuts described in][]{richards02} ($\sim$10\%);
 	\item fit a microlensing event model and subtract the model, then recompute $\eta$ for all the selected light curves;
@@
 We search SIMBAD\footnote{This research has made use of the SIMBAD database, operated at CDS, Strasbourg, France.} and remove any candidate among these with a known extragalactic counterpart within 10\asec.
-For those in the SDSS footprint, we examine the SDSS images \citep{sdss_images} to remove extended objects.
+For those in the SDSS footprint, we examine the SDSS images \citep{sdss_images} to identify extended objects.
 Removing extragalactic objects in this manner from our candidate list reduces contamination by, e.g., supernovae.
@@
 PTF1J0618 and PTF1J1728 have counterparts in 2MASS, and we can therefore use the \citet{kev07} $J$-band absolute magnitudes for stars of the relevant spectral types to estimate rough distances to these stars, assuming they have solar metallicity (see below for discussion of how we obtain spectral types for these stars).
-For PTF1J0618, this results in a distance that ranges from 1.1 kpc (if it is a M1 star) to 0.8 kpc (if it is a M3 star), corresponding to heights above the Galactic Plane 43 and 31 pc, respectively.
+For PTF1J0618, this results in a distance that ranges from 1.1 kpc (if it is a M1 star) to 0.8 kpc (if it is a M3 star), corresponding to heights above the Galactic Plane of 43 and 31 pc, respectively.
 For PTF1J1728, assuming it is a K0-K5 star results in distances ranging from 1.3 to 0.7 kpc, and heights of 0.9 to 0.5~kpc.
@@
 We lack a spectrum for PTF1J1615, so the distance estimate to this star is even more uncertain.
-Still, we compare the PTF $R$ magnitude to the stellar library compiled by \citet{Pickles1998} to estimate a distance, assuming (from its photometry) that it is a K5-M2 star\footnote{This ignores differences between Mould and Cousins $R$ filters.}.
-This results in distances ranging from 4.2 to 1.9 kpc and heights from 2.3 to 1.0~kpc.
+Still, we compare the PTF $R$ magnitude to the stellar library compiled by \citet{Pickles1998} to estimate a distance, assuming (from its photometry) that it is a K5-M2 star.\footnote{This ignores differences between Mould and Cousins $R$ filters.} This results in distances ranging from 4.2 to 1.9 kpc and heights from 2.3 to 1.0~kpc.
 The estimated heights above the Galactic Plane suggest that PTF1J0618 is a thin disk star, while the other two, PTF1J1615 and PT1J1728, are plausible thick disk members \citep{Bochanski2010}.
@@
 From 6300-8800~\AA, the grating was 158 line~mm$^{-1}$, was blazed at 7500~\AA, and gave a resolution of 2.5~\AA.
-The data were reduced using standard IRAF routines\footnote{IRAF is distributed by the National Optical Astronomy Observatories, which are operated by the Association of Universities for Research in Astronomy, Inc., under cooperative agreement with the National Science Foundation.}.
+The data were reduced using standard IRAF routines.\footnote{IRAF is distributed by the National Optical Astronomy Observatories, which are operated by the Association of Universities for Research in Astronomy, Inc., under cooperative agreement with the National Science Foundation.}
 
@@
 We further note that those outbursts typically last $\sim$years \citep[e.g.,][]{tomov2013}.
-We conclude that this is a good candidate microlensing event.
+We conclude that this is a plausible candidate microlensing event.
 
@@
 PTF1J1615 is one of two candidates that fall in the SDSS footprint; its counterpart, SDSS J161502.43$+$540053.6, has a proper motion of 2.66 mas/year, consistent with its automated SDSS classification as a star.
-The star has $u = 22.9$ mag\footnote{These are SDSS PSF magnitudes.
-Typically, PSF fitting provides better estimates of isolated star magnitudes; see \citet{stoughton02}.}, beyond the $u = 22.0$ 95\% completeness limit for SDSS \citep{stoughton02}, and we therefore focus on its $griz$ photometry.
+The star has $u = 22.9$ mag,\footnote{These are SDSS PSF magnitudes.
+Typically, PSF fitting provides better estimates of isolated star magnitudes; see \citet{stoughton02}.} beyond the $u = 22.0$ 95\% completeness limit for SDSS \citep{stoughton02}, and we therefore focus on its $griz$ photometry.
 With $(g-r) = 1.10\pm0.05$, $(r-i) = 1.00\pm0.03$, and $(i-z) = 0.56\pm0.04$, the star's colors are consistent with those of a late K/early M star \citep{kev07}.
@@
 However, the event detected by PTF had a duration $\sim$20 days, and so is highly unlikely to have been a flare.
-We conclude that this is a good candidate microlensing event.
+We conclude that this is a plausible candidate microlensing event.
 
@@
 
-E.~Bowsher and S.~Douglas obtained two 1200-s spectra of PTF1J1728 with the MDM Observatory Modular Spectrograph (MODspec) on the Hiltner 2.4-m telescope at MDM Observatory, Kitt Peak, AZ, on 2012 Nov 14\footnote{The MDM Observatory is operated by Dartmouth College, Columbia University, Ohio State University, Ohio University, and the University of Michigan.}.
-MODspec was configured to provide coverage from 4500 to 7500 \AA~with $\sim$1.8 \AA~sampling and a spectral resolution of $\sim$3300.
+E.~Bowsher and S.~Douglas obtained two 1200-s spectra of PTF1J1728 with the MDM Observatory Modular Spectrograph (MODspec) on the Hiltner 2.4-m telescope at MDM Observatory, Kitt Peak, AZ, on 2012 Nov 14.\footnote{The MDM Observatory is operated by Dartmouth College, Columbia University, Ohio State University, Ohio University, and the University of Michigan.} MODspec was configured to provide coverage from 4500 to 7500 \AA~with $\sim$1.8 \AA~sampling and a spectral resolution of $\sim$3300.
 The data reduction was performed by S.~Douglas using standard IRAF routines.
@@
 While the spectrum is noisy, PTF1J1728 appears to be a mid-K star; its HAMMER spectral type is K3.
-We conclude that this is a good candidate microlensing event.
+We conclude that this is a plausible candidate microlensing event.
 
@@
 Because it is only detected in the $W1$ and $W2$ bands, the WISE color does not help much: it is consistent with that of stars and of compact galaxies \citep[cf.\ Figure 14 in][]{yan2013}.
-We tentatively conclude that this not a good microlensing candidate.
+We tentatively conclude that this not a plausible microlensing candidate.
 
@@
 The $(W1-W2)$ color places the object well away from the stellar locus \citep[cf.\ Figure 14 in][]{yan2013}.
-We tentatively conclude that this is a likely extragalactic source and therefore not a good microlensing candidate.
+We tentatively conclude that this is a likely extragalactic source and therefore not a plausible microlensing candidate.
 
@@
 Its $(W1-W2)$ and $(W2-W3)$ colors are consistent with this object being a QSO \citep[cf.\ Figure 14 in][]{yan2013}.
-We tentatively conclude that this not a good microlensing candidate.
+We tentatively conclude that this not a plausible microlensing candidate.
+
 \subsubsection*{PTF1J174736.51$+$300506.9} % (g)}
@@
 %Visual comparison of the PTF and WISE images suggests that there may be some blending of objects in the WISE data, however the WISE photometric quality flags are excellent for each band and suggest that the source is not blended.
-We tentatively conclude that this not a good microlensing candidate.
+We tentatively conclude that this not a plausible microlensing candidate.
 
@@
 This interpretation is strengthened by a visual inspection of the PTF and WISE images, which show that the object is elongated in the Northeast-Southwest direction.
-We tentatively conclude that this not a good microlensing candidate.
+We tentatively conclude that this not a plausible microlensing candidate.
 
@@
 
-We found that the von Neumann ratio, $\eta$ \citep{von_neumann1941}, performs better than previously used statistics in recovering injected microlensing events in non-uniformly sampled data.
+We found that the von Neumann ratio, $\eta$ \citep[or Durbin-Watson statistic;][]{von_neumann1941, durbin50}, performs better than previously used statistics in recovering injected microlensing events in non-uniformly sampled data.
 We used $\eta$ to develop a selection procedure for extracting microlensing event candidates from the PTF light-curve archive.
@@
 \includegraphics[width=2.1\columnwidth] {fig14_2.eps}
-\caption{Full light curves (top) and zooms around the transient maximum (bottom) for the rejected microlensing event candidates. Bad (flagged) data points have been removed.}\label{fig:not_candidates}
+\caption{Full light curves (top) and zooms around the transient maximum (bottom) for the rejected microlensing event candidates. Flagged data points have been removed.}\label{fig:not_candidates}
 \end{figure*}
@@
 Rigorously classifying and cataloging these sources was outside of the scope of this paper, but many  presented interesting forms of variability.
-Though it is not necessarily scalable for future surveys to look at all light curves selected by using our algorithm, these $\sim$2000 light curves, once carefully classified, could serve as a training set for a more sophisticated machine-learning approach \cite[as discussed by, e.g.,][]{bloomPasp} to classify the light curves that survive our selection criteria.
+Though it is not necessarily scalable for future surveys to look at all light curves selected by using our algorithm, these $\sim$2000 light curves, once carefully classified, could serve as a training set for a more sophisticated machine-learning approach \cite[e.g.,][]{bloomPasp,brink2013} to classify the light curves that survive our selection criteria.
 Approaches that combine statistical cuts such as those described in this paper and some type of machine learning will be essential to making the most of the future time-domain data such as that collected by the Large Synoptic Survey Telescope.
@@
 We thank E.~Bowsher, S.~Douglas, and S.~Tang for their assistance in obtaining spectra for our candidates, and M.~Modjaz for help in interpreting several of these spectra.
+We are grateful to the anonymous referee and to our editor, Eric Feigelson, for comments that improved the paper.
 
@@
 %\setlength{\baselineskip}{0.6\baselineskip}
-%\bibliography{references_3}
+\bibliography{references_3}
 %\setlength{\baselineskip}{1.667\baselineskip}
 
-\begin{thebibliography}{53}
-\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
-
-\bibitem[{{Ag{\"u}eros} {et~al.}(2011){Ag{\"u}eros}, {Covey}, {Lemonias},
-  {Law}, {Kraus}, {Batalha}, {Bloom}, {Cenko}, {Kasliwal}, {Kulkarni},
-  {Nugent}, {Ofek}, {Poznanski}, \& {Quimby}}]{agueros11}
-{Ag{\"u}eros}, M.~A., {Covey}, K.~R., {Lemonias}, J.~J., {et~al.} 2011, \apj,
-  740, 110
-
-\bibitem[{{Ahn} {et~al.}(2012){Ahn}, {Alexandroff}, {Allende Prieto},
-  {Anderson}, {Anderton}, {Andrews}, {Aubourg}, {Bailey}, {Balbinot}, {Barnes},
-  \& et~al.}]{dr9paper}
-{Ahn}, C.~P., {Alexandroff}, R., {Allende Prieto}, C., {et~al.} 2012, \apjs,
-  203, 21
-
-\bibitem[{{Alard} \& {Lupton}(1998)}]{alard1998}
-{Alard}, C., \& {Lupton}, R.~H. 1998, \apj, 503, 325
-
-\bibitem[{{Alcock} {et~al.}(1993){Alcock}, {Allsman}, {Axelrod}, {Bennett},
-  {Cook}, {Park}, {Marshall}, {Stubbs}, {Griest}, {Perlmutter}, {Sutherland},
-  {Freeman}, {Peterson}, {Quinn}, \& {Rodgers}}]{original_macho}
-{Alcock}, C., {Allsman}, R.~A., {Axelrod}, T.~S., {et~al.} 1993, in
-  Astronomical Society of the Pacific Conference Series, Vol.~43, Sky Surveys.
-  Protostars to Protogalaxies, ed. {B.~T.~Soifer}, 291
-
-\bibitem[{{Alcock} {et~al.}(2000){Alcock}, {Allsman}, {Alves}, {Axelrod},
-  {Becker}, {Bennett}, {Cook}, {Drake}, {Freeman}, {Geha}, {Griest}, {Lehner},
-  {Marshall}, {Minniti}, {Nelson}, {Peterson}, {Popowski}, {Pratt}, {Quinn},
-  {Stubbs}, {Sutherland}, {Tomaney}, {Vandehei}, {Welch}, \& {MACHO
-  Collaboration}}]{alcock2000}
-{Alcock}, C., {Allsman}, R.~A., {Alves}, D.~R., {et~al.} 2000, \apj, 541, 734
-
-\bibitem[{{Alcock} {et~al.}(2001){Alcock}, {Allsman}, {Alves}, {Axelrod},
-  {Becker}, {Bennett}, {Cook}, {Drake}, {Freeman}, {Geha}, {Griest}, {Lehner},
-  {Marshall}, {Minniti}, {Nelson}, {Peterson}, {Popowski}, {Pratt}, {Quinn},
-  {Stubbs}, {Sutherland}, {Tomaney}, {Vandehei}, \&
-  {Welch}}]{macho_detection_efficiency}
-{Alcock}, C., {Allsman}, R.~A., {Alves}, D., {et~al.} 2001, \apjs, 136, 439
-
-\bibitem[{{Aubourg} {et~al.}(1993){Aubourg}, {Bareyre}, {Brehin}, {Gros},
-  {Lachieze-Rey}, {Laurent}, {Lesquoy}, {Magneville}, {Milsztajn}, {Moscoso},
-  {Queinnec}, {Rich}, {Spiro}, {Vigroux}, {Zylberajch}, {Ansari}, {Cavalier},
-  {Moniez}, {Beaulieu}, {Ferlet}, {Grison}, {Vidal-Madjar}, {Guibert},
-  {Moreau}, {Tajahmady}, {Maurice}, {Prevot}, \& {Gry}}]{eros_original}
-{Aubourg}, E., {Bareyre}, P., {Brehin}, S., {et~al.} 1993, The Messenger, 72,
-  20
-
-\bibitem[{{Bertin} \& {Arnouts}(1996)}]{bertin96}
-{Bertin}, E., \& {Arnouts}, S. 1996, \aaps, 117, 393
-
-\bibitem[{{Binney} {et~al.}(2000){Binney}, {Bissantz}, \&
-  {Gerhard}}]{binney2000}
-{Binney}, J., {Bissantz}, N., \& {Gerhard}, O. 2000, \apjl, 537, L99
-
-\bibitem[{{Bloom} {et~al.}(2012){Bloom}, {Richards}, {Nugent}, {Quimby},
-  {Kasliwal}, {Starr}, {Poznanski}, {Ofek}, {Cenko}, {Butler}, {Kulkarni},
-  {Gal-Yam}, \& {Law}}]{bloomPasp}
-{Bloom}, J.~S., {Richards}, J.~W., {Nugent}, P.~E., {et~al.} 2012, \pasp, 124,
-  1175
-
-\bibitem[{{Bochanski} {et~al.}(2010){Bochanski}, {Hawley}, {Covey}, {West},
-  {Reid}, {Golimowski}, \& {Ivezi{\'c}}}]{Bochanski2010}
-{Bochanski}, J.~J., {Hawley}, S.~L., {Covey}, K.~R., {et~al.} 2010, \aj, 139,
-  2679
-
-\bibitem[{{Covey}(2006)}]{kev06}
-{Covey}, K.~R. 2006, PhD thesis, University of Washington, Washington, USA
-
-\bibitem[{{Covey} {et~al.}(2007){Covey}, {Ivezi{\'c}}, {Schlegel},
-  {Finkbeiner}, {Padmanabhan}, {Lupton}, {Ag{\"u}eros}, {Bochanski}, {Hawley},
-  {West}, {Seth}, {Kimball}, {Gogarten}, {Claire}, {Haggard}, {Kaib},
-  {Schneider}, \& {Sesar}}]{kev07}
-{Covey}, K.~R., {Ivezi{\'c}}, {\v Z}., {Schlegel}, D., {et~al.} 2007, \aj, 134,
-  2398
-
-\bibitem[{{Crotts} \& {Tomaney}(1996)}]{crotts1996}
-{Crotts}, A.~P.~S., \& {Tomaney}, A.~B. 1996, \apjl, 473, L87
-
-\bibitem[{{Cutri} {et~al.}(2003)}]{cutri03}
-{Cutri}, R.~M., {et~al.} 2003, {2MASS All Sky Catalog of point sources.} (The
-  IRSA 2MASS All-Sky Point Source Catalog, NASA/IPAC Infrared Science
-  Archive.~http://irsa.ipac.caltech.edu/applications/Gator/)
-
-\bibitem[{{Di Stefano}(2008)}]{mesolensing}
-{Di Stefano}, R. 2008, \apj, 684, 46
-
-\bibitem[{{Foreman-Mackey} {et~al.}(2012){Foreman-Mackey}, {Hogg}, {Lang}, \&
-  {Goodman}}]{dfm}
-{Foreman-Mackey}, D., {Hogg}, D.~W., {Lang}, D., \& {Goodman}, J. 2012, ArXiv
-  e-prints
-
-\bibitem[{{Fukui} {et~al.}(2007){Fukui}, {Abe}, {Ayani}, {Fujii}, {Iizuka},
-  {Itow}, {Kabumoto}, {Kamiya}, {Kawabata}, {Kawanomoto}, {Kinugasa}, {Koff},
-  {Krajci}, {Naito}, {Nogami}, {Narusawa}, {Ohishi}, {Ohnishi}, {Sumi}, \&
-  {Tsumuraya}}]{fukui2007}
-{Fukui}, A., {Abe}, F., {Ayani}, K., {et~al.} 2007, \apj, 670, 423
-
-\bibitem[{{Gaudi}(2011)}]{gaudi2011}
-{Gaudi}, B.~S. 2011, {Microlensing by Exoplanets}, ed. S.~{Piper}, 79--110
-
-\bibitem[{{Gaudi} {et~al.}(2008){Gaudi}, {Patterson}, {Spiegel}, {Krajci},
-  {Koff}, {Pojma{\'n}ski}, {Dong}, {Gould}, {Prieto}, {Blake}, {Roming},
-  {Bennett}, {Bloom}, {Boyd}, {Eyler}, {de Ponthi{\`e}re}, {Mirabal}, {Morgan},
-  {Remillard}, {Vanmunster}, {Wagner}, \& {Watson}}]{gaudi2008}
-{Gaudi}, B.~S., {Patterson}, J., {Spiegel}, D.~S., {et~al.} 2008, \apj, 677,
-  1268
-
-\bibitem[{{Goodman} \& {Weare}(2010)}]{goodman}
-{Goodman}, J., \& {Weare}, J. 2010, Comm. App. Math. Comp. Sci., 5
-
-\bibitem[{{Hamadache} {et~al.}(2006){Hamadache}, {Le Guillou}, {Tisserand},
-  {Afonso}, {Albert}, {Andersen}, {Ansari}, {Aubourg}, {Bareyre}, {Beaulieu},
-  {Charlot}, {Coutures}, {Ferlet}, {Fouqu{\'e}}, {Glicenstein}, {Goldman},
-  {Gould}, {Graff}, {Gros}, {Haissinski}, {de Kat}, {Lesquoy}, {Loup},
-  {Magneville}, {Marquette}, {Maurice}, {Maury}, {Milsztajn}, {Moniez},
-  {Palanque-Delabrouille}, {Perdereau}, {Rahal}, {Rich}, {Spiro},
-  {Vidal-Madjar}, {Vigroux}, \& {Zylberajch}}]{hamadache2009}
-{Hamadache}, C., {Le Guillou}, L., {Tisserand}, P., {et~al.} 2006, \aap, 454,
-  185
-
-\bibitem[{{Han}(2008)}]{han2008}
-{Han}, C. 2008, \apj, 681, 806
-
-\bibitem[{{Kowalski} {et~al.}(2010){Kowalski}, {Hawley}, {Holtzman},
-  {Wisniewski}, \& {Hilton}}]{kowal2010}
-{Kowalski}, A.~F., {Hawley}, S.~L., {Holtzman}, J.~A., {Wisniewski}, J.~P., \&
-  {Hilton}, E.~J. 2010, \apjl, 714, L98
-
-\bibitem[{{Law} {et~al.}(2009){Law}, {Kulkarni}, {Dekany}, {Ofek}, {Quimby},
-  {Nugent}, {Surace}, {Grillmair}, {Bloom}, {Kasliwal}, {Bildsten}, {Brown},
-  {Cenko}, {Ciardi}, {Croner}, {Djorgovski}, {van Eyken}, {Filippenko}, {Fox},
-  {Gal-Yam}, {Hale}, {Hamam}, {Helou}, {Henning}, {Howell}, {Jacobsen},
-  {Laher}, {Mattingly}, {McKenna}, {Pickles}, {Poznanski}, {Rahmer}, {Rau},
-  {Rosing}, {Shara}, {Smith}, {Starr}, {Sullivan}, {Velur}, {Walters}, \&
-  {Zolkower}}]{nick2009}
-{Law}, N.~M., {Kulkarni}, S.~R., {Dekany}, R.~G., {et~al.} 2009, \pasp, 121,
-  1395
-
-\bibitem[{{Law} {et~al.}(2010){Law}, {Dekany}, {Rahmer}, {Hale}, {Smith},
-  {Quimby}, {Ofek}, {Kasliwal}, {Zolkower}, {Velur}, {Henning}, {Bui},
-  {McKenna}, {Nugent}, {Jacobsen}, {Walters}, {Bloom}, {Surace}, {Grillmair},
-  {Laher}, {Mattingly}, \& {Kulkarni}}]{nick2010}
-{Law}, N.~M., {Dekany}, R.~G., {Rahmer}, G., {et~al.} 2010, in Society of
-  Photo-Optical Instrumentation Engineers (SPIE) Conference Series, Vol. 7735
-
-\bibitem[{{Levitan} {et~al.}(2011){Levitan}, {Fulton}, {Groot}, {Kulkarni},
-  {Ofek}, {Prince}, {Shporer}, {Bloom}, {Cenko}, {Kasliwal}, {Law}, {Nugent},
-  {Poznanski}, {Quimby}, {Horesh}, {Sesar}, \& {Sternberg}}]{levitan2011}
-{Levitan}, D., {Fulton}, B.~J., {Groot}, P.~J., {et~al.} 2011, \apj, 739, 68
-
-\bibitem[{{Nieto--Santisteban} {et~al.}(2004)}]{sdss_images}
-{Nieto--Santisteban}, M.~A., {et~al.} 2004, in ASP Conf. Ser. 314: Astronomical
-  Data Analysis Software and Systems (ADASS) XIII, ed. F.~{Ochsenbein}, M.~G.
-  {Allen}, \& D.~{Egret}, 666--+
-
-\bibitem[{{Ofek} {et~al.}(2012){Ofek}, {Laher}, {Surace}, {Levitan}, {Sesar},
-  {Horesh}, {Law}, {van Eyken}, {Kulkarni}, {Prince}, {Nugent}, {Sullivan},
-  {Yaron}, {Pickles}, {Ag{\"u}eros}, {Arcavi}, {Bildsten}, {Bloom}, {Cenko},
-  {Gal-Yam}, {Grillmair}, {Helou}, {Kasliwal}, {Poznanski}, \&
-  {Quimby}}]{ofek2012}
-{Ofek}, E.~O., {Laher}, R., {Surace}, J., {et~al.} 2012, \pasp, 124, 854
-
-\bibitem[{{Oke} \& {Gunn}(1982)}]{oke_gunn}
-{Oke}, J.~B., \& {Gunn}, J.~E. 1982, \pasp, 94, 586
-
-\bibitem[{{Os{\l}owski} {et~al.}(2008){Os{\l}owski}, {Moderski}, {Bulik}, \&
-  {Belczynski}}]{oslowski2008}
-{Os{\l}owski}, S., {Moderski}, R., {Bulik}, T., \& {Belczynski}, K. 2008, \aap,
-  478, 429
-
-\bibitem[{{Paczynski}(1986)}]{paczynski1986}
-{Paczynski}, B. 1986, \apj, 304, 1
-
-\bibitem[{{Pickles}(1998)}]{Pickles1998}
-{Pickles}, A.~J. 1998, \pasp, 110, 863
-
-\bibitem[{{Polishook} {et~al.}(2012){Polishook}, {Ofek}, {Waszczak},
-  {Kulkarni}, {Gal-Yam}, {Aharonson}, {Laher}, {Surace}, {Klein}, {Bloom},
-  {Brosch}, {Prialnik}, {Grillmair}, {Cenko}, {Kasliwal}, {Law}, {Levitan},
-  {Nugent}, {Poznanski}, \& {Quimby}}]{polishook2012}
-{Polishook}, D., {Ofek}, E.~O., {Waszczak}, A., {et~al.} 2012, \mnras, 421,
-  2094
-
-\bibitem[{{Rahmer} {et~al.}(2008){Rahmer}, {Smith}, {Velur}, {Hale}, {Law},
-  {Bui}, {Petrie}, \& {Dekany}}]{rahmer2008}
-{Rahmer}, G., {Smith}, R., {Velur}, V., {et~al.} 2008, in Society of
-  Photo-Optical Instrumentation Engineers (SPIE) Conference Series, Vol. 7014
-
-\bibitem[{{Rattenbury} {et~al.}(2005){Rattenbury}, {Abe}, {Bennett}, {Bond},
-  {Calitz}, {Claret}, {Cook}, {Furuta}, {Gal-Yam}, {Glicenstein}, {Hearnshaw},
-  {Hauschildt}, {Kilmartin}, {Kurata}, {Masuda}, {Maoz}, {Matsubara},
-  {Meintjes}, {Moniez}, {Muraki}, {Noda}, {Ofek}, {Okajima}, {Philpott},
-  {Rhie}, {Sako}, {Sullivan}, {Sumi}, {Terndrup}, {Tristram}, {Wood},
-  {Yanagisawa}, \& {Yock}}]{rattenbury2005}
-{Rattenbury}, N.~J., {Abe}, F., {Bennett}, D.~P., {et~al.} 2005, \aap, 439, 645
-
-\bibitem[{{Rau} {et~al.}(2009){Rau}, {Kulkarni}, {Law}, {Bloom}, {Ciardi},
-  {Djorgovski}, {Fox}, {Gal-Yam}, {Grillmair}, {Kasliwal}, {Nugent}, {Ofek},
-  {Quimby}, {Reach}, {Shara}, {Bildsten}, {Cenko}, {Drake}, {Filippenko},
-  {Helfand}, {Helou}, {Howell}, {Poznanski}, \& {Sullivan}}]{rau2009}
-{Rau}, A., {Kulkarni}, S.~R., {Law}, N.~M., {et~al.} 2009, \pasp, 121, 1334
-
-\bibitem[{{Richards} {et~al.}(2002)}]{richards02}
-{Richards}, G.~T., {et~al.} 2002, \aj, 123, 2945
-
-\bibitem[{{Sartore} \& {Treves}(2010)}]{sartore2010}
-{Sartore}, N., \& {Treves}, A. 2010, \aap, 523, A33
-
-\bibitem[{{Shin} {et~al.}(2009){Shin}, {Sekora}, \& {Byun}}]{shin2009}
-{Shin}, M.-S., {Sekora}, M., \& {Byun}, Y.-I. 2009, \mnras, 400, 1897
-
-\bibitem[{{Stetson}(1996)}]{stetson1996}
-{Stetson}, P.~B. 1996, \pasp, 108, 851
-
-\bibitem[{{Stoughton} {et~al.}(2002){Stoughton}, {Lupton}, {Bernardi},
-  {Blanton}, {Burles}, {Castander}, {Connolly}, {Eisenstein}, {Frieman},
-  {Hennessy}, {Hindsley}, {Ivezi{\'c}}, {Kent}, {Kunszt}, {Lee}, {Meiksin},
-  {Munn}, {Newberg}, {Nichol}, {Nicinski}, {Pier}, {Richards}, {Richmond},
-  {Schlegel}, {Smith}, {Strauss}, {SubbaRao}, {Szalay}, {Thakar}, {Tucker},
-  {Vanden Berk}, {Yanny}, {Adelman}, {Anderson}, {Anderson}, {Annis},
-  {Bahcall}, {Bakken}, {Bartelmann}, {Bastian}, {Bauer}, {Berman},
-  {B{\"o}hringer}, {Boroski}, {Bracker}, {Briegel}, {Briggs}, {Brinkmann},
-  {Brunner}, {Carey}, {Carr}, {Chen}, {Christian}, {Colestock}, {Crocker},
-  {Csabai}, {Czarapata}, {Dalcanton}, {Davidsen}, {Davis}, {Dehnen},
-  {Dodelson}, {Doi}, {Dombeck}, {Donahue}, {Ellman}, {Elms}, {Evans}, {Eyer},
-  {Fan}, {Federwitz}, {Friedman}, {Fukugita}, {Gal}, {Gillespie}, {Glazebrook},
-  {Gray}, {Grebel}, {Greenawalt}, {Greene}, {Gunn}, {de Haas}, {Haiman},
-  {Haldeman}, {Hall}, {Hamabe}, {Hansen}, {Harris}, {Harris}, {Harvanek},
-  {Hawley}, {Hayes}, {Heckman}, {Helmi}, {Henden}, {Hogan}, {Hogg}, {Holmgren},
-  {Holtzman}, {Huang}, {Hull}, {Ichikawa}, {Ichikawa}, {Johnston}, {Kauffmann},
-  {Kim}, {Kimball}, {Kinney}, {Klaene}, {Kleinman}, {Klypin}, {Knapp},
-  {Korienek}, {Krolik}, {Kron}, {Krzesi{\'n}ski}, {Lamb}, {Leger},
-  {Limmongkol}, {Lindenmeyer}, {Long}, {Loomis}, {Loveday}, {MacKinnon},
-  {Mannery}, {Mantsch}, {Margon}, {McGehee}, {McKay}, {McLean}, {Menou},
-  {Merelli}, {Mo}, {Monet}, {Nakamura}, {Narayanan}, {Nash}, {Neilsen},
-  {Newman}, {Nitta}, {Odenkirchen}, {Okada}, {Okamura}, {Ostriker}, {Owen},
-  {Pauls}, {Peoples}, {Peterson}, {Petravick}, {Pope}, {Pordes}, {Postman},
-  {Prosapio}, {Quinn}, {Rechenmacher}, {Rivetta}, {Rix}, {Rockosi}, {Rosner},
-  {Ruthmansdorfer}, {Sandford}, {Schneider}, {Scranton}, {Sekiguchi}, {Sergey},
-  {Sheth}, {Shimasaku}, {Smee}, {Snedden}, {Stebbins}, {Stubbs}, {Szapudi},
-  {Szkody}, {Szokoly}, {Tabachnik}, {Tsvetanov}, {Uomoto}, {Vogeley}, {Voges},
-  {Waddell}, {Walterbos}, {Wang}, {Watanabe}, {Weinberg}, {White}, {White},
-  {Wilhite}, {Wolfe}, {Yasuda}, {York}, {Zehavi}, \& {Zheng}}]{stoughton02}
-{Stoughton}, C., {Lupton}, R.~H., {Bernardi}, M., {et~al.} 2002, \aj, 123, 485
-
-\bibitem[{{Sumi} {et~al.}(2011){Sumi}, {Kamiya}, {Bennett}, {Bond}, {Abe},
-  {Botzler}, {Fukui}, {Furusawa}, {Hearnshaw}, {Itow}, {Kilmartin}, {Korpela},
-  {Lin}, {Ling}, {Masuda}, {Matsubara}, {Miyake}, {Motomura}, {Muraki},
-  {Nagaya}, {Nakamura}, {Ohnishi}, {Okumura}, {Perrott}, {Rattenbury}, {Saito},
-  {Sako}, {Sullivan}, {Sweatman}, {Tristram}, {Udalski}, {Szyma{\'n}ski},
-  {Kubiak}, {Pietrzy{\'n}ski}, {Poleski}, {Soszy{\'n}ski}, {Wyrzykowski},
-  {Ulaczyk}, \& {Microlensing Observations in Astrophysics (MOA)
-  Collaboration}}]{sumi2011}
-{Sumi}, T., {Kamiya}, K., {Bennett}, D.~P., {et~al.} 2011, \nat, 473, 349
-
-\bibitem[{{Tomov} {et~al.}(2013){Tomov}, {Tomova}, \& {Bisikalo}}]{tomov2013}
-{Tomov}, N.~A., {Tomova}, M.~T., \& {Bisikalo}, D.~V. 2013, ArXiv e-prints
-
-\bibitem[{{Udalski}(2003)}]{udalski03}
-{Udalski}, A. 2003, \actaa, 53, 291
-
-\bibitem[{{Udalski} {et~al.}(1992){Udalski}, {Szymanski}, {Kaluzny}, {Kubiak},
-  \& {Mateo}}]{original_ogle}
-{Udalski}, A., {Szymanski}, M., {Kaluzny}, J., {Kubiak}, M., \& {Mateo}, M.
-  1992, \actaa, 42, 253
-
-\bibitem[{{Udalski} {et~al.}(1994){Udalski}, {Szymanski}, {Stanek}, {Kaluzny},
-  {Kubiak}, {Mateo}, {Krzeminski}, {Paczynski}, \&
-  {Venkat}}]{ogle_optical_depth}
-{Udalski}, A., {Szymanski}, M., {Stanek}, K.~Z., {et~al.} 1994, \actaa, 44, 165
-
-\bibitem[{{von Neumann} {et~al.}(1941){von Neumann}, {Kent}, {Bellinson}, \&
-  {Hart}}]{von_neumann1941}
-{von Neumann}, J., {Kent}, R.~H., {Bellinson}, H.~R., \& {Hart}, B.~I. 1941,
-  The Annals of Mathematical Statistics, 12, 153
-
-\bibitem[{{Wozniak}(2000)}]{con_idx}
-{Wozniak}, P.~R. 2000, \actaa, 50, 421
-
-\bibitem[{{Wright} {et~al.}(2010){Wright}, {Eisenhardt}, {Mainzer}, {Ressler},
-  {Cutri}, {Jarrett}, {Kirkpatrick}, {Padgett}, {McMillan}, {Skrutskie},
-  {Stanford}, {Cohen}, {Walker}, {Mather}, {Leisawitz}, {Gautier}, {McLean},
-  {Benford}, {Lonsdale}, {Blain}, {Mendez}, {Irace}, {Duval}, {Liu}, {Royer},
-  {Heinrichsen}, {Howard}, {Shannon}, {Kendall}, {Walsh}, {Larsen}, {Cardon},
-  {Schick}, {Schwalm}, {Abid}, {Fabinsky}, {Naes}, \& {Tsai}}]{WISE}
-{Wright}, E.~L., {Eisenhardt}, P.~R.~M., {Mainzer}, A.~K., {et~al.} 2010, \aj,
-  140, 1868
-
-\bibitem[{{Wyrzykowski} {et~al.}(2009){Wyrzykowski}, {Koz{\l}owski}, {Skowron},
-  {Belokurov}, {Smith}, {Udalski}, {Szyma{\'n}ski}, {Kubiak},
-  {Pietrzy{\'n}ski}, {Soszy{\'n}ski}, {Szewczyk}, \&
-  {{\.Z}ebru{\'n}}}]{wyrzykowski2009}
-{Wyrzykowski}, {\L}., {Koz{\l}owski}, S., {Skowron}, J., {et~al.} 2009, \mnras,
-  397, 1228
-
-\bibitem[{Yan {et~al.}(2013)Yan, Donoso, Tsai, Stern, Assef, Eisenhardt, Blain,
-  Cutri, Jarrett, Stanford, Wright, Bridge, \& Riechers}]{yan2013}
-Yan, L., Donoso, E., Tsai, C.-W., {et~al.} 2013, \aj, 145, 55
-
-\bibitem[{{York} {et~al.}(2000)}]{york00}
-{York}, D.~G., {et~al.} 2000, \aj, 120, 1579
-
-\end{thebibliography}
+%\begin{thebibliography}{54}
+%\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+%
+%\bibitem[{{Ag{\"u}eros} {et~al.}(2011){Ag{\"u}eros}, {Covey}, {Lemonias},
+%  {Law}, {Kraus}, {Batalha}, {Bloom}, {Cenko}, {Kasliwal}, {Kulkarni},
+%  {Nugent}, {Ofek}, {Poznanski}, \& {Quimby}}]{agueros11}
+%{Ag{\"u}eros}, M.~A., {Covey}, K.~R., {Lemonias}, J.~J., {et~al.} 2011, \apj,
+%  740, 110
+%  
+%\bibitem[{{Ahn} {et~al.}(2012){Ahn}, {Alexandroff}, {Allende Prieto},
+%  {Anderson}, {Anderton}, {Andrews}, {Aubourg}, {Bailey}, {Balbinot}, {Barnes},
+%  \& et~al.}]{dr9paper}
+%{Ahn}, C.~P., {Alexandroff}, R., {Allende Prieto}, C., {et~al.} 2012, \apjs,
+%  203, 21
+%
+%\bibitem[{{Alard} \& {Lupton}(1998)}]{alard1998}
+%{Alard}, C., \& {Lupton}, R.~H. 1998, \apj, 503, 325
+%
+%\bibitem[{{Alcock} {et~al.}(1993){Alcock}, {Allsman}, {Axelrod}, {Bennett},
+%  {Cook}, {Park}, {Marshall}, {Stubbs}, {Griest}, {Perlmutter}, {Sutherland},
+%  {Freeman}, {Peterson}, {Quinn}, \& {Rodgers}}]{original_macho}
+%{Alcock}, C., {Allsman}, R.~A., {Axelrod}, T.~S., {et~al.} 1993, in
+%  Astronomical Society of the Pacific Conference Series, Vol.~43, Sky Surveys.
+%  Protostars to Protogalaxies, ed. {B.~T.~Soifer}, 291
+%
+%\bibitem[{{Alcock} {et~al.}(2000){Alcock}, {Allsman}, {Alves}, {Axelrod},
+%  {Becker}, {Bennett}, {Cook}, {Drake}, {Freeman}, {Geha}, {Griest}, {Lehner},
+%  {Marshall}, {Minniti}, {Nelson}, {Peterson}, {Popowski}, {Pratt}, {Quinn},
+%  {Stubbs}, {Sutherland}, {Tomaney}, {Vandehei}, {Welch}, \& {MACHO
+%  Collaboration}}]{alcock2000}
+%{Alcock}, C., {Allsman}, R.~A., {Alves}, D.~R., {et~al.} 2000, \apj, 541, 734
+%
+%\bibitem[{{Alcock} {et~al.}(2001){Alcock}, {Allsman}, {Alves}, {Axelrod},
+%  {Becker}, {Bennett}, {Cook}, {Drake}, {Freeman}, {Geha}, {Griest}, {Lehner},
+%  {Marshall}, {Minniti}, {Nelson}, {Peterson}, {Popowski}, {Pratt}, {Quinn},
+%  {Stubbs}, {Sutherland}, {Tomaney}, {Vandehei}, \&
+%  {Welch}}]{macho_detection_efficiency}
+%{Alcock}, C., {Allsman}, R.~A., {Alves}, D., {et~al.} 2001, \apjs, 136, 439
+%
+%\bibitem[{{Aubourg} {et~al.}(1993){Aubourg}, {Bareyre}, {Brehin}, {Gros},
+%  {Lachieze-Rey}, {Laurent}, {Lesquoy}, {Magneville}, {Milsztajn}, {Moscoso},
+%  {Queinnec}, {Rich}, {Spiro}, {Vigroux}, {Zylberajch}, {Ansari}, {Cavalier},
+%  {Moniez}, {Beaulieu}, {Ferlet}, {Grison}, {Vidal-Madjar}, {Guibert},
+%  {Moreau}, {Tajahmady}, {Maurice}, {Prevot}, \& {Gry}}]{eros_original}
+%{Aubourg}, E., {Bareyre}, P., {Brehin}, S., {et~al.} 1993, The Messenger, 72,
+%  20
+%
+%\bibitem[{{Bertin} \& {Arnouts}(1996)}]{bertin96}
+%{Bertin}, E., \& {Arnouts}, S. 1996, \aaps, 117, 393
+%
+%\bibitem[{{Binney} {et~al.}(2000){Binney}, {Bissantz}, \&
+%  {Gerhard}}]{binney2000}
+%{Binney}, J., {Bissantz}, N., \& {Gerhard}, O. 2000, \apjl, 537, L99
+%
+%\bibitem[{{Bloom} {et~al.}(2012){Bloom}, {Richards}, {Nugent}, {Quimby},
+%  {Kasliwal}, {Starr}, {Poznanski}, {Ofek}, {Cenko}, {Butler}, {Kulkarni},
+%  {Gal-Yam}, \& {Law}}]{bloomPasp}
+%{Bloom}, J.~S., {Richards}, J.~W., {Nugent}, P.~E., {et~al.} 2012, \pasp, 124,
+%  1175
+%
+%\bibitem[{{Bochanski} {et~al.}(2010){Bochanski}, {Hawley}, {Covey}, {West},
+%  {Reid}, {Golimowski}, \& {Ivezi{\'c}}}]{Bochanski2010}
+%{Bochanski}, J.~J., {Hawley}, S.~L., {Covey}, K.~R., {et~al.} 2010, \aj, 139,
+%  2679
+%
+%\bibitem[{{Covey}(2006)}]{kev06}
+%{Covey}, K.~R. 2006, PhD thesis, University of Washington, Washington, USA
+%
+%\bibitem[{{Covey} {et~al.}(2007){Covey}, {Ivezi{\'c}}, {Schlegel},
+%  {Finkbeiner}, {Padmanabhan}, {Lupton}, {Ag{\"u}eros}, {Bochanski}, {Hawley},
+%  {West}, {Seth}, {Kimball}, {Gogarten}, {Claire}, {Haggard}, {Kaib},
+%  {Schneider}, \& {Sesar}}]{kev07}
+%{Covey}, K.~R., {Ivezi{\'c}}, {\v Z}., {Schlegel}, D., {et~al.} 2007, \aj, 134,
+%  2398
+%
+%\bibitem[{{Crotts} \& {Tomaney}(1996)}]{crotts1996}
+%{Crotts}, A.~P.~S., \& {Tomaney}, A.~B. 1996, \apjl, 473, L87
+%
+%\bibitem[{{Cutri} {et~al.}(2003)}]{cutri03}
+%{Cutri}, R.~M., {et~al.} 2003, {2MASS All Sky Catalog of point sources.} (The
+%  IRSA 2MASS All-Sky Point Source Catalog, NASA/IPAC Infrared Science
+%  Archive.~http://irsa.ipac.caltech.edu/applications/Gator/)
+%
+%\bibitem[{{Di Stefano}(2008)}]{mesolensing}
+%{Di Stefano}, R. 2008, \apj, 684, 46
+%
+%\bibitem[Durbin \& Watson(1950)]{durbin50}
+%J.~Durbin and G.~S. Watson.
+%\newblock Testing for serial correlation in least squares regression: I.
+%\newblock {\em Biometrika}, 37(3/4):pp. 409--428, 1950.
+%
+%\bibitem[{{Foreman-Mackey} {et~al.}(2012){Foreman-Mackey}, {Hogg}, {Lang}, \&
+%  {Goodman}}]{dfm}
+%{Foreman-Mackey}, D., {Hogg}, D.~W., {Lang}, D., \& {Goodman}, J. 2012, ArXiv
+%  e-prints
+%
+%\bibitem[{{Fukui} {et~al.}(2007){Fukui}, {Abe}, {Ayani}, {Fujii}, {Iizuka},
+%  {Itow}, {Kabumoto}, {Kamiya}, {Kawabata}, {Kawanomoto}, {Kinugasa}, {Koff},
+%  {Krajci}, {Naito}, {Nogami}, {Narusawa}, {Ohishi}, {Ohnishi}, {Sumi}, \&
+%  {Tsumuraya}}]{fukui2007}
+%{Fukui}, A., {Abe}, F., {Ayani}, K., {et~al.} 2007, \apj, 670, 423
+%
+%\bibitem[{{Gaudi}(2011)}]{gaudi2011}
+%{Gaudi}, B.~S. 2011, {Microlensing by Exoplanets}, ed. S.~{Piper}, 79--110
+%
+%\bibitem[{{Gaudi} {et~al.}(2008){Gaudi}, {Patterson}, {Spiegel}, {Krajci},
+%  {Koff}, {Pojma{\'n}ski}, {Dong}, {Gould}, {Prieto}, {Blake}, {Roming},
+%  {Bennett}, {Bloom}, {Boyd}, {Eyler}, {de Ponthi{\`e}re}, {Mirabal}, {Morgan},
+%  {Remillard}, {Vanmunster}, {Wagner}, \& {Watson}}]{gaudi2008}
+%{Gaudi}, B.~S., {Patterson}, J., {Spiegel}, D.~S., {et~al.} 2008, \apj, 677,
+%  1268
+%
+%\bibitem[{{Goodman} \& {Weare}(2010)}]{goodman}
+%{Goodman}, J., \& {Weare}, J. 2010, Comm. App. Math. Comp. Sci., 5
+%
+%\bibitem[{{Hamadache} {et~al.}(2006){Hamadache}, {Le Guillou}, {Tisserand},
+%  {Afonso}, {Albert}, {Andersen}, {Ansari}, {Aubourg}, {Bareyre}, {Beaulieu},
+%  {Charlot}, {Coutures}, {Ferlet}, {Fouqu{\'e}}, {Glicenstein}, {Goldman},
+%  {Gould}, {Graff}, {Gros}, {Haissinski}, {de Kat}, {Lesquoy}, {Loup},
+%  {Magneville}, {Marquette}, {Maurice}, {Maury}, {Milsztajn}, {Moniez},
+%  {Palanque-Delabrouille}, {Perdereau}, {Rahal}, {Rich}, {Spiro},
+%  {Vidal-Madjar}, {Vigroux}, \& {Zylberajch}}]{hamadache2009}
+%{Hamadache}, C., {Le Guillou}, L., {Tisserand}, P., {et~al.} 2006, \aap, 454,
+%  185
+%
+%\bibitem[{{Han}(2008)}]{han2008}
+%{Han}, C. 2008, \apj, 681, 806
+%
+%\bibitem[{{Kowalski} {et~al.}(2010){Kowalski}, {Hawley}, {Holtzman},
+%  {Wisniewski}, \& {Hilton}}]{kowal2010}
+%{Kowalski}, A.~F., {Hawley}, S.~L., {Holtzman}, J.~A., {Wisniewski}, J.~P., \&
+%  {Hilton}, E.~J. 2010, \apjl, 714, L98
+%
+%\bibitem[{{Law} {et~al.}(2009){Law}, {Kulkarni}, {Dekany}, {Ofek}, {Quimby},
+%  {Nugent}, {Surace}, {Grillmair}, {Bloom}, {Kasliwal}, {Bildsten}, {Brown},
+%  {Cenko}, {Ciardi}, {Croner}, {Djorgovski}, {van Eyken}, {Filippenko}, {Fox},
+%  {Gal-Yam}, {Hale}, {Hamam}, {Helou}, {Henning}, {Howell}, {Jacobsen},
+%  {Laher}, {Mattingly}, {McKenna}, {Pickles}, {Poznanski}, {Rahmer}, {Rau},
+%  {Rosing}, {Shara}, {Smith}, {Starr}, {Sullivan}, {Velur}, {Walters}, \&
+%  {Zolkower}}]{nick2009}
+%{Law}, N.~M., {Kulkarni}, S.~R., {Dekany}, R.~G., {et~al.} 2009, \pasp, 121,
+%  1395
+%
+%\bibitem[{{Law} {et~al.}(2010){Law}, {Dekany}, {Rahmer}, {Hale}, {Smith},
+%  {Quimby}, {Ofek}, {Kasliwal}, {Zolkower}, {Velur}, {Henning}, {Bui},
+%  {McKenna}, {Nugent}, {Jacobsen}, {Walters}, {Bloom}, {Surace}, {Grillmair},
+%  {Laher}, {Mattingly}, \& {Kulkarni}}]{nick2010}
+%{Law}, N.~M., {Dekany}, R.~G., {Rahmer}, G., {et~al.} 2010, in Society of
+%  Photo-Optical Instrumentation Engineers (SPIE) Conference Series, Vol. 7735
+%
+%\bibitem[{{Levitan} {et~al.}(2011){Levitan}, {Fulton}, {Groot}, {Kulkarni},
+%  {Ofek}, {Prince}, {Shporer}, {Bloom}, {Cenko}, {Kasliwal}, {Law}, {Nugent},
+%  {Poznanski}, {Quimby}, {Horesh}, {Sesar}, \& {Sternberg}}]{levitan2011}
+%{Levitan}, D., {Fulton}, B.~J., {Groot}, P.~J., {et~al.} 2011, \apj, 739, 68
+%
+%\bibitem[{{Nieto--Santisteban} {et~al.}(2004)}]{sdss_images}
+%{Nieto--Santisteban}, M.~A., {et~al.} 2004, in ASP Conf. Ser. 314: Astronomical
+%  Data Analysis Software and Systems (ADASS) XIII, ed. F.~{Ochsenbein}, M.~G.
+%  {Allen}, \& D.~{Egret}, 666--+
+%
+%\bibitem[{{Ofek} {et~al.}(2012){Ofek}, {Laher}, {Surace}, {Levitan}, {Sesar},
+%  {Horesh}, {Law}, {van Eyken}, {Kulkarni}, {Prince}, {Nugent}, {Sullivan},
+%  {Yaron}, {Pickles}, {Ag{\"u}eros}, {Arcavi}, {Bildsten}, {Bloom}, {Cenko},
+%  {Gal-Yam}, {Grillmair}, {Helou}, {Kasliwal}, {Poznanski}, \&
+%  {Quimby}}]{ofek2012}
+%{Ofek}, E.~O., {Laher}, R., {Surace}, J., {et~al.} 2012, \pasp, 124, 854
+%
+%\bibitem[{{Oke} \& {Gunn}(1982)}]{oke_gunn}
+%{Oke}, J.~B., \& {Gunn}, J.~E. 1982, \pasp, 94, 586
+%
+%\bibitem[{{Os{\l}owski} {et~al.}(2008){Os{\l}owski}, {Moderski}, {Bulik}, \&
+%  {Belczynski}}]{oslowski2008}
+%{Os{\l}owski}, S., {Moderski}, R., {Bulik}, T., \& {Belczynski}, K. 2008, \aap,
+%  478, 429
+%
+%\bibitem[{{Paczynski}(1986)}]{paczynski1986}
+%{Paczynski}, B. 1986, \apj, 304, 1
+%
+%\bibitem[{{Pickles}(1998)}]{Pickles1998}
+%{Pickles}, A.~J. 1998, \pasp, 110, 863
+%
+%\bibitem[{{Polishook} {et~al.}(2012){Polishook}, {Ofek}, {Waszczak},
+%  {Kulkarni}, {Gal-Yam}, {Aharonson}, {Laher}, {Surace}, {Klein}, {Bloom},
+%  {Brosch}, {Prialnik}, {Grillmair}, {Cenko}, {Kasliwal}, {Law}, {Levitan},
+%  {Nugent}, {Poznanski}, \& {Quimby}}]{polishook2012}
+%{Polishook}, D., {Ofek}, E.~O., {Waszczak}, A., {et~al.} 2012, \mnras, 421,
+%  2094
+%
+%\bibitem[{{Rahmer} {et~al.}(2008){Rahmer}, {Smith}, {Velur}, {Hale}, {Law},
+%  {Bui}, {Petrie}, \& {Dekany}}]{rahmer2008}
+%{Rahmer}, G., {Smith}, R., {Velur}, V., {et~al.} 2008, in Society of
+%  Photo-Optical Instrumentation Engineers (SPIE) Conference Series, Vol. 7014
+%
+%\bibitem[{{Rattenbury} {et~al.}(2005){Rattenbury}, {Abe}, {Bennett}, {Bond},
+%  {Calitz}, {Claret}, {Cook}, {Furuta}, {Gal-Yam}, {Glicenstein}, {Hearnshaw},
+%  {Hauschildt}, {Kilmartin}, {Kurata}, {Masuda}, {Maoz}, {Matsubara},
+%  {Meintjes}, {Moniez}, {Muraki}, {Noda}, {Ofek}, {Okajima}, {Philpott},
+%  {Rhie}, {Sako}, {Sullivan}, {Sumi}, {Terndrup}, {Tristram}, {Wood},
+%  {Yanagisawa}, \& {Yock}}]{rattenbury2005}
+%{Rattenbury}, N.~J., {Abe}, F., {Bennett}, D.~P., {et~al.} 2005, \aap, 439, 645
+%
+%\bibitem[{{Rau} {et~al.}(2009){Rau}, {Kulkarni}, {Law}, {Bloom}, {Ciardi},
+%  {Djorgovski}, {Fox}, {Gal-Yam}, {Grillmair}, {Kasliwal}, {Nugent}, {Ofek},
+%  {Quimby}, {Reach}, {Shara}, {Bildsten}, {Cenko}, {Drake}, {Filippenko},
+%  {Helfand}, {Helou}, {Howell}, {Poznanski}, \& {Sullivan}}]{rau2009}
+%{Rau}, A., {Kulkarni}, S.~R., {Law}, N.~M., {et~al.} 2009, \pasp, 121, 1334
+%
+%\bibitem[{{Richards} {et~al.}(2002)}]{richards02}
+%{Richards}, G.~T., {et~al.} 2002, \aj, 123, 2945
+%
+%\bibitem[{{Sartore} \& {Treves}(2010)}]{sartore2010}
+%{Sartore}, N., \& {Treves}, A. 2010, \aap, 523, A33
+%
+%\bibitem[{{Shin} {et~al.}(2009){Shin}, {Sekora}, \& {Byun}}]{shin2009}
+%{Shin}, M.-S., {Sekora}, M., \& {Byun}, Y.-I. 2009, \mnras, 400, 1897
+%
+%\bibitem[{{Stetson}(1996)}]{stetson1996}
+%{Stetson}, P.~B. 1996, \pasp, 108, 851
+%
+%\bibitem[{{Stoughton} {et~al.}(2002){Stoughton}, {Lupton}, {Bernardi},
+%  {Blanton}, {Burles}, {Castander}, {Connolly}, {Eisenstein}, {Frieman},
+%  {Hennessy}, {Hindsley}, {Ivezi{\'c}}, {Kent}, {Kunszt}, {Lee}, {Meiksin},
+%  {Munn}, {Newberg}, {Nichol}, {Nicinski}, {Pier}, {Richards}, {Richmond},
+%  {Schlegel}, {Smith}, {Strauss}, {SubbaRao}, {Szalay}, {Thakar}, {Tucker},
+%  {Vanden Berk}, {Yanny}, {Adelman}, {Anderson}, {Anderson}, {Annis},
+%  {Bahcall}, {Bakken}, {Bartelmann}, {Bastian}, {Bauer}, {Berman},
+%  {B{\"o}hringer}, {Boroski}, {Bracker}, {Briegel}, {Briggs}, {Brinkmann},
+%  {Brunner}, {Carey}, {Carr}, {Chen}, {Christian}, {Colestock}, {Crocker},
+%  {Csabai}, {Czarapata}, {Dalcanton}, {Davidsen}, {Davis}, {Dehnen},
+%  {Dodelson}, {Doi}, {Dombeck}, {Donahue}, {Ellman}, {Elms}, {Evans}, {Eyer},
+%  {Fan}, {Federwitz}, {Friedman}, {Fukugita}, {Gal}, {Gillespie}, {Glazebrook},
+%  {Gray}, {Grebel}, {Greenawalt}, {Greene}, {Gunn}, {de Haas}, {Haiman},
+%  {Haldeman}, {Hall}, {Hamabe}, {Hansen}, {Harris}, {Harris}, {Harvanek},
+%  {Hawley}, {Hayes}, {Heckman}, {Helmi}, {Henden}, {Hogan}, {Hogg}, {Holmgren},
+%  {Holtzman}, {Huang}, {Hull}, {Ichikawa}, {Ichikawa}, {Johnston}, {Kauffmann},
+%  {Kim}, {Kimball}, {Kinney}, {Klaene}, {Kleinman}, {Klypin}, {Knapp},
+%  {Korienek}, {Krolik}, {Kron}, {Krzesi{\'n}ski}, {Lamb}, {Leger},
+%  {Limmongkol}, {Lindenmeyer}, {Long}, {Loomis}, {Loveday}, {MacKinnon},
+%  {Mannery}, {Mantsch}, {Margon}, {McGehee}, {McKay}, {McLean}, {Menou},
+%  {Merelli}, {Mo}, {Monet}, {Nakamura}, {Narayanan}, {Nash}, {Neilsen},
+%  {Newman}, {Nitta}, {Odenkirchen}, {Okada}, {Okamura}, {Ostriker}, {Owen},
+%  {Pauls}, {Peoples}, {Peterson}, {Petravick}, {Pope}, {Pordes}, {Postman},
+%  {Prosapio}, {Quinn}, {Rechenmacher}, {Rivetta}, {Rix}, {Rockosi}, {Rosner},
+%  {Ruthmansdorfer}, {Sandford}, {Schneider}, {Scranton}, {Sekiguchi}, {Sergey},
+%  {Sheth}, {Shimasaku}, {Smee}, {Snedden}, {Stebbins}, {Stubbs}, {Szapudi},
+%  {Szkody}, {Szokoly}, {Tabachnik}, {Tsvetanov}, {Uomoto}, {Vogeley}, {Voges},
+%  {Waddell}, {Walterbos}, {Wang}, {Watanabe}, {Weinberg}, {White}, {White},
+%  {Wilhite}, {Wolfe}, {Yasuda}, {York}, {Zehavi}, \& {Zheng}}]{stoughton02}
+%{Stoughton}, C., {Lupton}, R.~H., {Bernardi}, M., {et~al.} 2002, \aj, 123, 485
+%
+%\bibitem[{{Sumi} {et~al.}(2011){Sumi}, {Kamiya}, {Bennett}, {Bond}, {Abe},
+%  {Botzler}, {Fukui}, {Furusawa}, {Hearnshaw}, {Itow}, {Kilmartin}, {Korpela},
+%  {Lin}, {Ling}, {Masuda}, {Matsubara}, {Miyake}, {Motomura}, {Muraki},
+%  {Nagaya}, {Nakamura}, {Ohnishi}, {Okumura}, {Perrott}, {Rattenbury}, {Saito},
+%  {Sako}, {Sullivan}, {Sweatman}, {Tristram}, {Udalski}, {Szyma{\'n}ski},
+%  {Kubiak}, {Pietrzy{\'n}ski}, {Poleski}, {Soszy{\'n}ski}, {Wyrzykowski},
+%  {Ulaczyk}, \& {Microlensing Observations in Astrophysics (MOA)
+%  Collaboration}}]{sumi2011}
+%{Sumi}, T., {Kamiya}, K., {Bennett}, D.~P., {et~al.} 2011, \nat, 473, 349
+%
+%\bibitem[{{Tomov} {et~al.}(2013){Tomov}, {Tomova}, \& {Bisikalo}}]{tomov2013}
+%{Tomov}, N.~A., {Tomova}, M.~T., \& {Bisikalo}, D.~V. 2013, ArXiv e-prints
+%
+%\bibitem[{{Udalski}(2003)}]{udalski03}
+%{Udalski}, A. 2003, \actaa, 53, 291
+%
+%\bibitem[{{Udalski} {et~al.}(1992){Udalski}, {Szymanski}, {Kaluzny}, {Kubiak},
+%  \& {Mateo}}]{original_ogle}
+%{Udalski}, A., {Szymanski}, M., {Kaluzny}, J., {Kubiak}, M., \& {Mateo}, M.
+%  1992, \actaa, 42, 253
+%
+%\bibitem[{{Udalski} {et~al.}(1994){Udalski}, {Szymanski}, {Stanek}, {Kaluzny},
+%  {Kubiak}, {Mateo}, {Krzeminski}, {Paczynski}, \&
+%  {Venkat}}]{ogle_optical_depth}
+%{Udalski}, A., {Szymanski}, M., {Stanek}, K.~Z., {et~al.} 1994, \actaa, 44, 165
+%
+%\bibitem[{{von Neumann} {et~al.}(1941){von Neumann}, {Kent}, {Bellinson}, \&
+%  {Hart}}]{von_neumann1941}
+%{von Neumann}, J., {Kent}, R.~H., {Bellinson}, H.~R., \& {Hart}, B.~I. 1941,
+%  The Annals of Mathematical Statistics, 12, 153
+%
+%\bibitem[{{Wozniak}(2000)}]{con_idx}
+%{Wozniak}, P.~R. 2000, \actaa, 50, 421
+%
+%\bibitem[{{Wright} {et~al.}(2010){Wright}, {Eisenhardt}, {Mainzer}, {Ressler},
+%  {Cutri}, {Jarrett}, {Kirkpatrick}, {Padgett}, {McMillan}, {Skrutskie},
+%  {Stanford}, {Cohen}, {Walker}, {Mather}, {Leisawitz}, {Gautier}, {McLean},
+%  {Benford}, {Lonsdale}, {Blain}, {Mendez}, {Irace}, {Duval}, {Liu}, {Royer},
+%  {Heinrichsen}, {Howard}, {Shannon}, {Kendall}, {Walsh}, {Larsen}, {Cardon},
+%  {Schick}, {Schwalm}, {Abid}, {Fabinsky}, {Naes}, \& {Tsai}}]{WISE}
+%{Wright}, E.~L., {Eisenhardt}, P.~R.~M., {Mainzer}, A.~K., {et~al.} 2010, \aj,
+%  140, 1868
+%
+%\bibitem[{{Wyrzykowski} {et~al.}(2009){Wyrzykowski}, {Koz{\l}owski}, {Skowron},
+%  {Belokurov}, {Smith}, {Udalski}, {Szyma{\'n}ski}, {Kubiak},
+%  {Pietrzy{\'n}ski}, {Soszy{\'n}ski}, {Szewczyk}, \&
+%  {{\.Z}ebru{\'n}}}]{wyrzykowski2009}
+%{Wyrzykowski}, {\L}., {Koz{\l}owski}, S., {Skowron}, J., {et~al.} 2009, \mnras,
+%  397, 1228
+%
+%\bibitem[{Yan {et~al.}(2013)Yan, Donoso, Tsai, Stern, Assef, Eisenhardt, Blain,
+%  Cutri, Jarrett, Stanford, Wright, Bridge, \& Riechers}]{yan2013}
+%Yan, L., Donoso, E., Tsai, C.-W., {et~al.} 2013, \aj, 145, 55
+%
+%\bibitem[{{York} {et~al.}(2000)}]{york00}
+%{York}, D.~G., {et~al.} 2000, \aj, 120, 1579
+%
+%\end{thebibliography}
 

### 2013-10-25 text/paper/ms.tex
added postage stamps, rosanne citation

@@
 Such an analysis is usually not possible for events in high-density fields where the lenses are, on average, much farther away and blending makes photometry --- and, therefore, mass measurements --- challenging.
-By contrast, high-Galactic-latitude events probe lenses with distances $\lesssim$1~kpc; the Einstein radii, therefore, tend to be larger ($\sim$milliarcseconds) and may cause detectable astrometric signatures \citep[][]{han2008}.
+By contrast, high-Galactic-latitude events probe lenses with distances $\lesssim$1~kpc; the Einstein radii, therefore, tend to be larger ($\sim$milliarcseconds) and may cause detectable astrometric signatures \citep[][]{han2008, mesolensing, mesolensing2}.
 A measurement of the astrometric shift due to microlensing will break the distance-mass degeneracy in the lensing parameters and enable a direct measurement of the lens mass.
 Finding even a handful of nearby lenses that can be studied in the same kind of detail as the Tago event would, therefore, be very valuable.
-
-% TODO: Han citation above, also mentioned in one of Rosanne's papers
 
@@
 While $\sim$10\% of the light curves pass the quality cut described in Section~\ref{sec:detection_eff}, only $\sim$0.1-0.5\% of these pass the FPR cut determined from the Monte Carlo simulations;
-	\item for fields with SDSS coverage, we use Data Release 9 photometry \citep{dr9paper} to remove sources typed as Galaxy or QSO (using \texttt{objc\_type}\footnote{{\tt http://www.sdss3.org/dr9/algorithms/classify.php}}) and flag candidate quasars \citep[using the cuts described in][]{richards02} ($\sim$10\%);
+	\item for fields with SDSS coverage, we use Data Release 9 photometry \citep{dr9paper} to remove sources typed as Galaxy or QSO (using \texttt{objc\_type}\footnote{{\tt http://www.sdss3.org/dr9/algorithms/classify.php}}) and flag candidate quasars \citep[using the cuts described in][]{richards02}; this eliminates $\sim$90\% of the remaining light curves;
 	\item fit a microlensing event model and subtract the model, then recompute $\eta$ for all the selected light curves;
-	\item reject the light curves for which the new value of $\eta$ still passes the cut\footnote{These are probably periodic variables.} ($\sim$5-10\%);
-	\item retain any remaining light curves for further inspection.
+	\item reject the light curves for which the new value of $\eta$ still passes the cut\footnote{These are probably periodic variables.} --- this step eliminates $\sim$90-95\% of the remaining light curves;
+	\item retain any surviving light curves for further inspection. Typically, there is $\lapprox$1 such light curve per field.
 \end{enumerate}

### 2013-10-31 scripts/candidate_pipeline.py
overhaul

@@ -23,12 +23,6 @@ warnings.simplefilter("ignore")
 # Third-party
 import numpy as np
 import scipy.optimize as so
-try:
-    from apwlib.globals import greenText, yellowText, redText
-except ImportError:
-    raise ImportError("apwlib not found! \nDo: 'git clone "
-        "git@github.com:adrn/apwlib.git' and run 'python setup.py "
-        "install' to install.")
 
 # Project
 import ptf.db.photometric_database as pdb
@@ -52,9 +46,14 @@ def select_candidates(field, selection_criteria, num_fit_attempts=10):
 
     light_curves = []
     for ccd in field.ccds.values():
-        logger.info(greenText("Starting with CCD {}".format(ccd.id)))
+        logger.info("Starting with CCD {}".format(ccd.id))
         chip = ccd.read()
-
+        ####### APW @ MDM
+        #print("Total:", len(chip.sources.read(field="matchedSourceID")))
+        #cdtn = "(ngoodobs > 10) " #& ((ngoodobs/nobs) > 0.5)"
+        #print("Condition:", len(chip.sources.readWhere(cdtn, field="matchedSourceID")))
+        #continue
+        ########################
         cdtn = ("(ngoodobs > {}) & (vonNeumannRatio > 0.0) & "
                 "(vonNeumannRatio < {}) & ((ngoodobs/nobs) > 0.5)")
         cdtn = cdtn.format(min_number_of_good_observations, eta_cut)
@@ -62,7 +61,6 @@ def select_candidates(field, selection_criteria, num_fit_attempts=10):
 
         logger.info("\tSelected {} pre-candidates from PDB"\
                     .format(len(source_ids)))
-
         for source_id in source_ids:
             # APW: TODO -- this is still the biggest time hog!!! It turns 
             #   out it's still faster than reading the whole thing into 
@@ -72,7 +70,8 @@ def select_candidates(field, selection_criteria, num_fit_attempts=10):
 
             # If light curve doesn't have enough clean observations, skip it
             if light_curve != None and \
-                len(light_curve) < min_number_of_good_observations: continue
+                len(light_curve) < min_number_of_good_observations: 
+                continue
 
             # Compute the variability indices for the cleaned light curve
             try:
@@ -99,7 +98,7 @@ def select_candidates(field, selection_criteria, num_fit_attempts=10):
             qso_status = richards_qso(sdss_colors)
             if sdss_colors != None and qso_status:
[…]

### 2013-10-31 scripts/paper_dataset.py

@@ -10,6 +10,7 @@ import os, sys
 import logging
 import random
 import cPickle as pickle
+from collections import defaultdict
 
 # Third-party
 import numpy as np
@@ -20,81 +21,43 @@ import ptf.db.photometric_database as pdb
 from ptf.globals import all_fields, data_path
 
 np.random.seed(42)
-dict_file = os.path.join(data_path, "publish_field_ccds.pickle")
 
-def make_lookup_dict(Nlightcurves=10000):
+cache = defaultdict(dict)
+def get_lcs(Nlightcurves=1000, seed=42):
     Nfields = len(all_fields)
     Nselected = 0
 
     ccd_ids = range(13)
     ccd_ids.pop(3)
-    
-    field_ccd = dict()
+
+    np.random.seed(seed)
+    lcs = []
     while Nselected < Nlightcurves:
         field_row = all_fields[np.random.randint(Nfields)]
         field_id = field_row['id']
-    
+        
         field = pdb.Field(field_id, 'R')
+        
         if len(field.ccds) == 0:
             continue
+        
+        if not cache[field_id].has_key(ccd.id):
+            ccd = random.choice(field.ccds.values())
+            chip = ccd.read()
+            sources = chip.sources.readWhere("ngoodobs > 10")
+            cache[field_id][ccd.id] = sources
+        
+        np.random.shuffle(cache[field_id][ccd.id])
+        
+        for source in cache[field_id][ccd.id]:
+            lc = ccd.light_curve(source["matchedSourceID"], clean=True, barebones=True)
+            if len(lc) > 10:
+                lcs.append(lc)
+    
+    return lcs
 
-        ccd = random.choice(field.ccds.values())
-
-        try:
-            field_ccd[field.id]
-        except KeyError:
-            field_ccd[field.id] = dict()
-
-        try:
-            field_ccd[field_id][ccd.id]
-        except KeyError:
-            field_ccd[field_id][ccd.id] = 0
-
-            field_ccd[field.id][ccd.id] += 1
-            Nselected += 1
-
-    with open(dict_file, 'w') as f:
-        pickle.dump(field_ccd, f)
-   
-def random_light_curves(ccd, N):
-    chip = ccd.read()
-    sources = chip.sources.readWhere("ngoodobs > 10")
-    np.random.shuffle(sources)
-
-    lcs = []
-    for source in sources:
-        lc = ccd.light_curve(source["matchedSourceID"], clean=True, barebones=True)
[…]

### 2013-10-31 text/paper/ms.tex

@@
 \section{Searching for events in the full PTF dataset}\label{sec:search}
-To identify candidate microlensing events in the $1.1\times10^9$ light curves with $>$10 $R$-band observations, we:
+To identify candidate microlensing events we: %in the $1.1\times10^9$ light curves with $>$10 $R$-band observations, we:
 %We apply the 1\% FPR cut to the 2097 PTF fields ($\sim$15,224 deg$^2$) with $>$10 $R$-band observations using the variability index $\eta$. Our procedure is as follows:
 \begin{enumerate}
-	\item select a field from the 2097 fields with $>$10 good $R$-band observations and run Monte Carlo simulations to compute the 1\% FPR limiting values for $\eta$, $\Delta \chi^2$, and $J$ for that field;
+	\item select the light curves from one of the 2097 fields with $>$10 $R$-band observations;
+	\item identify those light curves with more than 50\% good data points and $>$10 good observations (see Section~\ref{sec:detection_eff}) and run Monte Carlo simulations to compute the 1\% FPR limiting values for $\eta$, $\Delta \chi^2$, and $J$ for that field;
 	\item compute $\eta$ for all the light curves in the field.
-While $\sim$10\% of the light curves pass the quality cut described in Section~\ref{sec:detection_eff}, only $\sim$0.1-0.5\% of these pass the FPR cut determined from the Monte Carlo simulations;
+While $\sim$10\% of the light curves pass the quality cut described above, only $\sim$0.1-0.5\% of these survive the FPR cut determined from the Monte Carlo simulations;
 	\item for fields with SDSS coverage, we use Data Release 9 photometry \citep{dr9paper} to remove sources typed as Galaxy or QSO (using \texttt{objc\_type}\footnote{{\tt http://www.sdss3.org/dr9/algorithms/classify.php}}) and flag candidate quasars \citep[using the cuts described in][]{richards02}; this eliminates $\sim$90\% of the remaining light curves;

### 2013-10-31 scripts/paper_dataset.py
bork

@@ -35,47 +35,43 @@ def get_lcs(Nlightcurves=1000, seed=42):
     while Nselected < Nlightcurves:
         field_row = all_fields[np.random.randint(Nfields)]
         field_id = field_row['id']
-        
+
         field = pdb.Field(field_id, 'R')
-        
+
         if len(field.ccds) == 0:
             continue
-        
+
         if not cache[field_id].has_key(ccd.id):
             ccd = random.choice(field.ccds.values())
             chip = ccd.read()
             sources = chip.sources.readWhere("ngoodobs > 10")
             cache[field_id][ccd.id] = sources
-        
+
         np.random.shuffle(cache[field_id][ccd.id])
-        
+
         for source in cache[field_id][ccd.id]:
             lc = ccd.light_curve(source["matchedSourceID"], clean=True, barebones=True)
             if len(lc) > 10:
                 lcs.append(lc)
-    
+
     return lcs
 
 def store_light_curves(lcs):
+    data = None
     for lc in lcs:
-                n = len(lc.mjd)
-                f = [lc.field_id]*n
-                c = [lc.ccd_id]*n
-                s = [lc.source_id]*n
-                
-                if data is None:
-                    data = np.array([f, c, s, lc.mjd, lc.mag, lc.error])
-                else:
-                    data = np.hstack((data, np.array([f, c, s, lc.mjd, lc.mag, lc.error])))
+        n = len(lc.mjd)
+        f = [lc.field_id]*n
+        c = [lc.ccd_id]*n
+        s = [lc.source_id]*n
 
-            print(data.shape)
-            sys.exit(0)
+        if data is None:
+            data = np.array([f, c, s, lc.mjd, lc.mag, lc.error])
+        else:
+            data = np.hstack((data, np.array([f, c, s, lc.mjd, lc.mag, lc.error])))
 
-            all_lightcurves = []
-
+    print(data.shape)
+    sys.exit(0)
 
 if __name__ == "__main__":
-    if not os.path.exists(dict_file):
-        make_lookup_dict(Nlightcurves=10000)
-
-    store_light_curves()
+    lcs = get_lcs(100, seed=42)
+    store_light_curves(lcs)

### 2013-11-04 scripts/paper_dataset.py
attempt 2

@@ -21,43 +21,34 @@ from astropy.io import fits
 import ptf.db.photometric_database as pdb
 from ptf.globals import all_fields, data_path
 
-np.random.seed(42)
+def get_lcs(field, N_per_field=5, seed=42):
+    np.random.random(seed)
 
-cache = defaultdict(dict)
-def get_lcs(Nlightcurves=1000, seed=42):
-    Nfields = len(all_fields)
-
-    ccd_ids = range(13)
-    ccd_ids.pop(3)
-
-    np.random.seed(seed)
+    cache = dict()
     lcs = []
-    while len(lcs) < Nlightcurves:
-        field_row = all_fields[np.random.randint(Nfields)]
-        field_id = field_row['id']
+    for ii in range(N_per_field):
+        ccd = random.choice(field.ccds.values)
 
-        field = pdb.Field(field_id, 'R')
-
-        if len(field.ccds) == 0:
-            continue
-
-        ccd = random.choice(field.ccds.values())
-        if not cache[field_id].has_key(ccd.id):
+        if not cache.has_key(ccd.id):
             chip = ccd.read()
             sources = chip.sources.readWhere("ngoodobs > 10")
-            cache[field_id][ccd.id] = sources
+            cache[ccd.id] = sources
 
-        np.random.shuffle(cache[field_id][ccd.id])
+        sources = cache[ccd.id]
+        np.random.shuffle(sources)
 
-        for source in cache[field_id][ccd.id]:
-            lc = ccd.light_curve(source["matchedSourceID"], clean=True, barebones=True)
+        for source in sources:
+            lc = ccd.light_curve(source["matchedSourceID"], clean=True,
+                                 barebones=True)
             if len(lc) > 10:
                 lcs.append(lc)
                 break
 
+    field.close()
+
     return lcs
 
-def store_light_curves(lcs):
+def lcsToArray(lcs):
     data = None
     for lc in lcs:
         n = len(lc.mjd)
@@ -69,22 +60,37 @@ def store_light_curves(lcs):
             data = np.array([f, c, s, lc.mjd, lc.mag, lc.error])
         else:
             data = np.hstack((data, np.array([f, c, s, lc.mjd, lc.mag, lc.error])))
-    
+
     dtype = [("field", int), ("ccd", int), ("id", int), \
              ("mjd", float), ("R_mag", float), ("mag_err", float)]
-    
+
[…]

### 2013-11-04 scripts/paper_dataset.py
attempt 3

@@ -76,6 +76,8 @@ if __name__ == "__main__":
     Nfields = len(all_fields)
     field_subset = all_fields[ii*Nfields//subdiv:ii*Nfields//subdiv+Nfields//subdiv]
 
+    fn = os.path.join(path, "{0}.fits".format(ii))
+
     for f in field_subset:
         field = pdb.Field(f["id"], "R")
         if len(field.ccds) == 0:

### 2013-11-04 scripts/paper_figures.py

@@ -50,7 +50,7 @@ tick_font_size = 14
 
 ptfid_to_longname = {"PTFS1203am" : "PTF1J033545.80-041849.1",
                      "PTFS1206i" : "PTF1J061800.25+203142.5",
-                     "PTFS1212ak" : "PTF1J120642.60-192016.4", 
+                     "PTFS1212ak" : "PTF1J120642.60-192016.4",
                      "PTFS1213ap" : "PTF1J131500.09+715032.5",
                      "PTFS1215aj" : "PTF1J153202.91+674825.1",
                      "PTFS1216bt" : "PTF1J161502.39+540053.8",
@@ -252,7 +252,7 @@ def maximum_outlier_indices_plot(field_id):
 
             sig,md = np.std(best_outlier_lightcurve.mag), np.median(best_outlier_lightcurve.mag)
             ix = (best_outlier_lightcurve.mag < (md+8*sig)) & (best_outlier_lightcurve.mag > (md-8*sig))
-            best_outlier_lightcurve = PTFLightCurve(best_outlier_lightcurve.mjd[ix], 
+            best_outlier_lightcurve = PTFLightCurve(best_outlier_lightcurve.mjd[ix],
                                                     best_outlier_lightcurve.mag[ix],
                                                     best_outlier_lightcurve.error[ix])
             try:
@@ -271,9 +271,9 @@ def maximum_outlier_indices_plot(field_id):
         axes[ii].set_yticks([round(yt,2) for yt in yticks])
         for ticklabel in axes[ii].get_yticklabels():
             ticklabel.set_fontsize(22)
-        
+
         bbox_props = dict(boxstyle="round,pad=0.3", fc="white", ec="k", lw=1)
-        axes[ii].text(55975, yticks[2], "{0}({1})".format(min_max[index], pu.index_to_label(index)), 
+        axes[ii].text(55975, yticks[2], "{0}({1})".format(min_max[index], pu.index_to_label(index)),
                       fontsize=30,
                       bbox=bbox_props)
 
@@ -725,50 +725,51 @@ def ml_parameter_distributions(overwrite=False):
 
 def lightcurve_from_filename(filename):
     import astropy.io.ascii as ascii
-    
+
     with open(filename) as f:
         data_start = 0
         for line in f:
             if line.startswith("#"):
[…]

### 2013-11-04 scripts/detection_efficiency.py
hackz

@@ -40,11 +40,14 @@ matplotlib.use("Agg")
 import matplotlib.pyplot as plt
 
 # Project
-from ptf.simulation.simulatedlightcurve import SimulatedLightCurve
-import ptf.photometricdatabase as pdb
+from ptf.lightcurve import SimulatedLightCurve
+import ptf.db.photometric_database as pdb
 import ptf.analyze.analyze as analyze
-from ptf.globals import index_to_label
+#from ptf.globals import index_to_label
+import ptf.util as pu
 import ptf.variability_indices as vi
+
+index_to_label = pu.index_to_label
 
 best = False
 
@@ -103,6 +106,8 @@ def detection_efficiency_for_field(field, ccds=range(12), config=dict(), overwri
         os.remove(pickle_filename)
         logger.debug("Data file deleted...")
     
+    #print(pickle_filename, os.path.exists(pickle_filename))
+
     # If the cache pickle file doesn't exist, generate the data
     if not os.path.exists(pickle_filename):
         logger.info("Data file {} not found. Generating data...".format(pickle_filename))
@@ -299,7 +304,7 @@ def plot_distributions(selected_distributions, simulated_microlensing_statistics
             plt.setp(axes[jj,ii].get_yticklabels(), visible=False)
             
             if ii == 0:
-                axes[jj,ii].set_ylabel(index_to_label[index], fontsize=24, rotation=0)
+                axes[jj,ii].set_ylabel(index_to_label(index), fontsize=24, rotation=0)
             
             if ii == 0:
                 ylim = axes[jj,ii].get_ylim()
@@ -321,6 +326,8 @@ def plot_distributions(selected_distributions, simulated_microlensing_statistics
     fig.savefig(filename)
 
 def get_best_light_curve(field, ccd_id):
+    print(field.ccds)
+    sys.exit(0)
     ccd = field.ccds[ccd_id]
     chip = ccd.read()
     sources = chip.sources.readWhere("ngoodobs > 100")
@@ -341,7 +348,7 @@ def plot_best_light_curve(field_id, ccd_id):
     from pylab import ScalarFormatter
     ax.yaxis.set_major_formatter(ScalarFormatter(False))
     ax.set_xlabel("MJD", fontsize=20)
-    ax.set_ylabel("$R$ (mag)", fontsize=20)
+    ax.set_ylabel("$R$ [mag]", fontsize=20)
[…]

### 2013-11-04 text/paper/ms.tex

@@
 \shortauthors{Price-Whelan et al.}
-
 \bibliographystyle{apj}
@@
 To examine the probability that microlensing events can be recovered in these data, we test previous statistics used on uniformly sampled data to identify variables and transients.
-We find that one such statistic, the von Neumann ratio, also known as the Durbin-Watson statistic, performs best for identifying simulated microlensing events in our data.
+We find that one such statistic, the von Neumann ratio (also known as the Durbin-Watson statistic), performs best for identifying simulated microlensing events in our data.
 We develop a selection method using this statistic and apply it to data from all PTF fields with $>$10 $R$-band observations --- a total of $1.1\times10^9$ light curves --- to uncover three candidate sparse-field microlensing events.
@@
 Over the past 20 years, microlensing, in which gravitational lensing causes a transient increase in the flux from a background point source, has been used to search for dark and compact objects \citep{original_macho, oslowski2008, sartore2010}, to study Galactic structure and kinematics \citep{binney2000}, to determine the shape of stars \citep{rattenbury2005}, and to identify extrasolar planets \citep[][and references therein]{gaudi2011}.
-Such studies were once limited by the small number of detected events, but thanks to advances in CCD technology and the development of dedicated microlensing surveys, a few thousand events are now observed each year.\footnote{e.g., {\tt http://ogle.astrouw.edu.pl/ogle4/ews/ews.html}} In the coming decade, this number is expected to increase as the next generation of photometric, time-domain surveys comes online, providing an opportunity for the precise study of otherwise hard-to-characterize objects, such as low-mass stars, sub-stellar objects, and isolated neutron stars.
+Such studies were once limited by the small number of detected events, but thanks to advances in CCD technology and the development of dedicated microlensing surveys, a few thousand events are now observed each year.\footnote{e.g., \url{http://ogle.astrouw.edu.pl/ogle4/ews/ews.html}.} In the coming decade, this number is expected to increase as the next generation of photometric, time-domain surveys comes online, providing an opportunity for the precise study of otherwise hard-to-characterize objects, such as low-mass stars, sub-stellar objects, and isolated neutron stars.
 
@@
 Indeed, lensing events away from these fields are potentially very interesting.
-While the microlensing event rate in low-density stellar regions will be smaller, compared to events found by modern microlensing surveys, these are expected to involve closer sources and lenses, and less crowded backgrounds \citep{mesolensing}.
+While the microlensing event rate in low-density stellar regions will be smaller, compared to events found by modern microlensing surveys, these are expected to involve closer sources and lenses, and less crowded backgrounds \citep{mesolensing1}.
 Thus, such lenses offer more opportunities to constrain the properties of the events.
@@
 Such an analysis is usually not possible for events in high-density fields where the lenses are, on average, much farther away and blending makes photometry --- and, therefore, mass measurements --- challenging.
-By contrast, high-Galactic-latitude events probe lenses with distances $\lesssim$1~kpc; the Einstein radii, therefore, tend to be larger ($\sim$milliarcseconds) and may cause detectable astrometric signatures \citep[][]{han2008, mesolensing, mesolensing2}.
+By contrast, high-Galactic-latitude events probe lenses with distances $\lesssim$1~kpc; the Einstein radii, therefore, tend to be larger ($\sim$milliarcseconds) and may cause detectable astrometric signatures \citep[][]{han2008, mesolensing1, mesolensing2}.
 A measurement of the astrometric shift due to microlensing will break the distance-mass degeneracy in the lensing parameters and enable a direct measurement of the lens mass.
@@
 \section{PTF: Survey Design and Consequences for Microlensing Searches}\label{sec:ptf} 
-PTF data are collected using the former Canada-France-Hawaii Telescope 12K$\times$8K mosaic camera, which has 11 working chips, $0.92\times10^8$ pixels, and a 7.26 deg$^2$ field-of-view \citep{rahmer2008}, mounted on the 48-nch Oschin Schmidt Telescope (the P48) at Palomar Observatory, CA.
+PTF data are collected using the former Canada-France-Hawaii Telescope 12K$\times$8K mosaic camera, which has 11 working chips, $0.92\times10^8$ pixels, and a 7.26 deg$^2$ field-of-view \citep{rahmer2008}, mounted on the 48-inch Oschin Schmidt Telescope (the P48) at Palomar Observatory, CA.
 Under median seeing conditions (2.2$\arcsec$), observations in Mould $R$ or Sloan Digital Sky Survey \citep[SDSS;][]{york00} $g$ achieve 2.0$\arcsec$ full-width half-maximum images and reach 5$\sigma$ magnitudes of $R \approx 21.0$ and $g \approx 21.3$ mag in each 60~s exposure \citep{nick2009,rau2009,nick2010}.
@@
 While $\sim$10\% of the light curves pass the quality cut described above, only $\sim$0.1-0.5\% of these survive the FPR cut determined from the Monte Carlo simulations;
-	\item for fields with SDSS coverage, we use Data Release 9 photometry \citep{dr9paper} to remove sources typed as Galaxy or QSO (using \texttt{objc\_type}\footnote{{\tt http://www.sdss3.org/dr9/algorithms/classify.php}}) and flag candidate quasars \citep[using the cuts described in][]{richards02}; this eliminates $\sim$90\% of the remaining light curves;
+	\item for fields with SDSS coverage, we use Data Release 9 photometry \citep{dr9paper} to remove sources typed as Galaxy or QSO (using \texttt{objc\_type}\footnote{\url{http://www.sdss3.org/dr9/algorithms/classify.php}}) and flag candidate quasars \citep[using the cuts described in][]{richards02}; this eliminates $\sim$90\% of the remaining light curves;
 	\item fit a microlensing event model and subtract the model, then recompute $\eta$ for all the selected light curves;
@@
 We thank E.~Bowsher, S.~Douglas, and S.~Tang for their assistance in obtaining spectra for our candidates, and M.~Modjaz for help in interpreting several of these spectra.
-We are grateful to the anonymous referee and to our editor, Eric Feigelson, for comments that improved the paper.
+We are grateful to the anonymous referee and to our editor, E.~Feigelson, for comments that improved the paper.
 
@@
 Department of Energy Office of Science.
-The SDSS-III web site is http://www.sdss3.org/.
+The SDSS-III web site is \url{http://www.sdss3.org/}.
 
@@
 %\bigskip
-%\setlength{\baselineskip}{0.6\baselineskip}
+\setlength{\baselineskip}{0.6\baselineskip}
 \bibliography{references_3}
-%\setlength{\baselineskip}{1.667\baselineskip}
+\setlength{\baselineskip}{1.667\baselineskip}
 

### 2013-11-04 text/paper/ms.tex
marcel comments

@@
 We use the existing PTF data to study the trade-off between searching for microlensing events in a survey whose footprint is much larger than that of typical microlensing searches, but that has far-from-optimal time sampling.
-To examine the probability that microlensing events can be recovered in these data, we test previous statistics used on uniformly sampled data to identify variables and transients.
+To examine the probability that microlensing events can be recovered in these data, we test statistics previously used on uniformly sampled data to identify variables and transients.
 We find that one such statistic, the von Neumann ratio (also known as the Durbin-Watson statistic), performs best for identifying simulated microlensing events in our data.
@@
 Such an analysis is usually not possible for events in high-density fields where the lenses are, on average, much farther away and blending makes photometry --- and, therefore, mass measurements --- challenging.
-By contrast, high-Galactic-latitude events probe lenses with distances $\lesssim$1~kpc; the Einstein radii, therefore, tend to be larger ($\sim$milliarcseconds) and may cause detectable astrometric signatures \citep[][]{han2008, mesolensing1, mesolensing2}.
+By contrast, high-Galactic-latitude events probe lenses with distances $\lesssim$1~kpc; the Einstein radii, therefore, tend to be larger ($\sim$milliarcseconds) and may cause detectable astrometric signatures \citep[][]{han2008, mesolensing2, mesolensing1}.
 A measurement of the astrometric shift due to microlensing will break the distance-mass degeneracy in the lensing parameters and enable a direct measurement of the lens mass.
@@
 Each field has a unique sampling pattern determined by which of the PTF sub-surveys it belongs to, what time of the year it is visible, and how high a priority it is given by the scheduler.
-The imaging cadences range from 1-5 days, but there is a higher-cadence campaign of a field in Orion to find transiting planets around young stars \citep{nick2009}.
+The imaging cadences range from 1-5 days, but other cadences are possible: for example, there was a higher-cadence campaign of a field in Orion to find transiting planets around young stars \citep{nick2009}.
 The result is that PTF light curves often contain gaps and regions of high-cadence observations and/or of low-cadence observations, so that the archive is a massive dataset of irregularly-sampled, time-domain photometry.
 Figure~\ref{fig:sampling} shows six randomly selected light curves and illustrates the varying cadences and coverage that different fields may have over the same one-year period.
-While the PTF database is not yet public, we have made $10^5$ light curves available at {\tt }.
+While the PTF database is not yet public, we have made $10^4$ randomly selected light curves available at \url{http:// TODO }.
 These light curves meet the quality criteria described in Section~\ref{sec:detection_eff} and can be used to test the statistical methods described here (or others!).
@@
 We choose five statistical measures of variability compiled by \cite{shin2009} that have been previously applied to the classification and discovery of periodic variables.
-These variability indices are $\sigma/\mu$, $Con$, $\eta$, $J$, and $K$.\footnote{We do not implement the sixth index described by \cite{shin2009}, $AoVM$, because it mainly helps identify periodic sources.} $\sigma/\mu$ is the ratio of the sample standard deviation ($\sigma$) to the sample mean,
+These variability indices are $\sigma/\mu$, $Con$, $\eta$, $J$, and $K$.\footnote{We do not implement the sixth index described by \cite{shin2009}, $AoVM$, because it mainly helps identify periodic sources.} $\sigma/\mu$ is the ratio of the sample standard deviation ($\sigma$) to the sample mean ($\mu$),
 \begin{align}
@@
 $K$ is a measure of the kurtosis of the distribution of data points.
+We add one more index, $\Delta \chi^2$, the difference in $\chi^2$ between fitting a Gaussian model and fitting a linear model to a light curve.
 
-We add one more index, $\Delta \chi^2$, the difference in $\chi^2$ between fitting a Gaussian model and fitting a linear model to a light curve.
 Specifically, we use a Levenberg-Marquardt optimizer to perform a least-squares fit with each of these models to the light curves and then compute $\Delta\chi^2=\chi_{linear} - \chi_{gaussian}$.
@@
 As a result, the number of data points and number of model parameters are constant, and we therefore do not include an Akaike or Bayesian information criterion term.
+For the Gaussian fit, the optimizer is initialized with a $\sigma$ of 10 days, centered on the brightest data point.
 
-For the Gaussian fit, the optimizer is initialized with a $\sigma$ of 10 days, centered on the brightest data point.
 This is a standard statistical test used by microlensing surveys, and allows us to compare the relative performance of the (slightly modified) \cite{shin2009} indices and of this approach.
 
-Figure~\ref{fig:indices_examples} shows maximally outlying light curves for each variability statistic selected from $\sim$20,000 light curves for objects on a single CCD in PTF field 100101.
+Figure~\ref{fig:indices_examples} shows maximally outlying light curves for each variability statistic selected from $\sim$20,000 light curves for objects on a single CCD in PTF field 3756.
 %, centered at (J2000) 17:21:31.44 $+$61:07:46.6.
@@
 	 }
-\caption{Same as \figurename~\ref{fig:detection_efficiency_4327}, for PTF field 4588.}\label{fig:detection_efficiency_4588}
+\caption{Same as Figure~\ref{fig:detection_efficiency_4327}, for PTF field 4588.}\label{fig:detection_efficiency_4588}
 \end{figure}
@@
 	 }
-\caption{Same as \figurename~\ref{fig:detection_efficiency_4327}, for PTF field 100152.}\label{fig:detection_efficiency_100152}
+\caption{Same as Figure~\ref{fig:detection_efficiency_4327}, for PTF field 100152.}\label{fig:detection_efficiency_100152}
 \end{figure}
@@
 	\item select the light curves from one of the 2097 fields with $>$10 $R$-band observations;
-	\item identify those light curves with more than 50\% good data points and $>$10 good observations (see Section~\ref{sec:detection_eff}) and run Monte Carlo simulations to compute the 1\% FPR limiting values for $\eta$, $\Delta \chi^2$, and $J$ for that field;
+	\item identify those light curves with more than 50\% good data points and $>$10 good observations and run Monte Carlo simulations to compute the 1\% FPR limiting values for $\eta$, $\Delta \chi^2$, and $J$ for that field;
 	\item compute $\eta$ for all the light curves in the field.
@@
 We emphasize, however, that lacking simultaneous, multi-color photometry during the events, we cannot claim these as microlensing events with any confidence.
-These three candidates are shown in Figure~\ref{fig:candidates}, and PTF cutouts of the associated sources at quiescence and near peak brightness are shown in Figure~\ref{fig:postage_stamps}.
-Light curves for the other eight events are shown in Figure~\ref{fig:not_candidates}.
+These three candidates are shown in Figure~\ref{fig:candidates}, and PTF cutouts of the associated sources at quiescence and near peak brightness are shown in Figure~\ref{fig:postage_stamps} (light curves for the other eight events are shown in Figure~\ref{fig:not_candidates}).
 
@@
 \startdata
-J0618 & $0.178\pm0.019$ & $22.8\pm1.4$ & $55879.975\pm0.248$ & $20.226\pm0.026$\\
-J1615 & $0.388\pm0.015$ & $32.5\pm2.3$  & $55895.797\pm1.956$ & $20.016\pm0.005$\\
-J1728 & $1.338\pm0.001$ & $156.6\pm1.2$ & $55194.313\pm0.702$ & $16.259\pm0.001$ 
+J0618 & 0.178$\pm$0.011 & 23.7$\pm$1.6 & 55880.0$\pm$0.4 & 20.222$\pm$0.034 \\
+J1615 & 0.376$\pm$0.016 & 31.4$\pm$1.1 & 55895.5$\pm$2.1 & 20.019$\pm$0.004 \\
+J1728 & 1.338$\pm$0.002 & 161.1$\pm$4.9 & 55192.2$\pm$1.4 & 16.260$\pm$0.002 
 \enddata
@@
 \setlength{\baselineskip}{0.6\baselineskip}
-\bibliography{references_3}
+%\bibliography{references_3}
 \setlength{\baselineskip}{1.667\baselineskip}
 
-%\begin{thebibliography}{54}
-%\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
-%
-%\bibitem[{{Ag{\"u}eros} {et~al.}(2011){Ag{\"u}eros}, {Covey}, {Lemonias},
-%  {Law}, {Kraus}, {Batalha}, {Bloom}, {Cenko}, {Kasliwal}, {Kulkarni},
-%  {Nugent}, {Ofek}, {Poznanski}, \& {Quimby}}]{agueros11}
-%{Ag{\"u}eros}, M.~A., {Covey}, K.~R., {Lemonias}, J.~J., {et~al.} 2011, \apj,
-%  740, 110
-%  
-%\bibitem[{{Ahn} {et~al.}(2012){Ahn}, {Alexandroff}, {Allende Prieto},
-%  {Anderson}, {Anderton}, {Andrews}, {Aubourg}, {Bailey}, {Balbinot}, {Barnes},
-%  \& et~al.}]{dr9paper}
-%{Ahn}, C.~P., {Alexandroff}, R., {Allende Prieto}, C., {et~al.} 2012, \apjs,
-%  203, 21
-%
-%\bibitem[{{Alard} \& {Lupton}(1998)}]{alard1998}
-%{Alard}, C., \& {Lupton}, R.~H. 1998, \apj, 503, 325
-%
-%\bibitem[{{Alcock} {et~al.}(1993){Alcock}, {Allsman}, {Axelrod}, {Bennett},
-%  {Cook}, {Park}, {Marshall}, {Stubbs}, {Griest}, {Perlmutter}, {Sutherland},
-%  {Freeman}, {Peterson}, {Quinn}, \& {Rodgers}}]{original_macho}
-%{Alcock}, C., {Allsman}, R.~A., {Axelrod}, T.~S., {et~al.} 1993, in
-%  Astronomical Society of the Pacific Conference Series, Vol.~43, Sky Surveys.
-%  Protostars to Protogalaxies, ed. {B.~T.~Soifer}, 291
-%
-%\bibitem[{{Alcock} {et~al.}(2000){Alcock}, {Allsman}, {Alves}, {Axelrod},
-%  {Becker}, {Bennett}, {Cook}, {Drake}, {Freeman}, {Geha}, {Griest}, {Lehner},
-%  {Marshall}, {Minniti}, {Nelson}, {Peterson}, {Popowski}, {Pratt}, {Quinn},
-%  {Stubbs}, {Sutherland}, {Tomaney}, {Vandehei}, {Welch}, \& {MACHO
-%  Collaboration}}]{alcock2000}
-%{Alcock}, C., {Allsman}, R.~A., {Alves}, D.~R., {et~al.} 2000, \apj, 541, 734
-%
-%\bibitem[{{Alcock} {et~al.}(2001){Alcock}, {Allsman}, {Alves}, {Axelrod},
-%  {Becker}, {Bennett}, {Cook}, {Drake}, {Freeman}, {Geha}, {Griest}, {Lehner},
-%  {Marshall}, {Minniti}, {Nelson}, {Peterson}, {Popowski}, {Pratt}, {Quinn},
-%  {Stubbs}, {Sutherland}, {Tomaney}, {Vandehei}, \&
-%  {Welch}}]{macho_detection_efficiency}
-%{Alcock}, C., {Allsman}, R.~A., {Alves}, D., {et~al.} 2001, \apjs, 136, 439
-%
-%\bibitem[{{Aubourg} {et~al.}(1993){Aubourg}, {Bareyre}, {Brehin}, {Gros},
-%  {Lachieze-Rey}, {Laurent}, {Lesquoy}, {Magneville}, {Milsztajn}, {Moscoso},
-%  {Queinnec}, {Rich}, {Spiro}, {Vigroux}, {Zylberajch}, {Ansari}, {Cavalier},
-%  {Moniez}, {Beaulieu}, {Ferlet}, {Grison}, {Vidal-Madjar}, {Guibert},
-%  {Moreau}, {Tajahmady}, {Maurice}, {Prevot}, \& {Gry}}]{eros_original}
-%{Aubourg}, E., {Bareyre}, P., {Brehin}, S., {et~al.} 1993, The Messenger, 72,
-%  20
-%
-%\bibitem[{{Bertin} \& {Arnouts}(1996)}]{bertin96}
-%{Bertin}, E., \& {Arnouts}, S. 1996, \aaps, 117, 393
-%
-%\bibitem[{{Binney} {et~al.}(2000){Binney}, {Bissantz}, \&
-%  {Gerhard}}]{binney2000}
-%{Binney}, J., {Bissantz}, N., \& {Gerhard}, O. 2000, \apjl, 537, L99
-%
-%\bibitem[{{Bloom} {et~al.}(2012){Bloom}, {Richards}, {Nugent}, {Quimby},
-%  {Kasliwal}, {Starr}, {Poznanski}, {Ofek}, {Cenko}, {Butler}, {Kulkarni},
-%  {Gal-Yam}, \& {Law}}]{bloomPasp}
-%{Bloom}, J.~S., {Richards}, J.~W., {Nugent}, P.~E., {et~al.} 2012, \pasp, 124,
-%  1175
-%
-%\bibitem[{{Bochanski} {et~al.}(2010){Bochanski}, {Hawley}, {Covey}, {West},
-%  {Reid}, {Golimowski}, \& {Ivezi{\'c}}}]{Bochanski2010}
-%{Bochanski}, J.~J., {Hawley}, S.~L., {Covey}, K.~R., {et~al.} 2010, \aj, 139,
-%  2679
-%
-%\bibitem[{{Covey}(2006)}]{kev06}
-%{Covey}, K.~R. 2006, PhD thesis, University of Washington, Washington, USA
-%
-%\bibitem[{{Covey} {et~al.}(2007){Covey}, {Ivezi{\'c}}, {Schlegel},
-%  {Finkbeiner}, {Padmanabhan}, {Lupton}, {Ag{\"u}eros}, {Bochanski}, {Hawley},
-%  {West}, {Seth}, {Kimball}, {Gogarten}, {Claire}, {Haggard}, {Kaib},
-%  {Schneider}, \& {Sesar}}]{kev07}
-%{Covey}, K.~R., {Ivezi{\'c}}, {\v Z}., {Schlegel}, D., {et~al.} 2007, \aj, 134,
-%  2398
-%
-%\bibitem[{{Crotts} \& {Tomaney}(1996)}]{crotts1996}
-%{Crotts}, A.~P.~S., \& {Tomaney}, A.~B. 1996, \apjl, 473, L87
-%
-%\bibitem[{{Cutri} {et~al.}(2003)}]{cutri03}
-%{Cutri}, R.~M., {et~al.} 2003, {2MASS All Sky Catalog of point sources.} (The
-%  IRSA 2MASS All-Sky Point Source Catalog, NASA/IPAC Infrared Science
-%  Archive.~http://irsa.ipac.caltech.edu/applications/Gator/)
-%
-%\bibitem[{{Di Stefano}(2008)}]{mesolensing}
-%{Di Stefano}, R. 2008, \apj, 684, 46
-%
-%\bibitem[Durbin \& Watson(1950)]{durbin50}
-%J.~Durbin and G.~S. Watson.
-%\newblock Testing for serial correlation in least squares regression: I.
-%\newblock {\em Biometrika}, 37(3/4):pp. 409--428, 1950.
-%
-%\bibitem[{{Foreman-Mackey} {et~al.}(2012){Foreman-Mackey}, {Hogg}, {Lang}, \&
-%  {Goodman}}]{dfm}
-%{Foreman-Mackey}, D., {Hogg}, D.~W., {Lang}, D., \& {Goodman}, J. 2012, ArXiv
-%  e-prints
-%
-%\bibitem[{{Fukui} {et~al.}(2007){Fukui}, {Abe}, {Ayani}, {Fujii}, {Iizuka},
-%  {Itow}, {Kabumoto}, {Kamiya}, {Kawabata}, {Kawanomoto}, {Kinugasa}, {Koff},
-%  {Krajci}, {Naito}, {Nogami}, {Narusawa}, {Ohishi}, {Ohnishi}, {Sumi}, \&
-%  {Tsumuraya}}]{fukui2007}
-%{Fukui}, A., {Abe}, F., {Ayani}, K., {et~al.} 2007, \apj, 670, 423
-%
-%\bibitem[{{Gaudi}(2011)}]{gaudi2011}
-%{Gaudi}, B.~S. 2011, {Microlensing by Exoplanets}, ed. S.~{Piper}, 79--110
-%
-%\bibitem[{{Gaudi} {et~al.}(2008){Gaudi}, {Patterson}, {Spiegel}, {Krajci},
-%  {Koff}, {Pojma{\'n}ski}, {Dong}, {Gould}, {Prieto}, {Blake}, {Roming},
-%  {Bennett}, {Bloom}, {Boyd}, {Eyler}, {de Ponthi{\`e}re}, {Mirabal}, {Morgan},
-%  {Remillard}, {Vanmunster}, {Wagner}, \& {Watson}}]{gaudi2008}
-%{Gaudi}, B.~S., {Patterson}, J., {Spiegel}, D.~S., {et~al.} 2008, \apj, 677,
-%  1268
-%
-%\bibitem[{{Goodman} \& {Weare}(2010)}]{goodman}
-%{Goodman}, J., \& {Weare}, J. 2010, Comm. App. Math. Comp. Sci., 5
-%
-%\bibitem[{{Hamadache} {et~al.}(2006){Hamadache}, {Le Guillou}, {Tisserand},
-%  {Afonso}, {Albert}, {Andersen}, {Ansari}, {Aubourg}, {Bareyre}, {Beaulieu},
-%  {Charlot}, {Coutures}, {Ferlet}, {Fouqu{\'e}}, {Glicenstein}, {Goldman},
-%  {Gould}, {Graff}, {Gros}, {Haissinski}, {de Kat}, {Lesquoy}, {Loup},
-%  {Magneville}, {Marquette}, {Maurice}, {Maury}, {Milsztajn}, {Moniez},
-%  {Palanque-Delabrouille}, {Perdereau}, {Rahal}, {Rich}, {Spiro},
-%  {Vidal-Madjar}, {Vigroux}, \& {Zylberajch}}]{hamadache2009}
-%{Hamadache}, C., {Le Guillou}, L., {Tisserand}, P., {et~al.} 2006, \aap, 454,
-%  185
-%
-%\bibitem[{{Han}(2008)}]{han2008}
-%{Han}, C. 2008, \apj, 681, 806
-%
-%\bibitem[{{Kowalski} {et~al.}(2010){Kowalski}, {Hawley}, {Holtzman},
-%  {Wisniewski}, \& {Hilton}}]{kowal2010}
-%{Kowalski}, A.~F., {Hawley}, S.~L., {Holtzman}, J.~A., {Wisniewski}, J.~P., \&
-%  {Hilton}, E.~J. 2010, \apjl, 714, L98
-%
-%\bibitem[{{Law} {et~al.}(2009){Law}, {Kulkarni}, {Dekany}, {Ofek}, {Quimby},
-%  {Nugent}, {Surace}, {Grillmair}, {Bloom}, {Kasliwal}, {Bildsten}, {Brown},
-%  {Cenko}, {Ciardi}, {Croner}, {Djorgovski}, {van Eyken}, {Filippenko}, {Fox},
-%  {Gal-Yam}, {Hale}, {Hamam}, {Helou}, {Henning}, {Howell}, {Jacobsen},
-%  {Laher}, {Mattingly}, {McKenna}, {Pickles}, {Poznanski}, {Rahmer}, {Rau},
-%  {Rosing}, {Shara}, {Smith}, {Starr}, {Sullivan}, {Velur}, {Walters}, \&
-%  {Zolkower}}]{nick2009}
-%{Law}, N.~M., {Kulkarni}, S.~R., {Dekany}, R.~G., {et~al.} 2009, \pasp, 121,
-%  1395
-%
-%\bibitem[{{Law} {et~al.}(2010){Law}, {Dekany}, {Rahmer}, {Hale}, {Smith},
-%  {Quimby}, {Ofek}, {Kasliwal}, {Zolkower}, {Velur}, {Henning}, {Bui},
-%  {McKenna}, {Nugent}, {Jacobsen}, {Walters}, {Bloom}, {Surace}, {Grillmair},
-%  {Laher}, {Mattingly}, \& {Kulkarni}}]{nick2010}
-%{Law}, N.~M., {Dekany}, R.~G., {Rahmer}, G., {et~al.} 2010, in Society of
-%  Photo-Optical Instrumentation Engineers (SPIE) Conference Series, Vol. 7735
-%
-%\bibitem[{{Levitan} {et~al.}(2011){Levitan}, {Fulton}, {Groot}, {Kulkarni},
-%  {Ofek}, {Prince}, {Shporer}, {Bloom}, {Cenko}, {Kasliwal}, {Law}, {Nugent},
-%  {Poznanski}, {Quimby}, {Horesh}, {Sesar}, \& {Sternberg}}]{levitan2011}
-%{Levitan}, D., {Fulton}, B.~J., {Groot}, P.~J., {et~al.} 2011, \apj, 739, 68
-%
-%\bibitem[{{Nieto--Santisteban} {et~al.}(2004)}]{sdss_images}
-%{Nieto--Santisteban}, M.~A., {et~al.} 2004, in ASP Conf. Ser. 314: Astronomical
-%  Data Analysis Software and Systems (ADASS) XIII, ed. F.~{Ochsenbein}, M.~G.
-%  {Allen}, \& D.~{Egret}, 666--+
-%
-%\bibitem[{{Ofek} {et~al.}(2012){Ofek}, {Laher}, {Surace}, {Levitan}, {Sesar},
-%  {Horesh}, {Law}, {van Eyken}, {Kulkarni}, {Prince}, {Nugent}, {Sullivan},
-%  {Yaron}, {Pickles}, {Ag{\"u}eros}, {Arcavi}, {Bildsten}, {Bloom}, {Cenko},
-%  {Gal-Yam}, {Grillmair}, {Helou}, {Kasliwal}, {Poznanski}, \&
-%  {Quimby}}]{ofek2012}
-%{Ofek}, E.~O., {Laher}, R., {Surace}, J., {et~al.} 2012, \pasp, 124, 854
-%
-%\bibitem[{{Oke} \& {Gunn}(1982)}]{oke_gunn}
-%{Oke}, J.~B., \& {Gunn}, J.~E. 1982, \pasp, 94, 586
-%
-%\bibitem[{{Os{\l}owski} {et~al.}(2008){Os{\l}owski}, {Moderski}, {Bulik}, \&
-%  {Belczynski}}]{oslowski2008}
-%{Os{\l}owski}, S., {Moderski}, R., {Bulik}, T., \& {Belczynski}, K. 2008, \aap,
-%  478, 429
-%
-%\bibitem[{{Paczynski}(1986)}]{paczynski1986}
-%{Paczynski}, B. 1986, \apj, 304, 1
-%
-%\bibitem[{{Pickles}(1998)}]{Pickles1998}
-%{Pickles}, A.~J. 1998, \pasp, 110, 863
-%
-%\bibitem[{{Polishook} {et~al.}(2012){Polishook}, {Ofek}, {Waszczak},
-%  {Kulkarni}, {Gal-Yam}, {Aharonson}, {Laher}, {Surace}, {Klein}, {Bloom},
-%  {Brosch}, {Prialnik}, {Grillmair}, {Cenko}, {Kasliwal}, {Law}, {Levitan},
-%  {Nugent}, {Poznanski}, \& {Quimby}}]{polishook2012}
-%{Polishook}, D., {Ofek}, E.~O., {Waszczak}, A., {et~al.} 2012, \mnras, 421,
-%  2094
-%
-%\bibitem[{{Rahmer} {et~al.}(2008){Rahmer}, {Smith}, {Velur}, {Hale}, {Law},
-%  {Bui}, {Petrie}, \& {Dekany}}]{rahmer2008}
-%{Rahmer}, G., {Smith}, R., {Velur}, V., {et~al.} 2008, in Society of
-%  Photo-Optical Instrumentation Engineers (SPIE) Conference Series, Vol. 7014
-%
-%\bibitem[{{Rattenbury} {et~al.}(2005){Rattenbury}, {Abe}, {Bennett}, {Bond},
-%  {Calitz}, {Claret}, {Cook}, {Furuta}, {Gal-Yam}, {Glicenstein}, {Hearnshaw},
-%  {Hauschildt}, {Kilmartin}, {Kurata}, {Masuda}, {Maoz}, {Matsubara},
-%  {Meintjes}, {Moniez}, {Muraki}, {Noda}, {Ofek}, {Okajima}, {Philpott},
-%  {Rhie}, {Sako}, {Sullivan}, {Sumi}, {Terndrup}, {Tristram}, {Wood},
-%  {Yanagisawa}, \& {Yock}}]{rattenbury2005}
-%{Rattenbury}, N.~J., {Abe}, F., {Bennett}, D.~P., {et~al.} 2005, \aap, 439, 645
-%
-%\bibitem[{{Rau} {et~al.}(2009){Rau}, {Kulkarni}, {Law}, {Bloom}, {Ciardi},
-%  {Djorgovski}, {Fox}, {Gal-Yam}, {Grillmair}, {Kasliwal}, {Nugent}, {Ofek},
-%  {Quimby}, {Reach}, {Shara}, {Bildsten}, {Cenko}, {Drake}, {Filippenko},
-%  {Helfand}, {Helou}, {Howell}, {Poznanski}, \& {Sullivan}}]{rau2009}
-%{Rau}, A., {Kulkarni}, S.~R., {Law}, N.~M., {et~al.} 2009, \pasp, 121, 1334
-%
-%\bibitem[{{Richards} {et~al.}(2002)}]{richards02}
-%{Richards}, G.~T., {et~al.} 2002, \aj, 123, 2945
-%
-%\bibitem[{{Sartore} \& {Treves}(2010)}]{sartore2010}
-%{Sartore}, N., \& {Treves}, A. 2010, \aap, 523, A33
-%
-%\bibitem[{{Shin} {et~al.}(2009){Shin}, {Sekora}, \& {Byun}}]{shin2009}
-%{Shin}, M.-S., {Sekora}, M., \& {Byun}, Y.-I. 2009, \mnras, 400, 1897
-%
-%\bibitem[{{Stetson}(1996)}]{stetson1996}
-%{Stetson}, P.~B. 1996, \pasp, 108, 851
-%
-%\bibitem[{{Stoughton} {et~al.}(2002){Stoughton}, {Lupton}, {Bernardi},
-%  {Blanton}, {Burles}, {Castander}, {Connolly}, {Eisenstein}, {Frieman},
-%  {Hennessy}, {Hindsley}, {Ivezi{\'c}}, {Kent}, {Kunszt}, {Lee}, {Meiksin},
-%  {Munn}, {Newberg}, {Nichol}, {Nicinski}, {Pier}, {Richards}, {Richmond},
-%  {Schlegel}, {Smith}, {Strauss}, {SubbaRao}, {Szalay}, {Thakar}, {Tucker},
-%  {Vanden Berk}, {Yanny}, {Adelman}, {Anderson}, {Anderson}, {Annis},
-%  {Bahcall}, {Bakken}, {Bartelmann}, {Bastian}, {Bauer}, {Berman},
-%  {B{\"o}hringer}, {Boroski}, {Bracker}, {Briegel}, {Briggs}, {Brinkmann},
-%  {Brunner}, {Carey}, {Carr}, {Chen}, {Christian}, {Colestock}, {Crocker},
-%  {Csabai}, {Czarapata}, {Dalcanton}, {Davidsen}, {Davis}, {Dehnen},
-%  {Dodelson}, {Doi}, {Dombeck}, {Donahue}, {Ellman}, {Elms}, {Evans}, {Eyer},
-%  {Fan}, {Federwitz}, {Friedman}, {Fukugita}, {Gal}, {Gillespie}, {Glazebrook},
-%  {Gray}, {Grebel}, {Greenawalt}, {Greene}, {Gunn}, {de Haas}, {Haiman},
-%  {Haldeman}, {Hall}, {Hamabe}, {Hansen}, {Harris}, {Harris}, {Harvanek},
-%  {Hawley}, {Hayes}, {Heckman}, {Helmi}, {Henden}, {Hogan}, {Hogg}, {Holmgren},
-%  {Holtzman}, {Huang}, {Hull}, {Ichikawa}, {Ichikawa}, {Johnston}, {Kauffmann},
-%  {Kim}, {Kimball}, {Kinney}, {Klaene}, {Kleinman}, {Klypin}, {Knapp},
-%  {Korienek}, {Krolik}, {Kron}, {Krzesi{\'n}ski}, {Lamb}, {Leger},
-%  {Limmongkol}, {Lindenmeyer}, {Long}, {Loomis}, {Loveday}, {MacKinnon},
-%  {Mannery}, {Mantsch}, {Margon}, {McGehee}, {McKay}, {McLean}, {Menou},
-%  {Merelli}, {Mo}, {Monet}, {Nakamura}, {Narayanan}, {Nash}, {Neilsen},
-%  {Newman}, {Nitta}, {Odenkirchen}, {Okada}, {Okamura}, {Ostriker}, {Owen},
-%  {Pauls}, {Peoples}, {Peterson}, {Petravick}, {Pope}, {Pordes}, {Postman},
-%  {Prosapio}, {Quinn}, {Rechenmacher}, {Rivetta}, {Rix}, {Rockosi}, {Rosner},
-%  {Ruthmansdorfer}, {Sandford}, {Schneider}, {Scranton}, {Sekiguchi}, {Sergey},
-%  {Sheth}, {Shimasaku}, {Smee}, {Snedden}, {Stebbins}, {Stubbs}, {Szapudi},
-%  {Szkody}, {Szokoly}, {Tabachnik}, {Tsvetanov}, {Uomoto}, {Vogeley}, {Voges},
-%  {Waddell}, {Walterbos}, {Wang}, {Watanabe}, {Weinberg}, {White}, {White},
-%  {Wilhite}, {Wolfe}, {Yasuda}, {York}, {Zehavi}, \& {Zheng}}]{stoughton02}
-%{Stoughton}, C., {Lupton}, R.~H., {Bernardi}, M., {et~al.} 2002, \aj, 123, 485
-%
-%\bibitem[{{Sumi} {et~al.}(2011){Sumi}, {Kamiya}, {Bennett}, {Bond}, {Abe},
-%  {Botzler}, {Fukui}, {Furusawa}, {Hearnshaw}, {Itow}, {Kilmartin}, {Korpela},
-%  {Lin}, {Ling}, {Masuda}, {Matsubara}, {Miyake}, {Motomura}, {Muraki},
-%  {Nagaya}, {Nakamura}, {Ohnishi}, {Okumura}, {Perrott}, {Rattenbury}, {Saito},
-%  {Sako}, {Sullivan}, {Sweatman}, {Tristram}, {Udalski}, {Szyma{\'n}ski},
-%  {Kubiak}, {Pietrzy{\'n}ski}, {Poleski}, {Soszy{\'n}ski}, {Wyrzykowski},
-%  {Ulaczyk}, \& {Microlensing Observations in Astrophysics (MOA)
-%  Collaboration}}]{sumi2011}
-%{Sumi}, T., {Kamiya}, K., {Bennett}, D.~P., {et~al.} 2011, \nat, 473, 349
-%
-%\bibitem[{{Tomov} {et~al.}(2013){Tomov}, {Tomova}, \& {Bisikalo}}]{tomov2013}
-%{Tomov}, N.~A., {Tomova}, M.~T., \& {Bisikalo}, D.~V. 2013, ArXiv e-prints
-%
-%\bibitem[{{Udalski}(2003)}]{udalski03}
-%{Udalski}, A. 2003, \actaa, 53, 291
-%
-%\bibitem[{{Udalski} {et~al.}(1992){Udalski}, {Szymanski}, {Kaluzny}, {Kubiak},
-%  \& {Mateo}}]{original_ogle}
-%{Udalski}, A., {Szymanski}, M., {Kaluzny}, J., {Kubiak}, M., \& {Mateo}, M.
-%  1992, \actaa, 42, 253
-%
-%\bibitem[{{Udalski} {et~al.}(1994){Udalski}, {Szymanski}, {Stanek}, {Kaluzny},
-%  {Kubiak}, {Mateo}, {Krzeminski}, {Paczynski}, \&
-%  {Venkat}}]{ogle_optical_depth}
-%{Udalski}, A., {Szymanski}, M., {Stanek}, K.~Z., {et~al.} 1994, \actaa, 44, 165
-%
-%\bibitem[{{von Neumann} {et~al.}(1941){von Neumann}, {Kent}, {Bellinson}, \&
-%  {Hart}}]{von_neumann1941}
-%{von Neumann}, J., {Kent}, R.~H., {Bellinson}, H.~R., \& {Hart}, B.~I. 1941,
-%  The Annals of Mathematical Statistics, 12, 153
-%
-%\bibitem[{{Wozniak}(2000)}]{con_idx}
-%{Wozniak}, P.~R. 2000, \actaa, 50, 421
-%
-%\bibitem[{{Wright} {et~al.}(2010){Wright}, {Eisenhardt}, {Mainzer}, {Ressler},
-%  {Cutri}, {Jarrett}, {Kirkpatrick}, {Padgett}, {McMillan}, {Skrutskie},
-%  {Stanford}, {Cohen}, {Walker}, {Mather}, {Leisawitz}, {Gautier}, {McLean},
-%  {Benford}, {Lonsdale}, {Blain}, {Mendez}, {Irace}, {Duval}, {Liu}, {Royer},
-%  {Heinrichsen}, {Howard}, {Shannon}, {Kendall}, {Walsh}, {Larsen}, {Cardon},
-%  {Schick}, {Schwalm}, {Abid}, {Fabinsky}, {Naes}, \& {Tsai}}]{WISE}
-%{Wright}, E.~L., {Eisenhardt}, P.~R.~M., {Mainzer}, A.~K., {et~al.} 2010, \aj,
-%  140, 1868
-%
-%\bibitem[{{Wyrzykowski} {et~al.}(2009){Wyrzykowski}, {Koz{\l}owski}, {Skowron},
-%  {Belokurov}, {Smith}, {Udalski}, {Szyma{\'n}ski}, {Kubiak},
-%  {Pietrzy{\'n}ski}, {Soszy{\'n}ski}, {Szewczyk}, \&
-%  {{\.Z}ebru{\'n}}}]{wyrzykowski2009}
-%{Wyrzykowski}, {\L}., {Koz{\l}owski}, S., {Skowron}, J., {et~al.} 2009, \mnras,
-%  397, 1228
-%
-%\bibitem[{Yan {et~al.}(2013)Yan, Donoso, Tsai, Stern, Assef, Eisenhardt, Blain,
-%  Cutri, Jarrett, Stanford, Wright, Bridge, \& Riechers}]{yan2013}
-%Yan, L., Donoso, E., Tsai, C.-W., {et~al.} 2013, \aj, 145, 55
-%
-%\bibitem[{{York} {et~al.}(2000)}]{york00}
-%{York}, D.~G., {et~al.} 2000, \aj, 120, 1579
-%
-%\end{thebibliography}
+\begin{thebibliography}{56}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+
+\bibitem[{{Ag{\"u}eros} {et~al.}(2011){Ag{\"u}eros}, {Covey}, {Lemonias},
+  {Law}, {Kraus}, {Batalha}, {Bloom}, {Cenko}, {Kasliwal}, {Kulkarni},
+  {Nugent}, {Ofek}, {Poznanski}, \& {Quimby}}]{agueros11}
+{Ag{\"u}eros}, M.~A., {Covey}, K.~R., {Lemonias}, J.~J., {et~al.} 2011, \apj,
+  740, 110
+
+\bibitem[{{Ahn} {et~al.}(2012){Ahn}, {Alexandroff}, {Allende Prieto},
+  {Anderson}, {Anderton}, {Andrews}, {Aubourg}, {Bailey}, {Balbinot}, {Barnes},
+  \& et~al.}]{dr9paper}
+{Ahn}, C.~P., {Alexandroff}, R., {Allende Prieto}, C., {et~al.} 2012, \apjs,
+  203, 21
+
+\bibitem[{{Alard} \& {Lupton}(1998)}]{alard1998}
+{Alard}, C., \& {Lupton}, R.~H. 1998, \apj, 503, 325
+
+\bibitem[{{Alcock} {et~al.}(1993){Alcock}, {Allsman}, {Axelrod}, {Bennett},
+  {Cook}, {Park}, {Marshall}, {Stubbs}, {Griest}, {Perlmutter}, {Sutherland},
+  {Freeman}, {Peterson}, {Quinn}, \& {Rodgers}}]{original_macho}
+{Alcock}, C., {Allsman}, R.~A., {Axelrod}, T.~S., {et~al.} 1993, in
+  Astronomical Society of the Pacific Conference Series, Vol.~43, Sky Surveys.
+  Protostars to Protogalaxies, ed. {B.~T.~Soifer}, 291
+
+\bibitem[{{Alcock} {et~al.}(2000){Alcock}, {Allsman}, {Alves}, {Axelrod},
+  {Becker}, {Bennett}, {Cook}, {Drake}, {Freeman}, {Geha}, {Griest}, {Lehner},
+  {Marshall}, {Minniti}, {Nelson}, {Peterson}, {Popowski}, {Pratt}, {Quinn},
+  {Stubbs}, {Sutherland}, {Tomaney}, {Vandehei}, {Welch}, \& {MACHO
+  Collaboration}}]{alcock2000}
+{Alcock}, C., {Allsman}, R.~A., {Alves}, D.~R., {et~al.} 2000, \apj, 541, 734
+
+\bibitem[{{Alcock} {et~al.}(2001){Alcock}, {Allsman}, {Alves}, {Axelrod},
+  {Becker}, {Bennett}, {Cook}, {Drake}, {Freeman}, {Geha}, {Griest}, {Lehner},
+  {Marshall}, {Minniti}, {Nelson}, {Peterson}, {Popowski}, {Pratt}, {Quinn},
+  {Stubbs}, {Sutherland}, {Tomaney}, {Vandehei}, \&
+  {Welch}}]{macho_detection_efficiency}
+{Alcock}, C., {Allsman}, R.~A., {Alves}, D., {et~al.} 2001, \apjs, 136, 439
+
+\bibitem[{{Aubourg} {et~al.}(1993){Aubourg}, {Bareyre}, {Brehin}, {Gros},
+  {Lachieze-Rey}, {Laurent}, {Lesquoy}, {Magneville}, {Milsztajn}, {Moscoso},
+  {Queinnec}, {Rich}, {Spiro}, {Vigroux}, {Zylberajch}, {Ansari}, {Cavalier},
+  {Moniez}, {Beaulieu}, {Ferlet}, {Grison}, {Vidal-Madjar}, {Guibert},
+  {Moreau}, {Tajahmady}, {Maurice}, {Prevot}, \& {Gry}}]{eros_original}
+{Aubourg}, E., {Bareyre}, P., {Brehin}, S., {et~al.} 1993, The Messenger, 72,
+  20
+
+\bibitem[{{Bertin} \& {Arnouts}(1996)}]{bertin96}
+{Bertin}, E., \& {Arnouts}, S. 1996, \aaps, 117, 393
+
+\bibitem[{{Binney} {et~al.}(2000){Binney}, {Bissantz}, \&
+  {Gerhard}}]{binney2000}
+{Binney}, J., {Bissantz}, N., \& {Gerhard}, O. 2000, \apjl, 537, L99
+
+\bibitem[{{Bloom} {et~al.}(2012){Bloom}, {Richards}, {Nugent}, {Quimby},
+  {Kasliwal}, {Starr}, {Poznanski}, {Ofek}, {Cenko}, {Butler}, {Kulkarni},
+  {Gal-Yam}, \& {Law}}]{bloomPasp}
+{Bloom}, J.~S., {Richards}, J.~W., {Nugent}, P.~E., {et~al.} 2012, \pasp, 124,
+  1175
+
+\bibitem[{{Bochanski} {et~al.}(2010){Bochanski}, {Hawley}, {Covey}, {West},
+  {Reid}, {Golimowski}, \& {Ivezi{\'c}}}]{Bochanski2010}
+{Bochanski}, J.~J., {Hawley}, S.~L., {Covey}, K.~R., {et~al.} 2010, \aj, 139,
+  2679
+
+\bibitem[{{Brink} {et~al.}(2013){Brink}, {Richards}, {Poznanski}, {Bloom},
+  {Rice}, {Negahban}, \& {Wainwright}}]{brink2013}
+{Brink}, H., {Richards}, J.~W., {Poznanski}, D., {et~al.} 2013, \mnras, 435,
+  1047
+
+\bibitem[{{Covey}(2006)}]{kev06}
+{Covey}, K.~R. 2006, PhD thesis, University of Washington, Washington, USA
+
+\bibitem[{{Covey} {et~al.}(2007){Covey}, {Ivezi{\'c}}, {Schlegel},
+  {Finkbeiner}, {Padmanabhan}, {Lupton}, {Ag{\"u}eros}, {Bochanski}, {Hawley},
+  {West}, {Seth}, {Kimball}, {Gogarten}, {Claire}, {Haggard}, {Kaib},
+  {Schneider}, \& {Sesar}}]{kev07}
+{Covey}, K.~R., {Ivezi{\'c}}, {\v Z}., {Schlegel}, D., {et~al.} 2007, \aj, 134,
+  2398
+
+\bibitem[{{Crotts} \& {Tomaney}(1996)}]{crotts1996}
+{Crotts}, A.~P.~S., \& {Tomaney}, A.~B. 1996, \apjl, 473, L87
+
+\bibitem[{{Cutri} {et~al.}(2003)}]{cutri03}
+{Cutri}, R.~M., {et~al.} 2003, {2MASS All Sky Catalog of point sources.} (The
+  IRSA 2MASS All-Sky Point Source Catalog, NASA/IPAC Infrared Science
+  Archive.~http://irsa.ipac.caltech.edu/applications/Gator/)
+
+\bibitem[{{Di Stefano}(2008{\natexlab{a}})}]{mesolensing2}
+{Di Stefano}, R. 2008{\natexlab{a}}, \apj, 684, 59
+
+\bibitem[{{Di Stefano}(2008{\natexlab{b}})}]{mesolensing1}
+---. 2008{\natexlab{b}}, \apj, 684, 46
+
+\bibitem[Durbin \& Watson(1950)]{durbin50}
+J.~Durbin and G.~S. Watson.
+\newblock Testing for serial correlation in least squares regression: I.
+\newblock {\em Biometrika}, 37(3/4):pp. 409--428, 1950.
+
+\bibitem[{{Foreman-Mackey} {et~al.}(2012){Foreman-Mackey}, {Hogg}, {Lang}, \&
+  {Goodman}}]{dfm}
+{Foreman-Mackey}, D., {Hogg}, D.~W., {Lang}, D., \& {Goodman}, J. 2012, ArXiv
+  e-prints
+
+\bibitem[{{Fukui} {et~al.}(2007){Fukui}, {Abe}, {Ayani}, {Fujii}, {Iizuka},
+  {Itow}, {Kabumoto}, {Kamiya}, {Kawabata}, {Kawanomoto}, {Kinugasa}, {Koff},
+  {Krajci}, {Naito}, {Nogami}, {Narusawa}, {Ohishi}, {Ohnishi}, {Sumi}, \&
+  {Tsumuraya}}]{fukui2007}
+{Fukui}, A., {Abe}, F., {Ayani}, K., {et~al.} 2007, \apj, 670, 423
+
+\bibitem[{{Gaudi}(2011)}]{gaudi2011}
+{Gaudi}, B.~S. 2011, {Microlensing by Exoplanets}, ed. S.~{Piper}, 79--110
+
+\bibitem[{{Gaudi} {et~al.}(2008){Gaudi}, {Patterson}, {Spiegel}, {Krajci},
+  {Koff}, {Pojma{\'n}ski}, {Dong}, {Gould}, {Prieto}, {Blake}, {Roming},
+  {Bennett}, {Bloom}, {Boyd}, {Eyler}, {de Ponthi{\`e}re}, {Mirabal}, {Morgan},
+  {Remillard}, {Vanmunster}, {Wagner}, \& {Watson}}]{gaudi2008}
+{Gaudi}, B.~S., {Patterson}, J., {Spiegel}, D.~S., {et~al.} 2008, \apj, 677,
+  1268
+
+\bibitem[{{Goodman} \& {Weare}(2010)}]{goodman}
+{Goodman}, J., \& {Weare}, J. 2010, Comm. App. Math. Comp. Sci., 5
+
+\bibitem[{{Hamadache} {et~al.}(2006){Hamadache}, {Le Guillou}, {Tisserand},
+  {Afonso}, {Albert}, {Andersen}, {Ansari}, {Aubourg}, {Bareyre}, {Beaulieu},
+  {Charlot}, {Coutures}, {Ferlet}, {Fouqu{\'e}}, {Glicenstein}, {Goldman},
+  {Gould}, {Graff}, {Gros}, {Haissinski}, {de Kat}, {Lesquoy}, {Loup},
+  {Magneville}, {Marquette}, {Maurice}, {Maury}, {Milsztajn}, {Moniez},
+  {Palanque-Delabrouille}, {Perdereau}, {Rahal}, {Rich}, {Spiro},
+  {Vidal-Madjar}, {Vigroux}, \& {Zylberajch}}]{hamadache2009}
+{Hamadache}, C., {Le Guillou}, L., {Tisserand}, P., {et~al.} 2006, \aap, 454,
+  185
+
+\bibitem[{{Han}(2008)}]{han2008}
+{Han}, C. 2008, \apj, 681, 806
+
+\bibitem[{{Kowalski} {et~al.}(2010){Kowalski}, {Hawley}, {Holtzman},
+  {Wisniewski}, \& {Hilton}}]{kowal2010}
+{Kowalski}, A.~F., {Hawley}, S.~L., {Holtzman}, J.~A., {Wisniewski}, J.~P., \&
+  {Hilton}, E.~J. 2010, \apjl, 714, L98
+
+\bibitem[{{Law} {et~al.}(2009){Law}, {Kulkarni}, {Dekany}, {Ofek}, {Quimby},
+  {Nugent}, {Surace}, {Grillmair}, {Bloom}, {Kasliwal}, {Bildsten}, {Brown},
+  {Cenko}, {Ciardi}, {Croner}, {Djorgovski}, {van Eyken}, {Filippenko}, {Fox},
+  {Gal-Yam}, {Hale}, {Hamam}, {Helou}, {Henning}, {Howell}, {Jacobsen},
+  {Laher}, {Mattingly}, {McKenna}, {Pickles}, {Poznanski}, {Rahmer}, {Rau},
+  {Rosing}, {Shara}, {Smith}, {Starr}, {Sullivan}, {Velur}, {Walters}, \&
+  {Zolkower}}]{nick2009}
+{Law}, N.~M., {Kulkarni}, S.~R., {Dekany}, R.~G., {et~al.} 2009, \pasp, 121,
+  1395
+
+\bibitem[{{Law} {et~al.}(2010){Law}, {Dekany}, {Rahmer}, {Hale}, {Smith},
+  {Quimby}, {Ofek}, {Kasliwal}, {Zolkower}, {Velur}, {Henning}, {Bui},
+  {McKenna}, {Nugent}, {Jacobsen}, {Walters}, {Bloom}, {Surace}, {Grillmair},
+  {Laher}, {Mattingly}, \& {Kulkarni}}]{nick2010}
+{Law}, N.~M., {Dekany}, R.~G., {Rahmer}, G., {et~al.} 2010, in Society of
+  Photo-Optical Instrumentation Engineers (SPIE) Conference Series, Vol. 7735
+
+\bibitem[{{Levitan} {et~al.}(2011){Levitan}, {Fulton}, {Groot}, {Kulkarni},
+  {Ofek}, {Prince}, {Shporer}, {Bloom}, {Cenko}, {Kasliwal}, {Law}, {Nugent},
+  {Poznanski}, {Quimby}, {Horesh}, {Sesar}, \& {Sternberg}}]{levitan2011}
+{Levitan}, D., {Fulton}, B.~J., {Groot}, P.~J., {et~al.} 2011, \apj, 739, 68
+
+\bibitem[{{Nieto--Santisteban} {et~al.}(2004)}]{sdss_images}
+{Nieto--Santisteban}, M.~A., {et~al.} 2004, in ASP Conf. Ser. 314: Astronomical
+  Data Analysis Software and Systems (ADASS) XIII, ed. F.~{Ochsenbein}, M.~G.
+  {Allen}, \& D.~{Egret}, 666--+
+
+\bibitem[{{Ofek} {et~al.}(2012){Ofek}, {Laher}, {Surace}, {Levitan}, {Sesar},
+  {Horesh}, {Law}, {van Eyken}, {Kulkarni}, {Prince}, {Nugent}, {Sullivan},
+  {Yaron}, {Pickles}, {Ag{\"u}eros}, {Arcavi}, {Bildsten}, {Bloom}, {Cenko},
+  {Gal-Yam}, {Grillmair}, {Helou}, {Kasliwal}, {Poznanski}, \&
+  {Quimby}}]{ofek2012}
+{Ofek}, E.~O., {Laher}, R., {Surace}, J., {et~al.} 2012, \pasp, 124, 854
+
+\bibitem[{{Oke} \& {Gunn}(1982)}]{oke_gunn}
+{Oke}, J.~B., \& {Gunn}, J.~E. 1982, \pasp, 94, 586
+
+\bibitem[{{Os{\l}owski} {et~al.}(2008){Os{\l}owski}, {Moderski}, {Bulik}, \&
+  {Belczynski}}]{oslowski2008}
+{Os{\l}owski}, S., {Moderski}, R., {Bulik}, T., \& {Belczynski}, K. 2008, \aap,
+  478, 429
+
+\bibitem[{{Paczynski}(1986)}]{paczynski1986}
+{Paczynski}, B. 1986, \apj, 304, 1
+
+\bibitem[{{Pickles}(1998)}]{Pickles1998}
+{Pickles}, A.~J. 1998, \pasp, 110, 863
+
+\bibitem[{{Polishook} {et~al.}(2012){Polishook}, {Ofek}, {Waszczak},
+  {Kulkarni}, {Gal-Yam}, {Aharonson}, {Laher}, {Surace}, {Klein}, {Bloom},
+  {Brosch}, {Prialnik}, {Grillmair}, {Cenko}, {Kasliwal}, {Law}, {Levitan},
+  {Nugent}, {Poznanski}, \& {Quimby}}]{polishook2012}
+{Polishook}, D., {Ofek}, E.~O., {Waszczak}, A., {et~al.} 2012, \mnras, 421,
+  2094
+
+\bibitem[{{Rahmer} {et~al.}(2008){Rahmer}, {Smith}, {Velur}, {Hale}, {Law},
+  {Bui}, {Petrie}, \& {Dekany}}]{rahmer2008}
+{Rahmer}, G., {Smith}, R., {Velur}, V., {et~al.} 2008, in Society of
+  Photo-Optical Instrumentation Engineers (SPIE) Conference Series, Vol. 7014
+
+\bibitem[{{Rattenbury} {et~al.}(2005){Rattenbury}, {Abe}, {Bennett}, {Bond},
+  {Calitz}, {Claret}, {Cook}, {Furuta}, {Gal-Yam}, {Glicenstein}, {Hearnshaw},
+  {Hauschildt}, {Kilmartin}, {Kurata}, {Masuda}, {Maoz}, {Matsubara},
+  {Meintjes}, {Moniez}, {Muraki}, {Noda}, {Ofek}, {Okajima}, {Philpott},
+  {Rhie}, {Sako}, {Sullivan}, {Sumi}, {Terndrup}, {Tristram}, {Wood},
+  {Yanagisawa}, \& {Yock}}]{rattenbury2005}
+{Rattenbury}, N.~J., {Abe}, F., {Bennett}, D.~P., {et~al.} 2005, \aap, 439, 645
+
+\bibitem[{{Rau} {et~al.}(2009){Rau}, {Kulkarni}, {Law}, {Bloom}, {Ciardi},
+  {Djorgovski}, {Fox}, {Gal-Yam}, {Grillmair}, {Kasliwal}, {Nugent}, {Ofek},
+  {Quimby}, {Reach}, {Shara}, {Bildsten}, {Cenko}, {Drake}, {Filippenko},
+  {Helfand}, {Helou}, {Howell}, {Poznanski}, \& {Sullivan}}]{rau2009}
+{Rau}, A., {Kulkarni}, S.~R., {Law}, N.~M., {et~al.} 2009, \pasp, 121, 1334
+
+\bibitem[{{Richards} {et~al.}(2002)}]{richards02}
+{Richards}, G.~T., {et~al.} 2002, \aj, 123, 2945
+
+\bibitem[{{Sartore} \& {Treves}(2010)}]{sartore2010}
+{Sartore}, N., \& {Treves}, A. 2010, \aap, 523, A33
+
+\bibitem[{{Shin} {et~al.}(2009){Shin}, {Sekora}, \& {Byun}}]{shin2009}
+{Shin}, M.-S., {Sekora}, M., \& {Byun}, Y.-I. 2009, \mnras, 400, 1897
+
+\bibitem[{{Stetson}(1996)}]{stetson1996}
+{Stetson}, P.~B. 1996, \pasp, 108, 851
+
+\bibitem[{{Stoughton} {et~al.}(2002){Stoughton}, {Lupton}, {Bernardi},
+  {Blanton}, {Burles}, {Castander}, {Connolly}, {Eisenstein}, {Frieman},
+  {Hennessy}, {Hindsley}, {Ivezi{\'c}}, {Kent}, {Kunszt}, {Lee}, {Meiksin},
+  {Munn}, {Newberg}, {Nichol}, {Nicinski}, {Pier}, {Richards}, {Richmond},
+  {Schlegel}, {Smith}, {Strauss}, {SubbaRao}, {Szalay}, {Thakar}, {Tucker},
+  {Vanden Berk}, {Yanny}, {Adelman}, {Anderson}, {Anderson}, {Annis},
+  {Bahcall}, {Bakken}, {Bartelmann}, {Bastian}, {Bauer}, {Berman},
+  {B{\"o}hringer}, {Boroski}, {Bracker}, {Briegel}, {Briggs}, {Brinkmann},
+  {Brunner}, {Carey}, {Carr}, {Chen}, {Christian}, {Colestock}, {Crocker},
+  {Csabai}, {Czarapata}, {Dalcanton}, {Davidsen}, {Davis}, {Dehnen},
+  {Dodelson}, {Doi}, {Dombeck}, {Donahue}, {Ellman}, {Elms}, {Evans}, {Eyer},
+  {Fan}, {Federwitz}, {Friedman}, {Fukugita}, {Gal}, {Gillespie}, {Glazebrook},
+  {Gray}, {Grebel}, {Greenawalt}, {Greene}, {Gunn}, {de Haas}, {Haiman},
+  {Haldeman}, {Hall}, {Hamabe}, {Hansen}, {Harris}, {Harris}, {Harvanek},
+  {Hawley}, {Hayes}, {Heckman}, {Helmi}, {Henden}, {Hogan}, {Hogg}, {Holmgren},
+  {Holtzman}, {Huang}, {Hull}, {Ichikawa}, {Ichikawa}, {Johnston}, {Kauffmann},
+  {Kim}, {Kimball}, {Kinney}, {Klaene}, {Kleinman}, {Klypin}, {Knapp},
+  {Korienek}, {Krolik}, {Kron}, {Krzesi{\'n}ski}, {Lamb}, {Leger},
+  {Limmongkol}, {Lindenmeyer}, {Long}, {Loomis}, {Loveday}, {MacKinnon},
+  {Mannery}, {Mantsch}, {Margon}, {McGehee}, {McKay}, {McLean}, {Menou},
+  {Merelli}, {Mo}, {Monet}, {Nakamura}, {Narayanan}, {Nash}, {Neilsen},
+  {Newman}, {Nitta}, {Odenkirchen}, {Okada}, {Okamura}, {Ostriker}, {Owen},
+  {Pauls}, {Peoples}, {Peterson}, {Petravick}, {Pope}, {Pordes}, {Postman},
+  {Prosapio}, {Quinn}, {Rechenmacher}, {Rivetta}, {Rix}, {Rockosi}, {Rosner},
+  {Ruthmansdorfer}, {Sandford}, {Schneider}, {Scranton}, {Sekiguchi}, {Sergey},
+  {Sheth}, {Shimasaku}, {Smee}, {Snedden}, {Stebbins}, {Stubbs}, {Szapudi},
+  {Szkody}, {Szokoly}, {Tabachnik}, {Tsvetanov}, {Uomoto}, {Vogeley}, {Voges},
+  {Waddell}, {Walterbos}, {Wang}, {Watanabe}, {Weinberg}, {White}, {White},
+  {Wilhite}, {Wolfe}, {Yasuda}, {York}, {Zehavi}, \& {Zheng}}]{stoughton02}
+{Stoughton}, C., {Lupton}, R.~H., {Bernardi}, M., {et~al.} 2002, \aj, 123, 485
+
+\bibitem[{{Sumi} {et~al.}(2011){Sumi}, {Kamiya}, {Bennett}, {Bond}, {Abe},
+  {Botzler}, {Fukui}, {Furusawa}, {Hearnshaw}, {Itow}, {Kilmartin}, {Korpela},
+  {Lin}, {Ling}, {Masuda}, {Matsubara}, {Miyake}, {Motomura}, {Muraki},
+  {Nagaya}, {Nakamura}, {Ohnishi}, {Okumura}, {Perrott}, {Rattenbury}, {Saito},
+  {Sako}, {Sullivan}, {Sweatman}, {Tristram}, {Udalski}, {Szyma{\'n}ski},
+  {Kubiak}, {Pietrzy{\'n}ski}, {Poleski}, {Soszy{\'n}ski}, {Wyrzykowski},
+  {Ulaczyk}, \& {Microlensing Observations in Astrophysics (MOA)
+  Collaboration}}]{sumi2011}
+{Sumi}, T., {Kamiya}, K., {Bennett}, D.~P., {et~al.} 2011, \nat, 473, 349
+
+\bibitem[{{Tomov} {et~al.}(2013){Tomov}, {Tomova}, \& {Bisikalo}}]{tomov2013}
+{Tomov}, N.~A., {Tomova}, M.~T., \& {Bisikalo}, D.~V. 2013, ArXiv e-prints
+
+\bibitem[{{Udalski}(2003)}]{udalski03}
+{Udalski}, A. 2003, \actaa, 53, 291
+
+\bibitem[{{Udalski} {et~al.}(1992){Udalski}, {Szymanski}, {Kaluzny}, {Kubiak},
+  \& {Mateo}}]{original_ogle}
+{Udalski}, A., {Szymanski}, M., {Kaluzny}, J., {Kubiak}, M., \& {Mateo}, M.
+  1992, \actaa, 42, 253
+
+\bibitem[{{Udalski} {et~al.}(1994){Udalski}, {Szymanski}, {Stanek}, {Kaluzny},
+  {Kubiak}, {Mateo}, {Krzeminski}, {Paczynski}, \&
+  {Venkat}}]{ogle_optical_depth}
+{Udalski}, A., {Szymanski}, M., {Stanek}, K.~Z., {et~al.} 1994, \actaa, 44, 165
+
+\bibitem[{{von Neumann} {et~al.}(1941){von Neumann}, {Kent}, {Bellinson}, \&
+  {Hart}}]{von_neumann1941}
+{von Neumann}, J., {Kent}, R.~H., {Bellinson}, H.~R., \& {Hart}, B.~I. 1941,
+  The Annals of Mathematical Statistics, 12, 153
+
+\bibitem[{{Wozniak}(2000)}]{con_idx}
+{Wozniak}, P.~R. 2000, \actaa, 50, 421
+
+\bibitem[{{Wright} {et~al.}(2010){Wright}, {Eisenhardt}, {Mainzer}, {Ressler},
+  {Cutri}, {Jarrett}, {Kirkpatrick}, {Padgett}, {McMillan}, {Skrutskie},
+  {Stanford}, {Cohen}, {Walker}, {Mather}, {Leisawitz}, {Gautier}, {McLean},
+  {Benford}, {Lonsdale}, {Blain}, {Mendez}, {Irace}, {Duval}, {Liu}, {Royer},
+  {Heinrichsen}, {Howard}, {Shannon}, {Kendall}, {Walsh}, {Larsen}, {Cardon},
+  {Schick}, {Schwalm}, {Abid}, {Fabinsky}, {Naes}, \& {Tsai}}]{WISE}
+{Wright}, E.~L., {Eisenhardt}, P.~R.~M., {Mainzer}, A.~K., {et~al.} 2010, \aj,
+  140, 1868
+
+\bibitem[{{Wyrzykowski} {et~al.}(2009){Wyrzykowski}, {Koz{\l}owski}, {Skowron},
+  {Belokurov}, {Smith}, {Udalski}, {Szyma{\'n}ski}, {Kubiak},
+  {Pietrzy{\'n}ski}, {Soszy{\'n}ski}, {Szewczyk}, \&
+  {{\.Z}ebru{\'n}}}]{wyrzykowski2009}
+{Wyrzykowski}, {\L}., {Koz{\l}owski}, S., {Skowron}, J., {et~al.} 2009, \mnras,
+  397, 1228
+
+\bibitem[{Yan {et~al.}(2013)Yan, Donoso, Tsai, Stern, Assef, Eisenhardt, Blain,
+  Cutri, Jarrett, Stanford, Wright, Bridge, \& Riechers}]{yan2013}
+Yan, L., Donoso, E., Tsai, C.-W., {et~al.} 2013, \aj, 145, 55
+
+\bibitem[{{York} {et~al.}(2000)}]{york00}
+{York}, D.~G., {et~al.} 2000, \aj, 120, 1579
+
+\end{thebibliography}
 

### 2013-11-07 text/paper/ms.tex
paper ready to resubmit

@@
 Each field has a unique sampling pattern determined by which of the PTF sub-surveys it belongs to, what time of the year it is visible, and how high a priority it is given by the scheduler.
-The imaging cadences range from 1-5 days, but other cadences are possible: for example, there was a higher-cadence campaign of a field in Orion to find transiting planets around young stars \citep{nick2009}.
+The imaging cadences range from 1-5 days, but other cadences are possible: for example, there was a higher-cadence campaign of a field in Orion to find transiting planets around young stars \citep{vaneyken2011}.
 The result is that PTF light curves often contain gaps and regions of high-cadence observations and/or of low-cadence observations, so that the archive is a massive dataset of irregularly-sampled, time-domain photometry.
 Figure~\ref{fig:sampling} shows six randomly selected light curves and illustrates the varying cadences and coverage that different fields may have over the same one-year period.
-While the PTF database is not yet public, we have made $10^4$ randomly selected light curves available at \url{http:// TODO }.
+While the PTF database is not yet public, we have made $\sim$$10^4$ randomly selected light curves available at \url{http://adrian.pw/PTF/lightcurves.tar.gz}.
 These light curves meet the quality criteria described in Section~\ref{sec:detection_eff} and can be used to test the statistical methods described here (or others!).
@@
 We add one more index, $\Delta \chi^2$, the difference in $\chi^2$ between fitting a Gaussian model and fitting a linear model to a light curve.
+This is a standard statistical test used by microlensing surveys, and allows us to compare the relative performance of the (slightly modified) \cite{shin2009} indices and of this approach.
 
@@
 Our tests with $\Delta \chi^2$ below compare the distribution of values over light curves on the same chip.
-As a result, the number of data points and number of model parameters are constant, and we therefore do not include an Akaike or Bayesian information criterion term.
+As a result, the number of data points and number of model parameters are constant, and we therefore do not include an Akaike or Bayesian information criterion (AIC or BIC) term.
 For the Gaussian fit, the optimizer is initialized with a $\sigma$ of 10 days, centered on the brightest data point.
-
-This is a standard statistical test used by microlensing surveys, and allows us to compare the relative performance of the (slightly modified) \cite{shin2009} indices and of this approach.
 
@@
 Rigorously classifying and cataloging these sources was outside of the scope of this paper, but many  presented interesting forms of variability.
-Though it is not necessarily scalable for future surveys to look at all light curves selected by using our algorithm, these $\sim$2000 light curves, once carefully classified, could serve as a training set for a more sophisticated machine-learning approach \cite[e.g.,][]{bloomPasp,brink2013} to classify the light curves that survive our selection criteria.
+Though it is not necessarily scalable for future surveys to look at all light curves selected by using our algorithm, these $\sim$2000 light curves, once carefully classified, could serve as a training set for a more sophisticated machine-learning approach \cite[e.g.,][]{bloomPasp, brink2013} to classify the light curves that survive our selection criteria.
 Approaches that combine statistical cuts such as those described in this paper and some type of machine learning will be essential to making the most of the future time-domain data such as that collected by the Large Synoptic Survey Telescope.
@@
 %\bigskip
-\setlength{\baselineskip}{0.6\baselineskip}
+%\setlength{\baselineskip}{0.6\baselineskip}
 %\bibliography{references_3}
-\setlength{\baselineskip}{1.667\baselineskip}
+%\setlength{\baselineskip}{1.667\baselineskip}
 
-\begin{thebibliography}{56}
+\begin{thebibliography}{57}
 \expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
@@
 
+\bibitem[{{van Eyken} {et~al.}(2011){van Eyken}, {Ciardi}, {Rebull},
+  {Stauffer}, {Akeson}, {Beichman}, {Boden}, {von Braun}, {Gelino}, {Hoard},
+  {Howell}, {Kane}, {Plavchan}, {Ram{\'{\i}}rez}, {Bloom}, {Cenko}, {Kasliwal},
+  {Kulkarni}, {Law}, {Nugent}, {Ofek}, {Poznanski}, {Quimby}, {Grillmair},
+  {Laher}, {Levitan}, {Mattingly}, \& {Surace}}]{vaneyken2011}
+{van Eyken}, J.~C., {Ciardi}, D.~R., {Rebull}, L.~M., {et~al.} 2011, \aj, 142,
+  60
+
 \bibitem[{{von Neumann} {et~al.}(1941){von Neumann}, {Kent}, {Bellinson}, \&

### 2013-11-14 text/paper/ms.tex
shorten abstract, update link

@@
 \begin{abstract}
-Many current photometric, time-domain surveys are driven by specific goals, such as searches for supernovae or transiting exoplanets, which set the cadence with which individual fields are re-imaged.
-In the case of the Palomar Transient Factory (PTF), several such sub-surveys are being conducted in parallel, leading to extremely non-uniform sampling over the survey's nearly $20,000~\mathrm{deg}^2$ footprint.
-While the median $7.26~\mathrm{deg}^2$ PTF field has been imaged $\sim$40 times in \textit{R}-band, $\sim$$2300~\mathrm{deg}^2$ have been observed more than 100 times.
-We use the existing PTF data to study the trade-off between searching for microlensing events in a survey whose footprint is much larger than that of typical microlensing searches, but that has far-from-optimal time sampling.
-To examine the probability that microlensing events can be recovered in these data, we test statistics previously used on uniformly sampled data to identify variables and transients.
-We find that one such statistic, the von Neumann ratio (also known as the Durbin-Watson statistic), performs best for identifying simulated microlensing events in our data.
-We develop a selection method using this statistic and apply it to data from all PTF fields with $>$10 $R$-band observations --- a total of $1.1\times10^9$ light curves --- to uncover three candidate sparse-field microlensing events.
-These candidates cannot be confirmed, as we lack simultaneous, multi-color photometry of the events.
-However, their number is consistent with predictions for the microlensing event rate in the PTF footprint over the survey's three years of operations, as estimated from near-field microlensing models.
-This work can help constrain all-sky event rate predictions and tests microlensing signal recovery in large data sets, both of which will be useful to future wide-field, time-domain surveys, such as that planned with the Large Synoptic Survey Telescope.
+Many photometric time-domain surveys are driven by specific goals, such as searches for supernovae or transiting exoplanets, which set the cadence with which fields are re-imaged.
+In the case of the Palomar Transient Factory (PTF), several sub-surveys are conducted in parallel, leading to non-uniform sampling over its $\sim$$20,000~\mathrm{deg}^2$ footprint.
+While the median $7.26~\mathrm{deg}^2$ PTF field has been imaged $\sim$40 times in \textit{R}-band, $\sim$$2300~\mathrm{deg}^2$ have been observed $>$100 times.
+We use PTF data to study the trade-off between searching for microlensing events in a survey whose footprint is much larger than that of typical microlensing searches, but with far-from-optimal time sampling.
+To examine the probability that microlensing events can be recovered in these data, we test statistics used on uniformly sampled data to identify variables and transients.
+We find that the von Neumann ratio performs best for identifying simulated microlensing events in our data.
+We develop a selection method using this statistic and apply it to data from fields with $>$10 $R$-band observations, $1.1\times10^9$ light curves, uncovering three candidate microlensing events.
+We lack simultaneous, multi-color photometry to confirm these as microlensing events.
+However, their number is consistent with predictions for the event rate in the PTF footprint over the survey's three years of operations, as estimated from near-field microlensing models.
+This work can help constrain all-sky event rate predictions and tests microlensing signal recovery in large data sets, which will be useful to future time-domain surveys, such as that planned with the Large Synoptic Survey Telescope.
 \end{abstract}
@@
 Figure~\ref{fig:sampling} shows six randomly selected light curves and illustrates the varying cadences and coverage that different fields may have over the same one-year period.
-While the PTF database is not yet public, we have made $\sim$$10^4$ randomly selected light curves available at \url{http://adrian.pw/PTF/lightcurves.tar.gz}.
+While the PTF database is not yet public, we have made $\sim$$10^4$ randomly selected light curves available at \url{http://adrian.pw/ptf}.
 These light curves meet the quality criteria described in Section~\ref{sec:detection_eff} and can be used to test the statistical methods described here (or others!).

### 2013-11-14 text/paper/ms.tex
added astropy reference

@@
 
-This research made use of Astropy (\url{http://www.astropy.org}), a community-developed core Python package for astronomy.
+This research made use of Astropy, a community-developed core Python package for astronomy \citep{astropy2013}.
 
@@
 
-\begin{thebibliography}{57}
+\begin{thebibliography}{58}
 \expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
@@
 {Alcock}, C., {Allsman}, R.~A., {Alves}, D., {et~al.} 2001, \apjs, 136, 439
+
+\bibitem[{{Astropy Collaboration} {et~al.}(2013){Astropy Collaboration},
+  {Robitaille}, {Tollerud}, {Greenfield}, {Droettboom}, {Bray}, {Aldcroft},
+  {Davis}, {Ginsburg}, {Price-Whelan}, {Kerzendorf}, {Conley}, {Crighton},
+  {Barbary}, {Muna}, {Ferguson}, {Grollier}, {Parikh}, {Nair}, {Unther},
+  {Deil}, {Woillez}, {Conseil}, {Kramer}, {Turner}, {Singer}, {Fox}, {Weaver},
+  {Zabalza}, {Edwards}, {Azalee Bostroem}, {Burke}, {Casey}, {Crawford},
+  {Dencheva}, {Ely}, {Jenness}, {Labrie}, {Lim}, {Pierfederici}, {Pontzen},
+  {Ptak}, {Refsdal}, {Servillat}, \& {Streicher}}]{astropy2013}
+{Astropy Collaboration}, {Robitaille}, T.~P., {Tollerud}, E.~J., {et~al.} 2013,
+  \aap, 558, A33
 
